### 2022-08-18 appendix.Rmd
add manuscript

---
output:
  pdf_document: default
  html_document: default
---
\renewcommand{\appendixname}{\textbf{Supplementary Materials}}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

#

## Supplementary Results

### Construct validity for measures of perceived descriptive and injunctive norms

To evaluate the construct validity of our measures of perceived descriptive and injunctive norms, at Time 7 we asked participants to rate the extent to which each perceived norm item provided descriptive and injunctive information. For each item, participants were asked whether the item provided information about what people are doing, and whether the item provided information about what people should be doing. Participants responded on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly. For a full list of questions, see Supplementary Table \@ref(tab:itemTable).

Results showed that participants did differentiate the perceived norm items as expected. Participants rated the perceived descriptive norm items as providing more descriptive information than the perceived injunctive norm items, _t_(`r printnum(constructValDes$parameter, digits = 0)`) = `r printnum(constructValDes$statistic)`, _p_ `r printp(constructValDes$p.value)` (mean descriptive items = `r printnum(mean((d$DescriptiveLearning3.7 + d$DescriptiveLearning4.7) / 2, na.rm = T))`; mean injunctive items = `r printnum(mean((d$DescriptiveLearning1.7 + d$DescriptiveLearning2.7) / 2, na.rm = T))`). By contrast, participants rated the perceived injunctive norm items as providing more injunctive information than the perceived descriptive norm items, _t_(`r printnum(constructValInj$parameter, digits = 0)`) = `r printnum(constructValInj$statistic)`, _p_ `r printp(constructValInj$p.value)` (mean descriptive items = `r printnum(mean((d$InjunctiveLearning3.7 + d$InjunctiveLearning4.7) / 2, na.rm = T))`; mean injunctive items = `r printnum(mean((d$InjunctiveLearning1.7 + d$InjunctiveLearning2.7) / 2, na.rm = T))`).

\newpage

## Supplementary Figures

(ref:plotUSMapCaption) _Map of the United States with participant zip code locations._

```{r plotUSMap, echo=F, fig.width=5, fig.height=3, fig.cap="(ref:plotUSMapCaption)"}
plot1
```

\newpage

(ref:plotAttritionCaption) _Attrition across the course of the study._

```{r plotAttrition, echo=F, fig.width=5, fig.height=3, fig.cap="(ref:plotAttritionCaption)"}
plot2
```

\newpage

(ref:plotDAGCaption) _Directed acyclic graph reflecting causal assumptions._ In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms cause mask wearing behavior (Mask). County-level COVID-19 prevalence (CovidPrev) and political leaning (CountyPol) are also posited to be common causes of perceptions of social norms and mask wearing behavior. Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived descriptive norms on mask wearing. Similarly, it is necessary to control for perceptions of descriptive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived injunctive norms on mask wearing.

```{r plotDAG, echo=F, fig.width=5.6, fig.height=4.5, fig.cap="(ref:plotDAGCaption)"}
plot4
```

\newpage

(ref:plotCorBehNormCaption) _Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables._ Lines are fixed effect regression lines from multilevel models, shaded areas are 95% confidence intervals.

```{r plotCorBehNorm, echo=F, fig.width=6.5, fig.height=3.8, fig.cap="(ref:plotCorBehNormCaption)"}
plot5
```

\newpage

(ref:plotCDCSensCaption) _Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations._ These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms. Time is included as a continuous linear predictor, scaled between 0 and 1, with two forced change points (dashed lines). The first dashed line (left) indicates when the CDC relaxed their mask wearing recommendations in March 2021, and the second dashed line (right) indicates when the CDC strengthened their mask wearing recommendations in July 2021. This results in the estimation of four fixed effect parameters: the initial intercept, the slope in the first window, the slope in the second window, and the slope in the third window. Bolded lines and shaded areas represent fixed effect regression lines from multilevel models and 95% confidence intervals, respectively.

```{r plotCDCSens, echo=F, fig.width=4, fig.height=6, fig.cap="(ref:plotCDCSensCaption)"}
plot6
```

\newpage

(ref:plotNormCompareCaption) _Predictions from multilevel model of the difference between the perceived strength of descriptive and injunctive social norms._ Points and error bars represent estimated marginal means and 95% confidence intervals, respectively.

```{r plotNormCompare, echo=F, fig.width=3.5, fig.height=3.5, fig.cap="(ref:plotNormCompareCaption)"}
plot7
```

\newpage

## Supplementary Tables

(ref:itemTableCaption) List of norm interpretation questions asked at Time 7. _These questions were preceded by the following text:_ "There may or may not be a difference between what people around you are doing and what they should be doing. You can learn about what people are doing and what they should be doing in different ways. For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both". _Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly._

```{r itemTable, echo=F, warning=F, message=F}
apa_table(
  itemTable,
  booktabs = TRUE,
  longtable = TRUE,
  align = c("m{2.5cm}", "m{2.5cm}", "m{9cm}"),
  caption = "(ref:itemTableCaption)"
)
```

\newpage

(ref:modelSummaryTable1Caption) Unstandardized fixed effect parameters from multilevel models: perceptions of social norm strength predicting self-reported mask wearing. _Standard errors are included in brackets._

```{r modelSummaryTable1, echo=F, warning=F, message=F, results="asis"}
export_summs(m1.1, m1.2, stars = NULL,
             coefs = c("Intercept" = "(Intercept)",
                       "Descriptive norms" = "DesNorms",
                       "Injunctive norms" = "InjNorms")) %>% 
  set_caption("(ref:modelSummaryTable1Caption)") %>%
  set_label("tab:modelSummaryTable1")
```

\newpage

(ref:changePointsTableCaption) Unstandardized fixed effect parameters from multilevel models: trends over time with change points at CDC events.

```{r changePointsTable, echo=F, warning=F, message=F}
apa_table(
  changePointsTable,
  booktabs = TRUE,
  longtable = TRUE,
  align = c("l", "c", "c", "c"),
  caption = "(ref:changePointsTableCaption)"
)
```

\newpage

(ref:modelSummaryTable2Caption) Unstandardized fixed effect parameters from multilevel model: difference between perceived strength of descriptive and injunctive social norms. _Standard errors are included in brackets._

```{r modelSummaryTable2, echo=F, warning=F, message=F, results="asis"}
export_summs(m3.1, stars = NULL,
             coefs = c("Intercept" = "(Intercept)",
                       "Difference between norm types" = "varInjNorms")) %>% 
  set_caption("(ref:modelSummaryTable2Caption)") %>%
  set_label("tab:modelSummaryTable2")
```

\newpage

(ref:lavaanTableCaption) Standardised autoregressive and cross-lagged parameters from random-intercept cross-lagged panel model. _Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms. Variable name suffixes indicate time points. Arrows indicate the direction of prediction._

(ref:rightArrow) →

```{r lavaanTable, echo=F, warning=F, message=F}
apa_table(
  lavaanTable,
  booktabs = TRUE,
  longtable = TRUE,
  align = c("l", rep("c", 5)),
  caption = "(ref:lavaanTableCaption)"
)
```

\newpage

## Supplementary References

Pearl, J. (1995). Causal diagrams for empirical research. _Biometrika_, _82_(4), 669–688, https://doi.org/10.1093/biomet/82.4.669

### 2022-08-18 manuscript.Rmd
add manuscript

---
title             : "Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic"
shorttitle        : "Norms and mask wearing"

author: 
  - name          : "Samantha L. Heiman"
    affiliation   : "1"
    equal_contrib : yes
  - name          : "Scott Claessens"
    affiliation   : "2"
    equal_contrib : yes
  - name          : "Edward R. Hurt"
    affiliation   : "1"
  - name          : "Peter M. Todd"
    affiliation   : "1,3"
    corresponding : yes
    address       : "1101 E 10th St, Bloomington, IN 47405, United States"
    email         : "pmtodd@indiana.edu"

affiliation:
  - id            : "1"
    institution   : "Department of Psychological and Brain Sciences, Indiana University Bloomington, United States"
  - id            : "2"
    institution   : "School of Psychology, University of Auckland, New Zealand"
  - id            : "3"
    institution   : "Cognitive Science Program, Indiana University Bloomington, United States"

abstract: |
 Social norms allow humans to coordinate and cooperate in the face of existential threats. In particular, injunctive social norms prescribe what people _ought_ to do, whereas descriptive social norms inform what people _actually_ do. While previous experimental work has revealed people’s sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their influences on cooperative behavior over time. To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic. Leveraging a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; _n_ = `r nrow(tar_read(d))`), we tracked people’s reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded. Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior are tightly coupled and both change quickly in response to recommendations from public health authorities. In addition, a random-intercept cross-lagged panel model revealed that descriptive norms, but not injunctive norms, caused future increases in mask wearing behavior. These findings underline the relative importance of descriptive social norms in shaping cooperative behavior during uncertain times.

note              : "This working paper has not yet been peer-reviewed. This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation."
keywords          : "descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
wordcount         : "4826 words"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \usepackage{array}
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[normalem]{ulem}
  - \usepackage{colortbl}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage{adjustbox}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \raggedbottom
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1,small}}
  - \captionsetup[figure]{font={stretch=1,small}}
appendix          : "appendix.Rmd"
csl               : apa.csl
urlcolor          : blue
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
tar_load_everything()
# to ensure that the Rmd runs last, load appendix files again
tar_load(c(d, plot1, plot2, plot4, plot5, plot6, plot7,
           m1.1, m1.2, m3.1, itemTable, lavaanTable,
           changePointsTable, constructValDes, constructValInj))
```

Social norms are a key aspect of human sociality. Broadly defined as behavioral guidelines enforced by groups of people without the law [@VanKleef2019], social norms have attracted the attention of psychologists for decades. Since Asch's [-@Asch1956] early studies of normative conformity, evidence has accrued that humans are particularly attuned to social norms [@Tomasello2014]. From a young age, children begin to adhere to and enforce group-wide social norms [@Jensen2014; @Rakoczy2013; @Schmidt2012] and both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms [@Vaish2011; @Schaumberg2022]. This uniquely human sensitivity to social norms allows groups of unrelated people to cooperate and coordinate in the face of existential threats, such as resource scarcity, natural disasters, and infectious diseases [@Gelfand2011; @Henrich2015].

Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms [@Cialdini1991; @Deutsch1955]. Injunctive social norms are societal, indicating what others in the group tend to approve or disapprove of, and often involve social sanctions if violated. By contrast, descriptive norms are situational, simply describing what most people are doing in a given situation. Though these two kinds of social norms tend to align, they can also be in conflict with one another. For example, there may be an injunctive norm that cleaning up litter at a picnic site is the right thing to do: one _ought_ to behave this way. However, if an individual observes that most people are leaving their litter behind at the site, the descriptive norm is to not clean up. It is thus possible for injunctive and descriptive norms to have independent effects on behavior [@Schultz2007].

Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], social norms have remained an elusive concept in the behavioral sciences and several open questions remain [@Fehr2018a; @VanKleef2019]. In the current work, we focus on two such questions. First, how do injunctive and descriptive norms emerge naturally over time within a population? Second, how do evolving injunctive and descriptive norms influence behavior?

With regards to norm emergence, research in cultural evolution and behavioral economics have begun to illuminate how social norms emerge over time. In the longer term, cultural evolutionary models show that injunctive social norms can be vertically transmitted down through generations via imitation or teaching, or horizontally diffused from neighboring human populations [@Boyd1985; @Henrich2015]. For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership [@Kushnick2014] and post-marital residence [@Jordan2009]. However, much less is known about how social norms arise endogenously within populations in the shorter term. Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019] and require peer enforcement to become stable [@Fehr2018b]. But it remains unclear to what extent these findings generalize beyond the laboratory to real human populations.

With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the behavioral impact of social norms. For example, field experiments have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling [@Nigbur2010], paying taxes [@Larkin2018], and sustainably reusing towels in hotels [@Goldstein2008]. Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals’ attention on injunctive norms [@Schultz2007]. However, previous studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population. An alternative way to identify causality, whilst retaining ecological validity, is to follow perceptions of social norms and cooperative behavior over time amidst a real, unfolding social dilemma. This is particularly important because social norms are not static: they change dynamically over time through processes of deliberation and social interaction [@Smith2015].

To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, we studied mask wearing behavior during the COVID-19 pandemic. Before the pandemic, mask wearing was not a common behavior in the United States. In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a measure that people should take to minimize the spread of COVID-19. But mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., "flattening the curve" to protect at-risk individuals). Thus, the evolution of mask wearing in the United States throughout the COVID-19 pandemic allows us to study, on a short timescale within a single population, the emergence of novel injunctive and descriptive social norms and the causal effects of these norms on cooperative behavior.

Recent research has found positive relationships between perceptions of social norms and protective COVID-19 behaviors. In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure [@Macy2021], and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions [@Bokemper2021]. In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing [@Rudert2021]. These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between between-person and within-person change over time [@Hamaker2015], nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.

Here, we use a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; _n_ = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic. We aimed to answer two main research questions. First, how do descriptive and injunctive mask wearing norms emerge and evolve over time in the United States population? Second, how do descriptive and injunctive mask wearing norms affect mask wearing behavior over time?

# Methods

## Ethical approval

This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.

## Participants and sampling

Using the platform Prolific (https://www.prolific.co/), we distributed surveys to a representative sample of adults from the United States (_n_ = `r nrow(d)`, _M_~age~ = `r round(mean(d$Age.1, na.rm = TRUE))` years, `r round(mean(d$Ethnicity.1 == "White or Caucasian", na.rm = TRUE)*100)`% White, `r round(mean(d$Sex.1 == 2, na.rm = TRUE)*100)`% Women; see Supplementary Figure \@ref(fig:plotUSMap) for geographic distribution). From September 2020 to February 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors. This resulted in 15 unique time points of data collection throughout the pandemic. The first 12 time points were distributed monthly, while the last three time points were distributed every two months. `r sum(!is.na(d$StartDate.2))` of the initial `r nrow(d)` participants returned to complete the survey at Time 2, while `r sum(!is.na(d$StartDate.15))` participants continued through to Time 15 (see Supplementary Figure \@ref(fig:plotAttrition) for attrition rates across all time points).

## Measures

### Self-reported mask wearing behavior

At every time point, participants were asked about the number of in-person interactions they had in the last week. Following this question, participants self-reported their mask wearing behavior by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).

### Perceived descriptive and injunctive social norms

In 8 of the 15 time points (Time 2, 3, 5, 9, 11, 13, 14, and 15), we asked questions about perceived descriptive and injunctive mask wearing norms.

Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings. We measured perceived descriptive social norms as the mean average of the following two items: "_What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?_" and "_What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?_" These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).

Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules. We measured perceived injunctive social norms as the mean average of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.

To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items. We asked participants whether each of the four items informed them about what people _are_ doing or what people _should_ be doing (i.e., giving descriptive or injunctive information). Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \@ref(tab:itemTable)).

### Additional control variables

To account for potential confounding in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \@ref(fig:plotDAG)). In this causal model, we included two additional variables that could plausibly act as common causes of perceived social norms and mask wearing behavior: county-level COVID-19 prevalence and county-level political leaning. If COVID-19 is prevalent in a participant’s county, then there might be more mask wearing overall, from people in participants’ local area, respected individuals, and from participants themselves, causing increases in both self-reported mask wearing behavior and perceived social norms. Similarly, left-leaning counties might have increased levels of self-reported mask wearing behavior and perceived social norms, as the political context affects the proportion of individuals following COVID-19 protective measures [@BaxterKing2022]. It is therefore important to statistically control for these confounding variables.

We retrieved data on county-level COVID-19 prevalence from a CDC public health surveillance dataset [@CDC2022a]. In the CDC dataset, COVID-19 prevalence was operationalized as the total new COVID-19 cases per 100,000 population in counties over the past 7 days. Since this dataset was updated regularly starting from February 2022, and so each county had multiple entries, we averaged and logged this variable within each county to get the log average weekly new COVID-19 cases per 100,000 population. We retrieved data on county-level political leaning from the MIT Election Lab [@MITElectionLab]. For each county, the dataset stated how many votes the Republican, Democratic, Green Party, and other candidates received in the 2020 Presidential election. We used this information to determine which candidate won the vote in that county.

Participants reported their zip code rather than their county, and in a minority of cases the reported zip codes straddled multiple counties. In these cases, we averaged COVID-19 prevalence rates over the different counties within the zip codes, and used the political candidate that won in the majority of overlapping counties.

## Statistical analysis

To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models. First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points. Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor. These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations. Third, to compare the perceived strength of descriptive and injunctive norms, we regressed perceived norm strength on the type of norm, including random intercepts and slopes for participants and time points. We estimated these models using the _lme4_ R package [@Bates2015] and dealt with missing data via listwise deletion.

To quantify the within-person relationships between perceived descriptive norms, perceived injunctive norms, and mask wearing behavior simultaneously over time, we used a three-variable random-intercept cross-lagged panel model to analyze our longitudinal data [@Hamaker2015]. This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels. Positive cross-lagged effects from this model indicate that being above average on one variable at time~t-1~ predicts being above average in another variable at time~t~. These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015; @Granger1969]. We estimated this model using the _lavaan_ R package [@Rosseel2012]. We restricted the analysis to the eight time points with available data for all three variables, and used full information maximum likelihood estimation to deal with any other missing data across time points. Based on our causal model (see Supplementary Figure \@ref(fig:plotDAG)), we also included county-level COVID-19 prevalence and county-level political leaning variables as time-invariant controls. Due to varying time intervals between time points, we did not constrain any parameters to equality over time.

All analyses were conducted in R v4.1.1 [@RCoreTeam]. Visualizations were generated using the _cowplot_ [@Wilke2020] and _ggplot2_ [@Wickham2016] packages. The manuscript was reproducibly generated using the _targets_ [@Landau2021] and _papaja_ [@Aust2022] packages.

# Results

To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration. Figure \@ref(fig:plotTimeline) plots self-reported mask wearing behavior and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers. These events were obtained from the CDC Museum’s COVID-19 Timeline [@CDC2022b] and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).

(ref:plotTimelineCaption) _Timeline of self-reported mask wearing behavior and perceived social norms in the United States throughout the COVID-19 pandemic._ (a) Points and line ranges indicate mean averages and standard errors for the self-reported mask wearing item. This item was measured across all fifteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions. (b) Points and line ranges indicate mean averages and standard errors for perceived descriptive mask wearing norms (blue) and perceived injunctive mask wearing norms (green). These items were measured across eight time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms. (c) Smoothed data for new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; https://ourworldindata.org/). Across all panels, gray dashed lines represent significant pandemic-related events in the United States, such as vaccine approval from the Food and Drug Administration (FDA) and public health recommendations from the Centers for Disease Control and Prevention (CDC).

```{r plotTimeline, echo=F, fig.width=5, fig.height=6, fig.cap="(ref:plotTimelineCaption)"}
plot3
```

Three main observations can be made about the emergence and stability of social norms from these visualizations. First, social norms and behavior were tightly coupled over time. Although social norms are measured on fewer occasions than mask wearing behavior, we can see that as mask wearing behavior decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms. Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, though to lower levels than before. In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (_b_ = `r printnum(fixef(m1.1)["DesNorms"])`, 95% confidence interval [`r printnum(conf1.1["DesNorms","2.5 %"])` `r printnum(conf1.1["DesNorms","97.5 %"])`]) and between mask wearing behavior and perceived injunctive mask wearing norms (_b_ = `r printnum(fixef(m1.2)["InjNorms"])`, 95% CI [`r printnum(conf1.2["InjNorms","2.5 %"])` `r printnum(conf1.2["InjNorms","97.5 %"])`]) across individuals and time points (Supplementary Figure \@ref(fig:plotCorBehNorm); Supplementary Table \@ref(tab:modelSummaryTable1)).

Second, fluctuations in mask wearing behavior and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States. We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020. Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021. Impressively, shifts in mask wearing behavior and norms happened on the timescale of weeks. These trends were confirmed by a series of multilevel regression models with change points aligning with changes in CDC mask wearing recommendations (Supplementary Figure \@ref(fig:plotCDCSens); Supplementary Table \@ref(tab:changePointsTable)).

Third, across all eight time points with available data, injunctive mask wearing norms were consistently rated as stronger than descriptive mask wearing norms. Multilevel regression modeling revealed that, on average across individuals and time points, injunctive norms were rated as 0.6 points stronger than descriptive norms on a 7-point Likert scale (_b_ = `r printnum(fixef(m3.1)["varInjNorms"])`, 95% CI [`r printnum(conf3.1["varInjNorms","2.5 %"])` `r printnum(conf3.1["varInjNorms","97.5 %"])`]; Supplementary Figure \@ref(fig:plotNormCompare); Supplementary Table \@ref(tab:modelSummaryTable2)). In other words, participants in our study perceived the general encouragement of mask wearing and mask wearing behavior among trusted people to be higher than perceived mask wearing among people in their local area.

Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing behavior and perceived social norms over time. To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted an eight-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data. This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our three main variables: self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms (see Supplementary Table \@ref(tab:lavaanTable) for estimated model parameters). According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).

Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing behavior and perceived social norms. On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(standardizedSolution(riclpm)[187,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[187,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[187,"ci.upper"])`], _p_ `r printp(parameterEstimates(riclpm)[187,"pvalue"])`) and stronger injunctive mask wearing norms (_r_ = `r printnum(standardizedSolution(riclpm)[188,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[188,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[188,"ci.upper"])`], _p_ `r printp(parameterEstimates(riclpm)[188,"pvalue"])`). Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (_r_ = `r printnum(standardizedSolution(riclpm)[189,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[189,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[189,"ci.upper"])`], _p_ `r printp(parameterEstimates(riclpm)[189,"pvalue"])`).

Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for all three variables across the study duration. In random intercept cross-lagged panel models, autoregressive effects represent "persistence" or "inertia" in within-person fluctuations from stable trait levels. In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the "stability" of the measure over time, which is captured by the random intercepts in our model). By contrast, and most relevant for the current study, cross-lagged effects represent the effect of a within-person fluctuation in one measure on future within-person fluctuations in other measures. In other words, a positive cross-lagged effect indicates that being higher than average on one measure predicts being higher than average on _another_ measure in the following time point. Cross-lagged effects are thus used to infer within-person causal influences over time.

(ref:plotRICLPMCaption) _Results of eight-wave unconstrained random-intercept cross-lagged panel model._ Arrows represent within-person autoregressive and cross-lagged effects from the model, controlling for time invariant predictors and partitioning out stable between-person individual differences. Arrow thickness is scaled according to standardized effect size. Bolded arrows indicate significantly positive parameters, _p_ < 0.05. Gray arrows indicate non-significant parameters.

```{r plotRICLPM, echo=F, fig.width=7, fig.height=3, fig.cap="(ref:plotRICLPMCaption)"}
plot8
```

Earlier in the pandemic, we see less evidence of autoregressive and cross-lagged effects. Within-person fluctuations in descriptive and injunctive norms persist across only a few time points, and there is only one cross-lagged effect of mask wearing on future descriptive norms. However, once COVID-19 case numbers start rising again in June 2021, we begin to see more evidence of autoregressive and cross-lagged effects. From this point onwards, all three variables show evidence of persistence, with within-person fluctuations from trait levels predicting later within-person fluctuations in the same measures. But more importantly, several positive and significant cross-lagged effects emerge. On three occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior. According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (Time 9, $\beta$ = `r printnum(standardizedSolution(riclpm)[108,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[108,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[108,"ci.upper"])`], _p_ = `r printp(parameterEstimates(riclpm)[108,"pvalue"])`; Time 11, $\beta$ = `r printnum(standardizedSolution(riclpm)[111,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[111,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[111,"ci.upper"])`], _p_ = `r printp(parameterEstimates(riclpm)[111,"pvalue"])`; Time 13, $\beta$ = `r printnum(standardizedSolution(riclpm)[114,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[114,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[114,"ci.upper"])`], _p_ = `r printp(parameterEstimates(riclpm)[114,"pvalue"])`). We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms. Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.

However, the model reveals that, after controlling for perceived descriptive norms, within-person changes in perceived injunctive norms were unrelated to within-person changes in mask wearing behavior across the entire pandemic. All cross-lagged effects between the two variables, both from perceived injunctive norms to mask wearing behavior and vice versa, are non-significant. Any causal effect that perceived injunctive norms might have had on future mask wearing behavior appears to be fully mediated by perceived descriptive norms. For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing behavior. But aside from these indirect effects, perceived injunctive norms and mask wearing behavior were not directly causally related over time within individuals.

# Discussion

Using longitudinal data from the United States across the entire COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma. The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior are tightly coupled, and both change quickly in response to recommendations from public health authorities. The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior from June 2021 onwards. While injunctive norms were perceived as stronger than descriptive norms, we found that  injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.

Our finding that social norms and mask wearing behavior are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction [@Titlestad2019]. Moreover, the role of authorities, like the CDC, in shaping social norms and behavior supports the idea that institutions are part of the process by which culture and one’s own behaviors are mutually constructed [@Markus2010]. The CDC explicitly releases guidelines for the benefit of public health, one of many situations in which the function of authorities is to coordinate behavior on large scales. However, even when institutions do not intend to coordinate behavior, they still may introduce influential social norms into the population (e.g., through business practices or products produced).

We also found that injunctive norms were, on average, perceived as stronger than descriptive norms. This aligns with previous research that measured perceived injunctive and descriptive norms in the contexts of drinking [@Park2009] and marijuana use [@Montes2021]. What people ought to do is often perceived as stronger than what people _actually_ do. Since we operationalized injunctive norms in part by whether trusted individuals in the media were wearing masks, it is possible that our sample is accurately reporting that people in general were less adherent to mask wearing policies than government officials, newscasters, and other respected individuals. It may also have been easier for our participants to notice instances of non-compliance with mask wearing guidelines among the general population, simply because they have more opportunities to observe people’s behavior in public settings.

Despite perceiving stronger injunctive norms, it was _descriptive_ norms, not injunctive norms, that independently predicted future increases in mask wearing. In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects [@Rudert2021]. Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks [@Bokemper2021]. Why this specific effect of descriptive norms on mask wearing? Descriptive norms are particularly useful for fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic [@Gelfand2015]. During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms [@Hogg2008; @Wellen1998]. Supporting this situational-uncertainty explanation, our model revealed that descriptive norms only began to predict future mask wearing after the CDC flipped their mask wearing recommendation in March 2021. With changing recommendations from authorities increasing levels of uncertainty, people instead began to look to their neighbors and adapted their mask wearing behavior accordingly.

Our finding that injunctive norms do not predict future mask wearing behavior is at odds with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults [@Macy2021]. One possible explanation for these conflicting findings is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior [@Cialdini1991]. By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people’s behaviors, increasing the relative salience of injunctive norms. To test this idea, future research should expand our longitudinal cross-lagged approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).

We are limited in generalizing these findings due to the constraints of our sample and the variables considered. While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \@ref(fig:plotAttrition)). This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, different ethnicities, or those with different political ideologies. Since injunctive norm strength varies based on which group is seen as the source of the norm [@Neighbors2008], it would be interesting to learn whether different groups have different patterns of norm emergence over time. In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors [@BaxterKing2022]. Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large [@Remedios2022]. Our results also might not generalize to all social norms, behaviors, and social dilemmas. Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more abstract and remote than the COVID-19 pandemic. For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.

Regardless, in the case of mask wearing in the United States over the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and responded to both recommendations from authorities and current levels of cooperative behavior. Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing. Importantly, this key finding slices both ways. Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that _low_ local levels of mask wearing _discouraged_ future personal mask use. This echoes recent reports of people in the United States not wanting to be "singled out" by being the only one wearing a mask in their community [@Natanson2022]. Our work thus underscores the importance of consistent, visible community adherence for encouraging personal protective behaviors in response to global pandemics like COVID-19.

\newpage

# Acknowledgements

This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.

# Author Contributions

SLH, ERH, and PMT conceptualized the study. SLH and SC oversaw the data curation, investigation, and methodology of the study, and wrote the first draft of the paper. SC conducted the formal analysis and created all visualisations. ERH and PMT provided funding and supervision for the study. All authors reviewed and edited the final draft of the paper.

# Conflicts of Interest

There are no conflicts of interest to declare.

# Research Transparency and Reproducibility

All data and code to reproduce the statistical analyses in this manuscript can be found on GitHub: https://github.com/ScottClaessens/covidMaskWearing

\newpage

# References

\begingroup
\setlength{\parindent}{-1in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

\newpage

# (APPENDIX) Appendix {-}

```{r child = "appendix.Rmd"}
```

### 2022-08-18 manuscript.tex
add manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic}
\author{Samantha L. Heiman\textsuperscript{$\dagger{}$,1}, Scott Claessens\textsuperscript{$\dagger{}$,2}, Edward R. Hurt\textsuperscript{1}, \& Peter M. Todd\textsuperscript{1,3}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic},
  pdfauthor={Samantha L. Heiman,1, Scott Claessens,2, Edward R. Hurt1, \& Peter M. Todd1,3},
  pdflang={en-EN},
  pdfkeywords={descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 4826 words}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{array}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[normalem]{ulem}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{calc}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{wrapfig}
\usepackage{adjustbox}
\usepackage{hyperref}
\usepackage{setspace}
\raggedbottom
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\AtBeginEnvironment{tablenotes}{\doublespacing}
\captionsetup[table]{font={stretch=1,small}}
\captionsetup[figure]{font={stretch=1,small}}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\shorttitle{Norms and mask wearing}

\authornote{

\textsuperscript{$\dagger{}$} Samantha L. Heiman and Scott Claessens contributed equally to this work.

Correspondence concerning this article should be addressed to Peter M. Todd, 1101 E 10th St, Bloomington, IN 47405, United States. E-mail: \href{mailto:pmtodd@indiana.edu}{\nolinkurl{pmtodd@indiana.edu}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Psychological and Brain Sciences, Indiana University Bloomington, United States\\\textsuperscript{2} School of Psychology, University of Auckland, New Zealand\\\textsuperscript{3} Cognitive Science Program, Indiana University Bloomington, United States}

\note{This working paper has not yet been peer-reviewed. This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.}

\abstract{%
Social norms allow humans to coordinate and cooperate in the face of existential threats. In particular, injunctive social norms prescribe what people \emph{ought} to do, whereas descriptive social norms inform what people \emph{actually} do. While previous experimental work has revealed people's sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their influences on cooperative behavior over time. To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic. Leveraging a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; \emph{n} = 915), we tracked people's reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded. Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior are tightly coupled and both change quickly in response to recommendations from public health authorities. In addition, a random-intercept cross-lagged panel model revealed that descriptive norms, but not injunctive norms, caused future increases in mask wearing behavior. These findings underline the relative importance of descriptive social norms in shaping cooperative behavior during uncertain times.
}

\begin{document}
\maketitle

Social norms are a key aspect of human sociality. Broadly defined as behavioral guidelines enforced by groups of people without the law (van Kleef et al., 2019), social norms have attracted the attention of psychologists for decades. Since Asch's (1956) early studies of normative conformity, evidence has accrued that humans are particularly attuned to social norms (Tomasello, 2014). From a young age, children begin to adhere to and enforce group-wide social norms (Jensen et al., 2014; Rakoczy \& Schmidt, 2013; Schmidt \& Tomasello, 2012) and both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms (Schaumberg \& Skowronek, 2022; Vaish et al., 2011). This uniquely human sensitivity to social norms allows groups of unrelated people to cooperate and coordinate in the face of existential threats, such as resource scarcity, natural disasters, and infectious diseases (Gelfand et al., 2011; Henrich, 2015).

Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms (Cialdini et al., 1991; Deutsch \& Gerard, 1955). Injunctive social norms are societal, indicating what others in the group tend to approve or disapprove of, and often involve social sanctions if violated. By contrast, descriptive norms are situational, simply describing what most people are doing in a given situation. Though these two kinds of social norms tend to align, they can also be in conflict with one another. For example, there may be an injunctive norm that cleaning up litter at a picnic site is the right thing to do: one \emph{ought} to behave this way. However, if an individual observes that most people are leaving their litter behind at the site, the descriptive norm is to not clean up. It is thus possible for injunctive and descriptive norms to have independent effects on behavior (Schultz et al., 2007).

Despite decades of research on the causes and consequences of injunctive and descriptive norms (Cialdini et al., 1991; Cialdini, 2007; Cialdini \& Jacobson, 2021; Schultz et al., 2007), social norms have remained an elusive concept in the behavioral sciences and several open questions remain (Fehr \& Schurtenberger, 2018a; van Kleef et al., 2019). In the current work, we focus on two such questions. First, how do injunctive and descriptive norms emerge naturally over time within a population? Second, how do evolving injunctive and descriptive norms influence behavior?

With regards to norm emergence, research in cultural evolution and behavioral economics have begun to illuminate how social norms emerge over time. In the longer term, cultural evolutionary models show that injunctive social norms can be vertically transmitted down through generations via imitation or teaching, or horizontally diffused from neighboring human populations (Boyd, 1985; Henrich, 2015). For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership (Kushnick et al., 2014) and post-marital residence (Jordan et al., 2009). However, much less is known about how social norms arise endogenously within populations in the shorter term. Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (Titlestad et al., 2019) and require peer enforcement to become stable (Fehr \& Schurtenberger, 2018b). But it remains unclear to what extent these findings generalize beyond the laboratory to real human populations.

With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the behavioral impact of social norms. For example, field experiments have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (Nigbur et al., 2010), paying taxes (Larkin et al., 2018), and sustainably reusing towels in hotels (Goldstein et al., 2008). Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals' attention on injunctive norms (Schultz et al., 2007). However, previous studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population. An alternative way to identify causality, whilst retaining ecological validity, is to follow perceptions of social norms and cooperative behavior over time amidst a real, unfolding social dilemma. This is particularly important because social norms are not static: they change dynamically over time through processes of deliberation and social interaction (Smith et al., 2015).

To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, we studied mask wearing behavior during the COVID-19 pandemic. Before the pandemic, mask wearing was not a common behavior in the United States. In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a measure that people should take to minimize the spread of COVID-19. But mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., ``flattening the curve'' to protect at-risk individuals). Thus, the evolution of mask wearing in the United States throughout the COVID-19 pandemic allows us to study, on a short timescale within a single population, the emergence of novel injunctive and descriptive social norms and the causal effects of these norms on cooperative behavior.

Recent research has found positive relationships between perceptions of social norms and protective COVID-19 behaviors. In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (Macy et al., 2021), and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions (Bokemper et al., 2021). In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (Rudert \& Janke, 2021). These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between between-person and within-person change over time (Hamaker et al., 2015), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.

Here, we use a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; \emph{n} = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic. We aimed to answer two main research questions. First, how do descriptive and injunctive mask wearing norms emerge and evolve over time in the United States population? Second, how do descriptive and injunctive mask wearing norms affect mask wearing behavior over time?

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{ethical-approval}{%
\subsection{Ethical approval}\label{ethical-approval}}

This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.

\hypertarget{participants-and-sampling}{%
\subsection{Participants and sampling}\label{participants-and-sampling}}

Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution). From September 2020 to February 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors. This resulted in 15 unique time points of data collection throughout the pandemic. The first 12 time points were distributed monthly, while the last three time points were distributed every two months. 634 of the initial 915 participants returned to complete the survey at Time 2, while 379 participants continued through to Time 15 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).

\hypertarget{measures}{%
\subsection{Measures}\label{measures}}

\hypertarget{self-reported-mask-wearing-behavior}{%
\subsubsection{Self-reported mask wearing behavior}\label{self-reported-mask-wearing-behavior}}

At every time point, participants were asked about the number of in-person interactions they had in the last week. Following this question, participants self-reported their mask wearing behavior by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).

\hypertarget{perceived-descriptive-and-injunctive-social-norms}{%
\subsubsection{Perceived descriptive and injunctive social norms}\label{perceived-descriptive-and-injunctive-social-norms}}

In 8 of the 15 time points (Time 2, 3, 5, 9, 11, 13, 14, and 15), we asked questions about perceived descriptive and injunctive mask wearing norms.

Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings. We measured perceived descriptive social norms as the mean average of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).

Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules. We measured perceived injunctive social norms as the mean average of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.

To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items. We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information). Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).

\hypertarget{additional-control-variables}{%
\subsubsection{Additional control variables}\label{additional-control-variables}}

To account for potential confounding in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}). In this causal model, we included two additional variables that could plausibly act as common causes of perceived social norms and mask wearing behavior: county-level COVID-19 prevalence and county-level political leaning. If COVID-19 is prevalent in a participant's county, then there might be more mask wearing overall, from people in participants' local area, respected individuals, and from participants themselves, causing increases in both self-reported mask wearing behavior and perceived social norms. Similarly, left-leaning counties might have increased levels of self-reported mask wearing behavior and perceived social norms, as the political context affects the proportion of individuals following COVID-19 protective measures (Baxter-King et al., 2022). It is therefore important to statistically control for these confounding variables.

We retrieved data on county-level COVID-19 prevalence from a CDC public health surveillance dataset (Centers for Disease Control and Prevention, 2022b). In the CDC dataset, COVID-19 prevalence was operationalized as the total new COVID-19 cases per 100,000 population in counties over the past 7 days. Since this dataset was updated regularly starting from February 2022, and so each county had multiple entries, we averaged and logged this variable within each county to get the log average weekly new COVID-19 cases per 100,000 population. We retrieved data on county-level political leaning from the MIT Election Lab (MIT Election Data and Science Lab, 2018). For each county, the dataset stated how many votes the Republican, Democratic, Green Party, and other candidates received in the 2020 Presidential election. We used this information to determine which candidate won the vote in that county.

Participants reported their zip code rather than their county, and in a minority of cases the reported zip codes straddled multiple counties. In these cases, we averaged COVID-19 prevalence rates over the different counties within the zip codes, and used the political candidate that won in the majority of overlapping counties.

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models. First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points. Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor. These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations. Third, to compare the perceived strength of descriptive and injunctive norms, we regressed perceived norm strength on the type of norm, including random intercepts and slopes for participants and time points. We estimated these models using the \emph{lme4} R package (Bates et al., 2015) and dealt with missing data via listwise deletion.

To quantify the within-person relationships between perceived descriptive norms, perceived injunctive norms, and mask wearing behavior simultaneously over time, we used a three-variable random-intercept cross-lagged panel model to analyze our longitudinal data (Hamaker et al., 2015). This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels. Positive cross-lagged effects from this model indicate that being above average on one variable at time\textsubscript{t-1} predicts being above average in another variable at time\textsubscript{t}. These models are considered the gold standard for identifying Granger causality in longitudinal datasets (Granger, 1969; Hamaker et al., 2015). We estimated this model using the \emph{lavaan} R package (Rosseel, 2012). We restricted the analysis to the eight time points with available data for all three variables, and used full information maximum likelihood estimation to deal with any other missing data across time points. Based on our causal model (see Supplementary Figure \ref{fig:plotDAG}), we also included county-level COVID-19 prevalence and county-level political leaning variables as time-invariant controls. Due to varying time intervals between time points, we did not constrain any parameters to equality over time.

All analyses were conducted in R v4.1.1 (R Core Team, 2022). Visualizations were generated using the \emph{cowplot} (Wilke, 2020) and \emph{ggplot2} (Wickham, 2016) packages. The manuscript was reproducibly generated using the \emph{targets} (Landau, 2021) and \emph{papaja} (Aust \& Barth, 2022) packages.

\hypertarget{results}{%
\section{Results}\label{results}}

To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration. Figure \ref{fig:plotTimeline} plots self-reported mask wearing behavior and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers. These events were obtained from the CDC Museum's COVID-19 Timeline (Centers for Disease Control and Prevention, 2022a) and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotTimeline-1.pdf}
\caption{\label{fig:plotTimeline}\emph{Timeline of self-reported mask wearing behavior and perceived social norms in the United States throughout the COVID-19 pandemic.} (a) Points and line ranges indicate mean averages and standard errors for the self-reported mask wearing item. This item was measured across all fifteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions. (b) Points and line ranges indicate mean averages and standard errors for perceived descriptive mask wearing norms (blue) and perceived injunctive mask wearing norms (green). These items were measured across eight time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms. (c) Smoothed data for new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; \url{https://ourworldindata.org/}). Across all panels, gray dashed lines represent significant pandemic-related events in the United States, such as vaccine approval from the Food and Drug Administration (FDA) and public health recommendations from the Centers for Disease Control and Prevention (CDC).}
\end{figure}

Three main observations can be made about the emergence and stability of social norms from these visualizations. First, social norms and behavior were tightly coupled over time. Although social norms are measured on fewer occasions than mask wearing behavior, we can see that as mask wearing behavior decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms. Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, though to lower levels than before. In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (\emph{b} = 0.25, 95\% confidence interval {[}0.20 0.30{]}) and between mask wearing behavior and perceived injunctive mask wearing norms (\emph{b} = 0.25, 95\% CI {[}0.21 0.30{]}) across individuals and time points (Supplementary Figure \ref{fig:plotCorBehNorm}; Supplementary Table \ref{tab:modelSummaryTable1}).

Second, fluctuations in mask wearing behavior and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States. We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020. Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021. Impressively, shifts in mask wearing behavior and norms happened on the timescale of weeks. These trends were confirmed by a series of multilevel regression models with change points aligning with changes in CDC mask wearing recommendations (Supplementary Figure \ref{fig:plotCDCSens}; Supplementary Table \ref{tab:changePointsTable}).

Third, across all eight time points with available data, injunctive mask wearing norms were consistently rated as stronger than descriptive mask wearing norms. Multilevel regression modeling revealed that, on average across individuals and time points, injunctive norms were rated as 0.6 points stronger than descriptive norms on a 7-point Likert scale (\emph{b} = 0.60, 95\% CI {[}0.51 0.69{]}; Supplementary Figure \ref{fig:plotNormCompare}; Supplementary Table \ref{tab:modelSummaryTable2}). In other words, participants in our study perceived the general encouragement of mask wearing and mask wearing behavior among trusted people to be higher than perceived mask wearing among people in their local area.

Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing behavior and perceived social norms over time. To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted an eight-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data. This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our three main variables: self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms (see Supplementary Table \ref{tab:lavaanTable} for estimated model parameters). According to established fit statistics, this model fitted the data well (root mean square error of approximation = 0.037, 95\% CI {[}0.032 0.042{]}; standardized root mean squared residual = 0.056; comparative fit index = 0.961).

Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing behavior and perceived social norms. On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (\emph{r} = 0.45, 95\% CI {[}0.35 0.54{]}, \emph{p} \textless{} .001) and stronger injunctive mask wearing norms (\emph{r} = 0.29, 95\% CI {[}0.18 0.40{]}, \emph{p} \textless{} .001). Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (\emph{r} = 0.65, 95\% CI {[}0.58 0.71{]}, \emph{p} \textless{} .001).

Regarding within-person dynamics over time, Figure \ref{fig:plotRICLPM} displays autoregressive and cross-lagged effects for all three variables across the study duration. In random intercept cross-lagged panel models, autoregressive effects represent ``persistence'' or ``inertia'' in within-person fluctuations from stable trait levels. In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the ``stability'' of the measure over time, which is captured by the random intercepts in our model). By contrast, and most relevant for the current study, cross-lagged effects represent the effect of a within-person fluctuation in one measure on future within-person fluctuations in other measures. In other words, a positive cross-lagged effect indicates that being higher than average on one measure predicts being higher than average on \emph{another} measure in the following time point. Cross-lagged effects are thus used to infer within-person causal influences over time.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotRICLPM-1.pdf}
\caption{\label{fig:plotRICLPM}\emph{Results of eight-wave unconstrained random-intercept cross-lagged panel model.} Arrows represent within-person autoregressive and cross-lagged effects from the model, controlling for time invariant predictors and partitioning out stable between-person individual differences. Arrow thickness is scaled according to standardized effect size. Bolded arrows indicate significantly positive parameters, \emph{p} \textless{} 0.05. Gray arrows indicate non-significant parameters.}
\end{figure}

Earlier in the pandemic, we see less evidence of autoregressive and cross-lagged effects. Within-person fluctuations in descriptive and injunctive norms persist across only a few time points, and there is only one cross-lagged effect of mask wearing on future descriptive norms. However, once COVID-19 case numbers start rising again in June 2021, we begin to see more evidence of autoregressive and cross-lagged effects. From this point onwards, all three variables show evidence of persistence, with within-person fluctuations from trait levels predicting later within-person fluctuations in the same measures. But more importantly, several positive and significant cross-lagged effects emerge. On three occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior. According to recent effect size guidelines for cross-lagged panel models (Orth et al., 2022), the standardized beta coefficients for these cross-lagged effects were large (Time 9, \(\beta\) = 0.17, 95\% CI {[}0.03 0.31{]}, \emph{p} = .020; Time 11, \(\beta\) = 0.21, 95\% CI {[}0.06 0.36{]}, \emph{p} = .006; Time 13, \(\beta\) = 0.16, 95\% CI {[}0.01 0.31{]}, \emph{p} = .037). We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms. Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.

However, the model reveals that, after controlling for perceived descriptive norms, within-person changes in perceived injunctive norms were unrelated to within-person changes in mask wearing behavior across the entire pandemic. All cross-lagged effects between the two variables, both from perceived injunctive norms to mask wearing behavior and vice versa, are non-significant. Any causal effect that perceived injunctive norms might have had on future mask wearing behavior appears to be fully mediated by perceived descriptive norms. For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing behavior. But aside from these indirect effects, perceived injunctive norms and mask wearing behavior were not directly causally related over time within individuals.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Using longitudinal data from the United States across the entire COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma. The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior are tightly coupled, and both change quickly in response to recommendations from public health authorities. The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior from June 2021 onwards. While injunctive norms were perceived as stronger than descriptive norms, we found that injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.

Our finding that social norms and mask wearing behavior are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (Titlestad et al., 2019). Moreover, the role of authorities, like the CDC, in shaping social norms and behavior supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (Markus \& Kitayama, 2010). The CDC explicitly releases guidelines for the benefit of public health, one of many situations in which the function of authorities is to coordinate behavior on large scales. However, even when institutions do not intend to coordinate behavior, they still may introduce influential social norms into the population (e.g., through business practices or products produced).

We also found that injunctive norms were, on average, perceived as stronger than descriptive norms. This aligns with previous research that measured perceived injunctive and descriptive norms in the contexts of drinking (Park et al., 2009) and marijuana use (Montes et al., 2021). What people ought to do is often perceived as stronger than what people \emph{actually} do. Since we operationalized injunctive norms in part by whether trusted individuals in the media were wearing masks, it is possible that our sample is accurately reporting that people in general were less adherent to mask wearing policies than government officials, newscasters, and other respected individuals. It may also have been easier for our participants to notice instances of non-compliance with mask wearing guidelines among the general population, simply because they have more opportunities to observe people's behavior in public settings.

Despite perceiving stronger injunctive norms, it was \emph{descriptive} norms, not injunctive norms, that independently predicted future increases in mask wearing. In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects (Rudert \& Janke, 2021). Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (Bokemper et al., 2021). Why this specific effect of descriptive norms on mask wearing? Descriptive norms are particularly useful for fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (Gelfand \& Harrington, 2015). During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms (Hogg et al., 2008; Wellen et al., 1998). Supporting this situational-uncertainty explanation, our model revealed that descriptive norms only began to predict future mask wearing after the CDC flipped their mask wearing recommendation in March 2021. With changing recommendations from authorities increasing levels of uncertainty, people instead began to look to their neighbors and adapted their mask wearing behavior accordingly.

Our finding that injunctive norms do not predict future mask wearing behavior is at odds with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults (Macy et al., 2021). One possible explanation for these conflicting findings is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior (Cialdini et al., 1991). By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people's behaviors, increasing the relative salience of injunctive norms. To test this idea, future research should expand our longitudinal cross-lagged approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).

We are limited in generalizing these findings due to the constraints of our sample and the variables considered. While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \ref{fig:plotAttrition}). This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, different ethnicities, or those with different political ideologies. Since injunctive norm strength varies based on which group is seen as the source of the norm (Neighbors et al., 2008), it would be interesting to learn whether different groups have different patterns of norm emergence over time. In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors (Baxter-King et al., 2022). Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large (Remedios, 2022). Our results also might not generalize to all social norms, behaviors, and social dilemmas. Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more abstract and remote than the COVID-19 pandemic. For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.

Regardless, in the case of mask wearing in the United States over the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and responded to both recommendations from authorities and current levels of cooperative behavior. Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing. Importantly, this key finding slices both ways. Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that \emph{low} local levels of mask wearing \emph{discouraged} future personal mask use. This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community (Natanson, 2022). Our work thus underscores the importance of consistent, visible community adherence for encouraging personal protective behaviors in response to global pandemics like COVID-19.

\newpage

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

SLH, ERH, and PMT conceptualized the study. SLH and SC oversaw the data curation, investigation, and methodology of the study, and wrote the first draft of the paper. SC conducted the formal analysis and created all visualisations. ERH and PMT provided funding and supervision for the study. All authors reviewed and edited the final draft of the paper.

\hypertarget{conflicts-of-interest}{%
\section{Conflicts of Interest}\label{conflicts-of-interest}}

There are no conflicts of interest to declare.

\hypertarget{research-transparency-and-reproducibility}{%
\section{Research Transparency and Reproducibility}\label{research-transparency-and-reproducibility}}

All data and code to reproduce the statistical analyses in this manuscript can be found on GitHub: \url{https://github.com/ScottClaessens/covidMaskWearing}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-1in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Asch1956}{}}%
Asch, S. E. (1956). Studies of independence and conformity: I. A minority of one against a unanimous majority. \emph{Psychological Monographs: General and Applied}, \emph{70}(9), 1. \url{https://doi.org/10.1037/h0093718}

\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
Aust, F., \& Barth, M. (2022). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://github.com/crsh/papaja}

\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting linear mixed-effects models using lme4. \emph{Journal of Statistical Software}, \emph{67}(1), 1--48. \url{https://doi.org/10.18637/jss.v067.i01}

\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
Baxter-King, R., Brown, J. R., Enos, R. D., Naeim, A., \& Vavreck, L. (2022). How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}. \emph{Proceedings of the National Academy of Sciences}, \emph{119}(21), e2116311119. \url{https://doi.org/10.1073/pnas.2116311119}

\leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
Bokemper, S. E., Cucciniello, M., Rotesi, T., Pin, P., Malik, A. A., Willebrand, K., Paintsil, E. E., Omer, S. B., Huber, G. A., \& Melegaro, A. (2021). Experimental evidence that changing beliefs about mask efficacy and social norms increase mask wearing for {COVID-19} risk reduction: Results from the {United States} and {Italy}. \emph{PLOS ONE}, \emph{16}(10), e0258282. \url{https://doi.org/10.1371/journal.pone.0258282}

\leavevmode\vadjust pre{\hypertarget{ref-Boyd1985}{}}%
Boyd, \&. R., Robert. (1985). \emph{Culture and the evolutionary process}. University of Chicago Press.

\leavevmode\vadjust pre{\hypertarget{ref-CDC2022b}{}}%
Centers for Disease Control and Prevention. (2022a). \emph{{CDC Museum COVID-19 Timeline}}. \url{https://www.cdc.gov/museum/timeline/covid19.html}

\leavevmode\vadjust pre{\hypertarget{ref-CDC2022a}{}}%
Centers for Disease Control and Prevention. (2022b). \emph{{United States COVID-19 Community Levels by County}} {[}Data set{]}. \url{https://data.cdc.gov/Public-Health-Surveillance/United-States-COVID-19-Community-Levels-by-County/3nnm-4jni}

\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2007}{}}%
Cialdini, R. B. (2007). Descriptive social norms as underappreciated sources of social control. \emph{Psychometrika}, \emph{72}(2), 263--268. \url{https://doi.org/10.1007/s11336-006-1560-6}

\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2021}{}}%
Cialdini, R. B., \& Jacobson, R. P. (2021). Influences of social norms on climate change-related behaviors. \emph{Current Opinion in Behavioral Sciences}, \emph{42}, 1--8. \url{https://doi.org/10.1016/j.cobeha.2021.01.005}

\leavevmode\vadjust pre{\hypertarget{ref-Cialdini1991}{}}%
Cialdini, R. B., Kallgren, C. A., \& Reno, R. R. (1991). \emph{A focus theory of normative conduct: A theoretical refinement and reevaluation of the role of norms in human behavior} (M. P. Zanna, Ed.; Vol. 24, pp. 201--234). Academic Press. \url{https://doi.org/10.1016/S0065-2601(08)60330-5}

\leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
Deutsch, M., \& Gerard, H. B. (1955). A study of normative and informational social influences upon individual judgment. \emph{The Journal of Abnormal and Social Psychology}, \emph{51}(3), 629--636. \url{https://doi.org/10.1037/h0046408}

\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018a}{}}%
Fehr, E., \& Schurtenberger, I. (2018a). Normative foundations of human cooperation. \emph{Nature Human Behaviour}, \emph{2}(7), 458--468. \url{https://doi.org/10.1038/s41562-018-0385-5}

\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018b}{}}%
Fehr, E., \& Schurtenberger, I. (2018b). \emph{The dynamics of norm formation and norm decay} (Working Paper Series). Department of Economics 278, University of Zurich. \url{https://doi.org/10.5167/uzh-147925}

\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
Gelfand, M. J., \& Harrington, J. R. (2015). The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior? \emph{Journal of Cross-Cultural Psychology}, \emph{46}(10), 1273--1278. \url{https://doi.org/10.1177/0022022115600796}

\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2011}{}}%
Gelfand, M. J., Raver, J. L., Nishii, L., Leslie, L. M., Lun, J., Lim, B. C., Duan, L., Almaliach, A., Ang, S., Arnadottir, J., Aycan, Z., Boehnke, K., Boski, P., Cabecinhas, R., Chan, D., Chhokar, J., D'Amato, A., Ferrer, M., Fischlmayr, I. C., \ldots{} Yamaguchi, S. (2011). Differences between tight and loose cultures: A 33-nation study. \emph{Science}, \emph{332}(6033), 1100--1104. \url{https://doi.org/10.1126/science.1197754}

\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
Goldstein, N. J., Cialdini, R. B., \& Griskevicius, V. (2008). A room with a viewpoint: Using social norms to motivate environmental conservation in hotels. \emph{Journal of Consumer Research}, \emph{35}(3), 472--482. \url{https://doi.org/10.1086/586910}

\leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
Granger, C. W. J. (1969). Investigating causal relations by econometric models and cross-spectral methods. \emph{Econometrica}, \emph{37}(3), 424--438. \url{https://doi.org/10.2307/1912791}

\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
Hamaker, E. L., Kuiper, R. M., \& Grasman, R. P. P. P. (2015). A critique of the cross-lagged panel model. \emph{Psychological Methods}, \emph{20}(1), 102. \url{https://doi.org/10.1037/a0038889}

\leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
Henrich, J. (2015). \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}. Princeton University Press.

\leavevmode\vadjust pre{\hypertarget{ref-Hogg2008}{}}%
Hogg, M. A., Hohman, Z. P., \& Rivera, J. E. (2008). Why do people join groups? Three motivational accounts from social psychology. \emph{Social and Personality Psychology Compass}, \emph{2}(3), 1269--1280. \url{https://doi.org/10.1111/j.1751-9004.2008.00099.x}

\leavevmode\vadjust pre{\hypertarget{ref-Jensen2014}{}}%
Jensen, K., Vaish, A., \& Schmidt, M. F. H. (2014). The emergence of human prosociality: Aligning with others through feelings, concerns, and norms. \emph{Frontiers in Psychology}, \emph{5}, 822. \url{https://doi.org/10.3389/fpsyg.2014.00822}

\leavevmode\vadjust pre{\hypertarget{ref-Jordan2009}{}}%
Jordan, F. M., Gray, R. D., Greenhill, S. J., \& Mace, R. (2009). Matrilocal residence is ancestral in austronesian societies. \emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{276}(1664), 1957--1964. \url{https://doi.org/10.1098/rspb.2009.0088}

\leavevmode\vadjust pre{\hypertarget{ref-Kushnick2014}{}}%
Kushnick, G., Gray, R. D., \& Jordan, F. M. (2014). The sequential evolution of land tenure norms. \emph{Evolution and Human Behavior}, \emph{35}(4), 309--318. \url{https://doi.org/10.1016/j.evolhumbehav.2014.03.001}

\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
Landau, W. M. (2021). The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing. \emph{Journal of Open Source Software}, \emph{6}(57), 2959. \url{https://doi.org/10.21105/joss.02959}

\leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
Larkin, C., Sanders, M., Andresen, I., \& Algate, F. (2018). \emph{Testing local descriptive norms and salience of enforcement action: A field experiment to increase tax collection} (SSRN Scholarly Paper No. 3167575). \url{https://doi.org/10.2139/ssrn.3167575}

\leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
Macy, J. T., Owens, C., Mullis, K., \& Middlestadt, S. E. (2021). The role of self-efficacy and injunctive norms in helping older adults decide to stay home during the {COVID-19} pandemic. \emph{Frontiers in Public Health}, \emph{9}, 660813. \url{https://doi.org/10.3389/fpubh.2021.660813}

\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
Markus, H. R., \& Kitayama, S. (2010). Cultures and selves: A cycle of mutual constitution. \emph{Perspectives on Psychological Science}, \emph{5}(4), 420--430. \url{https://doi.org/10.1177/1745691610375557}

\leavevmode\vadjust pre{\hypertarget{ref-MITElectionLab}{}}%
MIT Election Data and Science Lab. (2018). \emph{{County Presidential Election Returns 2000-2020}} (Version V10) {[}Data set{]}. {Harvard Dataverse}. \url{https://doi.org/10.7910/DVN/VOQCHQ}

\leavevmode\vadjust pre{\hypertarget{ref-Montes2021}{}}%
Montes, K. S., Richards, D. K., \& Pearson, M. R. (2021). A novel approach to assess descriptive and injunctive norms for college student marijuana use. \emph{Addictive Behaviors}, \emph{117}, 106755. \url{https://doi.org/10.1016/j.addbeh.2020.106755}

\leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
Natanson, H. (2022). Peer pressure is ending mask usage in schools. \emph{The Washingston Post}. \url{https://www.washingtonpost.com/education/2022/02/25/peer-pressure-mask-optional-schools/}

\leavevmode\vadjust pre{\hypertarget{ref-Neighbors2008}{}}%
Neighbors, C., O'Connor, R. M., Lewis, M. A., Chawla, N., Lee, C. M., \& Fossos, N. (2008). The relative impact of injunctive norms on college student drinking: The role of reference group. \emph{Psychology of Addictive Behaviors}, \emph{22}(4), 576--581. \url{https://doi.org/10.1037/a0013043}

\leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
Nigbur, D., Lyons, E., \& Uzzell, D. (2010). Attitudes, norms, identity and environmental behaviour: Using an expanded theory of planned behaviour to predict participation in a kerbside recycling programme. \emph{British Journal of Social Psychology}, \emph{49}(2), 259--284. \url{https://doi.org/10.1348/014466609X449395}

\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
Orth, U., Meier, L. L., Bühler, J. L., Dapp, L. C., Krauss, S., Messerli, D., \& Robins, R. W. (2022). Effect size guidelines for cross-lagged effects. \emph{Psychological Methods}. \url{https://doi.org/10.1037/met0000499}

\leavevmode\vadjust pre{\hypertarget{ref-Park2009}{}}%
Park, H. S., Klein, K. A., Smith, S., \& Martell, D. (2009). Separating subjective norms, university descriptive and injunctive norms, and {U.S.} Descriptive and injunctive norms for drinking behavior intentions. \emph{Health Communication}, \emph{24}(8), 746--751. \url{https://doi.org/10.1080/10410230903265912}

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
R Core Team. (2022). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-Rakoczy2013}{}}%
Rakoczy, H., \& Schmidt, M. F. H. (2013). The early ontogeny of social norms. \emph{Child Development Perspectives}, \emph{7}(1), 17--21. \url{https://doi.org/10.1111/cdep.12010}

\leavevmode\vadjust pre{\hypertarget{ref-Remedios2022}{}}%
Remedios, J. D. (2022). Psychology must grapple with whiteness. \emph{Nature Reviews Psychology}, \emph{1}(3), 125--126. \url{https://doi.org/10.1038/s44159-022-00024-4}

\leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
Rosseel, Y. (2012). {lavaan}: An {R} package for structural equation modeling. \emph{Journal of Statistical Software}, \emph{48}(2), 1--36. \url{https://doi.org/10.18637/jss.v048.i02}

\leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
Rudert, S. C., \& Janke, S. (2021). Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic. \emph{Group Processes \& Intergroup Relations}, 13684302211023562. \url{https://doi.org/10.1177/13684302211023562}

\leavevmode\vadjust pre{\hypertarget{ref-Schaumberg2022}{}}%
Schaumberg, R. L., \& Skowronek, S. E. (2022). Shame broadcasts social norms: The positive social effects of shame on norm acquisition and normative behavior. \emph{Psychological Science}, 09567976221075303. \url{https://doi.org/10.1177/09567976221075303}

\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2012}{}}%
Schmidt, M. F. H., \& Tomasello, M. (2012). Young children enforce social norms. \emph{Current Directions in Psychological Science}, \emph{21}(4), 232--236. \url{https://doi.org/10.1177/0963721412448659}

\leavevmode\vadjust pre{\hypertarget{ref-Schultz2007}{}}%
Schultz, P. W., Nolan, J. M., Cialdini, R. B., Goldstein, N. J., \& Griskevicius, V. (2007). The constructive, destructive, and reconstructive power of social norms. \emph{Psychological Science}, \emph{18}(5), 429--434. \url{https://doi.org/10.1111/j.1467-9280.2007.01917.x}

\leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
Smith, L. G. E., Thomas, E. F., \& McGarty, C. (2015). "We must be the change we want to see in the world": Integrating norms and identities through social interaction. \emph{Political Psychology}, \emph{36}(5), 543--557. \url{https://doi.org/10.1111/pops.12180}

\leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
Titlestad, K., Snijders, T. A. B., Durrheim, K., Quayle, M., \& Postmes, T. (2019). The dynamic emergence of cooperative norms in a social dilemma. \emph{Journal of Experimental Social Psychology}, \emph{84}, 103799. \url{https://doi.org/10.1016/j.jesp.2019.03.010}

\leavevmode\vadjust pre{\hypertarget{ref-Tomasello2014}{}}%
Tomasello, M. (2014). The ultra-social animal. \emph{European Journal of Social Psychology}, \emph{44}(3), 187--194. \url{https://doi.org/10.1002/ejsp.2015}

\leavevmode\vadjust pre{\hypertarget{ref-Vaish2011}{}}%
Vaish, A., Carpenter, M., \& Tomasello, M. (2011). Young children's responses to guilt displays. \emph{Developmental Psychology}, \emph{47}(5), 1248--1262. \url{https://doi.org/10.1037/a0024462}

\leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
van Kleef, G. A., Gelfand, M. J., \& Jetten, J. (2019). The dynamic nature of social norms: New perspectives on norm development, impact, violation, and enforcement. \emph{Journal of Experimental Social Psychology}, \emph{84}, 103814. \url{https://doi.org/10.1016/j.jesp.2019.05.002}

\leavevmode\vadjust pre{\hypertarget{ref-Wellen1998}{}}%
Wellen, J. M., Hogg, M. A., \& Terry, D. J. (1998). Group norms and attitude-behavior consistency: The role of group salience and mood. \emph{Group Dynamics: Theory, Research, and Practice}, \emph{2}(1), 48--56. \url{https://doi.org/10.1037/1089-2699.2.1.48}

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}

\leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
Wilke, C. O. (2020). \emph{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}. \url{https://CRAN.R-project.org/package=cowplot}

\end{CSLReferences}

\endgroup

\newpage

\hypertarget{appendix-appendix}{%
\appendix}

\renewcommand{\appendixname}{\textbf{Supplementary Materials}}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

\hypertarget{section}{%
\section{}\label{section}}

\hypertarget{supplementary-results}{%
\subsection{Supplementary Results}\label{supplementary-results}}

\hypertarget{construct-validity-for-measures-of-perceived-descriptive-and-injunctive-norms}{%
\subsubsection{Construct validity for measures of perceived descriptive and injunctive norms}\label{construct-validity-for-measures-of-perceived-descriptive-and-injunctive-norms}}

To evaluate the construct validity of our measures of perceived descriptive and injunctive norms, at Time 7 we asked participants to rate the extent to which each perceived norm item provided descriptive and injunctive information. For each item, participants were asked whether the item provided information about what people are doing, and whether the item provided information about what people should be doing. Participants responded on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly. For a full list of questions, see Supplementary Table \ref{tab:itemTable}.

Results showed that participants did differentiate the perceived norm items as expected. Participants rated the perceived descriptive norm items as providing more descriptive information than the perceived injunctive norm items, \emph{t}(442) = -7.28, \emph{p} \textless{} .001 (mean descriptive items = 4.75; mean injunctive items = 4.25). By contrast, participants rated the perceived injunctive norm items as providing more injunctive information than the perceived descriptive norm items, \emph{t}(444) = 7.15, \emph{p} \textless{} .001 (mean descriptive items = 5.11; mean injunctive items = 5.54).

\newpage

\hypertarget{supplementary-figures}{%
\subsection{Supplementary Figures}\label{supplementary-figures}}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotUSMap-1.pdf}
\caption{\label{fig:plotUSMap}\emph{Map of the United States with participant zip code locations.}}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotAttrition-1.pdf}
\caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotDAG-1.pdf}
\caption{\label{fig:plotDAG}\emph{Directed acyclic graph reflecting causal assumptions.} In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms cause mask wearing behavior (Mask). County-level COVID-19 prevalence (CovidPrev) and political leaning (CountyPol) are also posited to be common causes of perceptions of social norms and mask wearing behavior. Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived descriptive norms on mask wearing. Similarly, it is necessary to control for perceptions of descriptive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotCorBehNorm-1.pdf}
\caption{\label{fig:plotCorBehNorm}\emph{Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables.} Lines are fixed effect regression lines from multilevel models, shaded areas are 95\% confidence intervals.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotCDCSens-1.pdf}
\caption{\label{fig:plotCDCSens}\emph{Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations.} These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms. Time is included as a continuous linear predictor, scaled between 0 and 1, with two forced change points (dashed lines). The first dashed line (left) indicates when the CDC relaxed their mask wearing recommendations in March 2021, and the second dashed line (right) indicates when the CDC strengthened their mask wearing recommendations in July 2021. This results in the estimation of four fixed effect parameters: the initial intercept, the slope in the first window, the slope in the second window, and the slope in the third window. Bolded lines and shaded areas represent fixed effect regression lines from multilevel models and 95\% confidence intervals, respectively.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotNormCompare-1.pdf}
\caption{\label{fig:plotNormCompare}\emph{Predictions from multilevel model of the difference between the perceived strength of descriptive and injunctive social norms.} Points and error bars represent estimated marginal means and 95\% confidence intervals, respectively.}
\end{figure}

\newpage

\hypertarget{supplementary-tables}{%
\subsection{Supplementary Tables}\label{supplementary-tables}}

\begin{center}
\begin{ThreePartTable}

\begin{longtable}{m{2.5cm}m{2.5cm}m{9cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:itemTable}List of norm interpretation questions asked at Time 7. \emph{These questions were preceded by the following text:} ``There may or may not be a difference between what people around you are doing and what they should be doing. You can learn about what people are doing and what they should be doing in different ways. For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both.'' \emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
\toprule
Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:itemTable} continued}}\\
\toprule
Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
\midrule
\endhead
Provides descriptive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone is doing?\\
 &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone is doing?\\
 & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone is doing?\\
 &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone is doing?\\
Provides injunctive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone should be doing?\\
 &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone should be doing?\\
 & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone should be doing?\\
 &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone should be doing?\\
\bottomrule
\end{longtable}

\end{ThreePartTable}
\end{center}

\newpage

  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

\begin{table}[ht]
\begin{centerbox}
\begin{threeparttable}
\captionsetup{justification=centering,singlelinecheck=off}
\caption{Unstandardized fixed effect parameters from multilevel models: perceptions of social norm strength predicting self-reported mask wearing. \emph{Standard errors are included in brackets.}}
 \label{tab:modelSummaryTable1}
\setlength{\tabcolsep}{0pt}
\begin{tabular}{l l l}

\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Model 1 \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Model 2 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Intercept \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.86\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.67\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.17) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.16) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Descriptive norms \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.25\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.03) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Injunctive norms \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.25\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.02) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3781\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3790\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (id) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     778\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     778\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (time) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}       8\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}       8\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} AIC \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 11995.16\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 12031.84\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} BIC \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 12051.30\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 12088.00\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (fixed) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.07\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.08\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (total) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.42\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.42\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
\arrayrulecolor{black}
\end{tabular}
\end{threeparttable}\par\end{centerbox}

\end{table}

\newpage

\begin{center}
\begin{ThreePartTable}

\begin{longtable}{lccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:changePointsTable}Unstandardized fixed effect parameters from multilevel models: trends over time with change points at CDC events.}\\
\toprule
  & \multicolumn{1}{c}{Mask wearing} & \multicolumn{1}{c}{Descriptive norms} & \multicolumn{1}{c}{Injunctive norms}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:changePointsTable} continued}}\\
\toprule
  & \multicolumn{1}{c}{Mask wearing} & \multicolumn{1}{c}{Descriptive norms} & \multicolumn{1}{c}{Injunctive norms}\\
\midrule
\endhead
Intercept & 4.13, 95\% CI [4.06 4.21] & 5.01, 95\% CI [4.93 5.10] & 5.65, 95\% CI [5.56 5.75]\\
Slope1 & 0.71, 95\% CI [0.40 1.04] & 1.26, 95\% CI [0.76 1.74] & 0.41, 95\% CI [-0.06 0.90]\\
Slope2 & -3.67, 95\% CI [-4.13 -3.24] & -7.26, 95\% CI [-7.78 -6.75] & -5.82, 95\% CI [-6.37 -5.31]\\
Slope3 & 0.99, 95\% CI [0.69 1.31] & 0.86, 95\% CI [0.56 1.16] & 0.51, 95\% CI [0.21 0.81]\\
N & 7493 & 3837 & 3843\\
R2 (fixed) & 0.05 & 0.26 & 0.21\\
R2 (total) & 0.44 & 0.72 & 0.71\\
\bottomrule
\end{longtable}

\end{ThreePartTable}
\end{center}

\newpage

  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

\begin{table}[ht]
\begin{centerbox}
\begin{threeparttable}
\captionsetup{justification=centering,singlelinecheck=off}
\caption{Unstandardized fixed effect parameters from multilevel model: difference between perceived strength of descriptive and injunctive social norms. \emph{Standard errors are included in brackets.}}
 \label{tab:modelSummaryTable2}
\setlength{\tabcolsep}{0pt}
\begin{tabular}{l l}

\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Model 1 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Intercept \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 4.28\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.29) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Difference between norm types \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.60\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.04) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 7680\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (id) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     779\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (time) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}       8\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} AIC \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 22409.98\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} BIC \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 22472.50\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (fixed) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.04\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{}
\arrayrulecolor{black}

\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (total) \hspace{6pt}\huxbpad{6pt}} &
\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.65\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]

\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
\arrayrulecolor{black}
\end{tabular}
\end{threeparttable}\par\end{centerbox}

\end{table}

\newpage

\begin{center}
\begin{ThreePartTable}

\begin{longtable}{lcccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:lavaanTable}Standardised autoregressive and cross-lagged parameters from random-intercept cross-lagged panel model. \emph{Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms. Variable name suffixes indicate time points. Arrows indicate the direction of prediction.}}\\
\toprule
Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:lavaanTable} continued}}\\
\toprule
Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%}\\
\midrule
\endhead
Des\_02 → Mask\_03 & -0.01 & 0.06 & -0.14 & 0.11\\
Des\_02 → Inj\_03 & 0.08 & 0.07 & -0.05 & 0.21\\
Des\_02 → Des\_03 & 0.20 & 0.07 & 0.07 & 0.33\\
Des\_03 → Mask\_05 & 0.11 & 0.06 & -0.01 & 0.24\\
Des\_03 → Inj\_05 & 0.10 & 0.07 & -0.03 & 0.23\\
Des\_03 → Des\_05 & 0.32 & 0.06 & 0.19 & 0.44\\
Des\_05 → Mask\_09 & 0.11 & 0.07 & -0.03 & 0.24\\
Des\_05 → Inj\_09 & 0.11 & 0.07 & -0.02 & 0.25\\
Des\_05 → Des\_09 & 0.06 & 0.07 & -0.08 & 0.20\\
Des\_09 → Mask\_11 & 0.17 & 0.07 & 0.03 & 0.31\\
Des\_09 → Inj\_11 & 0.28 & 0.07 & 0.15 & 0.41\\
Des\_09 → Des\_11 & 0.27 & 0.07 & 0.14 & 0.40\\
Des\_11 → Mask\_13 & 0.21 & 0.08 & 0.06 & 0.36\\
Des\_11 → Inj\_13 & 0.15 & 0.07 & 0.01 & 0.30\\
Des\_11 → Des\_13 & 0.27 & 0.07 & 0.13 & 0.41\\
Des\_13 → Mask\_14 & 0.16 & 0.08 & 0.01 & 0.31\\
Des\_13 → Inj\_14 & 0.21 & 0.06 & 0.09 & 0.33\\
Des\_13 → Des\_14 & 0.32 & 0.06 & 0.20 & 0.45\\
Des\_14 → Mask\_15 & 0.15 & 0.08 & -0.02 & 0.31\\
Des\_14 → Inj\_15 & 0.11 & 0.07 & -0.03 & 0.26\\
Des\_14 → Des\_15 & 0.35 & 0.07 & 0.22 & 0.48\\
Inj\_02 → Mask\_03 & 0.10 & 0.06 & -0.02 & 0.23\\
Inj\_02 → Inj\_03 & 0.07 & 0.07 & -0.06 & 0.21\\
Inj\_02 → Des\_03 & 0.03 & 0.06 & -0.09 & 0.15\\
Inj\_03 → Mask\_05 & 0.03 & 0.06 & -0.09 & 0.16\\
Inj\_03 → Inj\_05 & 0.14 & 0.07 & 0.01 & 0.28\\
Inj\_03 → Des\_05 & 0.01 & 0.06 & -0.11 & 0.14\\
Inj\_05 → Mask\_09 & -0.11 & 0.07 & -0.24 & 0.02\\
Inj\_05 → Inj\_09 & -0.07 & 0.07 & -0.21 & 0.07\\
Inj\_05 → Des\_09 & -0.05 & 0.07 & -0.19 & 0.08\\
Inj\_09 → Mask\_11 & 0.04 & 0.08 & -0.11 & 0.19\\
Inj\_09 → Inj\_11 & 0.13 & 0.07 & 0.00 & 0.27\\
Inj\_09 → Des\_11 & 0.07 & 0.07 & -0.06 & 0.21\\
Inj\_11 → Mask\_13 & 0.05 & 0.07 & -0.09 & 0.19\\
Inj\_11 → Inj\_13 & 0.32 & 0.07 & 0.19 & 0.45\\
Inj\_11 → Des\_13 & 0.22 & 0.07 & 0.08 & 0.35\\
Inj\_13 → Mask\_14 & -0.02 & 0.07 & -0.17 & 0.12\\
Inj\_13 → Inj\_14 & 0.42 & 0.06 & 0.31 & 0.54\\
Inj\_13 → Des\_14 & 0.23 & 0.06 & 0.10 & 0.35\\
Inj\_14 → Mask\_15 & 0.03 & 0.08 & -0.13 & 0.18\\
Inj\_14 → Inj\_15 & 0.47 & 0.07 & 0.34 & 0.60\\
Inj\_14 → Des\_15 & 0.31 & 0.06 & 0.19 & 0.44\\
Mask\_02 → Mask\_03 & -0.02 & 0.06 & -0.14 & 0.10\\
Mask\_02 → Inj\_03 & 0.00 & 0.06 & -0.13 & 0.12\\
Mask\_02 → Des\_03 & 0.01 & 0.06 & -0.11 & 0.12\\
Mask\_03 → Mask\_05 & 0.02 & 0.06 & -0.10 & 0.13\\
Mask\_03 → Inj\_05 & 0.04 & 0.06 & -0.08 & 0.16\\
Mask\_03 → Des\_05 & 0.12 & 0.06 & 0.00 & 0.23\\
Mask\_05 → Mask\_09 & 0.08 & 0.07 & -0.04 & 0.21\\
Mask\_05 → Inj\_09 & 0.10 & 0.07 & -0.03 & 0.23\\
Mask\_05 → Des\_09 & -0.01 & 0.07 & -0.15 & 0.12\\
Mask\_09 → Mask\_11 & -0.03 & 0.07 & -0.16 & 0.10\\
Mask\_09 → Inj\_11 & 0.01 & 0.06 & -0.11 & 0.14\\
Mask\_09 → Des\_11 & 0.10 & 0.06 & -0.02 & 0.22\\
Mask\_11 → Mask\_13 & 0.20 & 0.06 & 0.07 & 0.32\\
Mask\_11 → Inj\_13 & 0.05 & 0.06 & -0.07 & 0.16\\
Mask\_11 → Des\_13 & 0.10 & 0.06 & -0.02 & 0.21\\
Mask\_13 → Mask\_14 & 0.30 & 0.06 & 0.18 & 0.42\\
Mask\_13 → Inj\_14 & 0.10 & 0.05 & -0.01 & 0.20\\
Mask\_13 → Des\_14 & 0.20 & 0.05 & 0.09 & 0.30\\
Mask\_14 → Mask\_15 & 0.29 & 0.06 & 0.17 & 0.41\\
Mask\_14 → Inj\_15 & 0.04 & 0.06 & -0.07 & 0.16\\
Mask\_14 → Des\_15 & 0.11 & 0.05 & 0.01 & 0.21\\
\bottomrule
\end{longtable}

\end{ThreePartTable}
\end{center}

\newpage

\hypertarget{supplementary-references}{%
\subsection{Supplementary References}\label{supplementary-references}}

Pearl, J. (1995). Causal diagrams for empirical research. \emph{Biometrika}, \emph{82}(4), 669--688, \url{https://doi.org/10.1093/biomet/82.4.669}

\end{document}

### 2022-08-18 R/functions.R

@@ -32,14 +32,14 @@ loadData <- function(fileData, fileFIPS, fileElec, filePrev) {
   # load county-level political election results
   elec <-
     read.csv(fileElec) %>%
-    select(county_fips, winningparty) %>%
+    dplyr::select(county_fips, winningparty) %>%
     unique() %>%
     mutate(winningparty = ifelse(winningparty == "REPUBLICAN", 1,
                                  ifelse(winningparty == "DEMOCRAT", 0, NA)))
   # attach data to fips (county) and zip codes
   zip <-
     read.csv(fileFIPS) %>%
-    select(ZIP, STCOUNTYFP) %>%
+    dplyr::select(ZIP, STCOUNTYFP) %>%
     # join county-level COVID-19 prevalence data
     left_join(prev, by = c("STCOUNTYFP" = "county_fips")) %>%
     # join county-level political election results
@@ -55,7 +55,7 @@ loadData <- function(fileData, fileFIPS, fileElec, filePrev) {
   # join to main data frame
   out <- left_join(out, zip, by = c("Zip.1" = "ZIP"))
   # add further data on zip codes
-  zip <- reverse_zipcode(unique(out$Zip.1)) %>% mutate(zipcode = as.numeric(zipcode))
+  zip <- zipcodeR::reverse_zipcode(unique(out$Zip.1)) %>% mutate(zipcode = as.numeric(zipcode))
   out <- left_join(out, zip, by = c("Zip.1" = "zipcode"))
   return(out)
 }
@@ -67,12 +67,22 @@ loadOwid <- function(fileOwid) {
     mutate(date = dmy(date))
 }
 
+# t-test: construct validity
+testConstructValidity <- function(d, type) {
+  # how much information do the descriptive items give about X type of norms?
+  y1 <- (d[,paste0(type, "Learning1.7")] + d[,paste0(type, "Learning2.7")]) / 2
+  # how much information do the injunctive items give about X type of norms?
+  y2 <- (d[,paste0(type, "Learning3.7")] + d[,paste0(type, "Learning4.7")]) / 2
+  # paired t-test
+  t.test(y1[,1], y2[,1], paired = TRUE)
+}
+
 # correlations between behaviour and norms
 fitCorBehNorm <- function(d, predictor) {
   # get data in long format
   dLong <-
     d %>%
-    select(starts_with("DesNorms") | starts_with("InjNorms") | starts_with("ContactsMask")) %>%
[…]

### 2022-08-18 _targets.R

@@ -3,9 +3,11 @@ library(targets)
 library(tarchetypes)
 source("R/functions.R")
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("cowplot", "dagitty", "ggdag", "ggeffects", "ggraph", 
-                            "lavaan", "lme4", "lubridate", "rnaturalearth", 
-                            "rnaturalearthdata", "scales", "sf", "zipcodeR"))
+tar_option_set(packages = c("arm", "cowplot", "dagitty", "ggdag", "ggeffects", 
+                            "ggraph", "huxtable", "jtools", "knitr", "kableExtra", 
+                            "lavaan", "lme4", "lubridate", "MuMIn", "papaja", 
+                            "rnaturalearth", "rnaturalearthdata", "scales", 
+                            "sf", "zipcodeR"))
 # workflow
 list(
   # files
@@ -18,6 +20,9 @@ list(
   tar_target(d, loadData(fileData, fileFIPS, fileElec, filePrev)),
   # load US covid counts from our world in data
   tar_target(owid, loadOwid(fileOwid)),
+  # t-tests: construct validity
+  tar_target(constructValDes, testConstructValidity(d, type = "Descriptive")),
+  tar_target(constructValInj, testConstructValidity(d, type = "Injunctive")),
   # multilevel modelling
   # correlations between behaviour and norms
   tar_target(m1.1, fitCorBehNorm(d, predictor = "DesNorms")),
@@ -47,6 +52,12 @@ list(
   tar_target(plot6, plotCDCSens(m2.1, m2.2, m2.3)),
   tar_target(plot7, plotNormCompare(m3.1)),
   tar_target(plot8, plotRICLPM(riclpm)),
+  # tables
+  tar_target(itemTable, makeItemTable()),
+  tar_target(changePointsTable, makeChangePointsTable(m2.1, m2.2, m2.3)),
+  tar_target(lavaanTable, makeLavaanTable(riclpm)),
+  # manuscript
+  tar_render(manuscript, "manuscript.Rmd"),
   # print session info for reproducibility
   tar_target(sessionInfo, writeLines(capture.output(sessionInfo()), "sessionInfo.txt"))
 )

### 2022-10-13 appendix.Rmd
update analysis with t16, t17, and t18

@@
 
-```{r plotAttrition, echo=F, fig.width=5, fig.height=3, fig.cap="(ref:plotAttritionCaption)"}
+```{r plotAttrition, echo=F, fig.width=5.5, fig.height=3, fig.cap="(ref:plotAttritionCaption)"}
 plot2
@@
 plot6
-```
-
-\newpage
-
-(ref:plotNormCompareCaption) _Predictions from multilevel model of the difference between the perceived strength of descriptive and injunctive social norms._ Points and error bars represent estimated marginal means and 95% confidence intervals, respectively.
-
-```{r plotNormCompare, echo=F, fig.width=3.5, fig.height=3.5, fig.cap="(ref:plotNormCompareCaption)"}
-plot7
 ```
@@
 
-(ref:modelSummaryTable2Caption) Unstandardized fixed effect parameters from multilevel model: difference between perceived strength of descriptive and injunctive social norms.
-_Standard errors are included in brackets._
-
-```{r modelSummaryTable2, echo=F, warning=F, message=F, results="asis"}
-export_summs(m3.1, stars = NULL,
-             coefs = c("Intercept" = "(Intercept)",
-                       "Difference between norm types" = "varInjNorms")) %>% 
-  set_caption("(ref:modelSummaryTable2Caption)") %>%
-  set_label("tab:modelSummaryTable2")
-```
-
-\newpage
-
 (ref:lavaanTableCaption) Standardised autoregressive and cross-lagged parameters from random-intercept cross-lagged panel model.

### 2022-10-13 R/functions.R

@@ -1,60 +1,68 @@
 # custom functions
 
-loadData <- function(fileData, fileFIPS, fileElec, filePrev) {
+loadData <- function(fileData) {
   # load data
   out <-
-    read.csv(file = fileData, sep = '\t') %>%
-    as_tibble() %>%
-    # create composite measures of injunctive and descriptive norms
-    mutate(InjNorms.2  = (MaskRespect.2    + MaskEncouraged.2 ) / 2,
-           InjNorms.3  = (MaskRespect.3    + MaskEncouraged.3 ) / 2,
-           InjNorms.5  = (MaskRespect.5    + MaskEncouraged.5 ) / 2,
-           InjNorms.9  = (MaskRespect.9    + MaskEncouraged.9 ) / 2,
-           InjNorms.11 = (MaskRespect.11   + MaskEncouraged.11) / 2,
-           InjNorms.13 = (MaskRespect.13   + MaskEncouraged.13) / 2,
-           InjNorms.14 = (MaskRespect.14   + MaskEncouraged.14) / 2,
-           InjNorms.15 = (MaskRespect.15   + MaskEncouraged.15) / 2,
-           DesNorms.2  = (NeighborMask1.2  + NeighborMask2.2  ) / 2,
-           DesNorms.3  = (NeighborMask1.3  + NeighborMask2.3  ) / 2,
-           DesNorms.5  = (NeighborMask1.5  + NeighborMask2.5  ) / 2,
-           DesNorms.9  = (NeighborMask1.9  + NeighborMask2.9  ) / 2,
-           DesNorms.11 = (NeighborMask1.11 + NeighborMask2.11 ) / 2,
-           DesNorms.13 = (NeighborMask1.13 + NeighborMask2.13 ) / 2,
-           DesNorms.14 = (NeighborMask1.14 + NeighborMask2.14 ) / 2,
-           DesNorms.15 = (NeighborMask1.15 + NeighborMask2.15 ) / 2
+    read_csv(file = fileData) %>%
+    # fix some mask motives items (7 coded as 8 in qualtrics)
+    mutate_at(vars(contains("MaskMotives")), function(x) ifelse(x == 8, 7, x)) %>%
+    mutate_at(vars(contains("MaskMotivates")), function(x) ifelse(x == 8, 7, x)) %>%
+    # fix mask legality items
[…]

### 2022-10-13 _targets.R

@@ -7,17 +7,14 @@ tar_option_set(packages = c("arm", "cowplot", "dagitty", "ggdag", "ggeffects",
                             "ggraph", "huxtable", "jtools", "knitr", "kableExtra", 
                             "lavaan", "lme4", "lubridate", "MuMIn", "papaja", 
                             "rnaturalearth", "rnaturalearthdata", "scales", 
-                            "sf", "zipcodeR"))
+                            "semTools", "sf", "zipcodeR"))
 # workflow
 list(
   # files
-  tar_target(fileData, "data/cus_t1-15_noIDs_280422av2337.tsv", format = "file"),
-  tar_target(fileFIPS, "data/ZIP-COUNTY-FIPS_2018-03.csv", format = "file"),
-  tar_target(fileElec, "data/countypres_2020_HC.csv", format = "file"),
-  tar_target(filePrev, "data/United_States_COVID-19_Community_Levels_by_County.csv", format = "file"),
+  tar_target(fileData, "data/usCohortCleanData_t1-t18.csv", format = "file"),
   tar_target(fileOwid, "data/owid-covid-data.csv", format = "file"),
   # load data
-  tar_target(d, loadData(fileData, fileFIPS, fileElec, filePrev)),
+  tar_target(d, loadData(fileData)),
   # load US covid counts from our world in data
   tar_target(owid, loadOwid(fileOwid)),
   # t-tests: construct validity
@@ -31,17 +28,15 @@ list(
   tar_target(m2.1, fitCDCSens(d, outcome = "ContactsMask")),
   tar_target(m2.2, fitCDCSens(d, outcome = "DesNorms")),
   tar_target(m2.3, fitCDCSens(d, outcome = "InjNorms")),
-  # difference between levels of descriptive and injunctive norms
-  tar_target(m3.1, fitNormCompare(d)),
   # confidence intervals
   tar_target(conf1.1, confint(m1.1, method = "Wald")),
   tar_target(conf1.2, confint(m1.2, method = "Wald")),
   tar_target(conf2.1, confint(m2.1, method = "Wald")),
   tar_target(conf2.2, confint(m2.2, method = "Wald")),
   tar_target(conf2.3, confint(m2.3, method = "Wald")),
-  tar_target(conf3.1, confint(m3.1, method = "Wald")),
[…]

### 2022-10-13 appendix.Rmd
first draft manuscript revision

@@
 
-(ref:plotDAGCaption) _Directed acyclic graph reflecting causal assumptions._ In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms cause mask wearing behavior (Mask).
-County-level COVID-19 prevalence (CovidPrev) and political leaning (CountyPol) are also posited to be common causes of perceptions of social norms and mask wearing behavior.
-Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived descriptive norms on mask wearing.
-Similarly, it is necessary to control for perceptions of descriptive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived injunctive norms on mask wearing.
+(ref:plotDAGCaption) _Directed acyclic graph reflecting causal assumptions._ In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing behavior (Mask).
+Perceptions of descriptive and injunctive norms also indirectly cause mask wearing behavior through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
+Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
+Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.
 
@@
 (ref:lavaanTableCaption) Standardised autoregressive and cross-lagged parameters from random-intercept cross-lagged panel model.
-_Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms.
+_Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms, Fact = factual beliefs, Pers = personal normative beliefs.
 Variable name suffixes indicate time points.
@@
   longtable = TRUE,
+  font_size = "footnotesize",
   align = c("l", rep("c", 5)),

### 2022-10-13 manuscript.Rmd
Update manuscript.Rmd
first draft manuscript revision

@@
 To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
-Leveraging a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; _n_ = `r nrow(tar_read(d))`), we tracked people’s reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
+Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = `r nrow(tar_read(d))`), we tracked people’s reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
 Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior are tightly coupled and both change quickly in response to recommendations from public health authorities.
@@
 keywords          : "descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
-wordcount         : "4826 words"
+wordcount         : "xxxx words"
 bibliography      : ["r-references.bib"]
@@
 appendix          : "appendix.Rmd"
-csl               : apa.csl
-urlcolor          : blue
+csl               : pnas.csl
 ---
@@
 # to ensure that the Rmd runs last, load appendix files again
-tar_load(c(d, plot1, plot2, plot4, plot5, plot6, plot7,
-           m1.1, m1.2, m3.1, itemTable, lavaanTable,
+tar_load(c(d, plot1, plot2, plot4, plot5, plot6,
+           m1.1, m1.2, itemTable, lavaanTable,
            changePointsTable, constructValDes, constructValInj))
@@
 Social norms are a key aspect of human sociality.
-Broadly defined as behavioral guidelines enforced by groups of people without the law [@VanKleef2019], social norms have attracted the attention of psychologists for decades.
-Since Asch's [-@Asch1956] early studies of normative conformity, evidence has accrued that humans are particularly attuned to social norms [@Tomasello2014].
+Broadly defined as commonly known behavioral guidelines enforced by groups of people, social norms have attracted the attention of behavioral scientists for decades [@Bicchieri2009; @Cialdini1991; @Fehr2018a; @Legros2020; @VanKleef2019].
+Since Asch's early studies of normative conformity [@Asch1956], evidence has accrued that humans are particularly attuned to social norms [@Tomasello2014].
 From a young age, children begin to adhere to and enforce group-wide social norms [@Jensen2014; @Rakoczy2013; @Schmidt2012] and both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms [@Vaish2011; @Schaumberg2022].
@@
 
-Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms [@Cialdini1991; @Deutsch1955].
-Injunctive social norms are societal, indicating what others in the group tend to approve or disapprove of, and often involve social sanctions if violated.
+Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms [@Bicchieri2009; @Cialdini1991; @Deutsch1955].
+Injunctive social norms indicate what others in the group tend to approve or disapprove of and often involve social sanctions if violated.
 By contrast, descriptive norms are situational, simply describing what most people are doing in a given situation.
@@
 
-Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], social norms have remained an elusive concept in the behavioral sciences and several open questions remain [@Fehr2018a; @VanKleef2019].
+Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Bicchieri2009; @Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], social norms have remained an elusive concept in the behavioral sciences and several open questions remain [@Fehr2018a; @VanKleef2019].
 In the current work, we focus on two such questions.
@@
 Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals’ attention on injunctive norms [@Schultz2007].
-However, previous studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population.
+However, these cross-sectional studies have two main limitations.
+First, studies have not adequately controlled for other potential non-social influences on behavior, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
+For example, descriptive social norms might only affect cooperative behavior indirectly by influencing perceptions of the effectiveness of the behavior (i.e., factual beliefs) or individuals' moral preferences (i.e., personal normative beliefs).
+Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population.
 An alternative way to identify causality, whilst retaining ecological validity, is to follow perceptions of social norms and cooperative behavior over time amidst a real, unfolding social dilemma.
@@
 Before the pandemic, mask wearing was not a common behavior in the United States.
-In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a measure that people should take to minimize the spread of COVID-19.
+In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a cooperative behavior that people should adopt to minimize the spread of COVID-19.
 But mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., "flattening the curve" to protect at-risk individuals).
@@
 These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between between-person and within-person change over time [@Hamaker2015], nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
+These studies also do not control for the influences of non-social beliefs, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
 
-Here, we use a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; _n_ = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
+Here, we use two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
 We aimed to answer two main research questions.
@@
 Second, how do descriptive and injunctive mask wearing norms affect mask wearing behavior over time?
-
-# Methods
-
-## Ethical approval
-
-This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
-
-## Participants and sampling
-
-Using the platform Prolific (https://www.prolific.co/), we distributed surveys to a representative sample of adults from the United States (_n_ = `r nrow(d)`, _M_~age~ = `r round(mean(d$Age.1, na.rm = TRUE))` years, `r round(mean(d$Ethnicity.1 == "White or Caucasian", na.rm = TRUE)*100)`% White, `r round(mean(d$Sex.1 == 2, na.rm = TRUE)*100)`% Women; see Supplementary Figure \@ref(fig:plotUSMap) for geographic distribution).
-From September 2020 to February 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
-This resulted in 15 unique time points of data collection throughout the pandemic.
-The first 12 time points were distributed monthly, while the last three time points were distributed every two months.
-`r sum(!is.na(d$StartDate.2))` of the initial `r nrow(d)` participants returned to complete the survey at Time 2, while `r sum(!is.na(d$StartDate.15))` participants continued through to Time 15 (see Supplementary Figure \@ref(fig:plotAttrition) for attrition rates across all time points).
-
-## Measures
-
-### Self-reported mask wearing behavior
-
-At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing behavior by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
-
-### Perceived descriptive and injunctive social norms
-
-In 8 of the 15 time points (Time 2, 3, 5, 9, 11, 13, 14, and 15), we asked questions about perceived descriptive and injunctive mask wearing norms.
-
-Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
-We measured perceived descriptive social norms as the mean average of the following two items: "_What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?_" and "_What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?_" These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
-
-Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules.
-We measured perceived injunctive social norms as the mean average of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
-
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items.
-We asked participants whether each of the four items informed them about what people _are_ doing or what people _should_ be doing (i.e., giving descriptive or injunctive information).
-Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \@ref(tab:itemTable)).
-
-### Additional control variables
-
-To account for potential confounding in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \@ref(fig:plotDAG)).
-In this causal model, we included two additional variables that could plausibly act as common causes of perceived social norms and mask wearing behavior: county-level COVID-19 prevalence and county-level political leaning.
-If COVID-19 is prevalent in a participant’s county, then there might be more mask wearing overall, from people in participants’ local area, respected individuals, and from participants themselves, causing increases in both self-reported mask wearing behavior and perceived social norms.
-Similarly, left-leaning counties might have increased levels of self-reported mask wearing behavior and perceived social norms, as the political context affects the proportion of individuals following COVID-19 protective measures [@BaxterKing2022].
-It is therefore important to statistically control for these confounding variables.
-
-We retrieved data on county-level COVID-19 prevalence from a CDC public health surveillance dataset [@CDC2022a].
-In the CDC dataset, COVID-19 prevalence was operationalized as the total new COVID-19 cases per 100,000 population in counties over the past 7 days.
-Since this dataset was updated regularly starting from February 2022, and so each county had multiple entries, we averaged and logged this variable within each county to get the log average weekly new COVID-19 cases per 100,000 population.
-We retrieved data on county-level political leaning from the MIT Election Lab [@MITElectionLab].
-For each county, the dataset stated how many votes the Republican, Democratic, Green Party, and other candidates received in the 2020 Presidential election.
-We used this information to determine which candidate won the vote in that county.
-
-Participants reported their zip code rather than their county, and in a minority of cases the reported zip codes straddled multiple counties.
-In these cases, we averaged COVID-19 prevalence rates over the different counties within the zip codes, and used the political candidate that won in the majority of overlapping counties.
-
-## Statistical analysis
-
-To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models.
-First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
-These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-Third, to compare the perceived strength of descriptive and injunctive norms, we regressed perceived norm strength on the type of norm, including random intercepts and slopes for participants and time points.
-We estimated these models using the _lme4_ R package [@Bates2015] and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between perceived descriptive norms, perceived injunctive norms, and mask wearing behavior simultaneously over time, we used a three-variable random-intercept cross-lagged panel model to analyze our longitudinal data [@Hamaker2015].
-This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time~t-1~ predicts being above average in another variable at time~t~.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015; @Granger1969].
-We estimated this model using the _lavaan_ R package [@Rosseel2012].
-We restricted the analysis to the eight time points with available data for all three variables, and used full information maximum likelihood estimation to deal with any other missing data across time points.
-Based on our causal model (see Supplementary Figure \@ref(fig:plotDAG)), we also included county-level COVID-19 prevalence and county-level political leaning variables as time-invariant controls.
-Due to varying time intervals between time points, we did not constrain any parameters to equality over time.
-
-All analyses were conducted in R v4.1.1 [@RCoreTeam].
-Visualizations were generated using the _cowplot_ [@Wilke2020] and _ggplot2_ [@Wickham2016] packages.
-The manuscript was reproducibly generated using the _targets_ [@Landau2021] and _papaja_ [@Aust2022] packages.
 
@@
 (ref:plotTimelineCaption) _Timeline of self-reported mask wearing behavior and perceived social norms in the United States throughout the COVID-19 pandemic._ (a) Points and line ranges indicate mean averages and standard errors for the self-reported mask wearing item.
-This item was measured across all fifteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
+This item was measured across all eighteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
 (b) Points and line ranges indicate mean averages and standard errors for perceived descriptive mask wearing norms (blue) and perceived injunctive mask wearing norms (green).
-These items were measured across eight time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
+These items were measured across eleven time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
 (c) Smoothed data for new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; https://ourworldindata.org/).
@@
 
-```{r plotTimeline, echo=F, fig.width=5, fig.height=6, fig.cap="(ref:plotTimelineCaption)"}
+```{r plotTimeline, echo=F, fig.width=6, fig.height=6, fig.cap="(ref:plotTimelineCaption)"}
 plot3
@@
 
-Three main observations can be made about the emergence and stability of social norms from these visualizations.
+Two main observations can be made about the emergence and stability of social norms from these visualizations.
 First, social norms and behavior were tightly coupled over time.
@@
 Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021.
+Finally, perceived social norms and mask wearing declined again after the CDC eased restrictions in March 2022.
 Impressively, shifts in mask wearing behavior and norms happened on the timescale of weeks.
@@
 
-Third, across all eight time points with available data, injunctive mask wearing norms were consistently rated as stronger than descriptive mask wearing norms.
-Multilevel regression modeling revealed that, on average across individuals and time points, injunctive norms were rated as 0.6 points stronger than descriptive norms on a 7-point Likert scale (_b_ = `r printnum(fixef(m3.1)["varInjNorms"])`, 95% CI [`r printnum(conf3.1["varInjNorms","2.5 %"])` `r printnum(conf3.1["varInjNorms","97.5 %"])`]; Supplementary Figure \@ref(fig:plotNormCompare); Supplementary Table \@ref(tab:modelSummaryTable2)).
-In other words, participants in our study perceived the general encouragement of mask wearing and mask wearing behavior among trusted people to be higher than perceived mask wearing among people in their local area.
+Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing behavior and perceived social norms over time.
+To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
+This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables and controls: perceived descriptive mask wearing norms, perceived injunctive mask wearing norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior (see Supplementary Figure \@ref(fig:plotDAG) for justification of control variables).
+According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).
+Since we are primarily interested in the causal effects of social norms on behavior, we focus on these variables in what follows (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
 
-Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing behavior and perceived social norms over time.
-To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted an eight-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our three main variables: self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms (see Supplementary Table \@ref(tab:lavaanTable) for estimated model parameters).
-According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).
+```{r, echo = FALSE}
+sc <- standardizedSolution(riclpm)
+pe <- parameterEstimates(riclpm)
+```
 
 Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing behavior and perceived social norms.
-On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(standardizedSolution(riclpm)[187,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[187,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[187,"ci.upper"])`], _p_ `r printp(parameterEstimates(riclpm)[187,"pvalue"])`) and stronger injunctive mask wearing norms (_r_ = `r printnum(standardizedSolution(riclpm)[188,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[188,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[188,"ci.upper"])`], _p_ `r printp(parameterEstimates(riclpm)[188,"pvalue"])`).
-Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (_r_ = `r printnum(standardizedSolution(riclpm)[189,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[189,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[189,"ci.upper"])`], _p_ `r printp(parameterEstimates(riclpm)[189,"pvalue"])`).
+On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`) and stronger injunctive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri2","pvalue"])`).
+Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (_r_ = `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri2" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`).
 
-Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for all three variables across the study duration.
+Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing behavior across the study duration, controlling for non-social beliefs and political orientation.
 In random intercept cross-lagged panel models, autoregressive effects represent "persistence" or "inertia" in within-person fluctuations from stable trait levels.
@@
 
-(ref:plotRICLPMCaption) _Results of eight-wave unconstrained random-intercept cross-lagged panel model._ Arrows represent within-person autoregressive and cross-lagged effects from the model, controlling for time invariant predictors and partitioning out stable between-person individual differences.
+(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Arrows represent within-person autoregressive and cross-lagged effects from the model, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
@@
 
-```{r plotRICLPM, echo=F, fig.width=7, fig.height=3, fig.cap="(ref:plotRICLPMCaption)"}
-plot8
+```{r plotRICLPM, echo=F, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotRICLPMCaption)"}
+plot7
 ```
 
-Earlier in the pandemic, we see less evidence of autoregressive and cross-lagged effects.
-Within-person fluctuations in descriptive and injunctive norms persist across only a few time points, and there is only one cross-lagged effect of mask wearing on future descriptive norms.
-However, once COVID-19 case numbers start rising again in June 2021, we begin to see more evidence of autoregressive and cross-lagged effects.
-From this point onwards, all three variables show evidence of persistence, with within-person fluctuations from trait levels predicting later within-person fluctuations in the same measures.
-But more importantly, several positive and significant cross-lagged effects emerge.
-On three occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior.
-According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (Time 9, $\beta$ = `r printnum(standardizedSolution(riclpm)[108,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[108,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[108,"ci.upper"])`], _p_ = `r printp(parameterEstimates(riclpm)[108,"pvalue"])`; Time 11, $\beta$ = `r printnum(standardizedSolution(riclpm)[111,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[111,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[111,"ci.upper"])`], _p_ = `r printp(parameterEstimates(riclpm)[111,"pvalue"])`; Time 13, $\beta$ = `r printnum(standardizedSolution(riclpm)[114,"est.std"])`, 95% CI [`r printnum(standardizedSolution(riclpm)[114,"ci.lower"])` `r printnum(standardizedSolution(riclpm)[114,"ci.upper"])`], _p_ = `r printp(parameterEstimates(riclpm)[114,"pvalue"])`).
+In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing behavior.
+On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior.
+According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (Time 2, $\beta$ = `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.lower"])` `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_05" & pe$op == "~" & pe$rhs == "w3_02","pvalue"])`; Time 5, $\beta$ = `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.lower"])` `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_09" & pe$op == "~" & pe$rhs == "w3_05","pvalue"])`; Time 11, $\beta$ = `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.lower"])` `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_13" & pe$op == "~" & pe$rhs == "w3_11","pvalue"])`; Time 13, $\beta$ = `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.lower"])` `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_14" & pe$op == "~" & pe$rhs == "w3_13","pvalue"])`).
+These cross-lagged effects from descriptive norms to mask wearing diminish in 2022.
 We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms.
@@
 
-However, the model reveals that, after controlling for perceived descriptive norms, within-person changes in perceived injunctive norms were unrelated to within-person changes in mask wearing behavior across the entire pandemic.
-All cross-lagged effects between the two variables, both from perceived injunctive norms to mask wearing behavior and vice versa, are non-significant.
+However, the model reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did not predict future within-person changes in mask wearing behavior across the entire pandemic.
+All cross-lagged effects from perceived injunctive norms to mask wearing behavior are non-significant.
 Any causal effect that perceived injunctive norms might have had on future mask wearing behavior appears to be fully mediated by perceived descriptive norms.
 For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing behavior.
-But aside from these indirect effects, perceived injunctive norms and mask wearing behavior were not directly causally related over time within individuals.
+But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing behavior over time within individuals.
 
@@
 The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior are tightly coupled, and both change quickly in response to recommendations from public health authorities.
-The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior from June 2021 onwards.
-While injunctive norms were perceived as stronger than descriptive norms, we found that  injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.
+The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior at the beginning of the pandemic.
+By contrast, injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.
 
@@
 
-We also found that injunctive norms were, on average, perceived as stronger than descriptive norms.
-This aligns with previous research that measured perceived injunctive and descriptive norms in the contexts of drinking [@Park2009] and marijuana use [@Montes2021].
-What people ought to do is often perceived as stronger than what people _actually_ do.
-Since we operationalized injunctive norms in part by whether trusted individuals in the media were wearing masks, it is possible that our sample is accurately reporting that people in general were less adherent to mask wearing policies than government officials, newscasters, and other respected individuals.
-It may also have been easier for our participants to notice instances of non-compliance with mask wearing guidelines among the general population, simply because they have more opportunities to observe people’s behavior in public settings.
-
-Despite perceiving stronger injunctive norms, it was _descriptive_ norms, not injunctive norms, that independently predicted future increases in mask wearing.
+In our longitudinal analysis, we found that _descriptive_ norms, not injunctive norms, independently predicted future increases in mask wearing.
 In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects [@Rudert2021].
@@
 During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms [@Hogg2008; @Wellen1998].
-Supporting this situational-uncertainty explanation, our model revealed that descriptive norms only began to predict future mask wearing after the CDC flipped their mask wearing recommendation in March 2021.
-With changing recommendations from authorities increasing levels of uncertainty, people instead began to look to their neighbors and adapted their mask wearing behavior accordingly.
+Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its height.
+Unsure of how to respond to a novel threatening situation, people began to look to their neighbors and adapted their mask wearing behavior accordingly.
+However, as the uncertainty surrounding the pandemic lessened in 2022, reliance on descriptive norms appeared to decrease.
 
 Our finding that injunctive norms do not predict future mask wearing behavior is at odds with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults [@Macy2021].
-One possible explanation for these conflicting findings is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior [@Cialdini1991].
+One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous correlations.
+Another explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior [@Cialdini1991].
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people’s behaviors, increasing the relative salience of injunctive norms.
@@
 Our work thus underscores the importance of consistent, visible community adherence for encouraging personal protective behaviors in response to global pandemics like COVID-19.
+
+# Methods
+
+## Ethical approval
+
+This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
+
+## Participants and sampling
+
+Using the platform Prolific (https://www.prolific.co/), we distributed surveys to a representative sample of adults from the United States (_n_ = `r nrow(d)`, _M_~age~ = `r round(mean(d$Age.1, na.rm = TRUE))` years, `r round(mean(d$Ethnicity.1 == "White or Caucasian", na.rm = TRUE)*100)`% White, `r round(mean(d$Sex.1 == 2, na.rm = TRUE)*100)`% Women; see Supplementary Figure \@ref(fig:plotUSMap) for geographic distribution).
+From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
+This resulted in 18 unique time points of data collection throughout the pandemic.
+The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
+`r sum(!is.na(d$StartDate.2))` of the initial `r nrow(d)` participants returned to complete the survey at Time 2, while `r sum(!is.na(d$StartDate.18))` participants continued through to Time 18 (see Supplementary Figure \@ref(fig:plotAttrition) for attrition rates across all time points).
+
+## Measures
+
+### Self-reported mask wearing behavior
+
+At every time point, participants were asked about the number of in-person interactions they had in the last week.
+Following this question, participants self-reported their mask wearing behavior by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+
+### Perceived descriptive and injunctive social norms
+
+In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
+
+Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
+We measured perceived descriptive social norms as the mean average of the following two items: "_What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?_" and "_What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?_" These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
+
+Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules.
+We measured perceived injunctive social norms as the mean average of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
+
+To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items.
+We asked participants whether each of the four items informed them about what people _are_ doing or what people _should_ be doing (i.e., giving descriptive or injunctive information).
+Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \@ref(tab:itemTable)).
+
+### Additional control variables
+
+To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \@ref(fig:plotDAG)).
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research [@Bicchieri2014]: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the morally right thing to do).
+These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing behavior.
+In addition, we also included political orientation as a common cause of all other variables.
+This is justified by evidence showing that mask wearing was heavily politicized in the United States throughout the pandemic [@BaxterKing2022].
+Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
+
+Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
+Factual beliefs were measured as the mean average of the following two items: "_I wear a face mask when going out in public to keep myself from getting sick_" and "_I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet_".
+Personal normative beliefs were measured with a single item: "_Wearing a face mask when going out in public is the right thing to do_".
+These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
+
+Political orientation was measured in the first time point only.
+We measured political orientation as the mean average of the following two items: "_How would you describe your political orientation with regard to social issues?_" and "_How would you describe your political orientation with regard to economic issues?_".
+These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
+
+## Statistical analysis
+
+To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models.
+First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
+Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
+These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
+We estimated these models using the _lme4_ R package [@Bates2015] and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data [@Hamaker2015].
+This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time~t-1~ predicts being above average in another variable at time~t~.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015; @Granger1969].
+
+We estimated the random-intercept cross-lagged panel model using the _lavaan_ R package [@Rosseel2012].
+In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included five main variables in the model: perceived descriptive norms, perceived injunctive norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior.
+For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
+We modelled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We restricted the analysis to the ten time points with available data for all five variables.
+Full information maximum likelihood estimation was used to deal with missing data.
+
+All analyses were conducted in R v4.1.1 [@RCoreTeam].
+Visualizations were generated using the _cowplot_ [@Wilke2020] and _ggplot2_ [@Wickham2016] packages.
+The manuscript was reproducibly generated using the _targets_ [@Landau2021] and _papaja_ [@Aust2022] packages.
 
@@
 \begingroup
-\setlength{\parindent}{-1in}
-\setlength{\leftskip}{0.5in}
 

### 2022-10-13 manuscript.tex
first draft manuscript revision

@@
 \PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
 %
 \documentclass[
-  english,
   man,floatsintext]{apa6}
-\title{Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic}
-\author{Samantha L. Heiman\textsuperscript{$\dagger{}$,1}, Scott Claessens\textsuperscript{$\dagger{}$,2}, Edward R. Hurt\textsuperscript{1}, \& Peter M. Todd\textsuperscript{1,3}}
-\date{}
-
 \usepackage{amsmath,amssymb}
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic},
-  pdfauthor={Samantha L. Heiman,1, Scott Claessens,2, Edward R. Hurt1, \& Peter M. Todd1,3},
-  pdflang={en-EN},
-  pdfkeywords={descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation},
-  colorlinks=true,
-  linkcolor={Maroon},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={blue},
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 % Manuscript styling
@@
 {}{\PatchFailed}
-\keywords{descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 4826 words}
+\keywords{descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: xxxx words}
 \usepackage{lineno}
@@
 \captionsetup[figure]{font={stretch=1,small}}
-\ifXeTeX
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
 \ifLuaTeX
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic},
+  pdfauthor={Samantha L. Heiman,1, Scott Claessens,2, Edward R. Hurt1, \& Peter M. Todd1,3},
+  pdflang={en-EN},
+  pdfkeywords={descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+
+\title{Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic}
+\author{Samantha L. Heiman\textsuperscript{$\dagger{}$,1}, Scott Claessens\textsuperscript{$\dagger{}$,2}, Edward R. Hurt\textsuperscript{1}, \& Peter M. Todd\textsuperscript{1,3}}
+\date{}
 
@@
 To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
-Leveraging a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; \emph{n} = 915), we tracked people's reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
+Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 915), we tracked people's reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
 Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior are tightly coupled and both change quickly in response to recommendations from public health authorities.
@@
 Social norms are a key aspect of human sociality.
-Broadly defined as behavioral guidelines enforced by groups of people without the law (van Kleef et al., 2019), social norms have attracted the attention of psychologists for decades.
-Since Asch's (1956) early studies of normative conformity, evidence has accrued that humans are particularly attuned to social norms (Tomasello, 2014).
-From a young age, children begin to adhere to and enforce group-wide social norms (Jensen et al., 2014; Rakoczy \& Schmidt, 2013; Schmidt \& Tomasello, 2012) and both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms (Schaumberg \& Skowronek, 2022; Vaish et al., 2011).
-This uniquely human sensitivity to social norms allows groups of unrelated people to cooperate and coordinate in the face of existential threats, such as resource scarcity, natural disasters, and infectious diseases (Gelfand et al., 2011; Henrich, 2015).
-
-Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms (Cialdini et al., 1991; Deutsch \& Gerard, 1955).
-Injunctive social norms are societal, indicating what others in the group tend to approve or disapprove of, and often involve social sanctions if violated.
+Broadly defined as commonly known behavioral guidelines enforced by groups of people, social norms have attracted the attention of behavioral scientists for decades (1--5).
+Since Asch's early studies of normative conformity (6), evidence has accrued that humans are particularly attuned to social norms (7).
+From a young age, children begin to adhere to and enforce group-wide social norms (8--10) and both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms (11, 12).
+This uniquely human sensitivity to social norms allows groups of unrelated people to cooperate and coordinate in the face of existential threats, such as resource scarcity, natural disasters, and infectious diseases (13, 14).
+
+Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms (1, 2, 15).
+Injunctive social norms indicate what others in the group tend to approve or disapprove of and often involve social sanctions if violated.
 By contrast, descriptive norms are situational, simply describing what most people are doing in a given situation.
@@
 However, if an individual observes that most people are leaving their litter behind at the site, the descriptive norm is to not clean up.
-It is thus possible for injunctive and descriptive norms to have independent effects on behavior (Schultz et al., 2007).
-
-Despite decades of research on the causes and consequences of injunctive and descriptive norms (Cialdini et al., 1991; Cialdini, 2007; Cialdini \& Jacobson, 2021; Schultz et al., 2007), social norms have remained an elusive concept in the behavioral sciences and several open questions remain (Fehr \& Schurtenberger, 2018a; van Kleef et al., 2019).
+It is thus possible for injunctive and descriptive norms to have independent effects on behavior (16).
+
+Despite decades of research on the causes and consequences of injunctive and descriptive norms (1, 2, 16--18), social norms have remained an elusive concept in the behavioral sciences and several open questions remain (3, 5).
 In the current work, we focus on two such questions.
@@
 With regards to norm emergence, research in cultural evolution and behavioral economics have begun to illuminate how social norms emerge over time.
-In the longer term, cultural evolutionary models show that injunctive social norms can be vertically transmitted down through generations via imitation or teaching, or horizontally diffused from neighboring human populations (Boyd, 1985; Henrich, 2015).
-For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership (Kushnick et al., 2014) and post-marital residence (Jordan et al., 2009).
+In the longer term, cultural evolutionary models show that injunctive social norms can be vertically transmitted down through generations via imitation or teaching, or horizontally diffused from neighboring human populations (14, 19).
+For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership (20) and post-marital residence (21).
 However, much less is known about how social norms arise endogenously within populations in the shorter term.
-Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (Titlestad et al., 2019) and require peer enforcement to become stable (Fehr \& Schurtenberger, 2018b).
+Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (22) and require peer enforcement to become stable (23).
 But it remains unclear to what extent these findings generalize beyond the laboratory to real human populations.
@@
 With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the behavioral impact of social norms.
-For example, field experiments have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (Nigbur et al., 2010), paying taxes (Larkin et al., 2018), and sustainably reusing towels in hotels (Goldstein et al., 2008).
-Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals' attention on injunctive norms (Schultz et al., 2007).
-However, previous studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population.
+For example, field experiments have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (24), paying taxes (25), and sustainably reusing towels in hotels (26).
+Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals' attention on injunctive norms (16).
+However, these cross-sectional studies have two main limitations.
+First, studies have not adequately controlled for other potential non-social influences on behavior, such as factual beliefs and personal normative beliefs (27).
+For example, descriptive social norms might only affect cooperative behavior indirectly by influencing perceptions of the effectiveness of the behavior (i.e., factual beliefs) or individuals' moral preferences (i.e., personal normative beliefs).
+Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population.
 An alternative way to identify causality, whilst retaining ecological validity, is to follow perceptions of social norms and cooperative behavior over time amidst a real, unfolding social dilemma.
-This is particularly important because social norms are not static: they change dynamically over time through processes of deliberation and social interaction (Smith et al., 2015).
+This is particularly important because social norms are not static: they change dynamically over time through processes of deliberation and social interaction (28).
 
@@
 Before the pandemic, mask wearing was not a common behavior in the United States.
-In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a measure that people should take to minimize the spread of COVID-19.
+In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a cooperative behavior that people should adopt to minimize the spread of COVID-19.
 But mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., ``flattening the curve'' to protect at-risk individuals).
@@
 Recent research has found positive relationships between perceptions of social norms and protective COVID-19 behaviors.
-In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (Macy et al., 2021), and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions (Bokemper et al., 2021).
-In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (Rudert \& Janke, 2021).
-These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between between-person and within-person change over time (Hamaker et al., 2015), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
-
-Here, we use a year and a half of longitudinal data from a representative sample of adults in the United States (15 time points; \emph{n} = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
+In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (29), and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions (30).
+In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (31).
+These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between between-person and within-person change over time (32), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
+These studies also do not control for the influences of non-social beliefs, such as factual beliefs and personal normative beliefs (27).
+
+Here, we use two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
 We aimed to answer two main research questions.
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{ethical-approval}{%
-\subsection{Ethical approval}\label{ethical-approval}}
-
-This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
-
-\hypertarget{participants-and-sampling}{%
-\subsection{Participants and sampling}\label{participants-and-sampling}}
-
-Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution).
-From September 2020 to February 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
-This resulted in 15 unique time points of data collection throughout the pandemic.
-The first 12 time points were distributed monthly, while the last three time points were distributed every two months.
-634 of the initial 915 participants returned to complete the survey at Time 2, while 379 participants continued through to Time 15 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
-
-\hypertarget{measures}{%
-\subsection{Measures}\label{measures}}
-
-\hypertarget{self-reported-mask-wearing-behavior}{%
-\subsubsection{Self-reported mask wearing behavior}\label{self-reported-mask-wearing-behavior}}
-
-At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing behavior by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
-
-\hypertarget{perceived-descriptive-and-injunctive-social-norms}{%
-\subsubsection{Perceived descriptive and injunctive social norms}\label{perceived-descriptive-and-injunctive-social-norms}}
-
-In 8 of the 15 time points (Time 2, 3, 5, 9, 11, 13, 14, and 15), we asked questions about perceived descriptive and injunctive mask wearing norms.
-
-Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
-We measured perceived descriptive social norms as the mean average of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
-
-Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules.
-We measured perceived injunctive social norms as the mean average of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
-
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items.
-We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
-Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).
-
-\hypertarget{additional-control-variables}{%
-\subsubsection{Additional control variables}\label{additional-control-variables}}
-
-To account for potential confounding in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
-In this causal model, we included two additional variables that could plausibly act as common causes of perceived social norms and mask wearing behavior: county-level COVID-19 prevalence and county-level political leaning.
-If COVID-19 is prevalent in a participant's county, then there might be more mask wearing overall, from people in participants' local area, respected individuals, and from participants themselves, causing increases in both self-reported mask wearing behavior and perceived social norms.
-Similarly, left-leaning counties might have increased levels of self-reported mask wearing behavior and perceived social norms, as the political context affects the proportion of individuals following COVID-19 protective measures (Baxter-King et al., 2022).
-It is therefore important to statistically control for these confounding variables.
-
-We retrieved data on county-level COVID-19 prevalence from a CDC public health surveillance dataset (Centers for Disease Control and Prevention, 2022b).
-In the CDC dataset, COVID-19 prevalence was operationalized as the total new COVID-19 cases per 100,000 population in counties over the past 7 days.
-Since this dataset was updated regularly starting from February 2022, and so each county had multiple entries, we averaged and logged this variable within each county to get the log average weekly new COVID-19 cases per 100,000 population.
-We retrieved data on county-level political leaning from the MIT Election Lab (MIT Election Data and Science Lab, 2018).
-For each county, the dataset stated how many votes the Republican, Democratic, Green Party, and other candidates received in the 2020 Presidential election.
-We used this information to determine which candidate won the vote in that county.
-
-Participants reported their zip code rather than their county, and in a minority of cases the reported zip codes straddled multiple counties.
-In these cases, we averaged COVID-19 prevalence rates over the different counties within the zip codes, and used the political candidate that won in the majority of overlapping counties.
-
-\hypertarget{statistical-analysis}{%
-\subsection{Statistical analysis}\label{statistical-analysis}}
-
-To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models.
-First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
-These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-Third, to compare the perceived strength of descriptive and injunctive norms, we regressed perceived norm strength on the type of norm, including random intercepts and slopes for participants and time points.
-We estimated these models using the \emph{lme4} R package (Bates et al., 2015) and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between perceived descriptive norms, perceived injunctive norms, and mask wearing behavior simultaneously over time, we used a three-variable random-intercept cross-lagged panel model to analyze our longitudinal data (Hamaker et al., 2015).
-This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time\textsubscript{t-1} predicts being above average in another variable at time\textsubscript{t}.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets (Granger, 1969; Hamaker et al., 2015).
-We estimated this model using the \emph{lavaan} R package (Rosseel, 2012).
-We restricted the analysis to the eight time points with available data for all three variables, and used full information maximum likelihood estimation to deal with any other missing data across time points.
-Based on our causal model (see Supplementary Figure \ref{fig:plotDAG}), we also included county-level COVID-19 prevalence and county-level political leaning variables as time-invariant controls.
-Due to varying time intervals between time points, we did not constrain any parameters to equality over time.
-
-All analyses were conducted in R v4.1.1 (R Core Team, 2022).
-Visualizations were generated using the \emph{cowplot} (Wilke, 2020) and \emph{ggplot2} (Wickham, 2016) packages.
-The manuscript was reproducibly generated using the \emph{targets} (Landau, 2021) and \emph{papaja} (Aust \& Barth, 2022) packages.
-
 \hypertarget{results}{%
@@
 Figure \ref{fig:plotTimeline} plots self-reported mask wearing behavior and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum's COVID-19 Timeline (Centers for Disease Control and Prevention, 2022a) and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).
+These events were obtained from the CDC Museum's COVID-19 Timeline (33) and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).
 
@@
 \caption{\label{fig:plotTimeline}\emph{Timeline of self-reported mask wearing behavior and perceived social norms in the United States throughout the COVID-19 pandemic.} (a) Points and line ranges indicate mean averages and standard errors for the self-reported mask wearing item.
-This item was measured across all fifteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
+This item was measured across all eighteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
 (b) Points and line ranges indicate mean averages and standard errors for perceived descriptive mask wearing norms (blue) and perceived injunctive mask wearing norms (green).
-These items were measured across eight time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
+These items were measured across eleven time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
 (c) Smoothed data for new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; \url{https://ourworldindata.org/}).
@@
 Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, though to lower levels than before.
-In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (\emph{b} = 0.25, 95\% confidence interval {[}0.20 0.30{]}) and between mask wearing behavior and perceived injunctive mask wearing norms (\emph{b} = 0.25, 95\% CI {[}0.21 0.30{]}) across individuals and time points (Supplementary Figure \ref{fig:plotCorBehNorm}; Supplementary Table \ref{tab:modelSummaryTable1}).
+In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (\emph{b} = 0.29, 95\% confidence interval {[}0.23 0.35{]}) and between mask wearing behavior and perceived injunctive mask wearing norms (\emph{b} = 0.26, 95\% CI {[}0.22 0.30{]}) across individuals and time points (Supplementary Figure \ref{fig:plotCorBehNorm}; Supplementary Table \ref{tab:modelSummaryTable1}).
 
@@
 
-Third, across all eight time points with available data, injunctive mask wearing norms were consistently rated as stronger than descriptive mask wearing norms.
-Multilevel regression modeling revealed that, on average across individuals and time points, injunctive norms were rated as 0.6 points stronger than descriptive norms on a 7-point Likert scale (\emph{b} = 0.60, 95\% CI {[}0.51 0.69{]}; Supplementary Figure \ref{fig:plotNormCompare}; Supplementary Table \ref{tab:modelSummaryTable2}).
-In other words, participants in our study perceived the general encouragement of mask wearing and mask wearing behavior among trusted people to be higher than perceived mask wearing among people in their local area.
-
 Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing behavior and perceived social norms over time.
-To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted an eight-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our three main variables: self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms (see Supplementary Table \ref{tab:lavaanTable} for estimated model parameters).
-According to established fit statistics, this model fitted the data well (root mean square error of approximation = 0.037, 95\% CI {[}0.032 0.042{]}; standardized root mean squared residual = 0.056; comparative fit index = 0.961).
+To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
+This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables and controls: perceived descriptive mask wearing norms, perceived injunctive mask wearing norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior (see Supplementary Figure \ref{fig:plotDAG} for justification of control variables).
+According to established fit statistics, this model fitted the data well (root mean square error of approximation = 0.030, 95\% CI {[}0.028 0.033{]}; standardized root mean squared residual = 0.087; comparative fit index = 0.957).
+Since we are primarily interested in the causal effects of social norms on behavior, we focus on these variables in what follows (but see Supplementary Table \ref{tab:lavaanTable} for full list of estimated autoregressive and cross-lagged effects).
 
 Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing behavior and perceived social norms.
-On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (\emph{r} = 0.45, 95\% CI {[}0.35 0.54{]}, \emph{p} \textless{} .001) and stronger injunctive mask wearing norms (\emph{r} = 0.29, 95\% CI {[}0.18 0.40{]}, \emph{p} \textless{} .001).
-Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (\emph{r} = 0.65, 95\% CI {[}0.58 0.71{]}, \emph{p} \textless{} .001).
-
-Regarding within-person dynamics over time, Figure \ref{fig:plotRICLPM} displays autoregressive and cross-lagged effects for all three variables across the study duration.
+On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (\emph{r} = 0.19, 95\% CI {[}0.04 0.33{]}, \emph{p} = .019) and stronger injunctive mask wearing norms (\emph{r} = 0.27, 95\% CI {[}0.14 0.40{]}, \emph{p} \textless{} .001).
+Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (\emph{r} = 0.71, 95\% CI {[}0.65 0.78{]}, \emph{p} \textless{} .001).
+
+Regarding within-person dynamics over time, Figure \ref{fig:plotRICLPM} displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing behavior across the study duration, controlling for non-social beliefs and political orientation.
 In random intercept cross-lagged panel models, autoregressive effects represent ``persistence'' or ``inertia'' in within-person fluctuations from stable trait levels.
@@
 \includegraphics{manuscript_files/figure-latex/plotRICLPM-1.pdf}
-\caption{\label{fig:plotRICLPM}\emph{Results of eight-wave unconstrained random-intercept cross-lagged panel model.} Arrows represent within-person autoregressive and cross-lagged effects from the model, controlling for time invariant predictors and partitioning out stable between-person individual differences.
+\caption{\label{fig:plotRICLPM}\emph{Results of ten-wave unconstrained random-intercept cross-lagged panel model.} Arrows represent within-person autoregressive and cross-lagged effects from the model, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
@@
 
-Earlier in the pandemic, we see less evidence of autoregressive and cross-lagged effects.
-Within-person fluctuations in descriptive and injunctive norms persist across only a few time points, and there is only one cross-lagged effect of mask wearing on future descriptive norms.
-However, once COVID-19 case numbers start rising again in June 2021, we begin to see more evidence of autoregressive and cross-lagged effects.
-From this point onwards, all three variables show evidence of persistence, with within-person fluctuations from trait levels predicting later within-person fluctuations in the same measures.
-But more importantly, several positive and significant cross-lagged effects emerge.
-On three occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior.
-According to recent effect size guidelines for cross-lagged panel models (Orth et al., 2022), the standardized beta coefficients for these cross-lagged effects were large (Time 9, \(\beta\) = 0.17, 95\% CI {[}0.03 0.31{]}, \emph{p} = .020; Time 11, \(\beta\) = 0.21, 95\% CI {[}0.06 0.36{]}, \emph{p} = .006; Time 13, \(\beta\) = 0.16, 95\% CI {[}0.01 0.31{]}, \emph{p} = .037).
+In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing behavior.
+On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior.
+According to recent effect size guidelines for cross-lagged panel models (34), the standardized beta coefficients for these cross-lagged effects were large (Time 2, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; Time 5, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; Time 11, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; Time 13, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
+These cross-lagged effects from descriptive norms to mask wearing diminish in 2022.
 We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms.
@@
 The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior are tightly coupled, and both change quickly in response to recommendations from public health authorities.
-The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior from June 2021 onwards.
-While injunctive norms were perceived as stronger than descriptive norms, we found that injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.
-
-Our finding that social norms and mask wearing behavior are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (Titlestad et al., 2019).
-Moreover, the role of authorities, like the CDC, in shaping social norms and behavior supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (Markus \& Kitayama, 2010).
+The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior at the beginning of the pandemic.
+By contrast, injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.
+
+Our finding that social norms and mask wearing behavior are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (22).
+Moreover, the role of authorities, like the CDC, in shaping social norms and behavior supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (35).
 The CDC explicitly releases guidelines for the benefit of public health, one of many situations in which the function of authorities is to coordinate behavior on large scales.
@@
 
-We also found that injunctive norms were, on average, perceived as stronger than descriptive norms.
-This aligns with previous research that measured perceived injunctive and descriptive norms in the contexts of drinking (Park et al., 2009) and marijuana use (Montes et al., 2021).
-What people ought to do is often perceived as stronger than what people \emph{actually} do.
-Since we operationalized injunctive norms in part by whether trusted individuals in the media were wearing masks, it is possible that our sample is accurately reporting that people in general were less adherent to mask wearing policies than government officials, newscasters, and other respected individuals.
-It may also have been easier for our participants to notice instances of non-compliance with mask wearing guidelines among the general population, simply because they have more opportunities to observe people's behavior in public settings.
-
-Despite perceiving stronger injunctive norms, it was \emph{descriptive} norms, not injunctive norms, that independently predicted future increases in mask wearing.
-In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects (Rudert \& Janke, 2021).
-Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (Bokemper et al., 2021).
+In our longitudinal analysis, we found that \emph{descriptive} norms, not injunctive norms, independently predicted future increases in mask wearing.
+In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects (31).
+Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (30).
 Why this specific effect of descriptive norms on mask wearing?
-Descriptive norms are particularly useful for fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (Gelfand \& Harrington, 2015).
-During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms (Hogg et al., 2008; Wellen et al., 1998).
-Supporting this situational-uncertainty explanation, our model revealed that descriptive norms only began to predict future mask wearing after the CDC flipped their mask wearing recommendation in March 2021.
-With changing recommendations from authorities increasing levels of uncertainty, people instead began to look to their neighbors and adapted their mask wearing behavior accordingly.
-
-Our finding that injunctive norms do not predict future mask wearing behavior is at odds with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults (Macy et al., 2021).
-One possible explanation for these conflicting findings is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior (Cialdini et al., 1991).
+Descriptive norms are particularly useful for fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (36).
+During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms (37, 38).
+Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its height.
+Unsure of how to respond to a novel threatening situation, people began to look to their neighbors and adapted their mask wearing behavior accordingly.
+However, as the uncertainty surrounding the pandemic lessened in 2022, reliance on descriptive norms appeared to decrease.
+
+Our finding that injunctive norms do not predict future mask wearing behavior is at odds with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults (29).
+One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous correlations.
+Another explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior (2).
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people's behaviors, increasing the relative salience of injunctive norms.
@@
 This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, different ethnicities, or those with different political ideologies.
-Since injunctive norm strength varies based on which group is seen as the source of the norm (Neighbors et al., 2008), it would be interesting to learn whether different groups have different patterns of norm emergence over time.
-In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors (Baxter-King et al., 2022).
-Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large (Remedios, 2022).
+Since injunctive norm strength varies based on which group is seen as the source of the norm (39), it would be interesting to learn whether different groups have different patterns of norm emergence over time.
+In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors (40).
+Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large (41).
 Our results also might not generalize to all social norms, behaviors, and social dilemmas.
@@
 Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that \emph{low} local levels of mask wearing \emph{discouraged} future personal mask use.
-This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community (Natanson, 2022).
+This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community (42).
 Our work thus underscores the importance of consistent, visible community adherence for encouraging personal protective behaviors in response to global pandemics like COVID-19.
+
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+\hypertarget{ethical-approval}{%
+\subsection{Ethical approval}\label{ethical-approval}}
+
+This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
+
+\hypertarget{participants-and-sampling}{%
+\subsection{Participants and sampling}\label{participants-and-sampling}}
+
+Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution).
+From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
+This resulted in 18 unique time points of data collection throughout the pandemic.
+The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
+634 of the initial 915 participants returned to complete the survey at Time 2, while 347 participants continued through to Time 18 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
+
+\hypertarget{measures}{%
+\subsection{Measures}\label{measures}}
+
+\hypertarget{self-reported-mask-wearing-behavior}{%
+\subsubsection{Self-reported mask wearing behavior}\label{self-reported-mask-wearing-behavior}}
+
+At every time point, participants were asked about the number of in-person interactions they had in the last week.
+Following this question, participants self-reported their mask wearing behavior by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+
+\hypertarget{perceived-descriptive-and-injunctive-social-norms}{%
+\subsubsection{Perceived descriptive and injunctive social norms}\label{perceived-descriptive-and-injunctive-social-norms}}
+
+In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
+
+Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
+We measured perceived descriptive social norms as the mean average of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
+
+Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules.
+We measured perceived injunctive social norms as the mean average of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
+
+To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items.
+We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
+Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).
+
+\hypertarget{additional-control-variables}{%
+\subsubsection{Additional control variables}\label{additional-control-variables}}
+
+To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research (27): factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the morally right thing to do).
+These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing behavior.
+In addition, we also included political orientation as a common cause of all other variables.
+This is justified by evidence showing that mask wearing was heavily politicized in the United States throughout the pandemic (40).
+Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
+
+Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
+Factual beliefs were measured as the mean average of the following two items: ``\emph{I wear a face mask when going out in public to keep myself from getting sick}'' and ``\emph{I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet}''.
+Personal normative beliefs were measured with a single item: ``\emph{Wearing a face mask when going out in public is the right thing to do}''.
+These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
+
+Political orientation was measured in the first time point only.
+We measured political orientation as the mean average of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
+These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
+
+\hypertarget{statistical-analysis}{%
+\subsection{Statistical analysis}\label{statistical-analysis}}
+
+To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models.
+First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
+Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
+These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
+We estimated these models using the \emph{lme4} R package (43) and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data (32).
+This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time\textsubscript{t-1} predicts being above average in another variable at time\textsubscript{t}.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets (32, 44).
+
+We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package (45).
+In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included five main variables in the model: perceived descriptive norms, perceived injunctive norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior.
+For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
+We modelled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We restricted the analysis to the ten time points with available data for all five variables.
+Full information maximum likelihood estimation was used to deal with missing data.
+
+All analyses were conducted in R v4.1.1 (46).
+Visualizations were generated using the \emph{cowplot} (47) and \emph{ggplot2} (48) packages.
+The manuscript was reproducibly generated using the \emph{targets} (49) and \emph{papaja} (50) packages.
 
@@
 \begingroup
-\setlength{\parindent}{-1in}
-\setlength{\leftskip}{0.5in}
 
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
+\begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2009}{}}%
+\CSLLeftMargin{1. }%
+\CSLRightInline{C. Bicchieri, E. Xiao, \href{https://doi.org/10.1002/bdm.621}{Do the right thing: But only if others do so}. \emph{Journal of Behavioral Decision Making} \textbf{22}, 191--208 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Cialdini1991}{}}%
+\CSLLeftMargin{2. }%
+\CSLRightInline{R.
+B.
+Cialdini, C.
+A.
+Kallgren, R.
+R.
+Reno, {``\href{https://doi.org/10.1016/S0065-2601(08)60330-5}{A focus theory of normative conduct: A theoretical refinement and reevaluation of the role of norms in human behavior}''} in Advances in experimental social psychology., M.
+P.
+Zanna, Ed.
+(Academic Press, 1991), pp.
+201--234.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018a}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{E. Fehr, I. Schurtenberger, \href{https://doi.org/10.1038/s41562-018-0385-5}{Normative foundations of human cooperation}. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Legros2020}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{S.
+Legros, B.
+Cislaghi, \href{https://doi.org/10.1177/1745691619866455}{Mapping the social-norms literature: An overview of reviews}.
+\emph{Perspectives on Psychological Science} \textbf{15}, 62--80 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{G.
+A.
+van Kleef, M.
+J.
+Gelfand, J.
+Jetten, \href{https://doi.org/10.1016/j.jesp.2019.05.002}{The dynamic nature of social norms: New perspectives on norm development, impact, violation, and enforcement}.
+\emph{Journal of Experimental Social Psychology} \textbf{84}, 103814 (2019).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Asch1956}{}}%
-Asch, S.
+\CSLLeftMargin{6. }%
+\CSLRightInline{S.
 E.
-(1956).
-Studies of independence and conformity: I.
-A minority of one against a unanimous majority.
-\emph{Psychological Monographs: General and Applied}, \emph{70}(9), 1.
-\url{https://doi.org/10.1037/h0093718}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-Aust, F., \& Barth, M. (2022). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://github.com/crsh/papaja}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
-Bates, D., Mächler, M., Bolker, B., \& Walker, S.
-(2015).
-Fitting linear mixed-effects models using lme4.
-\emph{Journal of Statistical Software}, \emph{67}(1), 1--48.
-\url{https://doi.org/10.18637/jss.v067.i01}
-
-\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
-Baxter-King, R., Brown, J.
-R., Enos, R.
-D., Naeim, A., \& Vavreck, L.
-(2022).
-How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}.
-\emph{Proceedings of the National Academy of Sciences}, \emph{119}(21), e2116311119.
-\url{https://doi.org/10.1073/pnas.2116311119}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
-Bokemper, S.
-E., Cucciniello, M., Rotesi, T., Pin, P., Malik, A.
-A., Willebrand, K., Paintsil, E.
-E., Omer, S.
-B., Huber, G.
-A., \& Melegaro, A.
-(2021).
-Experimental evidence that changing beliefs about mask efficacy and social norms increase mask wearing for {COVID-19} risk reduction: Results from the {United States} and {Italy}.
-\emph{PLOS ONE}, \emph{16}(10), e0258282.
-\url{https://doi.org/10.1371/journal.pone.0258282}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Boyd1985}{}}%
-Boyd, \&. R., Robert. (1985). \emph{Culture and the evolutionary process}. University of Chicago Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-CDC2022b}{}}%
-Centers for Disease Control and Prevention. (2022a). \emph{{CDC Museum COVID-19 Timeline}}. \url{https://www.cdc.gov/museum/timeline/covid19.html}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CDC2022a}{}}%
-Centers for Disease Control and Prevention.
-(2022b).
-\emph{{United States COVID-19 Community Levels by County}} {[}Data set{]}.
-\url{https://data.cdc.gov/Public-Health-Surveillance/United-States-COVID-19-Community-Levels-by-County/3nnm-4jni}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2007}{}}%
-Cialdini, R. B. (2007). Descriptive social norms as underappreciated sources of social control. \emph{Psychometrika}, \emph{72}(2), 263--268. \url{https://doi.org/10.1007/s11336-006-1560-6}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2021}{}}%
-Cialdini, R.
-B., \& Jacobson, R.
-P.
-(2021).
-Influences of social norms on climate change-related behaviors.
-\emph{Current Opinion in Behavioral Sciences}, \emph{42}, 1--8.
-\url{https://doi.org/10.1016/j.cobeha.2021.01.005}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cialdini1991}{}}%
-Cialdini, R.
-B., Kallgren, C.
-A., \& Reno, R.
-R.
-(1991).
-\emph{A focus theory of normative conduct: A theoretical refinement and reevaluation of the role of norms in human behavior} (M.
-P.
-Zanna, Ed.; Vol.
-24, pp.
-201--234).
-Academic Press.
-\url{https://doi.org/10.1016/S0065-2601(08)60330-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
-Deutsch, M., \& Gerard, H.
-B.
-(1955).
-A study of normative and informational social influences upon individual judgment.
-\emph{The Journal of Abnormal and Social Psychology}, \emph{51}(3), 629--636.
-\url{https://doi.org/10.1037/h0046408}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018a}{}}%
-Fehr, E., \& Schurtenberger, I. (2018a). Normative foundations of human cooperation. \emph{Nature Human Behaviour}, \emph{2}(7), 458--468. \url{https://doi.org/10.1038/s41562-018-0385-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018b}{}}%
-Fehr, E., \& Schurtenberger, I.
-(2018b).
-\emph{The dynamics of norm formation and norm decay} (Working Paper Series).
-Department of Economics 278, University of Zurich.
-\url{https://doi.org/10.5167/uzh-147925}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
-Gelfand, M.
-J., \& Harrington, J.
-R.
-(2015).
-The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?
-\emph{Journal of Cross-Cultural Psychology}, \emph{46}(10), 1273--1278.
-\url{https://doi.org/10.1177/0022022115600796}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2011}{}}%
-Gelfand, M.
-J., Raver, J.
-L., Nishii, L., Leslie, L.
-M., Lun, J., Lim, B.
-C., Duan, L., Almaliach, A., Ang, S., Arnadottir, J., Aycan, Z., Boehnke, K., Boski, P., Cabecinhas, R., Chan, D., Chhokar, J., D'Amato, A., Ferrer, M., Fischlmayr, I.
-C., \ldots{} Yamaguchi, S.
-(2011).
-Differences between tight and loose cultures: A 33-nation study.
-\emph{Science}, \emph{332}(6033), 1100--1104.
-\url{https://doi.org/10.1126/science.1197754}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
-Goldstein, N.
-J., Cialdini, R.
-B., \& Griskevicius, V.
-(2008).
-A room with a viewpoint: Using social norms to motivate environmental conservation in hotels.
-\emph{Journal of Consumer Research}, \emph{35}(3), 472--482.
-\url{https://doi.org/10.1086/586910}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
-Granger, C. W. J. (1969). Investigating causal relations by econometric models and cross-spectral methods. \emph{Econometrica}, \emph{37}(3), 424--438. \url{https://doi.org/10.2307/1912791}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
-Hamaker, E. L., Kuiper, R. M., \& Grasman, R. P. P. P. (2015). A critique of the cross-lagged panel model. \emph{Psychological Methods}, \emph{20}(1), 102. \url{https://doi.org/10.1037/a0038889}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
-Henrich, J. (2015). \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}. Princeton University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hogg2008}{}}%
-Hogg, M.
-A., Hohman, Z.
-P., \& Rivera, J.
-E.
-(2008).
-Why do people join groups?
-Three motivational accounts from social psychology.
-\emph{Social and Personality Psychology Compass}, \emph{2}(3), 1269--1280.
-\url{https://doi.org/10.1111/j.1751-9004.2008.00099.x}
+Asch, \href{https://doi.org/10.1037/h0093718}{Studies of independence and conformity: I.
+A minority of one against a unanimous majority}.
+\emph{Psychological monographs: General and applied} \textbf{70}, 1 (1956).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Tomasello2014}{}}%
+\CSLLeftMargin{7. }%
+\CSLRightInline{M. Tomasello, \href{https://doi.org/10.1002/ejsp.2015}{The ultra-social animal}. \emph{European Journal of Social Psychology} \textbf{44}, 187--194 (2014).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Jensen2014}{}}%
-Jensen, K., Vaish, A., \& Schmidt, M.
+\CSLLeftMargin{8. }%
+\CSLRightInline{K.
+Jensen, A.
+Vaish, M.
 F.
 H.
-(2014).
-The emergence of human prosociality: Aligning with others through feelings, concerns, and norms.
-\emph{Frontiers in Psychology}, \emph{5}, 822.
-\url{https://doi.org/10.3389/fpsyg.2014.00822}
+Schmidt, \href{https://doi.org/10.3389/fpsyg.2014.00822}{The emergence of human prosociality: Aligning with others through feelings, concerns, and norms}.
+\emph{Frontiers in Psychology} \textbf{5}, 822 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rakoczy2013}{}}%
+\CSLLeftMargin{9. }%
+\CSLRightInline{H. Rakoczy, M. F. H. Schmidt, \href{https://doi.org/10.1111/cdep.12010}{The early ontogeny of social norms}. \emph{Child Development Perspectives} \textbf{7}, 17--21 (2013).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2012}{}}%
+\CSLLeftMargin{10. }%
+\CSLRightInline{M.
+F.
+H.
+Schmidt, M.
+Tomasello, \href{https://doi.org/10.1177/0963721412448659}{Young children enforce social norms}.
+\emph{Current Directions in Psychological Science} \textbf{21}, 232--236 (2012).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Vaish2011}{}}%
+\CSLLeftMargin{11. }%
+\CSLRightInline{A.
+Vaish, M.
+Carpenter, M.
+Tomasello, \href{https://doi.org/10.1037/a0024462}{Young children's responses to guilt displays}.
+\emph{Developmental Psychology} \textbf{47}, 1248--1262 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schaumberg2022}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{R.
+L.
+Schaumberg, S.
+E.
+Skowronek, \href{https://doi.org/10.1177/09567976221075303}{Shame broadcasts social norms: The positive social effects of shame on norm acquisition and normative behavior}.
+\emph{Psychological Science}, 09567976221075303 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2011}{}}%
+\CSLLeftMargin{13. }%
+\CSLRightInline{M.
+J.
+Gelfand, \emph{et al.}, \href{https://doi.org/10.1126/science.1197754}{Differences between tight and loose cultures: A 33-nation study}.
+\emph{Science} \textbf{332}, 1100--1104 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
+\CSLLeftMargin{14. }%
+\CSLRightInline{J. Henrich, \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter} (Princeton University Press, 2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{M.
+Deutsch, H.
+B.
+Gerard, \href{https://doi.org/10.1037/h0046408}{A study of normative and informational social influences upon individual judgment}.
+\emph{The journal of abnormal and social psychology} \textbf{51}, 629--636 (1955).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schultz2007}{}}%
+\CSLLeftMargin{16. }%
+\CSLRightInline{P.
+W.
+Schultz, J.
+M.
+Nolan, R.
+B.
+Cialdini, N.
+J.
+Goldstein, V.
+Griskevicius, \href{https://doi.org/10.1111/j.1467-9280.2007.01917.x}{The constructive, destructive, and reconstructive power of social norms}.
+\emph{Psychological Science} \textbf{18}, 429--434 (2007).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2007}{}}%
+\CSLLeftMargin{17. }%
+\CSLRightInline{R.
+B.
+Cialdini, \href{https://doi.org/10.1007/s11336-006-1560-6}{Descriptive social norms as underappreciated sources of social control}.
+\emph{Psychometrika} \textbf{72}, 263--268 (2007).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2021}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{R.
+B.
+Cialdini, R.
+P.
+Jacobson, \href{https://doi.org/10.1016/j.cobeha.2021.01.005}{Influences of social norms on climate change-related behaviors}.
+\emph{Current Opinion in Behavioral Sciences} \textbf{42}, 1--8 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Boyd1985}{}}%
+\CSLLeftMargin{19. }%
+\CSLRightInline{\&. R. Boyd Robert, \emph{Culture and the evolutionary process} (University of Chicago Press, 1985).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kushnick2014}{}}%
+\CSLLeftMargin{20. }%
+\CSLRightInline{G.
+Kushnick, R.
+D.
+Gray, F.
+M.
+Jordan, \href{https://doi.org/10.1016/j.evolhumbehav.2014.03.001}{The sequential evolution of land tenure norms}.
+\emph{Evolution and Human Behavior} \textbf{35}, 309--318 (2014).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Jordan2009}{}}%
-Jordan, F.
-M., Gray, R.
-D., Greenhill, S.
-J., \& Mace, R.
-(2009).
-Matrilocal residence is ancestral in austronesian societies.
-\emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{276}(1664), 1957--1964.
-\url{https://doi.org/10.1098/rspb.2009.0088}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kushnick2014}{}}%
-Kushnick, G., Gray, R.
-D., \& Jordan, F.
+\CSLLeftMargin{21. }%
+\CSLRightInline{F.
 M.
-(2014).
-The sequential evolution of land tenure norms.
-\emph{Evolution and Human Behavior}, \emph{35}(4), 309--318.
-\url{https://doi.org/10.1016/j.evolhumbehav.2014.03.001}
+Jordan, R.
+D.
+Gray, S.
+J.
+Greenhill, R.
+Mace, \href{https://doi.org/10.1098/rspb.2009.0088}{Matrilocal residence is ancestral in austronesian societies}.
+\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{276}, 1957--1964 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
+\CSLLeftMargin{22. }%
+\CSLRightInline{K.
+Titlestad, T.
+A.
+B.
+Snijders, K.
+Durrheim, M.
+Quayle, T.
+Postmes, \href{https://doi.org/10.1016/j.jesp.2019.03.010}{The dynamic emergence of cooperative norms in a social dilemma}.
+\emph{Journal of Experimental Social Psychology} \textbf{84}, 103799 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018b}{}}%
+\CSLLeftMargin{23. }%
+\CSLRightInline{E.
+Fehr, I.
+Schurtenberger, {``The dynamics of norm formation and norm decay''} (Department of Economics 278, University of Zurich, 2018) https:/doi.org/\href{https://doi.org/10.5167/uzh-147925}{10.5167/uzh-147925}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
+\CSLLeftMargin{24. }%
+\CSLRightInline{D.
+Nigbur, E.
+Lyons, D.
+Uzzell, \href{https://doi.org/10.1348/014466609X449395}{Attitudes, norms, identity and environmental behaviour: Using an expanded theory of planned behaviour to predict participation in a kerbside recycling programme}.
+\emph{British Journal of Social Psychology} \textbf{49}, 259--284 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
+\CSLLeftMargin{25. }%
+\CSLRightInline{C.
+Larkin, M.
+Sanders, I.
+Andresen, F.
+Algate, {``Testing local descriptive norms and salience of enforcement action: A field experiment to increase tax collection''} (2018) https:/doi.org/\href{https://doi.org/10.2139/ssrn.3167575}{10.2139/ssrn.3167575}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
+\CSLLeftMargin{26. }%
+\CSLRightInline{N.
+J.
+Goldstein, R.
+B.
+Cialdini, V.
+Griskevicius, \href{https://doi.org/10.1086/586910}{A room with a viewpoint: Using social norms to motivate environmental conservation in hotels}.
+\emph{Journal of Consumer Research} \textbf{35}, 472--482 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
+\CSLLeftMargin{27. }%
+\CSLRightInline{C.
+Bicchieri, J.
+W.
+Lindemans, T.
+Jiang, \href{https://doi.org/10.3389/fpsyg.2014.01418}{A structured approach to a diagnostic of collective practices}.
+\emph{Frontiers in Psychology} \textbf{5}, 1418 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
+\CSLLeftMargin{28. }%
+\CSLRightInline{L.
+G.
+E.
+Smith, E.
+F.
+Thomas, C.
+McGarty, \href{https://doi.org/10.1111/pops.12180}{"We must be the change we want to see in the world": Integrating norms and identities through social interaction}.
+\emph{Political Psychology} \textbf{36}, 543--557 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
+\CSLLeftMargin{29. }%
+\CSLRightInline{J.
+T.
+Macy, C.
+Owens, K.
+Mullis, S.
+E.
+Middlestadt, \href{https://doi.org/10.3389/fpubh.2021.660813}{The role of self-efficacy and injunctive norms in helping older adults decide to stay home during the {COVID-19} pandemic}.
+\emph{Frontiers in Public Health} \textbf{9}, 660813 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{S.
+E.
+Bokemper, \emph{et al.}, \href{https://doi.org/10.1371/journal.pone.0258282}{Experimental evidence that changing beliefs about mask efficacy and social norms increase mask wearing for {COVID-19} risk reduction: Results from the {United States} and {Italy}}.
+\emph{PLOS ONE} \textbf{16}, e0258282 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
+\CSLLeftMargin{31. }%
+\CSLRightInline{S.
+C.
+Rudert, S.
+Janke, \href{https://doi.org/10.1177/13684302211023562}{Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic}.
+\emph{Group Processes \& Intergroup Relations}, 13684302211023562 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
+\CSLLeftMargin{32. }%
+\CSLRightInline{E.
+L.
+Hamaker, R.
+M.
+Kuiper, R.
+P.
+P.
+P.
+Grasman, \href{https://doi.org/10.1037/a0038889}{A critique of the cross-lagged panel model}.
+\emph{Psychological methods} \textbf{20}, 102 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CDC2022b}{}}%
+\CSLLeftMargin{33. }%
+\CSLRightInline{Centers for Disease Control and Prevention, \href{https://www.cdc.gov/museum/timeline/covid19.html}{{CDC Museum COVID-19 Timeline}} (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
+\CSLLeftMargin{34. }%
+\CSLRightInline{U.
+Orth, \emph{et al.}, Effect size guidelines for cross-lagged effects.
+\emph{Psychological Methods} (2022) https:/doi.org/\href{https://doi.org/10.1037/met0000499}{10.1037/met0000499}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
+\CSLLeftMargin{35. }%
+\CSLRightInline{H.
+R.
+Markus, S.
+Kitayama, \href{https://doi.org/10.1177/1745691610375557}{Cultures and selves: A cycle of mutual constitution}.
+\emph{Perspectives on Psychological Science} \textbf{5}, 420--430 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
+\CSLLeftMargin{36. }%
+\CSLRightInline{M.
+J.
+Gelfand, J.
+R.
+Harrington, \href{https://doi.org/10.1177/0022022115600796}{The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?} \emph{Journal of Cross-Cultural Psychology} \textbf{46}, 1273--1278 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hogg2008}{}}%
+\CSLLeftMargin{37. }%
+\CSLRightInline{M.
+A.
+Hogg, Z.
+P.
+Hohman, J.
+E.
+Rivera, \href{https://doi.org/10.1111/j.1751-9004.2008.00099.x}{Why do people join groups?
+Three motivational accounts from social psychology}.
+\emph{Social and Personality Psychology Compass} \textbf{2}, 1269--1280 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wellen1998}{}}%
+\CSLLeftMargin{38. }%
+\CSLRightInline{J.
+M.
+Wellen, M.
+A.
+Hogg, D.
+J.
+Terry, \href{https://doi.org/10.1037/1089-2699.2.1.48}{Group norms and attitude-behavior consistency: The role of group salience and mood}.
+\emph{Group Dynamics: Theory, Research, and Practice} \textbf{2}, 48--56 (1998).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Neighbors2008}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{C.
+Neighbors, \emph{et al.}, \href{https://doi.org/10.1037/a0013043}{The relative impact of injunctive norms on college student drinking: The role of reference group.} \emph{Psychology of Addictive Behaviors} \textbf{22}, 576--581 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
+\CSLLeftMargin{40. }%
+\CSLRightInline{R.
+Baxter-King, J.
+R.
+Brown, R.
+D.
+Enos, A.
+Naeim, L.
+Vavreck, \href{https://doi.org/10.1073/pnas.2116311119}{How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{119}, e2116311119 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Remedios2022}{}}%
+\CSLLeftMargin{41. }%
+\CSLRightInline{J. D. Remedios, \href{https://doi.org/10.1038/s44159-022-00024-4}{Psychology must grapple with whiteness}. \emph{Nature Reviews Psychology} \textbf{1}, 125--126 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{H.
+Natanson, \href{https://www.washingtonpost.com/education/2022/02/25/peer-pressure-mask-optional-schools/}{Peer pressure is ending mask usage in schools}.
+\emph{The Washingston Post} (2022) (August 18, 2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
+\CSLLeftMargin{43. }%
+\CSLRightInline{D.
+Bates, M.
+Mächler, B.
+Bolker, S.
+Walker, \href{https://doi.org/10.18637/jss.v067.i01}{Fitting linear mixed-effects models using lme4}.
+\emph{Journal of Statistical Software} \textbf{67}, 1--48 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
+\CSLLeftMargin{44. }%
+\CSLRightInline{C.
+W.
+J.
+Granger, \href{https://doi.org/10.2307/1912791}{Investigating causal relations by econometric models and cross-spectral methods}.
+\emph{Econometrica} \textbf{37}, 424--438 (1969).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
+\CSLLeftMargin{45. }%
+\CSLRightInline{Y. Rosseel, \href{https://doi.org/10.18637/jss.v048.i02}{{lavaan}: An {R} package for structural equation modeling}. \emph{Journal of Statistical Software} \textbf{48}, 1--36 (2012).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
+\CSLLeftMargin{46. }%
+\CSLRightInline{R Core Team, \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}} (R Foundation for Statistical Computing, 2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
+\CSLLeftMargin{47. }%
+\CSLRightInline{C. O. Wilke, \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}} (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
+\CSLLeftMargin{48. }%
+\CSLRightInline{H. Wickham, \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}} (Springer-Verlag New York, 2016).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-Landau, W.
+\CSLLeftMargin{49. }%
+\CSLRightInline{W.
 M.
-(2021).
-The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing.
-\emph{Journal of Open Source Software}, \emph{6}(57), 2959.
-\url{https://doi.org/10.21105/joss.02959}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
-Larkin, C., Sanders, M., Andresen, I., \& Algate, F.
-(2018).
-\emph{Testing local descriptive norms and salience of enforcement action: A field experiment to increase tax collection} (SSRN Scholarly Paper No.
-3167575).
-\url{https://doi.org/10.2139/ssrn.3167575}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
-Macy, J.
-T., Owens, C., Mullis, K., \& Middlestadt, S.
-E.
-(2021).
-The role of self-efficacy and injunctive norms in helping older adults decide to stay home during the {COVID-19} pandemic.
-\emph{Frontiers in Public Health}, \emph{9}, 660813.
-\url{https://doi.org/10.3389/fpubh.2021.660813}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
-Markus, H.
-R., \& Kitayama, S.
-(2010).
-Cultures and selves: A cycle of mutual constitution.
-\emph{Perspectives on Psychological Science}, \emph{5}(4), 420--430.
-\url{https://doi.org/10.1177/1745691610375557}
-
-\leavevmode\vadjust pre{\hypertarget{ref-MITElectionLab}{}}%
-MIT Election Data and Science Lab. (2018). \emph{{County Presidential Election Returns 2000-2020}} (Version V10) {[}Data set{]}. {Harvard Dataverse}. \url{https://doi.org/10.7910/DVN/VOQCHQ}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Montes2021}{}}%
-Montes, K.
-S., Richards, D.
-K., \& Pearson, M.
-R.
-(2021).
-A novel approach to assess descriptive and injunctive norms for college student marijuana use.
-\emph{Addictive Behaviors}, \emph{117}, 106755.
-\url{https://doi.org/10.1016/j.addbeh.2020.106755}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
-Natanson, H. (2022). Peer pressure is ending mask usage in schools. \emph{The Washingston Post}. \url{https://www.washingtonpost.com/education/2022/02/25/peer-pressure-mask-optional-schools/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Neighbors2008}{}}%
-Neighbors, C., O'Connor, R.
-M., Lewis, M.
-A., Chawla, N., Lee, C.
-M., \& Fossos, N.
-(2008).
-The relative impact of injunctive norms on college student drinking: The role of reference group.
-\emph{Psychology of Addictive Behaviors}, \emph{22}(4), 576--581.
-\url{https://doi.org/10.1037/a0013043}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
-Nigbur, D., Lyons, E., \& Uzzell, D.
-(2010).
-Attitudes, norms, identity and environmental behaviour: Using an expanded theory of planned behaviour to predict participation in a kerbside recycling programme.
-\emph{British Journal of Social Psychology}, \emph{49}(2), 259--284.
-\url{https://doi.org/10.1348/014466609X449395}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
-Orth, U., Meier, L.
-L., Bühler, J.
-L., Dapp, L.
-C., Krauss, S., Messerli, D., \& Robins, R.
-W.
-(2022).
-Effect size guidelines for cross-lagged effects.
-\emph{Psychological Methods}.
-\url{https://doi.org/10.1037/met0000499}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Park2009}{}}%
-Park, H.
-S., Klein, K.
-A., Smith, S., \& Martell, D.
-(2009).
-Separating subjective norms, university descriptive and injunctive norms, and {U.S.} Descriptive and injunctive norms for drinking behavior intentions.
-\emph{Health Communication}, \emph{24}(8), 746--751.
-\url{https://doi.org/10.1080/10410230903265912}
-
-\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-R Core Team. (2022). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rakoczy2013}{}}%
-Rakoczy, H., \& Schmidt, M. F. H. (2013). The early ontogeny of social norms. \emph{Child Development Perspectives}, \emph{7}(1), 17--21. \url{https://doi.org/10.1111/cdep.12010}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Remedios2022}{}}%
-Remedios, J. D. (2022). Psychology must grapple with whiteness. \emph{Nature Reviews Psychology}, \emph{1}(3), 125--126. \url{https://doi.org/10.1038/s44159-022-00024-4}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
-Rosseel, Y. (2012). {lavaan}: An {R} package for structural equation modeling. \emph{Journal of Statistical Software}, \emph{48}(2), 1--36. \url{https://doi.org/10.18637/jss.v048.i02}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
-Rudert, S.
-C., \& Janke, S.
-(2021).
-Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic.
-\emph{Group Processes \& Intergroup Relations}, 13684302211023562.
-\url{https://doi.org/10.1177/13684302211023562}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Schaumberg2022}{}}%
-Schaumberg, R.
-L., \& Skowronek, S.
-E.
-(2022).
-Shame broadcasts social norms: The positive social effects of shame on norm acquisition and normative behavior.
-\emph{Psychological Science}, 09567976221075303.
-\url{https://doi.org/10.1177/09567976221075303}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2012}{}}%
-Schmidt, M.
-F.
-H., \& Tomasello, M.
-(2012).
-Young children enforce social norms.
-\emph{Current Directions in Psychological Science}, \emph{21}(4), 232--236.
-\url{https://doi.org/10.1177/0963721412448659}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Schultz2007}{}}%
-Schultz, P.
-W., Nolan, J.
-M., Cialdini, R.
-B., Goldstein, N.
-J., \& Griskevicius, V.
-(2007).
-The constructive, destructive, and reconstructive power of social norms.
-\emph{Psychological Science}, \emph{18}(5), 429--434.
-\url{https://doi.org/10.1111/j.1467-9280.2007.01917.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
-Smith, L.
-G.
-E., Thomas, E.
-F., \& McGarty, C.
-(2015).
-"We must be the change we want to see in the world": Integrating norms and identities through social interaction.
-\emph{Political Psychology}, \emph{36}(5), 543--557.
-\url{https://doi.org/10.1111/pops.12180}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
-Titlestad, K., Snijders, T.
-A.
-B., Durrheim, K., Quayle, M., \& Postmes, T.
-(2019).
-The dynamic emergence of cooperative norms in a social dilemma.
-\emph{Journal of Experimental Social Psychology}, \emph{84}, 103799.
-\url{https://doi.org/10.1016/j.jesp.2019.03.010}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Tomasello2014}{}}%
-Tomasello, M. (2014). The ultra-social animal. \emph{European Journal of Social Psychology}, \emph{44}(3), 187--194. \url{https://doi.org/10.1002/ejsp.2015}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Vaish2011}{}}%
-Vaish, A., Carpenter, M., \& Tomasello, M. (2011). Young children's responses to guilt displays. \emph{Developmental Psychology}, \emph{47}(5), 1248--1262. \url{https://doi.org/10.1037/a0024462}
-
-\leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
-van Kleef, G.
-A., Gelfand, M.
-J., \& Jetten, J.
-(2019).
-The dynamic nature of social norms: New perspectives on norm development, impact, violation, and enforcement.
-\emph{Journal of Experimental Social Psychology}, \emph{84}, 103814.
-\url{https://doi.org/10.1016/j.jesp.2019.05.002}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wellen1998}{}}%
-Wellen, J.
-M., Hogg, M.
-A., \& Terry, D.
-J.
-(1998).
-Group norms and attitude-behavior consistency: The role of group salience and mood.
-\emph{Group Dynamics: Theory, Research, and Practice}, \emph{2}(1), 48--56.
-\url{https://doi.org/10.1037/1089-2699.2.1.48}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-Wilke, C. O. (2020). \emph{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}. \url{https://CRAN.R-project.org/package=cowplot}
+Landau, \href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
+\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
+\CSLLeftMargin{50. }%
+\CSLRightInline{F. Aust, M. Barth, \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}} (2022).}
 
@@
 \includegraphics{manuscript_files/figure-latex/plotDAG-1.pdf}
-\caption{\label{fig:plotDAG}\emph{Directed acyclic graph reflecting causal assumptions.} In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms cause mask wearing behavior (Mask).
-County-level COVID-19 prevalence (CovidPrev) and political leaning (CountyPol) are also posited to be common causes of perceptions of social norms and mask wearing behavior.
-Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived descriptive norms on mask wearing.
-Similarly, it is necessary to control for perceptions of descriptive norms, county-level COVID-19 prevalence, and county-level political leaning to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
+\caption{\label{fig:plotDAG}\emph{Directed acyclic graph reflecting causal assumptions.} In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing behavior (Mask).
+Perceptions of descriptive and injunctive norms also indirectly cause mask wearing behavior through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
+Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
+Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotNormCompare-1.pdf}
-\caption{\label{fig:plotNormCompare}\emph{Predictions from multilevel model of the difference between the perceived strength of descriptive and injunctive social norms.} Points and error bars represent estimated marginal means and 95\% confidence intervals, respectively.}
-\end{figure}
-
-\newpage
-
 \hypertarget{supplementary-tables}{%
@@
 You can learn about what people are doing and what they should be doing in different ways.
-For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both.'' \emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
+For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both''.
+\emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
 \toprule
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Intercept \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.86\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.67\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.54\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 2.45\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.17) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.16) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.20) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.18) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Descriptive norms \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.25\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.29\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
@@
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} \hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.25\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.26\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3781\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3790\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 4785\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 4798\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (id) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     778\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     778\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     783\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     783\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (time) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}       8\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}       8\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}      11\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}      11\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} AIC \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 11995.16\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 12031.84\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 15309.62\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 15411.28\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} BIC \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 12051.30\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 12088.00\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 15367.88\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 15469.57\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (fixed) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.07\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.10\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
 \multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.08\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
@@
 \multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (total) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.42\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.42\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.47\hphantom{0} \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.47\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
 
@@
 \endhead
-Intercept & 4.13, 95\% CI [4.06 4.21] & 5.01, 95\% CI [4.93 5.10] & 5.65, 95\% CI [5.56 5.75]\\
-Slope1 & 0.71, 95\% CI [0.40 1.04] & 1.26, 95\% CI [0.76 1.74] & 0.41, 95\% CI [-0.06 0.90]\\
-Slope2 & -3.67, 95\% CI [-4.13 -3.24] & -7.26, 95\% CI [-7.78 -6.75] & -5.82, 95\% CI [-6.37 -5.31]\\
-Slope3 & 0.99, 95\% CI [0.69 1.31] & 0.86, 95\% CI [0.56 1.16] & 0.51, 95\% CI [0.21 0.81]\\
-N & 7493 & 3837 & 3843\\
-R2 (fixed) & 0.05 & 0.26 & 0.21\\
-R2 (total) & 0.44 & 0.72 & 0.71\\
+Intercept & 4.13, 95\% CI [4.06 4.21] & 5.00, 95\% CI [4.89 5.10] & 5.64, 95\% CI [5.53 5.75]\\
+Slope1 & 0.99, 95\% CI [0.56 1.43] & 1.96, 95\% CI [1.23 2.72] & 0.78, 95\% CI [0.07 1.50]\\
+Slope2 & -5.24, 95\% CI [-5.78 -4.70] & -10.37, 95\% CI [-11.08 -9.67] & -8.34, 95\% CI [-9.02 -7.66]\\
+Slope3 & 1.81, 95\% CI [1.31 2.28] & 1.82, 95\% CI [1.36 2.27] & 1.14, 95\% CI [0.69 1.59]\\
+Slope4 & -4.17, 95\% CI [-4.62 -3.70] & -5.41, 95\% CI [-5.79 -5.01] & -4.66, 95\% CI [-5.05 -4.26]\\
+N & 8505 & 4851 & 4861\\
+R2 (fixed) & 0.11 & 0.4 & 0.34\\
+R2 (total) & 0.38 & 0.68 & 0.67\\
 \bottomrule
@@
 
-  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
-  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
-  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
-  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
-
-\begin{table}[ht]
-\begin{centerbox}
-\begin{threeparttable}
-\captionsetup{justification=centering,singlelinecheck=off}
-\caption{Unstandardized fixed effect parameters from multilevel model: difference between perceived strength of descriptive and injunctive social norms.
-\emph{Standard errors are included in brackets.}}
- \label{tab:modelSummaryTable2}
-\setlength{\tabcolsep}{0pt}
-\begin{tabular}{l l}
-
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Model 1 \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Intercept \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 4.28\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.29) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Difference between norm types \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.60\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (0.04) \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 7680\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (id) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}     779\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N (time) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt}       8\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} AIC \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 22409.98\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} BIC \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 22472.50\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (fixed) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.04\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{}
-\arrayrulecolor{black}
-
-\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 (total) \hspace{6pt}\huxbpad{6pt}} &
-\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.65\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
-
-\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
-\arrayrulecolor{black}
-\end{tabular}
-\end{threeparttable}\par\end{centerbox}
-
-\end{table}
-
-\newpage
-
 \begin{center}
@@
 
+\footnotesize{
+
 \begin{longtable}{lcccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
 \caption{\label{tab:lavaanTable}Standardised autoregressive and cross-lagged parameters from random-intercept cross-lagged panel model.
-\emph{Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms.
+\emph{Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms, Fact = factual beliefs, Pers = personal normative beliefs.
 Variable name suffixes indicate time points.
@@
 \endhead
-Des\_02 → Mask\_03 & -0.01 & 0.06 & -0.14 & 0.11\\
-Des\_02 → Inj\_03 & 0.08 & 0.07 & -0.05 & 0.21\\
-Des\_02 → Des\_03 & 0.20 & 0.07 & 0.07 & 0.33\\
-Des\_03 → Mask\_05 & 0.11 & 0.06 & -0.01 & 0.24\\
-Des\_03 → Inj\_05 & 0.10 & 0.07 & -0.03 & 0.23\\
-Des\_03 → Des\_05 & 0.32 & 0.06 & 0.19 & 0.44\\
-Des\_05 → Mask\_09 & 0.11 & 0.07 & -0.03 & 0.24\\
-Des\_05 → Inj\_09 & 0.11 & 0.07 & -0.02 & 0.25\\
-Des\_05 → Des\_09 & 0.06 & 0.07 & -0.08 & 0.20\\
-Des\_09 → Mask\_11 & 0.17 & 0.07 & 0.03 & 0.31\\
-Des\_09 → Inj\_11 & 0.28 & 0.07 & 0.15 & 0.41\\
-Des\_09 → Des\_11 & 0.27 & 0.07 & 0.14 & 0.40\\
-Des\_11 → Mask\_13 & 0.21 & 0.08 & 0.06 & 0.36\\
-Des\_11 → Inj\_13 & 0.15 & 0.07 & 0.01 & 0.30\\
-Des\_11 → Des\_13 & 0.27 & 0.07 & 0.13 & 0.41\\
-Des\_13 → Mask\_14 & 0.16 & 0.08 & 0.01 & 0.31\\
+Des\_02 → Mask\_05 & 0.17 & 0.05 & 0.06 & 0.28\\
+Des\_02 → Inj\_05 & 0.17 & 0.06 & 0.06 & 0.28\\
+Des\_02 → Des\_05 & 0.37 & 0.05 & 0.26 & 0.47\\
+Des\_02 → Fact\_05 & 0.09 & 0.06 & -0.04 & 0.21\\
+Des\_02 → Pers\_05 & 0.04 & 0.06 & -0.08 & 0.17\\
+Des\_05 → Mask\_09 & 0.21 & 0.06 & 0.08 & 0.34\\
+Des\_05 → Inj\_09 & 0.23 & 0.07 & 0.10 & 0.36\\
+Des\_05 → Des\_09 & 0.26 & 0.06 & 0.14 & 0.39\\
+Des\_05 → Fact\_09 & 0.16 & 0.07 & 0.02 & 0.30\\
+Des\_05 → Pers\_09 & 0.27 & 0.07 & 0.12 & 0.42\\
+Des\_09 → Mask\_11 & 0.04 & 0.07 & -0.09 & 0.18\\
+Des\_09 → Inj\_11 & 0.20 & 0.07 & 0.07 & 0.33\\
+Des\_09 → Des\_11 & 0.26 & 0.07 & 0.13 & 0.39\\
+Des\_09 → Fact\_11 & 0.03 & 0.06 & -0.09 & 0.16\\
+Des\_09 → Pers\_11 & 0.07 & 0.07 & -0.06 & 0.20\\
+Des\_11 → Mask\_13 & 0.15 & 0.07 & 0.01 & 0.30\\
+Des\_11 → Inj\_13 & 0.03 & 0.07 & -0.12 & 0.17\\
+Des\_11 → Des\_13 & 0.27 & 0.07 & 0.14 & 0.41\\
+Des\_11 → Fact\_13 & 0.07 & 0.07 & -0.07 & 0.21\\
+Des\_11 → Pers\_13 & 0.06 & 0.07 & -0.08 & 0.20\\
+Des\_13 → Mask\_14 & 0.16 & 0.07 & 0.02 & 0.29\\
 Des\_13 → Inj\_14 & 0.21 & 0.06 & 0.09 & 0.33\\
-Des\_13 → Des\_14 & 0.32 & 0.06 & 0.20 & 0.45\\
-Des\_14 → Mask\_15 & 0.15 & 0.08 & -0.02 & 0.31\\
-Des\_14 → Inj\_15 & 0.11 & 0.07 & -0.03 & 0.26\\
-Des\_14 → Des\_15 & 0.35 & 0.07 & 0.22 & 0.48\\
-Inj\_02 → Mask\_03 & 0.10 & 0.06 & -0.02 & 0.23\\
-Inj\_02 → Inj\_03 & 0.07 & 0.07 & -0.06 & 0.21\\
-Inj\_02 → Des\_03 & 0.03 & 0.06 & -0.09 & 0.15\\
-Inj\_03 → Mask\_05 & 0.03 & 0.06 & -0.09 & 0.16\\
-Inj\_03 → Inj\_05 & 0.14 & 0.07 & 0.01 & 0.28\\
-Inj\_03 → Des\_05 & 0.01 & 0.06 & -0.11 & 0.14\\
-Inj\_05 → Mask\_09 & -0.11 & 0.07 & -0.24 & 0.02\\
-Inj\_05 → Inj\_09 & -0.07 & 0.07 & -0.21 & 0.07\\
-Inj\_05 → Des\_09 & -0.05 & 0.07 & -0.19 & 0.08\\
-Inj\_09 → Mask\_11 & 0.04 & 0.08 & -0.11 & 0.19\\
-Inj\_09 → Inj\_11 & 0.13 & 0.07 & 0.00 & 0.27\\
-Inj\_09 → Des\_11 & 0.07 & 0.07 & -0.06 & 0.21\\
-Inj\_11 → Mask\_13 & 0.05 & 0.07 & -0.09 & 0.19\\
-Inj\_11 → Inj\_13 & 0.32 & 0.07 & 0.19 & 0.45\\
-Inj\_11 → Des\_13 & 0.22 & 0.07 & 0.08 & 0.35\\
-Inj\_13 → Mask\_14 & -0.02 & 0.07 & -0.17 & 0.12\\
-Inj\_13 → Inj\_14 & 0.42 & 0.06 & 0.31 & 0.54\\
-Inj\_13 → Des\_14 & 0.23 & 0.06 & 0.10 & 0.35\\
-Inj\_14 → Mask\_15 & 0.03 & 0.08 & -0.13 & 0.18\\
-Inj\_14 → Inj\_15 & 0.47 & 0.07 & 0.34 & 0.60\\
-Inj\_14 → Des\_15 & 0.31 & 0.06 & 0.19 & 0.44\\
-Mask\_02 → Mask\_03 & -0.02 & 0.06 & -0.14 & 0.10\\
-Mask\_02 → Inj\_03 & 0.00 & 0.06 & -0.13 & 0.12\\
-Mask\_02 → Des\_03 & 0.01 & 0.06 & -0.11 & 0.12\\
-Mask\_03 → Mask\_05 & 0.02 & 0.06 & -0.10 & 0.13\\
-Mask\_03 → Inj\_05 & 0.04 & 0.06 & -0.08 & 0.16\\
-Mask\_03 → Des\_05 & 0.12 & 0.06 & 0.00 & 0.23\\
-Mask\_05 → Mask\_09 & 0.08 & 0.07 & -0.04 & 0.21\\
-Mask\_05 → Inj\_09 & 0.10 & 0.07 & -0.03 & 0.23\\
-Mask\_05 → Des\_09 & -0.01 & 0.07 & -0.15 & 0.12\\
-Mask\_09 → Mask\_11 & -0.03 & 0.07 & -0.16 & 0.10\\
-Mask\_09 → Inj\_11 & 0.01 & 0.06 & -0.11 & 0.14\\
-Mask\_09 → Des\_11 & 0.10 & 0.06 & -0.02 & 0.22\\
-Mask\_11 → Mask\_13 & 0.20 & 0.06 & 0.07 & 0.32\\
-Mask\_11 → Inj\_13 & 0.05 & 0.06 & -0.07 & 0.16\\
-Mask\_11 → Des\_13 & 0.10 & 0.06 & -0.02 & 0.21\\
-Mask\_13 → Mask\_14 & 0.30 & 0.06 & 0.18 & 0.42\\
-Mask\_13 → Inj\_14 & 0.10 & 0.05 & -0.01 & 0.20\\
-Mask\_13 → Des\_14 & 0.20 & 0.05 & 0.09 & 0.30\\
-Mask\_14 → Mask\_15 & 0.29 & 0.06 & 0.17 & 0.41\\
-Mask\_14 → Inj\_15 & 0.04 & 0.06 & -0.07 & 0.16\\
-Mask\_14 → Des\_15 & 0.11 & 0.05 & 0.01 & 0.21\\
+Des\_13 → Des\_14 & 0.40 & 0.06 & 0.28 & 0.51\\
+Des\_13 → Fact\_14 & 0.03 & 0.06 & -0.09 & 0.14\\
+Des\_13 → Pers\_14 & 0.01 & 0.06 & -0.11 & 0.12\\
+Des\_14 → Mask\_15 & 0.05 & 0.08 & -0.09 & 0.20\\
+Des\_14 → Inj\_15 & -0.01 & 0.07 & -0.16 & 0.13\\
+Des\_14 → Des\_15 & 0.34 & 0.06 & 0.22 & 0.46\\
+Des\_14 → Fact\_15 & 0.12 & 0.07 & -0.01 & 0.25\\
+Des\_14 → Pers\_15 & 0.09 & 0.07 & -0.05 & 0.23\\
+Des\_15 → Mask\_16 & 0.03 & 0.07 & -0.11 & 0.18\\
+Des\_15 → Inj\_16 & 0.23 & 0.08 & 0.08 & 0.38\\
+Des\_15 → Des\_16 & 0.30 & 0.07 & 0.15 & 0.45\\
+Des\_15 → Fact\_16 & 0.13 & 0.07 & 0.00 & 0.26\\
+Des\_15 → Pers\_16 & 0.01 & 0.07 & -0.12 & 0.14\\
+Des\_16 → Mask\_17 & 0.06 & 0.08 & -0.10 & 0.21\\
+Des\_16 → Inj\_17 & 0.24 & 0.08 & 0.08 & 0.39\\
+Des\_16 → Des\_17 & 0.53 & 0.07 & 0.40 & 0.66\\
+Des\_16 → Fact\_17 & 0.06 & 0.07 & -0.08 & 0.20\\
+Des\_16 → Pers\_17 & 0.03 & 0.07 & -0.10 & 0.16\\
+Des\_17 → Mask\_18 & 0.08 & 0.07 & -0.06 & 0.21\\
+Des\_17 → Inj\_18 & 0.30 & 0.07 & 0.17 & 0.43\\
+Des\_17 → Des\_18 & 0.46 & 0.06 & 0.34 & 0.58\\
+Des\_17 → Fact\_18 & 0.12 & 0.06 & 0.00 & 0.24\\
+Des\_17 → Pers\_18 & 0.07 & 0.06 & -0.05 & 0.20\\
+Fact\_02 → Mask\_05 & 0.06 & 0.07 & -0.08 & 0.19\\
+Fact\_02 → Inj\_05 & -0.10 & 0.07 & -0.24 & 0.03\\
+Fact\_02 → Des\_05 & -0.02 & 0.07 & -0.15 & 0.12\\
+Fact\_02 → Fact\_05 & 0.22 & 0.08 & 0.07 & 0.38\\
+Fact\_02 → Pers\_05 & -0.08 & 0.08 & -0.23 & 0.08\\
+Fact\_05 → Mask\_09 & 0.15 & 0.08 & -0.01 & 0.31\\
+Fact\_05 → Inj\_09 & -0.07 & 0.08 & -0.23 & 0.09\\
+Fact\_05 → Des\_09 & -0.05 & 0.08 & -0.20 & 0.11\\
+Fact\_05 → Fact\_09 & 0.07 & 0.09 & -0.10 & 0.25\\
+Fact\_05 → Pers\_09 & -0.03 & 0.09 & -0.20 & 0.15\\
+Fact\_09 → Mask\_11 & 0.15 & 0.08 & -0.01 & 0.30\\
+Fact\_09 → Inj\_11 & 0.03 & 0.08 & -0.12 & 0.18\\
+Fact\_09 → Des\_11 & 0.10 & 0.08 & -0.05 & 0.24\\
+Fact\_09 → Fact\_11 & 0.26 & 0.07 & 0.12 & 0.40\\
+Fact\_09 → Pers\_11 & 0.14 & 0.07 & -0.01 & 0.28\\
+Fact\_11 → Mask\_13 & 0.18 & 0.09 & 0.00 & 0.35\\
+Fact\_11 → Inj\_13 & 0.05 & 0.09 & -0.13 & 0.22\\
+Fact\_11 → Des\_13 & -0.12 & 0.08 & -0.28 & 0.04\\
+Fact\_11 → Fact\_13 & 0.19 & 0.08 & 0.03 & 0.36\\
+Fact\_11 → Pers\_13 & 0.16 & 0.08 & 0.00 & 0.33\\
+Fact\_13 → Mask\_14 & 0.05 & 0.08 & -0.12 & 0.21\\
+Fact\_13 → Inj\_14 & 0.04 & 0.07 & -0.11 & 0.18\\
+Fact\_13 → Des\_14 & 0.01 & 0.08 & -0.14 & 0.16\\
+Fact\_13 → Fact\_14 & 0.25 & 0.07 & 0.11 & 0.39\\
+Fact\_13 → Pers\_14 & 0.19 & 0.07 & 0.06 & 0.33\\
+Fact\_14 → Mask\_15 & 0.32 & 0.08 & 0.16 & 0.48\\
+Fact\_14 → Inj\_15 & -0.06 & 0.08 & -0.22 & 0.10\\
+Fact\_14 → Des\_15 & 0.15 & 0.07 & 0.01 & 0.29\\
+Fact\_14 → Fact\_15 & 0.47 & 0.07 & 0.33 & 0.60\\
+Fact\_14 → Pers\_15 & 0.31 & 0.08 & 0.16 & 0.47\\
+Fact\_15 → Mask\_16 & 0.10 & 0.09 & -0.08 & 0.28\\
+Fact\_15 → Inj\_16 & 0.08 & 0.10 & -0.11 & 0.27\\
+Fact\_15 → Des\_16 & 0.10 & 0.10 & -0.09 & 0.29\\
+Fact\_15 → Fact\_16 & 0.39 & 0.08 & 0.23 & 0.55\\
+Fact\_15 → Pers\_16 & 0.10 & 0.08 & -0.06 & 0.27\\
+Fact\_16 → Mask\_17 & 0.21 & 0.09 & 0.03 & 0.39\\
+Fact\_16 → Inj\_17 & -0.01 & 0.09 & -0.19 & 0.18\\
+Fact\_16 → Des\_17 & -0.05 & 0.09 & -0.22 & 0.12\\
+Fact\_16 → Fact\_17 & 0.22 & 0.08 & 0.06 & 0.39\\
+Fact\_16 → Pers\_17 & 0.06 & 0.08 & -0.10 & 0.22\\
+Fact\_17 → Mask\_18 & 0.10 & 0.09 & -0.08 & 0.28\\
+Fact\_17 → Inj\_18 & -0.10 & 0.09 & -0.28 & 0.08\\
+Fact\_17 → Des\_18 & 0.08 & 0.09 & -0.10 & 0.25\\
+Fact\_17 → Fact\_18 & 0.37 & 0.08 & 0.21 & 0.53\\
+Fact\_17 → Pers\_18 & 0.48 & 0.08 & 0.32 & 0.64\\
+Inj\_02 → Mask\_05 & 0.01 & 0.05 & -0.10 & 0.11\\
+Inj\_02 → Inj\_05 & 0.28 & 0.05 & 0.17 & 0.38\\
+Inj\_02 → Des\_05 & 0.07 & 0.05 & -0.03 & 0.18\\
+Inj\_02 → Fact\_05 & 0.05 & 0.06 & -0.08 & 0.17\\
+Inj\_02 → Pers\_05 & -0.01 & 0.06 & -0.13 & 0.11\\
+Inj\_05 → Mask\_09 & -0.07 & 0.06 & -0.19 & 0.05\\
+Inj\_05 → Inj\_09 & 0.08 & 0.06 & -0.04 & 0.21\\
+Inj\_05 → Des\_09 & -0.02 & 0.06 & -0.14 & 0.11\\
+Inj\_05 → Fact\_09 & 0.02 & 0.07 & -0.11 & 0.16\\
+Inj\_05 → Pers\_09 & -0.04 & 0.07 & -0.18 & 0.10\\
+Inj\_09 → Mask\_11 & 0.08 & 0.08 & -0.07 & 0.23\\
+Inj\_09 → Inj\_11 & 0.11 & 0.07 & -0.03 & 0.26\\
+Inj\_09 → Des\_11 & -0.03 & 0.07 & -0.17 & 0.11\\
+Inj\_09 → Fact\_11 & 0.01 & 0.07 & -0.12 & 0.15\\
+Inj\_09 → Pers\_11 & 0.05 & 0.07 & -0.08 & 0.19\\
+Inj\_11 → Mask\_13 & -0.01 & 0.07 & -0.15 & 0.13\\
+Inj\_11 → Inj\_13 & 0.29 & 0.07 & 0.15 & 0.43\\
+Inj\_11 → Des\_13 & 0.21 & 0.07 & 0.08 & 0.34\\
+Inj\_11 → Fact\_13 & 0.12 & 0.07 & -0.01 & 0.26\\
+Inj\_11 → Pers\_13 & 0.09 & 0.07 & -0.04 & 0.23\\
+Inj\_13 → Mask\_14 & -0.05 & 0.07 & -0.19 & 0.08\\
+Inj\_13 → Inj\_14 & 0.40 & 0.06 & 0.28 & 0.52\\
+Inj\_13 → Des\_14 & 0.15 & 0.07 & 0.03 & 0.28\\
+Inj\_13 → Fact\_14 & 0.02 & 0.06 & -0.10 & 0.14\\
+Inj\_13 → Pers\_14 & 0.09 & 0.06 & -0.02 & 0.21\\
+Inj\_14 → Mask\_15 & 0.08 & 0.07 & -0.06 & 0.22\\
+Inj\_14 → Inj\_15 & 0.45 & 0.07 & 0.32 & 0.58\\
+Inj\_14 → Des\_15 & 0.29 & 0.06 & 0.16 & 0.41\\
+Inj\_14 → Fact\_15 & 0.10 & 0.06 & -0.02 & 0.22\\
+Inj\_14 → Pers\_15 & 0.06 & 0.07 & -0.07 & 0.20\\
+Inj\_15 → Mask\_16 & 0.14 & 0.07 & 0.00 & 0.28\\
+Inj\_15 → Inj\_16 & 0.21 & 0.07 & 0.06 & 0.35\\
+Inj\_15 → Des\_16 & 0.06 & 0.07 & -0.08 & 0.21\\
+Inj\_15 → Fact\_16 & 0.01 & 0.06 & -0.12 & 0.13\\
+Inj\_15 → Pers\_16 & 0.10 & 0.06 & -0.03 & 0.22\\
+Inj\_16 → Mask\_17 & -0.01 & 0.07 & -0.15 & 0.13\\
+Inj\_16 → Inj\_17 & 0.38 & 0.07 & 0.23 & 0.52\\
+Inj\_16 → Des\_17 & 0.13 & 0.07 & 0.00 & 0.27\\
+Inj\_16 → Fact\_17 & 0.00 & 0.07 & -0.14 & 0.13\\
+Inj\_16 → Pers\_17 & -0.03 & 0.06 & -0.16 & 0.09\\
+Inj\_17 → Mask\_18 & -0.02 & 0.07 & -0.15 & 0.11\\
+Inj\_17 → Inj\_18 & 0.45 & 0.06 & 0.33 & 0.57\\
+Inj\_17 → Des\_18 & 0.19 & 0.06 & 0.07 & 0.32\\
+Inj\_17 → Fact\_18 & 0.01 & 0.06 & -0.11 & 0.13\\
+Inj\_17 → Pers\_18 & 0.01 & 0.06 & -0.11 & 0.13\\
+Mask\_02 → Mask\_05 & 0.21 & 0.05 & 0.11 & 0.31\\
+Mask\_02 → Inj\_05 & 0.09 & 0.05 & -0.01 & 0.20\\
+Mask\_02 → Des\_05 & 0.04 & 0.05 & -0.07 & 0.14\\
+Mask\_02 → Fact\_05 & -0.05 & 0.06 & -0.17 & 0.07\\
+Mask\_02 → Pers\_05 & -0.05 & 0.06 & -0.17 & 0.06\\
+Mask\_05 → Mask\_09 & 0.19 & 0.06 & 0.07 & 0.30\\
+Mask\_05 → Inj\_09 & 0.13 & 0.06 & 0.01 & 0.26\\
+Mask\_05 → Des\_09 & 0.02 & 0.06 & -0.10 & 0.14\\
+Mask\_05 → Fact\_09 & 0.14 & 0.07 & 0.01 & 0.27\\
+Mask\_05 → Pers\_09 & 0.06 & 0.07 & -0.07 & 0.20\\
+Mask\_09 → Mask\_11 & -0.01 & 0.07 & -0.14 & 0.12\\
+Mask\_09 → Inj\_11 & 0.06 & 0.06 & -0.07 & 0.18\\
+Mask\_09 → Des\_11 & 0.16 & 0.06 & 0.04 & 0.28\\
+Mask\_09 → Fact\_11 & 0.19 & 0.06 & 0.08 & 0.31\\
+Mask\_09 → Pers\_11 & 0.16 & 0.06 & 0.05 & 0.28\\
+Mask\_11 → Mask\_13 & 0.07 & 0.07 & -0.06 & 0.21\\
+Mask\_11 → Inj\_13 & 0.06 & 0.07 & -0.07 & 0.19\\
+Mask\_11 → Des\_13 & 0.07 & 0.06 & -0.06 & 0.19\\
+Mask\_11 → Fact\_13 & 0.04 & 0.06 & -0.09 & 0.16\\
+Mask\_11 → Pers\_13 & 0.03 & 0.07 & -0.10 & 0.16\\
+Mask\_13 → Mask\_14 & 0.19 & 0.06 & 0.08 & 0.31\\
+Mask\_13 → Inj\_14 & 0.07 & 0.05 & -0.03 & 0.18\\
+Mask\_13 → Des\_14 & 0.12 & 0.06 & 0.01 & 0.23\\
+Mask\_13 → Fact\_14 & 0.07 & 0.05 & -0.03 & 0.17\\
+Mask\_13 → Pers\_14 & 0.01 & 0.05 & -0.09 & 0.11\\
+Mask\_14 → Mask\_15 & 0.21 & 0.06 & 0.09 & 0.33\\
+Mask\_14 → Inj\_15 & 0.06 & 0.06 & -0.06 & 0.18\\
+Mask\_14 → Des\_15 & 0.08 & 0.05 & -0.02 & 0.18\\
+Mask\_14 → Fact\_15 & 0.05 & 0.05 & -0.06 & 0.15\\
+Mask\_14 → Pers\_15 & -0.05 & 0.06 & -0.17 & 0.06\\
+Mask\_15 → Mask\_16 & 0.25 & 0.07 & 0.12 & 0.39\\
+Mask\_15 → Inj\_16 & 0.02 & 0.07 & -0.12 & 0.16\\
+Mask\_15 → Des\_16 & 0.01 & 0.07 & -0.13 & 0.15\\
+Mask\_15 → Fact\_16 & 0.10 & 0.06 & -0.03 & 0.22\\
+Mask\_15 → Pers\_16 & 0.09 & 0.06 & -0.03 & 0.22\\
+Mask\_16 → Mask\_17 & 0.33 & 0.07 & 0.20 & 0.46\\
+Mask\_16 → Inj\_17 & -0.04 & 0.07 & -0.18 & 0.10\\
+Mask\_16 → Des\_17 & 0.16 & 0.06 & 0.03 & 0.28\\
+Mask\_16 → Fact\_17 & 0.22 & 0.06 & 0.09 & 0.34\\
+Mask\_16 → Pers\_17 & 0.12 & 0.06 & 0.01 & 0.24\\
+Mask\_17 → Mask\_18 & 0.39 & 0.06 & 0.27 & 0.51\\
+Mask\_17 → Inj\_18 & -0.05 & 0.06 & -0.17 & 0.07\\
+Mask\_17 → Des\_18 & 0.02 & 0.06 & -0.10 & 0.13\\
+Mask\_17 → Fact\_18 & 0.13 & 0.06 & 0.02 & 0.24\\
+Mask\_17 → Pers\_18 & 0.04 & 0.06 & -0.07 & 0.16\\
+Pers\_02 → Mask\_05 & 0.05 & 0.07 & -0.09 & 0.18\\
+Pers\_02 → Inj\_05 & 0.09 & 0.07 & -0.05 & 0.22\\
+Pers\_02 → Des\_05 & 0.03 & 0.07 & -0.10 & 0.17\\
+Pers\_02 → Fact\_05 & 0.06 & 0.08 & -0.09 & 0.22\\
+Pers\_02 → Pers\_05 & 0.36 & 0.07 & 0.21 & 0.50\\
+Pers\_05 → Mask\_09 & -0.27 & 0.08 & -0.42 & -0.12\\
+Pers\_05 → Inj\_09 & -0.16 & 0.08 & -0.31 & 0.00\\
+Pers\_05 → Des\_09 & -0.06 & 0.08 & -0.21 & 0.10\\
+Pers\_05 → Fact\_09 & -0.21 & 0.08 & -0.37 & -0.05\\
+Pers\_05 → Pers\_09 & -0.21 & 0.09 & -0.38 & -0.04\\
+Pers\_09 → Mask\_11 & 0.04 & 0.08 & -0.11 & 0.20\\
+Pers\_09 → Inj\_11 & 0.08 & 0.08 & -0.07 & 0.23\\
+Pers\_09 → Des\_11 & 0.04 & 0.07 & -0.10 & 0.19\\
+Pers\_09 → Fact\_11 & 0.06 & 0.07 & -0.08 & 0.20\\
+Pers\_09 → Pers\_11 & 0.16 & 0.07 & 0.02 & 0.31\\
+Pers\_11 → Mask\_13 & 0.08 & 0.08 & -0.08 & 0.24\\
+Pers\_11 → Inj\_13 & 0.09 & 0.08 & -0.07 & 0.24\\
+Pers\_11 → Des\_13 & 0.12 & 0.08 & -0.03 & 0.27\\
+Pers\_11 → Fact\_13 & 0.18 & 0.08 & 0.03 & 0.33\\
+Pers\_11 → Pers\_13 & 0.20 & 0.08 & 0.05 & 0.35\\
+Pers\_13 → Mask\_14 & 0.24 & 0.08 & 0.08 & 0.40\\
+Pers\_13 → Inj\_14 & -0.07 & 0.07 & -0.21 & 0.07\\
+Pers\_13 → Des\_14 & -0.03 & 0.08 & -0.18 & 0.12\\
+Pers\_13 → Fact\_14 & 0.34 & 0.07 & 0.21 & 0.48\\
+Pers\_13 → Pers\_14 & 0.41 & 0.07 & 0.29 & 0.54\\
+Pers\_14 → Mask\_15 & -0.05 & 0.08 & -0.22 & 0.11\\
+Pers\_14 → Inj\_15 & 0.15 & 0.08 & -0.02 & 0.31\\
+Pers\_14 → Des\_15 & -0.07 & 0.07 & -0.21 & 0.07\\
+Pers\_14 → Fact\_15 & 0.02 & 0.07 & -0.13 & 0.16\\
+Pers\_14 → Pers\_15 & 0.14 & 0.08 & -0.02 & 0.30\\
+Pers\_15 → Mask\_16 & 0.11 & 0.09 & -0.05 & 0.28\\
+Pers\_15 → Inj\_16 & 0.08 & 0.09 & -0.10 & 0.25\\
+Pers\_15 → Des\_16 & 0.17 & 0.09 & 0.00 & 0.35\\
+Pers\_15 → Fact\_16 & 0.11 & 0.08 & -0.05 & 0.26\\
+Pers\_15 → Pers\_16 & 0.41 & 0.08 & 0.27 & 0.56\\
+Pers\_16 → Mask\_17 & 0.00 & 0.08 & -0.17 & 0.17\\
+Pers\_16 → Inj\_17 & 0.05 & 0.09 & -0.12 & 0.23\\
+Pers\_16 → Des\_17 & -0.02 & 0.08 & -0.18 & 0.14\\
+Pers\_16 → Fact\_17 & 0.26 & 0.08 & 0.11 & 0.41\\
+Pers\_16 → Pers\_17 & 0.56 & 0.07 & 0.42 & 0.69\\
+Pers\_17 → Mask\_18 & 0.09 & 0.09 & -0.08 & 0.26\\
+Pers\_17 → Inj\_18 & -0.01 & 0.08 & -0.17 & 0.15\\
+Pers\_17 → Des\_18 & -0.02 & 0.08 & -0.18 & 0.14\\
+Pers\_17 → Fact\_18 & 0.16 & 0.08 & 0.01 & 0.31\\
+Pers\_17 → Pers\_18 & 0.12 & 0.08 & -0.03 & 0.27\\
 \bottomrule
@@
 
+}
+
 \end{ThreePartTable}

### 2022-10-13 R/functions.R

@@ -553,7 +553,7 @@ plotTimeline <- function(d, owid) {
     annotate("text", x = ymd("2021-02-03"), y = 2.5, colour = "grey", size = 2.8, angle = 90, lineheight = 0.85,
              label = "CDC: fully vaccinated\npeople can gather\nindoors without masks") +
     annotate("text", x = ymd("2021-06-07"), y = 2.0, colour = "grey", size = 2.8, angle = 90, lineheight = 0.85,
-             label = "CDC: updated\nguidelines for\nindoor mask\nuse in high\nrisk areas") +
+             label = "CDC: guidelines\nreinstated for\nindoor mask\nuse in high\nrisk areas") +
     annotate("text", x = ymd("2021-10-26"), y = 2.2, colour = "grey", size = 2.8, angle = 90, lineheight = 0.85,
              label = "CDC: recommend\nbooster shots\nfor adults") +
     annotate("text", x = ymd("2022-01-22"), y = 2.2, colour = "grey", size = 2.8, angle = 90, lineheight = 0.85,
@@ -789,7 +789,7 @@ plotRICLPM <- function(model) {
       xend   = as.Date(dates[as.numeric(substr(lhs, 4, 5))]),
       y      = c(var1, var2, var3)[as.numeric(substr(rhs, 2, 2))],
       yend   = c(var1, var2, var3)[as.numeric(substr(lhs, 2, 2))],
-      size   = abs(est.std),
+      size   = ifelse(est.std >= 0, est.std, 0), # for visualisation, min effect = 0
       colour = ifelse(pvalue.y < 0.05, "black", "lightgrey")
     )
   # plot

### 2022-10-13 _targets.R

@@ -3,10 +3,10 @@ library(targets)
 library(tarchetypes)
 source("R/functions.R")
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("arm", "cowplot", "dagitty", "ggdag", "ggeffects", 
-                            "ggraph", "huxtable", "jtools", "knitr", "kableExtra", 
-                            "lavaan", "lme4", "lubridate", "MuMIn", "papaja", 
-                            "rnaturalearth", "rnaturalearthdata", "scales", 
+tar_option_set(packages = c("arm", "broom.mixed", "cowplot", "dagitty", "ggdag", 
+                            "ggeffects", "ggraph", "huxtable", "jtools", "knitr",
+                            "kableExtra", "lavaan", "lme4", "lubridate", "MuMIn",
+                            "papaja", "rnaturalearth", "rnaturalearthdata", "scales", 
                             "semTools", "sf", "zipcodeR"))
 # workflow
 list(
@@ -51,7 +51,7 @@ list(
   tar_target(changePointsTable, makeChangePointsTable(m2.1, m2.2, m2.3)),
   tar_target(lavaanTable, makeLavaanTable(riclpm)),
   # manuscript
-  #tar_render(manuscript, "manuscript.Rmd"),
+  tar_render(manuscript, "manuscript.Rmd"),
   # print session info for reproducibility
   tar_target(sessionInfo, writeLines(capture.output(sessionInfo()), "sessionInfo.txt"))
 )

### 2022-10-17 appendix.Rmd
second draft with edits

@@
 To evaluate the construct validity of our measures of perceived descriptive and injunctive norms, at Time 7 we asked participants to rate the extent to which each perceived norm item provided descriptive and injunctive information.
-For each item, participants were asked whether the item provided information about what people are doing, and whether the item provided information about what people should be doing.
+For each item, participants were asked whether the item provided information about what people _are_ doing, and whether the item provided information about what people _should_ be doing.
 Participants responded on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.
@@
 
-(ref:plotUSMapCaption) _Map of the United States with participant zip code locations._
+(ref:plotCorBehNormCaption) _Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables._ Models contain random intercepts for participants and time points.
+Lines are fixed effect regression lines from multilevel models, shaded areas are 95% confidence intervals.
 
-```{r plotUSMap, echo=F, fig.width=5, fig.height=3, fig.cap="(ref:plotUSMapCaption)"}
-plot1
+```{r plotCorBehNorm, echo=F, fig.width=6.5, fig.height=3.8, fig.cap="(ref:plotCorBehNormCaption)"}
+plot5
 ```
@@
 
-(ref:plotAttritionCaption) _Attrition across the course of the study._
+(ref:plotCDCSensCaption) _Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations._ These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
+Time is included as a continuous linear predictor, scaled between 0 and 1, with three forced change points (dashed lines).
+From the left, the first dashed line indicates when the CDC relaxed their mask wearing recommendations in March 2021, the second dashed line indicates when the CDC strengthened their mask wearing recommendations in July 2021, and the third dashed line indicates when the CDC updated their community levels and relaxed their mask wearing recommendations in March 2022.
+This results in the estimation of five fixed effect parameters: the initial intercept, the slope in the first window, the slope in the second window, the slope in the third window, and the slope is the fourth window.
+Bolded lines and shaded areas represent fixed effect regression lines from multilevel models and 95% confidence intervals, respectively.
 
-```{r plotAttrition, echo=F, fig.width=5.5, fig.height=3, fig.cap="(ref:plotAttritionCaption)"}
-plot2
+```{r plotCDCSens, echo=F, fig.width=4, fig.height=6, fig.cap="(ref:plotCDCSensCaption)"}
+plot6
 ```
@@
 
-(ref:plotCorBehNormCaption) _Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables._ Lines are fixed effect regression lines from multilevel models, shaded areas are 95% confidence intervals.
+(ref:plotAttritionCaption) _Attrition across the course of the study._
 
-```{r plotCorBehNorm, echo=F, fig.width=6.5, fig.height=3.8, fig.cap="(ref:plotCorBehNormCaption)"}
-plot5
+```{r plotAttrition, echo=F, fig.width=5.5, fig.height=3, fig.cap="(ref:plotAttritionCaption)"}
+plot2
 ```
@@
 
-(ref:plotCDCSensCaption) _Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations._ These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
-Time is included as a continuous linear predictor, scaled between 0 and 1, with two forced change points (dashed lines).
-The first dashed line (left) indicates when the CDC relaxed their mask wearing recommendations in March 2021, and the second dashed line (right) indicates when the CDC strengthened their mask wearing recommendations in July 2021.
-This results in the estimation of four fixed effect parameters: the initial intercept, the slope in the first window, the slope in the second window, and the slope in the third window.
-Bolded lines and shaded areas represent fixed effect regression lines from multilevel models and 95% confidence intervals, respectively.
+(ref:plotUSMapCaption) _Map of the United States with participant zip code locations._
 
-```{r plotCDCSens, echo=F, fig.width=4, fig.height=6, fig.cap="(ref:plotCDCSensCaption)"}
-plot6
+```{r plotUSMap, echo=F, fig.width=5, fig.height=3, fig.cap="(ref:plotUSMapCaption)"}
+plot1
 ```
@@
 ## Supplementary Tables
-
-(ref:itemTableCaption) List of norm interpretation questions asked at Time 7.
-_These questions were preceded by the following text:_ "There may or may not be a difference between what people around you are doing and what they should be doing.
-You can learn about what people are doing and what they should be doing in different ways.
-For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both".
-_Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly._
-
-```{r itemTable, echo=F, warning=F, message=F}
-apa_table(
-  itemTable,
-  booktabs = TRUE,
-  longtable = TRUE,
-  align = c("m{2.5cm}", "m{2.5cm}", "m{9cm}"),
-  caption = "(ref:itemTableCaption)"
-)
-```
-
-\newpage
 
@@
 
+(ref:itemTableCaption) List of norm interpretation questions asked at Time 7.
+_These questions were preceded by the following text:_ "There may or may not be a difference between what people around you are doing and what they should be doing.
+You can learn about what people are doing and what they should be doing in different ways.
+For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both".
+_Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly._
+
+```{r itemTable, echo=F, warning=F, message=F}
+apa_table(
+  itemTable,
+  booktabs = TRUE,
+  longtable = TRUE,
+  #align = c("m{2.5cm}", "m{2.5cm}", "m{9cm}"),
+  caption = "(ref:itemTableCaption)"
+)
+```
+
+\newpage
+
 ## Supplementary References

### 2022-10-17 manuscript.Rmd
second draft with edits

@@
     equal_contrib : yes
+    corresponding : yes
+    address       : "1101 E 10th St, Bloomington, IN 47405, United States"
+    email         : "slheiman@iu.edu"
   - name          : "Scott Claessens"
@@
     equal_contrib : yes
-  - name          : "Edward R. Hurt"
+  - name          : "Jessica D. Ayers"
+    affiliation   : "3"
+  - name          : "Diego Guevara Beltran"
+    affiliation   : "3"
+  - name          : "Andrew Van Horn"
+    affiliation   : "4"
+  - name          : "Edward R. Hirt"
     affiliation   : "1"
+  - name          : "Athena Aktipis"
+    affiliation   : "3"
   - name          : "Peter M. Todd"
-    affiliation   : "1,3"
-    corresponding : yes
-    address       : "1101 E 10th St, Bloomington, IN 47405, United States"
-    email         : "pmtodd@indiana.edu"
+    affiliation   : "1,5"
 
@@
   - id            : "3"
+    institution   : "Department of Psychology, Arizona State University, United States"
+  - id            : "4"
+    institution   : "Department of Anthropology, Rutgers University - New Brunswick, United States"
+  - id            : "5"
     institution   : "Cognitive Science Program, Indiana University Bloomington, United States"
@@
 In particular, injunctive social norms prescribe what people _ought_ to do, whereas descriptive social norms inform what people _actually_ do.
-While previous experimental work has revealed people’s sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their influences on cooperative behavior over time.
+While previous experimental work has revealed people’s sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
 To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
 Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = `r nrow(tar_read(d))`), we tracked people’s reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
-Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior are tightly coupled and both change quickly in response to recommendations from public health authorities.
+Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior were tightly coupled and both changed quickly in response to recommendations from public health authorities.
 In addition, a random-intercept cross-lagged panel model revealed that descriptive norms, but not injunctive norms, caused future increases in mask wearing behavior.
@@
 keywords          : "descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
-wordcount         : "xxxx words"
+wordcount         : "4683 words"
 bibliography      : ["r-references.bib"]
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
 header-includes: 
@@
 
-Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Bicchieri2009; @Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], social norms have remained an elusive concept in the behavioral sciences and several open questions remain [@Fehr2018a; @VanKleef2019].
+Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Bicchieri2009; @Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], social norms have remained an elusive concept in the behavioral sciences and several open questions remain [@Fehr2018a; @Legros2020; @VanKleef2019].
 In the current work, we focus on two such questions.
 First, how do injunctive and descriptive norms emerge naturally over time within a population?
-Second, how do evolving injunctive and descriptive norms influence behavior?
+Second, how do evolving injunctive and descriptive norms causally influence behavior over time?
 
@@
 First, studies have not adequately controlled for other potential non-social influences on behavior, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
-For example, descriptive social norms might only affect cooperative behavior indirectly by influencing perceptions of the effectiveness of the behavior (i.e., factual beliefs) or individuals' moral preferences (i.e., personal normative beliefs).
+For example, social norms might only affect cooperative behavior indirectly by influencing perceptions of the effectiveness of the behavior (i.e., factual beliefs) or individuals' moral preferences (i.e., personal normative beliefs).
 Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population.
@@
 Before the pandemic, mask wearing was not a common behavior in the United States.
-In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a cooperative behavior that people should adopt to minimize the spread of COVID-19.
-But mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., "flattening the curve" to protect at-risk individuals).
+In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a cooperative behavior that people should adopt to minimize the spread of COVID-19 [@CDC2022].
+Mask wearing has individual benefits, but the CDC emphasized the cooperative aspects of mask wearing as a tool for reducing community spread [@CDC2020].
+Indeed, mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., "flattening the curve" to protect at-risk individuals).
 Thus, the evolution of mask wearing in the United States throughout the COVID-19 pandemic allows us to study, on a short timescale within a single population, the emergence of novel injunctive and descriptive social norms and the causal effects of these norms on cooperative behavior.
@@
 In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing [@Rudert2021].
-These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between between-person and within-person change over time [@Hamaker2015], nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
-These studies also do not control for the influences of non-social beliefs, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
+These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time [@Hamaker2015], nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
+These studies also do not adequately distinguish between the effects of social norms and the effects of other non-social beliefs that are also predicted to shape behavior, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
 
 Here, we use two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
-We aimed to answer two main research questions.
+Between September 2020 and October 2022, we asked participants to report their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
+We used these data to answer two main research questions.
 First, how do descriptive and injunctive mask wearing norms emerge and evolve over time in the United States population?
-Second, how do descriptive and injunctive mask wearing norms affect mask wearing behavior over time?
+Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing behavior over time?
 
@@
 Figure \@ref(fig:plotTimeline) plots self-reported mask wearing behavior and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum’s COVID-19 Timeline [@CDC2022b] and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).
+These events were obtained from the CDC Museum’s COVID-19 Timeline [@CDC2022] and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).
 
@@
 This item was measured across all eighteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
-(b) Points and line ranges indicate mean averages and standard errors for perceived descriptive mask wearing norms (blue) and perceived injunctive mask wearing norms (green).
+(b) Points and line ranges indicate mean averages and standard errors for perceived injunctive mask wearing norms (green) and perceived descriptive mask wearing norms (blue).
 These items were measured across eleven time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
@@
 Although social norms are measured on fewer occasions than mask wearing behavior, we can see that as mask wearing behavior decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
-Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, though to lower levels than before.
-In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (_b_ = `r printnum(fixef(m1.1)["DesNorms"])`, 95% confidence interval [`r printnum(conf1.1["DesNorms","2.5 %"])` `r printnum(conf1.1["DesNorms","97.5 %"])`]) and between mask wearing behavior and perceived injunctive mask wearing norms (_b_ = `r printnum(fixef(m1.2)["InjNorms"])`, 95% CI [`r printnum(conf1.2["InjNorms","2.5 %"])` `r printnum(conf1.2["InjNorms","97.5 %"])`]) across individuals and time points (Supplementary Figure \@ref(fig:plotCorBehNorm); Supplementary Table \@ref(tab:modelSummaryTable1)).
+Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, before declining again in 2022.
+In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (_b_ = `r printnum(fixef(m1.1)["DesNorms"])`, 95% confidence interval [`r printnum(conf1.1["DesNorms","2.5 %"])` `r printnum(conf1.1["DesNorms","97.5 %"])`], _p_ `r printp(summary(m1.1)$coefficients["DesNorms","Pr(>|t|)"])`) and between mask wearing behavior and perceived injunctive mask wearing norms (_b_ = `r printnum(fixef(m1.2)["InjNorms"])`, 95% CI [`r printnum(conf1.2["InjNorms","2.5 %"])` `r printnum(conf1.2["InjNorms","97.5 %"])`], _p_ `r printp(summary(m1.2)$coefficients["InjNorms","Pr(>|t|)"])`) across individuals and time points (Supplementary Figure \@ref(fig:plotCorBehNorm); Supplementary Table \@ref(tab:modelSummaryTable1)).
 
@@
 We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020.
-Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021.
-Finally, perceived social norms and mask wearing declined again after the CDC eased restrictions in March 2022.
+Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation following widespread vaccine availability in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021.
+Finally, perceived social norms and mask wearing declined again after the CDC eased mask wearing guidelines in March 2022.
 Impressively, shifts in mask wearing behavior and norms happened on the timescale of weeks.
@@
 To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables and controls: perceived descriptive mask wearing norms, perceived injunctive mask wearing norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior (see Supplementary Figure \@ref(fig:plotDAG) for justification of control variables).
+This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables (self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
+In line with our proposed causal model (Supplementary Figure \@ref(fig:plotDAG)), we also included political orientation as an exogenous time-invariant control.
 According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).
-Since we are primarily interested in the causal effects of social norms on behavior, we focus on these variables in what follows (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
+Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing behavior, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
 
@@
 
-(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Arrows represent within-person autoregressive and cross-lagged effects from the model, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
+(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Arrows represent within-person autoregressive and cross-lagged effects for mask wearing behavior and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
@@
 According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (Time 2, $\beta$ = `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.lower"])` `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_05" & pe$op == "~" & pe$rhs == "w3_02","pvalue"])`; Time 5, $\beta$ = `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.lower"])` `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_09" & pe$op == "~" & pe$rhs == "w3_05","pvalue"])`; Time 11, $\beta$ = `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.lower"])` `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_13" & pe$op == "~" & pe$rhs == "w3_11","pvalue"])`; Time 13, $\beta$ = `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.lower"])` `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_14" & pe$op == "~" & pe$rhs == "w3_13","pvalue"])`).
-These cross-lagged effects from descriptive norms to mask wearing diminish in 2022.
+In 2022, these cross-lagged effects from descriptive norms to mask wearing behavior diminished.
 We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms.
@@
 
-In our longitudinal analysis, we found that _descriptive_ norms, not injunctive norms, independently predicted future increases in mask wearing.
+In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future increases in mask wearing across multiple time points.
+These cross-lagged effects were independent of the effects of non-social beliefs and political orientation.
 In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects [@Rudert2021].
@@
 During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms [@Hogg2008; @Wellen1998].
-Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its height.
+Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its peak.
 Unsure of how to respond to a novel threatening situation, people began to look to their neighbors and adapted their mask wearing behavior accordingly.
@@
 One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous correlations.
-Another explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior [@Cialdini1991].
+Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior [@Cialdini1991].
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people’s behaviors, increasing the relative salience of injunctive norms.
 To test this idea, future research should expand our longitudinal cross-lagged approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
+It is also possible that since injunctive norm perceptions did influence later descriptive norm perceptions, people use injunctive norm information as a signal that aids in interpreting or perceiving other people's behavior as a descriptive norm.
+Future studies could also manipulate injunctive norm presence and determine the effect of these norms on how people interpret behavioral information.
 
@@
 
-# Methods
+# Materials and Methods
 
@@
 
-Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules.
+Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
 We measured perceived injunctive social norms as the mean average of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
@@
 First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
+Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
 These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these models using the _lme4_ R package [@Bates2015] and dealt with missing data via listwise deletion.
+We estimated these multilevel regression models using the _lme4_ R package [@Bates2015] and dealt with missing data via listwise deletion.
 
@@
 We estimated the random-intercept cross-lagged panel model using the _lavaan_ R package [@Rosseel2012].
-In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included five main variables in the model: perceived descriptive norms, perceived injunctive norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior.
+In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included three main variables (self-reported mask wearing behavior, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
 For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
@@
 
-SLH, ERH, and PMT conceptualized the study.
-SLH and SC oversaw the data curation, investigation, and methodology of the study, and wrote the first draft of the paper.
+SLH, JDA, DGB, AVH, ERH, AA, and PMT conceptualized the study.
+SLH, SC, JDA, DGB, and AVH oversaw the data curation, investigation, and methodology of the study.
+SLH and SC wrote the first draft of the paper.
 SC conducted the formal analysis and created all visualisations.
-ERH and PMT provided funding and supervision for the study.
+ERH, AA, and PMT provided funding and supervision for the study.
 All authors reviewed and edited the final draft of the paper.

### 2022-10-17 manuscript.tex
second draft with edits

@@
   pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic},
-  pdfauthor={Samantha L. Heiman,1, Scott Claessens,2, Edward R. Hurt1, \& Peter M. Todd1,3},
+  pdfauthor={Samantha L. Heiman,1, Scott Claessens,2, Edward R. Hirt1, \& Peter M. Todd1,3},
   pdflang={en-EN},
@@
 \title{Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic}
-\author{Samantha L. Heiman\textsuperscript{$\dagger{}$,1}, Scott Claessens\textsuperscript{$\dagger{}$,2}, Edward R. Hurt\textsuperscript{1}, \& Peter M. Todd\textsuperscript{1,3}}
+\author{Samantha L. Heiman\textsuperscript{$\dagger{}$,1}, Scott Claessens\textsuperscript{$\dagger{}$,2}, Edward R. Hirt\textsuperscript{1}, \& Peter M. Todd\textsuperscript{1,3}}
 \date{}
@@
 In particular, injunctive social norms prescribe what people \emph{ought} to do, whereas descriptive social norms inform what people \emph{actually} do.
-While previous experimental work has revealed people's sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their influences on cooperative behavior over time.
+While previous experimental work has revealed people's sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
 To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
 Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 915), we tracked people's reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
-Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior are tightly coupled and both change quickly in response to recommendations from public health authorities.
+Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior were tightly coupled and both changed quickly in response to recommendations from public health authorities.
 In addition, a random-intercept cross-lagged panel model revealed that descriptive norms, but not injunctive norms, caused future increases in mask wearing behavior.
@@
 First, how do injunctive and descriptive norms emerge naturally over time within a population?
-Second, how do evolving injunctive and descriptive norms influence behavior?
+Second, how do evolving injunctive and descriptive norms causally influence behavior over time?
 
@@
 In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (31).
-These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between between-person and within-person change over time (32), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
-These studies also do not control for the influences of non-social beliefs, such as factual beliefs and personal normative beliefs (27).
+These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time (32), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
+These studies also do not adequately distinguish between the effects of social norms and the effects of other non-social beliefs that are also predicted to shape behavior, such as factual beliefs and personal normative beliefs (27).
 
 Here, we use two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
-We aimed to answer two main research questions.
+Between September 2020 and October 2022, we asked participants to report their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
+We used these data to answer two main research questions.
 First, how do descriptive and injunctive mask wearing norms emerge and evolve over time in the United States population?
-Second, how do descriptive and injunctive mask wearing norms affect mask wearing behavior over time?
+Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing behavior over time?
 
@@
 Although social norms are measured on fewer occasions than mask wearing behavior, we can see that as mask wearing behavior decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
-Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, though to lower levels than before.
-In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (\emph{b} = 0.29, 95\% confidence interval {[}0.23 0.35{]}) and between mask wearing behavior and perceived injunctive mask wearing norms (\emph{b} = 0.26, 95\% CI {[}0.22 0.30{]}) across individuals and time points (Supplementary Figure \ref{fig:plotCorBehNorm}; Supplementary Table \ref{tab:modelSummaryTable1}).
+Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, before declining again in 2022.
+In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (\emph{b} = 0.29, 95\% confidence interval {[}0.23 0.35{]}, \emph{p} \textless{} .001) and between mask wearing behavior and perceived injunctive mask wearing norms (\emph{b} = 0.26, 95\% CI {[}0.22 0.30{]}, \emph{p} \textless{} .001) across individuals and time points (Supplementary Figure \ref{fig:plotCorBehNorm}; Supplementary Table \ref{tab:modelSummaryTable1}).
 
@@
 Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021.
-Finally, perceived social norms and mask wearing declined again after the CDC eased restrictions in March 2022.
+Finally, perceived social norms and mask wearing declined again after the CDC eased mask wearing guidelines in March 2022.
 Impressively, shifts in mask wearing behavior and norms happened on the timescale of weeks.
@@
 To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables and controls: perceived descriptive mask wearing norms, perceived injunctive mask wearing norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior (see Supplementary Figure \ref{fig:plotDAG} for justification of control variables).
+This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables (self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
+In line with our proposed causal model (Supplementary Figure \ref{fig:plotDAG}), we also included political orientation as an exogenous time-invariant control.
 According to established fit statistics, this model fitted the data well (root mean square error of approximation = 0.030, 95\% CI {[}0.028 0.033{]}; standardized root mean squared residual = 0.087; comparative fit index = 0.957).
-Since we are primarily interested in the causal effects of social norms on behavior, we focus on these variables in what follows (but see Supplementary Table \ref{tab:lavaanTable} for full list of estimated autoregressive and cross-lagged effects).
+Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing behavior, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \ref{tab:lavaanTable} for full list of estimated autoregressive and cross-lagged effects).
 
@@
 \includegraphics{manuscript_files/figure-latex/plotRICLPM-1.pdf}
-\caption{\label{fig:plotRICLPM}\emph{Results of ten-wave unconstrained random-intercept cross-lagged panel model.} Arrows represent within-person autoregressive and cross-lagged effects from the model, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
+\caption{\label{fig:plotRICLPM}\emph{Results of ten-wave unconstrained random-intercept cross-lagged panel model.} Arrows represent within-person autoregressive and cross-lagged effects for mask wearing behavior and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
@@
 According to recent effect size guidelines for cross-lagged panel models (34), the standardized beta coefficients for these cross-lagged effects were large (Time 2, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; Time 5, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; Time 11, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; Time 13, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
-These cross-lagged effects from descriptive norms to mask wearing diminish in 2022.
+In 2022, these cross-lagged effects from descriptive norms to mask wearing behavior diminished.
 We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms.
@@
 
-In our longitudinal analysis, we found that \emph{descriptive} norms, not injunctive norms, independently predicted future increases in mask wearing.
+In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future increases in mask wearing across multiple time points.
+These cross-lagged effects were independent of the effects of non-social beliefs and political orientation.
 In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects (31).
@@
 During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms (37, 38).
-Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its height.
+Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its peak.
 Unsure of how to respond to a novel threatening situation, people began to look to their neighbors and adapted their mask wearing behavior accordingly.
@@
 One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous correlations.
-Another explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior (2).
+Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior (2).
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people's behaviors, increasing the relative salience of injunctive norms.
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
 
@@
 First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze changes over time, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
+Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
 These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these models using the \emph{lme4} R package (43) and dealt with missing data via listwise deletion.
+We estimated these multilevel regression models using the \emph{lme4} R package (43) and dealt with missing data via listwise deletion.
 
@@
 We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package (45).
-In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included five main variables in the model: perceived descriptive norms, perceived injunctive norms, factual beliefs, personal normative beliefs, and self-reported mask wearing behavior.
+In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing behavior, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
 For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
@@
 A minority of one against a unanimous majority}.
-\emph{Psychological monographs: General and applied} \textbf{70}, 1 (1956).}
+\emph{Psychological Monographs: General and Applied} \textbf{70}, 1 (1956).}
 
@@
 Gerard, \href{https://doi.org/10.1037/h0046408}{A study of normative and informational social influences upon individual judgment}.
-\emph{The journal of abnormal and social psychology} \textbf{51}, 629--636 (1955).}
+\emph{The Journal of Abnormal and Social Psychology} \textbf{51}, 629--636 (1955).}
 
@@
 \CSLLeftMargin{19. }%
-\CSLRightInline{\&. R. Boyd Robert, \emph{Culture and the evolutionary process} (University of Chicago Press, 1985).}
+\CSLRightInline{R. Boyd, P. J. Richerson, \emph{Culture and the evolutionary process} (University of Chicago Press, 1985).}
 
@@
 Greenhill, R.
-Mace, \href{https://doi.org/10.1098/rspb.2009.0088}{Matrilocal residence is ancestral in austronesian societies}.
+Mace, \href{https://doi.org/10.1098/rspb.2009.0088}{Matrilocal residence is ancestral in {A}ustronesian societies}.
 \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{276}, 1957--1964 (2009).}
@@
 Grasman, \href{https://doi.org/10.1037/a0038889}{A critique of the cross-lagged panel model}.
-\emph{Psychological methods} \textbf{20}, 102 (2015).}
+\emph{Psychological Methods} \textbf{20}, 102 (2015).}
 
@@
 To evaluate the construct validity of our measures of perceived descriptive and injunctive norms, at Time 7 we asked participants to rate the extent to which each perceived norm item provided descriptive and injunctive information.
-For each item, participants were asked whether the item provided information about what people are doing, and whether the item provided information about what people should be doing.
+For each item, participants were asked whether the item provided information about what people \emph{are} doing, and whether the item provided information about what people \emph{should} be doing.
 Participants responded on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/plotUSMap-1.pdf}
-\caption{\label{fig:plotUSMap}\emph{Map of the United States with participant zip code locations.}}
+\includegraphics{manuscript_files/figure-latex/plotCorBehNorm-1.pdf}
+\caption{\label{fig:plotCorBehNorm}\emph{Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables.} Models contain random intercepts for participants and time points.
+Lines are fixed effect regression lines from multilevel models, shaded areas are 95\% confidence intervals.}
 \end{figure}
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/plotAttrition-1.pdf}
-\caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
+\includegraphics{manuscript_files/figure-latex/plotCDCSens-1.pdf}
+\caption{\label{fig:plotCDCSens}\emph{Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations.} These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
+Time is included as a continuous linear predictor, scaled between 0 and 1, with three forced change points (dashed lines).
+From the left, the first dashed line indicates when the CDC relaxed their mask wearing recommendations in March 2021, the second dashed line indicates when the CDC strengthened their mask wearing recommendations in July 2021, and the third dashed line indicates when the CDC updated their community levels and relaxed their mask wearing recommendations in March 2022.
+This results in the estimation of five fixed effect parameters: the initial intercept, the slope in the first window, the slope in the second window, the slope in the third window, and the slope is the fourth window.
+Bolded lines and shaded areas represent fixed effect regression lines from multilevel models and 95\% confidence intervals, respectively.}
 \end{figure}
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/plotCorBehNorm-1.pdf}
-\caption{\label{fig:plotCorBehNorm}\emph{Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables.} Lines are fixed effect regression lines from multilevel models, shaded areas are 95\% confidence intervals.}
+\includegraphics{manuscript_files/figure-latex/plotAttrition-1.pdf}
+\caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
 \end{figure}
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/plotCDCSens-1.pdf}
-\caption{\label{fig:plotCDCSens}\emph{Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations.} These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
-Time is included as a continuous linear predictor, scaled between 0 and 1, with two forced change points (dashed lines).
-The first dashed line (left) indicates when the CDC relaxed their mask wearing recommendations in March 2021, and the second dashed line (right) indicates when the CDC strengthened their mask wearing recommendations in July 2021.
-This results in the estimation of four fixed effect parameters: the initial intercept, the slope in the first window, the slope in the second window, and the slope in the third window.
-Bolded lines and shaded areas represent fixed effect regression lines from multilevel models and 95\% confidence intervals, respectively.}
+\includegraphics{manuscript_files/figure-latex/plotUSMap-1.pdf}
+\caption{\label{fig:plotUSMap}\emph{Map of the United States with participant zip code locations.}}
 \end{figure}
@@
 \subsection{Supplementary Tables}\label{supplementary-tables}}
-
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{longtable}{m{2.5cm}m{2.5cm}m{9cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:itemTable}List of norm interpretation questions asked at Time 7.
-\emph{These questions were preceded by the following text:} ``There may or may not be a difference between what people around you are doing and what they should be doing.
-You can learn about what people are doing and what they should be doing in different ways.
-For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both''.
-\emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
-\toprule
-Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:itemTable} continued}}\\
-\toprule
-Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
-\midrule
-\endhead
-Provides descriptive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone is doing?\\
- &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone is doing?\\
- & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone is doing?\\
- &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone is doing?\\
-Provides injunctive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone should be doing?\\
- &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone should be doing?\\
- & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone should be doing?\\
- &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone should be doing?\\
-\bottomrule
-\end{longtable}
-
-\end{ThreePartTable}
-\end{center}
-
-\newpage
 
@@
 
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{longtable}{m{2.5cm}m{2.5cm}m{9cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:itemTable}List of norm interpretation questions asked at Time 7.
+\emph{These questions were preceded by the following text:} ``There may or may not be a difference between what people around you are doing and what they should be doing.
+You can learn about what people are doing and what they should be doing in different ways.
+For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both''.
+\emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
+\toprule
+Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:itemTable} continued}}\\
+\toprule
+Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
+\midrule
+\endhead
+Provides descriptive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone is doing?\\
+ &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone is doing?\\
+ & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone is doing?\\
+ &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone is doing?\\
+Provides injunctive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone should be doing?\\
+ &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone should be doing?\\
+ & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone should be doing?\\
+ &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone should be doing?\\
+\bottomrule
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
+
 \hypertarget{supplementary-references}{%

### 2022-10-17 R/functions.R

@@ -557,7 +557,7 @@ plotTimeline <- function(d, owid) {
     annotate("text", x = ymd("2021-10-26"), y = 2.2, colour = "grey", size = 2.8, angle = 90, lineheight = 0.85,
              label = "CDC: recommend\nbooster shots\nfor adults") +
     annotate("text", x = ymd("2022-01-22"), y = 2.2, colour = "grey", size = 2.8, angle = 90, lineheight = 0.85,
-             label = "CDC: update\ncommunity\nlevels and ease\nrestrictions") +
+             label = "CDC: update\ncommunity levels\nand ease mask\nguidelines") +
     annotate("text", x = ymd("2022-04-10"), y = 1.8, colour = "grey", size = 2.8, angle = 90, lineheight = 0.85,
              label = "US hits\n1,000,000\ndeaths") +
     # add main geoms

### 2022-10-17 _targets.R

@@ -5,9 +5,9 @@ source("R/functions.R")
 options(tidyverse.quiet = TRUE)
 tar_option_set(packages = c("arm", "broom.mixed", "cowplot", "dagitty", "ggdag", 
                             "ggeffects", "ggraph", "huxtable", "jtools", "knitr",
-                            "kableExtra", "lavaan", "lme4", "lubridate", "MuMIn",
-                            "papaja", "rnaturalearth", "rnaturalearthdata", "scales", 
-                            "semTools", "sf", "zipcodeR"))
+                            "kableExtra", "lavaan", "lme4", "lmerTest", "lubridate", 
+                            "MuMIn", "papaja", "rnaturalearth", "rnaturalearthdata", 
+                            "scales", "semTools", "sf", "zipcodeR"))
 # workflow
 list(
   # files

### 2022-12-19 appendix.Rmd
submission

@@
 ---
-\renewcommand{\appendixname}{\textbf{Supplementary Materials}}
+\renewcommand{\appendixname}{\textbf{Supplementary Material}}
 \renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
@@
 
-(ref:plotCorBehNormCaption) _Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables._ Models contain random intercepts for participants and time points.
+(ref:plotDAGCaption) _Directed acyclic graph reflecting causal assumptions._ In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing (Mask).
+Perceptions of descriptive and injunctive norms also indirectly cause mask wearing through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
+Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
+Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.
+
+```{r plotDAG, echo=F, fig.width=5.6, fig.height=4.5, fig.cap="(ref:plotDAGCaption)"}
+plot4
+```
+
+\newpage
+
+(ref:plotCorBehNormCaption) _Predictions from multilevel models with self-reported mask wearing as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables._ Models contain random intercepts for participants and time points.
 Lines are fixed effect regression lines from multilevel models, shaded areas are 95% confidence intervals.
@@
 
-(ref:plotCDCSensCaption) _Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations._ These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
+(ref:plotCDCSensCaption) _Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations._ These models track temporal changes in (a) self-reported mask wearing, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
 Time is included as a continuous linear predictor, scaled between 0 and 1, with three forced change points (dashed lines).
@@
 plot6
-```
-
-\newpage
-
-(ref:plotDAGCaption) _Directed acyclic graph reflecting causal assumptions._ In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing behavior (Mask).
-Perceptions of descriptive and injunctive norms also indirectly cause mask wearing behavior through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
-Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
-Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
-Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.
-
-```{r plotDAG, echo=F, fig.width=5.6, fig.height=4.5, fig.cap="(ref:plotDAGCaption)"}
-plot4
 ```
@@
   longtable = TRUE,
+  font_size = "small",
   align = c("l", "c", "c", "c"),
@@
   longtable = TRUE,
-  #align = c("m{2.5cm}", "m{2.5cm}", "m{9cm}"),
+  align = c("m{2.5cm}", "m{2.5cm}", "m{9cm}"),
   caption = "(ref:itemTableCaption)"

### 2022-12-19 manuscript.Rmd
submission

@@
 ---
-title             : "Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic"
+title             : "Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic"
 shorttitle        : "Norms and mask wearing"
@@
   - name          : "Samantha L. Heiman"
-    affiliation   : "1"
-    equal_contrib : yes
-    corresponding : yes
-    address       : "1101 E 10th St, Bloomington, IN 47405, United States"
-    email         : "slheiman@iu.edu"
+    affiliation   : "*,1"
+    #corresponding : yes
+    #address       : "1101 E 10th St, Bloomington, IN 47405, United States"
+    #email         : "slheiman@iu.edu"
   - name          : "Scott Claessens"
-    affiliation   : "2"
-    equal_contrib : yes
+    affiliation   : "*,2"
   - name          : "Jessica D. Ayers"
@@
   - name          : "Diego Guevara Beltran"
-    affiliation   : "3"
+    affiliation   : "4"
   - name          : "Andrew Van Horn"
-    affiliation   : "4"
+    affiliation   : "5,6"
   - name          : "Edward R. Hirt"
@@
   - name          : "Athena Aktipis"
-    affiliation   : "3"
+    affiliation   : "†,4"
   - name          : "Peter M. Todd"
-    affiliation   : "1,5"
+    affiliation   : "†,1,7"
 
@@
   - id            : "1"
-    institution   : "Department of Psychological and Brain Sciences, Indiana University Bloomington, United States"
+    institution   : "\\footnotesize Department of Psychological and Brain Sciences, Indiana University Bloomington, United States"
   - id            : "2"
-    institution   : "School of Psychology, University of Auckland, New Zealand"
+    institution   : "\\footnotesize School of Psychology, University of Auckland, New Zealand"
   - id            : "3"
-    institution   : "Department of Psychology, Arizona State University, United States"
+    institution   : "\\footnotesize Department of Psychological Science, Boise State University, United States"
   - id            : "4"
-    institution   : "Department of Anthropology, Rutgers University - New Brunswick, United States"
+    institution   : "\\footnotesize Department of Psychology, Arizona State University, United States"
   - id            : "5"
-    institution   : "Cognitive Science Program, Indiana University Bloomington, United States"
+    institution   : "\\footnotesize Department of Physics, Case Western Reserve University, United States"
+  - id            : "6"
+    institution   : "\\footnotesize Department of Art History, Case Western Reserve University, United States"  
+  - id            : "7"
+    institution   : "\\footnotesize Cognitive Science Program, Indiana University Bloomington, United States"
 
 abstract: |
- Social norms allow humans to coordinate and cooperate in the face of existential threats.
-In particular, injunctive social norms prescribe what people _ought_ to do, whereas descriptive social norms inform what people _actually_ do.
-While previous experimental work has revealed people’s sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
-To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
-Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = `r nrow(tar_read(d))`), we tracked people’s reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
+ Human sociality is governed by two types of social norms: injunctive social norms, which prescribe what people _ought_ to do, and descriptive social norms, which reflect what people _actually_ do.
+These norms enable people to cooperate in the face of group-wide challenges.
+While previous experimental work has shown that people’s behavior is influenced by social norms, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
+To understand how cooperative behavior emerges and is shaped by changing social norms in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
+Mask wearing has individual benefits, but it is also a cooperative behavior that provides collective benefits of reduced disease transmission in the community.
+Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = `r nrow(tar_read(d))`), we tracked people’s reported mask wearing and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
 Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior were tightly coupled and both changed quickly in response to recommendations from public health authorities.
-In addition, a random-intercept cross-lagged panel model revealed that descriptive norms, but not injunctive norms, caused future increases in mask wearing behavior.
-These findings underline the relative importance of descriptive social norms in shaping cooperative behavior during uncertain times.
+In addition, a random-intercept cross-lagged panel model revealed that perceived descriptive norms, but not perceived injunctive norms, caused future within-person increases in individuals' mask wearing.
+These findings show that, during uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
 
-note              : "This working paper has not yet been peer-reviewed.
-This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation."
-keywords          : "descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
-wordcount         : "4683 words"
+note              : "\\footnotesize \\raggedright \\mbox{* indicates shared first authorship, † indicates shared senior authorship} \\par Correspondence concerning this article should be addressed to Samantha L.
+Heiman, 1101 E 10th St, Bloomington, IN 47405, United States.
+E-mail: slheiman@iu.edu \\par This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
+\\par This working paper has not yet been peer-reviewed."
+keywords          : "social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
+wordcount         : "5276 words"
 bibliography      : ["r-references.bib"]
@@
 documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_word
+classoption       : "man, donotrepeattitle"
+output            : papaja::apa6_pdf
 header-includes: 
@@
 knitr::opts_chunk$set(warning = FALSE, message = FALSE)
-tar_load_everything()
-# to ensure that the Rmd runs last, load appendix files again
-tar_load(c(d, plot1, plot2, plot4, plot5, plot6,
-           m1.1, m1.2, itemTable, lavaanTable,
-           changePointsTable, constructValDes, constructValInj))
+tar_load(c(fileData, fileOwid, d, owid,
+           constructValDes, constructValInj,
+           m1.1, m1.2, m2.1, m2.2, m2.3,
+           conf1.1, conf1.2, conf2.1, conf2.2, conf2.3,
+           pars2.1, pars2.2, pars2.3, riclpm, fitMeasures,
+           plot1, plot2, plot3, plot4, plot5, plot6, plot7,
+           itemTable, changePointsTable, lavaanTable))
 ```
 
-Social norms are a key aspect of human sociality.
-Broadly defined as commonly known behavioral guidelines enforced by groups of people, social norms have attracted the attention of behavioral scientists for decades [@Bicchieri2009; @Cialdini1991; @Fehr2018a; @Legros2020; @VanKleef2019].
-Since Asch's early studies of normative conformity [@Asch1956], evidence has accrued that humans are particularly attuned to social norms [@Tomasello2014].
-From a young age, children begin to adhere to and enforce group-wide social norms [@Jensen2014; @Rakoczy2013; @Schmidt2012] and both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms [@Vaish2011; @Schaumberg2022].
-This uniquely human sensitivity to social norms allows groups of unrelated people to cooperate and coordinate in the face of existential threats, such as resource scarcity, natural disasters, and infectious diseases [@Gelfand2011; @Henrich2015].
+# \normalfont Significance Statement
 
-Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms [@Bicchieri2009; @Cialdini1991; @Deutsch1955].
+Social norms have been identified as important drivers of human cooperation, but the emergence of social norms within populations and their subsequent effects on cooperative behavior are not well understood.
+Here, we use mask wearing during the COVID-19 pandemic as a real-world setting in which to study the emergence of social norms.
+Over two years in the United States, we find that social norms and mask wearing are tightly coupled and change quickly in response to public health recommendations.
+Moreover, longitudinal modeling suggests that mask wearing is causally preceded by perceptions of descriptive norms (i.e.
+what people _are_ doing) but not injunctive norms (i.e.
+what people _ought_ to be doing).
+
+\newpage
+
+\begin{center}Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic\end{center}
+
+Social norms are a key aspect of human sociality [@Bicchieri2009; @Cialdini1991; @Fehr2018a; @VanKleef2019].
+Broadly, social norms are defined as commonly known behavioral guidelines enforced by groups of people [@Legros2020].
+Since Asch's early studies of normative conformity [@Asch1956], evidence has accrued that humans, unlike other species, are highly attuned to social norms [@Tomasello2014; @Riedl2012].
+From a young age, children begin to adhere to and enforce group-wide social norms [@Jensen2014; @Rakoczy2013; @Schmidt2012].
+Both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms [@Vaish2011; @Schaumberg2022].
+People also readily punish third-party norm violators [@Fehr2004].
+Emerging from this uniquely normative psychology, social norms specify the rules and practices that govern human social life, forming the "grammar of society" [@Bicchieri2006].
+
+By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as inter-group conflict, resource scarcity, high population density, natural disasters, and infectious diseases [@Gelfand2011; @Henrich2015; @Roos2015].
+In the Maasai people of East Africa, for example, social norms governing transfers of livestock enable individuals to collectively manage the risks associated with nomadic pastoralism (e.g., famine, droughts) [@Aktipis2011].
+In countries with high population densities, social norms provide the rules and regularities that allow large groups of people to effectively align their behavior and avoid chaos and conflict [@Gelfand2011].
+Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans [@Henrich2015; @Chudek2011].
+
+Previous research has distinguished between two primary types of social norms: injunctive norms and descriptive norms [@Bicchieri2009; @Cialdini1991; @Deutsch1955].
 Injunctive social norms indicate what others in the group tend to approve or disapprove of and often involve social sanctions if violated.
-By contrast, descriptive norms are situational, simply describing what most people are doing in a given situation.
+By contrast, descriptive norms simply describe what most people are doing in a given situation.
 Though these two kinds of social norms tend to align, they can also be in conflict with one another.
@@
 However, if an individual observes that most people are leaving their litter behind at the site, the descriptive norm is to not clean up.
-It is thus possible for injunctive and descriptive norms to have independent effects on behavior [@Schultz2007].
+It is thus possible for injunctive and descriptive norms to differentially affect behavior [@Schultz2007].
 
-Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Bicchieri2009; @Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], social norms have remained an elusive concept in the behavioral sciences and several open questions remain [@Fehr2018a; @Legros2020; @VanKleef2019].
-In the current work, we focus on two such questions.
-First, how do injunctive and descriptive norms emerge naturally over time within a population?
+Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Bicchieri2009; @Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], open questions remain regarding the emergence and causal influence of social norms [@Fehr2018a; @Legros2020; @VanKleef2019].
+First, how do injunctive and descriptive norms emerge over time within a population?
 Second, how do evolving injunctive and descriptive norms causally influence behavior over time?
 
-With regards to norm emergence, research in cultural evolution and behavioral economics have begun to illuminate how social norms emerge over time.
-In the longer term, cultural evolutionary models show that injunctive social norms can be vertically transmitted down through generations via imitation or teaching, or horizontally diffused from neighboring human populations [@Boyd1985; @Henrich2015].
+Research in cultural evolution and behavioral economics has investigated how social norms emerge in a population over time.
+In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring human populations [@Boyd1985; @Henrich2015].
 For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership [@Kushnick2014] and post-marital residence [@Jordan2009].
-However, much less is known about how social norms arise endogenously within populations in the shorter term.
+However, less is known about how social norms arise endogenously within populations in the short term.
 Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019] and require peer enforcement to become stable [@Fehr2018b].
-But it remains unclear to what extent these findings generalize beyond the laboratory to real human populations.
+But it remains unclear whether these findings generalize beyond the laboratory to real human populations.
 
-With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the behavioral impact of social norms.
-For example, field experiments have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling [@Nigbur2010], paying taxes [@Larkin2018], and sustainably reusing towels in hotels [@Goldstein2008].
-Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals’ attention on injunctive norms [@Schultz2007].
+With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the impact of social norms on behavior.
+For example, field experiments in the United Kingdom and United States have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling [@Nigbur2010], paying taxes [@Larkin2018], and sustainably reusing towels in hotels [@Goldstein2008].
+Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals’ attention on opposing injunctive norms (e.g., seeing that the litter is swept into piles, showcasing that it should be cleaned up) [@Cialdini1991; @Schultz2007].
+
 However, these cross-sectional studies have two main limitations.
-First, studies have not adequately controlled for other potential non-social influences on behavior, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
-For example, social norms might only affect cooperative behavior indirectly by influencing perceptions of the effectiveness of the behavior (i.e., factual beliefs) or individuals' moral preferences (i.e., personal normative beliefs).
-Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population.
-An alternative way to identify causality, whilst retaining ecological validity, is to follow perceptions of social norms and cooperative behavior over time amidst a real, unfolding social dilemma.
-This is particularly important because social norms are not static: they change dynamically over time through processes of deliberation and social interaction [@Smith2015].
+First, studies have not adequately controlled for other potential non-social influences on behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
+Previous research has labeled these non-social beliefs as "factual beliefs" and "personal normative beliefs" respectively and emphasized that they should be distinguished from injunctive and descriptive social norms [@Bicchieri2014].
+For example, willingness to sustainably reuse towels might be driven by perceptions that towel reuse has a positive impact on the environment and/or personal beliefs that towel reuse is the right thing to do, rather than by social norms.
+Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point.
+But social norms are not static: they change dynamically over time through processes of deliberation and interaction [@Smith2015].
+An alternative way to assess causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
+In response to novel social dilemmas, social norms and cooperative behaviors emerge as people discover the benefits and costs of different behaviors [@Ostrom1990], enabling researchers to take multiple snapshots over time and establish temporal patterns.
+However, despite the promise of this approach, previous research has not studied the longitudinal interplay between social norms and cooperative behavior.
 
-To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, we studied mask wearing behavior during the COVID-19 pandemic.
+To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, in the current work we focus on mask wearing during the COVID-19 pandemic.
 Before the pandemic, mask wearing was not a common behavior in the United States.
-In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a cooperative behavior that people should adopt to minimize the spread of COVID-19 [@CDC2022].
-Mask wearing has individual benefits, but the CDC emphasized the cooperative aspects of mask wearing as a tool for reducing community spread [@CDC2020].
-Indeed, mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., "flattening the curve" to protect at-risk individuals).
-Thus, the evolution of mask wearing in the United States throughout the COVID-19 pandemic allows us to study, on a short timescale within a single population, the emergence of novel injunctive and descriptive social norms and the causal effects of these norms on cooperative behavior.
+In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a protective behavior that people should adopt to minimize the spread of the disease [@CDC2022].
+Mask wearing has individual benefits, but the CDC also emphasized the collective benefits of the behavior in reducing the spread of the disease throughout the community [@CDC2020].
+Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., "flattening the curve" to protect at-risk individuals).
+Thus, the development of mask wearing in the United States during the COVID-19 pandemic allows us to study the emergence of novel injunctive and descriptive social norms and their causal effects on cooperative behavior over a short timescale within a single population.
 
-Recent research has found positive relationships between perceptions of social norms and protective COVID-19 behaviors.
+Recent research has found positive relationships between individuals' perceptions of social norms and their protective COVID-19 behaviors.
 In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure [@Macy2021], and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions [@Bokemper2021].
 In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing [@Rudert2021].
-These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time [@Hamaker2015], nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
-These studies also do not adequately distinguish between the effects of social norms and the effects of other non-social beliefs that are also predicted to shape behavior, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
+These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time [@Hamaker2015], nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the pandemic.
+Furthermore, these studies do not distinguish between the effects of social norms and the effects of other non-social beliefs that are predicted to shape behavior, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
+Several of the studies also do not control for political ideology, which is important to account for since mask wearing was highly politicized in the United States during the COVID-19 pandemic [@BaxterKing2022].
 
@@
 Between September 2020 and October 2022, we asked participants to report their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
-We used these data to answer two main research questions.
-First, how do descriptive and injunctive mask wearing norms emerge and evolve over time in the United States population?
-Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing behavior over time?
+Guided by a causal model of social norms and mask wearing (Supplementary Figure \@ref(fig:plotDAG)), we also asked participants about their factual beliefs (i.e., whether they believe mask wearing is effective), personal normative beliefs (i.e., whether they believed mask wearing is the right thing to do), and political ideology, and controlled for these factors.
+We used our longitudinal data to answer our two main research questions in a specific real-world context.
+First, how do descriptive and injunctive mask wearing norms emerge and develop over time in the United States population during the pandemic?
+Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing over time?
 
@@
 To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration.
-Figure \@ref(fig:plotTimeline) plots self-reported mask wearing behavior and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum’s COVID-19 Timeline [@CDC2022] and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).
+Figure \@ref(fig:plotTimeline) plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
+These events were obtained from the CDC Museum’s COVID-19 Timeline [@CDC2022].
 
-(ref:plotTimelineCaption) _Timeline of self-reported mask wearing behavior and perceived social norms in the United States throughout the COVID-19 pandemic._ (a) Points and line ranges indicate mean averages and standard errors for the self-reported mask wearing item.
+(ref:plotTimelineCaption) _Timeline of self-reported mask wearing and perceived social norms in the United States during the COVID-19 pandemic._ (a) Points and line ranges indicate means ± two standard errors for the self-reported mask wearing item.
 This item was measured across all eighteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
-(b) Points and line ranges indicate mean averages and standard errors for perceived injunctive mask wearing norms (green) and perceived descriptive mask wearing norms (blue).
+(b) Points and line ranges indicate means ± two standard errors for perceived injunctive mask wearing norms (green) and perceived descriptive mask wearing norms (blue).
 These items were measured across eleven time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
-(c) Smoothed data for new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; https://ourworldindata.org/).
+(c) Smoothed data for daily new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; https://ourworldindata.org/).
 Across all panels, gray dashed lines represent significant pandemic-related events in the United States, such as vaccine approval from the Food and Drug Administration (FDA) and public health recommendations from the Centers for Disease Control and Prevention (CDC).
@@
 First, social norms and behavior were tightly coupled over time.
-Although social norms are measured on fewer occasions than mask wearing behavior, we can see that as mask wearing behavior decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
-Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, before declining again in 2022.
-In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (_b_ = `r printnum(fixef(m1.1)["DesNorms"])`, 95% confidence interval [`r printnum(conf1.1["DesNorms","2.5 %"])` `r printnum(conf1.1["DesNorms","97.5 %"])`], _p_ `r printp(summary(m1.1)$coefficients["DesNorms","Pr(>|t|)"])`) and between mask wearing behavior and perceived injunctive mask wearing norms (_b_ = `r printnum(fixef(m1.2)["InjNorms"])`, 95% CI [`r printnum(conf1.2["InjNorms","2.5 %"])` `r printnum(conf1.2["InjNorms","97.5 %"])`], _p_ `r printp(summary(m1.2)$coefficients["InjNorms","Pr(>|t|)"])`) across individuals and time points (Supplementary Figure \@ref(fig:plotCorBehNorm); Supplementary Table \@ref(tab:modelSummaryTable1)).
+Although social norms are measured on fewer occasions than mask wearing, we can see that as mask wearing decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
+Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing and perceived social norms, before declining again in 2022.
+In line with these patterns, multilevel regression models revealed positive correlations between mask wearing and perceived descriptive mask wearing norms (_b_ = `r printnum(fixef(m1.1)["DesNorms"])`, 95% confidence interval [`r printnum(conf1.1["DesNorms","2.5 %"])` `r printnum(conf1.1["DesNorms","97.5 %"])`], _p_ `r printp(summary(m1.1)$coefficients["DesNorms","Pr(>|t|)"])`) and between mask wearing and perceived injunctive mask wearing norms (_b_ = `r printnum(fixef(m1.2)["InjNorms"])`, 95% CI [`r printnum(conf1.2["InjNorms","2.5 %"])` `r printnum(conf1.2["InjNorms","97.5 %"])`], _p_ `r printp(summary(m1.2)$coefficients["InjNorms","Pr(>|t|)"])`) across individuals and time points (Supplementary Figure \@ref(fig:plotCorBehNorm); Supplementary Table \@ref(tab:modelSummaryTable1)).
 
-Second, fluctuations in mask wearing behavior and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States.
+Second, fluctuations in mask wearing and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States.
 We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020.
-Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation following widespread vaccine availability in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021.
+Perceived social norms and mask wearing subsequently declined after the CDC rescinded their mask wearing recommendation following widespread vaccine availability in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high-risk areas in August 2021.
 Finally, perceived social norms and mask wearing declined again after the CDC eased mask wearing guidelines in March 2022.
-Impressively, shifts in mask wearing behavior and norms happened on the timescale of weeks.
 These trends were confirmed by a series of multilevel regression models with change points aligning with changes in CDC mask wearing recommendations (Supplementary Figure \@ref(fig:plotCDCSens); Supplementary Table \@ref(tab:changePointsTable)).
 
-Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing behavior and perceived social norms over time.
-To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables (self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
+Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing and perceived social norms over time.
+To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
+This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
 In line with our proposed causal model (Supplementary Figure \@ref(fig:plotDAG)), we also included political orientation as an exogenous time-invariant control.
 According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).
-Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing behavior, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
+Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
 
@@
 
-Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing behavior and perceived social norms.
+Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
 On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`) and stronger injunctive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri2","pvalue"])`).
@@
 
-Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing behavior across the study duration, controlling for non-social beliefs and political orientation.
+Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing across the study duration, controlling for non-social beliefs and political orientation.
 In random intercept cross-lagged panel models, autoregressive effects represent "persistence" or "inertia" in within-person fluctuations from stable trait levels.
-In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the "stability" of the measure over time, which is captured by the random intercepts in our model).
+In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the "stable trait level" over time, which is captured by the random intercepts in our model).
+For example, an autoregressive effect from mask wearing in February 2021 to future mask wearing in June 2021 would suggest that wearing masks more than average in February predicts wearing masks more than average in June.
 By contrast, and most relevant for the current study, cross-lagged effects represent the effect of a within-person fluctuation in one measure on future within-person fluctuations in other measures.
 In other words, a positive cross-lagged effect indicates that being higher than average on one measure predicts being higher than average on _another_ measure in the following time point.
+For example, a cross-lagged effect from descriptive norms in February 2021 to future mask wearing in June 2021 would suggest that perceiving descriptive norms as stronger than average in February predicts wearing masks more than average in June.
 Cross-lagged effects are thus used to infer within-person causal influences over time.
 
-(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Arrows represent within-person autoregressive and cross-lagged effects for mask wearing behavior and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
+(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
@@
 Gray arrows indicate non-significant parameters.
+There are no significant direct paths from injunctive norms to future mask wearing, showing that people’s beliefs about what they should be doing did not have any direct influences on future mask wearing.
+On the other hand, there are significant paths from descriptive norms to future mask wearing, meaning that people’s beliefs about what others are doing influenced their future mask wearing.
 
@@
 
-In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing behavior.
-On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior.
-According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (Time 2, $\beta$ = `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.lower"])` `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_05" & pe$op == "~" & pe$rhs == "w3_02","pvalue"])`; Time 5, $\beta$ = `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.lower"])` `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_09" & pe$op == "~" & pe$rhs == "w3_05","pvalue"])`; Time 11, $\beta$ = `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.lower"])` `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_13" & pe$op == "~" & pe$rhs == "w3_11","pvalue"])`; Time 13, $\beta$ = `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.lower"])` `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_14" & pe$op == "~" & pe$rhs == "w3_13","pvalue"])`).
-In 2022, these cross-lagged effects from descriptive norms to mask wearing behavior diminished.
-We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms.
+In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing.
+On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
+According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (first wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.lower"])` `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_05" & pe$op == "~" & pe$rhs == "w3_02","pvalue"])`; second wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.lower"])` `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_09" & pe$op == "~" & pe$rhs == "w3_05","pvalue"])`; fourth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.lower"])` `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_13" & pe$op == "~" & pe$rhs == "w3_11","pvalue"])`; fifth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.lower"])` `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_14" & pe$op == "~" & pe$rhs == "w3_13","pvalue"])`).
+In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
+We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive norms.
 Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.
 
-However, the model reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did not predict future within-person changes in mask wearing behavior across the entire pandemic.
-All cross-lagged effects from perceived injunctive norms to mask wearing behavior are non-significant.
-Any causal effect that perceived injunctive norms might have had on future mask wearing behavior appears to be fully mediated by perceived descriptive norms.
-For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing behavior.
-But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing behavior over time within individuals.
+However, the model also reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did _not_ predict future within-person changes in mask wearing across our study duration.
+All cross-lagged effects from perceived injunctive norms to mask wearing are non-significant.
+Any causal effect that perceived injunctive norms might have had on future mask wearing appears to be fully mediated by perceived descriptive norms.
+This means that believing that others think that mask wearing is the right thing to do influences one’s perception of what others are actually doing, which then influences future behavior.
+For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing.
+But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing over time within individuals.
 
@@
 
-Using longitudinal data from the United States across the entire COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma.
-The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior are tightly coupled, and both change quickly in response to recommendations from public health authorities.
-The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior at the beginning of the pandemic.
-By contrast, injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.
+Using longitudinal data from the United States across two years of the COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma.
+The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior were tightly coupled and both changed dynamically in response to recommendations from public health authorities.
+Moreover, the results of our cross-lagged panel model indicate that descriptive norms caused future increases in mask wearing in the first year and a half of the pandemic.
+By contrast, injunctive norms were not directly causally related to future mask wearing over the entire two years of data collection.
 
-Our finding that social norms and mask wearing behavior are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction [@Titlestad2019].
-Moreover, the role of authorities, like the CDC, in shaping social norms and behavior supports the idea that institutions are part of the process by which culture and one’s own behaviors are mutually constructed [@Markus2010].
-The CDC explicitly releases guidelines for the benefit of public health, one of many situations in which the function of authorities is to coordinate behavior on large scales.
-However, even when institutions do not intend to coordinate behavior, they still may introduce influential social norms into the population (e.g., through business practices or products produced).
+Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction [@Titlestad2019].
+Moreover, the fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one’s own behaviors are mutually constructed [@Markus2010].
+Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms [@Amato2018].
+While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our descriptive trends have shown, can unfold over measurement intervals as short as four to six weeks.
 
-In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future increases in mask wearing across multiple time points.
+In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing across multiple time points.
 These cross-lagged effects were independent of the effects of non-social beliefs and political orientation.
-In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects [@Rudert2021].
-Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks [@Bokemper2021].
-Why this specific effect of descriptive norms on mask wearing?
-Descriptive norms are particularly useful for fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic [@Gelfand2015].
-During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms [@Hogg2008; @Wellen1998].
-Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its peak.
-Unsure of how to respond to a novel threatening situation, people began to look to their neighbors and adapted their mask wearing behavior accordingly.
-However, as the uncertainty surrounding the pandemic lessened in 2022, reliance on descriptive norms appeared to decrease.
+In line with this finding, similar descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) during the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects as we have done here [@Rudert2021].
+Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (i.e., a descriptive norm) [@Bokemper2021].
 
-Our finding that injunctive norms do not predict future mask wearing behavior is at odds with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults [@Macy2021].
-One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous correlations.
-Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior [@Cialdini1991].
+Why have descriptive norms had these positive effects on protective COVID-19 behaviors like mask wearing?
+Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic [@Gelfand2015].
+During times of uncertainty, people look to others to get information quickly about how they should behave, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms [@Hogg2008; @Wellen1998].
+Moreover, humans tend to be conditional cooperators who adapt their levels of cooperation depending on the degree of cooperative behavior in the population [@Chaudhuri2011].
+Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
+
+Our longitudinal model revealed that, at the between-person level, people who perceived stronger injunctive norms on average also reported more frequent mask wearing on average.
+This correlation is in line with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults [@Macy2021].
+However, at the within-person level, we found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
+This finding is at odds with previous evidence that experimentally-induced injunctive social norms can cause increases in cooperative behavior [@Schultz2007].
+
+One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous experimental results.
+Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were made more salient than injunctive norms during the pandemic.
+According to the focus theory of normative conduct [@Cialdini1991], this difference in norm salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people’s behaviors, increasing the relative salience of injunctive norms.
 To test this idea, future research should expand our longitudinal cross-lagged approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
-It is also possible that since injunctive norm perceptions did influence later descriptive norm perceptions, people use injunctive norm information as a signal that aids in interpreting or perceiving other people's behavior as a descriptive norm.
-Future studies could also manipulate injunctive norm presence and determine the effect of these norms on how people interpret behavioral information.
+
+While perceived injunctive norms did not directly predict future mask wearing, on one occasion they did indirectly predict future mask wearing _through_ perceived descriptive norms.
+In other words, perceptions of what others think people ought to be doing predicted future perceptions of what people were actually doing, which then predicted future mask wearing.
+This suggests that people may use injunctive norm information as a signal that they should check what others are doing, which could then aid in interpreting or perceiving other people's behavior as a descriptive norm.
+Future studies could manipulate injunctive norm presence and determine the effect of these norms on how people attend to and interpret behavioral information.
 
@@
 While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \@ref(fig:plotAttrition)).
-This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, different ethnicities, or those with different political ideologies.
+This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
 Since injunctive norm strength varies based on which group is seen as the source of the norm [@Neighbors2008], it would be interesting to learn whether different groups have different patterns of norm emergence over time.
@@
 Our results also might not generalize to all social norms, behaviors, and social dilemmas.
-Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more abstract and remote than the COVID-19 pandemic.
+Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more remote than the COVID-19 pandemic.
 For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.
 
-Regardless, in the case of mask wearing in the United States over the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and responded to both recommendations from authorities and current levels of cooperative behavior.
+For the case of mask wearing in the United States during the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and tracked ongoing changes in both recommendations from authorities and current levels of cooperative behavior.
 Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
-Importantly, this key finding slices both ways.
+Importantly, this key finding slices two ways.
 Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that _low_ local levels of mask wearing _discouraged_ future personal mask use.
@@
 From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
-This resulted in 18 unique time points of data collection throughout the pandemic.
+This resulted in 18 unique time points of data collection during the pandemic.
 The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
-`r sum(!is.na(d$StartDate.2))` of the initial `r nrow(d)` participants returned to complete the survey at Time 2, while `r sum(!is.na(d$StartDate.18))` participants continued through to Time 18 (see Supplementary Figure \@ref(fig:plotAttrition) for attrition rates across all time points).
+Of the initial `r nrow(d)` participants, `r sum(!is.na(d$StartDate.2))` returned to complete the survey at Time 2, while `r sum(!is.na(d$StartDate.18))` participants continued through to Time 18 (see Supplementary Figure \@ref(fig:plotAttrition) for attrition rates across all time points).
 
@@
 
-### Self-reported mask wearing behavior
+### Self-reported mask wearing
 
 At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing behavior by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+Following this question, participants self-reported their mask wearing by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
 
@@
 Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
-We measured perceived descriptive social norms as the mean average of the following two items: "_What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?_" and "_What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?_" These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
+We measured perceived descriptive social norms as the mean of the following two items: "_What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?_" and "_What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?_" These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
 
 Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
-We measured perceived injunctive social norms as the mean average of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
+We measured perceived injunctive social norms as the mean of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
 
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items.
+To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
 We asked participants whether each of the four items informed them about what people _are_ doing or what people _should_ be doing (i.e., giving descriptive or injunctive information).
@@
 To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \@ref(fig:plotDAG)).
-In this causal model, we included two kinds of non-social beliefs highlighted by previous research [@Bicchieri2014]: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the morally right thing to do).
-These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing behavior.
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research [@Bicchieri2014]: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
+These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
 In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States throughout the pandemic [@BaxterKing2022].
+This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic [@BaxterKing2022].
 Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
@@
 Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
-Factual beliefs were measured as the mean average of the following two items: "_I wear a face mask when going out in public to keep myself from getting sick_" and "_I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet_".
+Factual beliefs were measured as the mean of the following two items: "_I wear a face mask when going out in public to keep myself from getting sick_" and "_I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet_".
 Personal normative beliefs were measured with a single item: "_Wearing a face mask when going out in public is the right thing to do_".
@@
 Political orientation was measured in the first time point only.
-We measured political orientation as the mean average of the following two items: "_How would you describe your political orientation with regard to social issues?_" and "_How would you describe your political orientation with regard to economic issues?_".
+We measured political orientation as the mean of the following two items: "_How would you describe your political orientation with regard to social issues?_" and "_How would you describe your political orientation with regard to economic issues?_".
 These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
@@
 
-To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models.
-First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
+To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
+First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
+Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
 These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
@@
 This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time~t-1~ predicts being above average in another variable at time~t~.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time _t-1_ predicts being above average in another variable at time _t_.
 These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015; @Granger1969].
@@
 We estimated the random-intercept cross-lagged panel model using the _lavaan_ R package [@Rosseel2012].
-In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included three main variables (self-reported mask wearing behavior, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
+In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
 For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
-We modelled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
 We restricted the analysis to the ten time points with available data for all five variables.
@@
 This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
+We would also like to thank the Language, Culture, and Cognition lab at the University of Auckland for providing feedback on a previous version of this manuscript.
 
@@
 
-SLH, JDA, DGB, AVH, ERH, AA, and PMT conceptualized the study.
-SLH, SC, JDA, DGB, and AVH oversaw the data curation, investigation, and methodology of the study.
+SLH, JDA, DGB, AV, ERH, AA, and PMT conceptualized the study.
+SLH, SC, JDA, DGB, and AV oversaw the data curation, investigation, and methodology of the study.
 SLH and SC wrote the first draft of the paper.
-SC conducted the formal analysis and created all visualisations.
+SC conducted the formal analysis and created all visualizations.
 ERH, AA, and PMT provided funding and supervision for the study.

### 2022-12-19 manuscript.tex
submission

@@
 \documentclass[
-  man,floatsintext]{apa6}
+  man, donotrepeattitle,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 {}{\PatchFailed}
-\keywords{descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: xxxx words}
+\keywords{social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 5276 words}
 \usepackage{lineno}
@@
 \hypersetup{
-  pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic},
-  pdfauthor={Samantha L. Heiman,1, Scott Claessens,2, Edward R. Hirt1, \& Peter M. Todd1,3},
+  pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic},
+  pdfauthor={Samantha L. Heiman*,1, Scott Claessens*,2, Jessica D. Ayers3, Diego Guevara Beltran4, Andrew Van Horn5,6, Edward R. Hirt1, Athena Aktipis†,4, \& Peter M. Todd†,1,7},
   pdflang={en-EN},
-  pdfkeywords={descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation},
+  pdfkeywords={social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation},
   hidelinks,
@@
 
-\title{Descriptive, not injunctive, social norms caused increases in mask wearing throughout the COVID-19 pandemic}
-\author{Samantha L. Heiman\textsuperscript{$\dagger{}$,1}, Scott Claessens\textsuperscript{$\dagger{}$,2}, Edward R. Hirt\textsuperscript{1}, \& Peter M. Todd\textsuperscript{1,3}}
+\title{Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic}
+\author{Samantha L.
+Heiman\textsuperscript{*,1}, Scott Claessens\textsuperscript{*,2}, Jessica D.
+Ayers\textsuperscript{3}, Diego Guevara Beltran\textsuperscript{4}, Andrew Van Horn\textsuperscript{5,6}, Edward R.
+Hirt\textsuperscript{1}, Athena Aktipis\textsuperscript{†,4}, \& Peter M.
+Todd\textsuperscript{†,1,7}}
 \date{}
@@
 
-\authornote{
-
-\textsuperscript{$\dagger{}$} Samantha L. Heiman and Scott Claessens contributed equally to this work.
-
-Correspondence concerning this article should be addressed to Peter M.
-Todd, 1101 E 10th St, Bloomington, IN 47405, United States.
-E-mail: \href{mailto:pmtodd@indiana.edu}{\nolinkurl{pmtodd@indiana.edu}}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} \footnotesize Department of Psychological and Brain Sciences, Indiana University Bloomington, United States\\\textsuperscript{2} \footnotesize School of Psychology, University of Auckland, New Zealand\\\textsuperscript{3} \footnotesize Department of Psychological Science, Boise State University, United States\\\textsuperscript{4} \footnotesize Department of Psychology, Arizona State University, United States\\\textsuperscript{5} \footnotesize Department of Physics, Case Western Reserve University, United States\\\textsuperscript{6} \footnotesize Department of Art History, Case Western Reserve University, United States\\\textsuperscript{7} \footnotesize Cognitive Science Program, Indiana University Bloomington, United States}
+
+\note{
+
+\footnotesize
+
+\raggedright
+
+\mbox{* indicates shared first authorship, † indicates shared senior authorship}
+
+\par
+
+Correspondence concerning this article should be addressed to Samantha L.
+Heiman, 1101 E 10th St, Bloomington, IN 47405, United States.
+E-mail: \href{mailto:slheiman@iu.edu}{\nolinkurl{slheiman@iu.edu}}
+
+\par
+
+This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
+
+\par
+
+This working paper has not yet been peer-reviewed.
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Psychological and Brain Sciences, Indiana University Bloomington, United States\\\textsuperscript{2} School of Psychology, University of Auckland, New Zealand\\\textsuperscript{3} Cognitive Science Program, Indiana University Bloomington, United States}
-
-\note{This working paper has not yet been peer-reviewed.
-This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.}
-
 \abstract{%
-Social norms allow humans to coordinate and cooperate in the face of existential threats.
-In particular, injunctive social norms prescribe what people \emph{ought} to do, whereas descriptive social norms inform what people \emph{actually} do.
-While previous experimental work has revealed people's sensitivity to normative influence, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
-To understand how social norms emerge and shape behavior in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
-Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 915), we tracked people's reported mask wearing behavior and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
+Human sociality is governed by two types of social norms: injunctive social norms, which prescribe what people \emph{ought} to do, and descriptive social norms, which reflect what people \emph{actually} do.
+These norms enable people to cooperate in the face of group-wide challenges.
+While previous experimental work has shown that people's behavior is influenced by social norms, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
+To understand how cooperative behavior emerges and is shaped by changing social norms in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
+Mask wearing has individual benefits, but it is also a cooperative behavior that provides collective benefits of reduced disease transmission in the community.
+Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 915), we tracked people's reported mask wearing and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
 Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior were tightly coupled and both changed quickly in response to recommendations from public health authorities.
-In addition, a random-intercept cross-lagged panel model revealed that descriptive norms, but not injunctive norms, caused future increases in mask wearing behavior.
-These findings underline the relative importance of descriptive social norms in shaping cooperative behavior during uncertain times.
+In addition, a random-intercept cross-lagged panel model revealed that perceived descriptive norms, but not perceived injunctive norms, caused future within-person increases in individuals' mask wearing.
+These findings show that, during uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
 }
@@
 
-Social norms are a key aspect of human sociality.
-Broadly defined as commonly known behavioral guidelines enforced by groups of people, social norms have attracted the attention of behavioral scientists for decades (1--5).
-Since Asch's early studies of normative conformity (6), evidence has accrued that humans are particularly attuned to social norms (7).
-From a young age, children begin to adhere to and enforce group-wide social norms (8--10) and both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms (11, 12).
-This uniquely human sensitivity to social norms allows groups of unrelated people to cooperate and coordinate in the face of existential threats, such as resource scarcity, natural disasters, and infectious diseases (13, 14).
-
-Previous research has distinguished between two primary kinds of social norms: injunctive norms and descriptive norms (1, 2, 15).
+\hypertarget{significance-statement}{%
+\section{\texorpdfstring{\normalfont Significance Statement}{Significance Statement}}\label{significance-statement}}
+
+Social norms have been identified as important drivers of human cooperation, but the emergence of social norms within populations and their subsequent effects on cooperative behavior are not well understood.
+Here, we use mask wearing during the COVID-19 pandemic as a real-world setting in which to study the emergence of social norms.
+Over two years in the United States, we find that social norms and mask wearing are tightly coupled and change quickly in response to public health recommendations.
+Moreover, longitudinal modeling suggests that mask wearing is causally preceded by perceptions of descriptive norms (i.e.~what people \emph{are} doing) but not injunctive norms (i.e.~what people \emph{ought} to be doing).
+
+\newpage
+
+\begin{center}Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic\end{center}
+
+Social norms are a key aspect of human sociality (1--4).
+Broadly, social norms are defined as commonly known behavioral guidelines enforced by groups of people (5).
+Since Asch's early studies of normative conformity (6), evidence has accrued that humans, unlike other species, are highly attuned to social norms (7, 8).
+From a young age, children begin to adhere to and enforce group-wide social norms (9--11).
+Both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms (12, 13).
+People also readily punish third-party norm violators (14).
+Emerging from this uniquely normative psychology, social norms specify the rules and practices that govern human social life, forming the ``grammar of society'' (15).
+
+By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as inter-group conflict, resource scarcity, high population density, natural disasters, and infectious diseases (16--18).
+In the Maasai people of East Africa, for example, social norms governing transfers of livestock enable individuals to collectively manage the risks associated with nomadic pastoralism (e.g., famine, droughts) (19).
+In countries with high population densities, social norms provide the rules and regularities that allow large groups of people to effectively align their behavior and avoid chaos and conflict (16).
+Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans (17, 20).
+
+Previous research has distinguished between two primary types of social norms: injunctive norms and descriptive norms (1, 2, 21).
 Injunctive social norms indicate what others in the group tend to approve or disapprove of and often involve social sanctions if violated.
-By contrast, descriptive norms are situational, simply describing what most people are doing in a given situation.
+By contrast, descriptive norms simply describe what most people are doing in a given situation.
 Though these two kinds of social norms tend to align, they can also be in conflict with one another.
@@
 However, if an individual observes that most people are leaving their litter behind at the site, the descriptive norm is to not clean up.
-It is thus possible for injunctive and descriptive norms to have independent effects on behavior (16).
-
-Despite decades of research on the causes and consequences of injunctive and descriptive norms (1, 2, 16--18), social norms have remained an elusive concept in the behavioral sciences and several open questions remain (3, 5).
-In the current work, we focus on two such questions.
-First, how do injunctive and descriptive norms emerge naturally over time within a population?
+It is thus possible for injunctive and descriptive norms to differentially affect behavior (22).
+
+Despite decades of research on the causes and consequences of injunctive and descriptive norms (1, 2, 22--24), open questions remain regarding the emergence and causal influence of social norms (3--5).
+First, how do injunctive and descriptive norms emerge over time within a population?
 Second, how do evolving injunctive and descriptive norms causally influence behavior over time?
 
-With regards to norm emergence, research in cultural evolution and behavioral economics have begun to illuminate how social norms emerge over time.
-In the longer term, cultural evolutionary models show that injunctive social norms can be vertically transmitted down through generations via imitation or teaching, or horizontally diffused from neighboring human populations (14, 19).
-For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership (20) and post-marital residence (21).
-However, much less is known about how social norms arise endogenously within populations in the shorter term.
-Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (22) and require peer enforcement to become stable (23).
-But it remains unclear to what extent these findings generalize beyond the laboratory to real human populations.
-
-With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the behavioral impact of social norms.
-For example, field experiments have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (24), paying taxes (25), and sustainably reusing towels in hotels (26).
-Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals' attention on injunctive norms (16).
+Research in cultural evolution and behavioral economics has investigated how social norms emerge in a population over time.
+In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring human populations (17, 25).
+For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership (26) and post-marital residence (27).
+However, less is known about how social norms arise endogenously within populations in the short term.
+Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (28) and require peer enforcement to become stable (29).
+But it remains unclear whether these findings generalize beyond the laboratory to real human populations.
+
+With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the impact of social norms on behavior.
+For example, field experiments in the United Kingdom and United States have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (30), paying taxes (31), and sustainably reusing towels in hotels (32).
+Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals' attention on opposing injunctive norms (e.g., seeing that the litter is swept into piles, showcasing that it should be cleaned up) (2, 22).
+
 However, these cross-sectional studies have two main limitations.
-First, studies have not adequately controlled for other potential non-social influences on behavior, such as factual beliefs and personal normative beliefs (27).
-For example, descriptive social norms might only affect cooperative behavior indirectly by influencing perceptions of the effectiveness of the behavior (i.e., factual beliefs) or individuals' moral preferences (i.e., personal normative beliefs).
-Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point, and thus do not allow cooperative social norms to emerge and affect behavior naturally over time within a population.
-An alternative way to identify causality, whilst retaining ecological validity, is to follow perceptions of social norms and cooperative behavior over time amidst a real, unfolding social dilemma.
-This is particularly important because social norms are not static: they change dynamically over time through processes of deliberation and social interaction (28).
-
-To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, we studied mask wearing behavior during the COVID-19 pandemic.
+First, studies have not adequately controlled for other potential non-social influences on behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
+Previous research has labeled these non-social beliefs as ``factual beliefs'' and ``personal normative beliefs'' respectively and emphasized that they should be distinguished from injunctive and descriptive social norms (33).
+For example, willingness to sustainably reuse towels might be driven by perceptions that towel reuse has a positive impact on the environment and/or personal beliefs that towel reuse is the right thing to do, rather than by social norms.
+Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point.
+But social norms are not static: they change dynamically over time through processes of deliberation and interaction (34).
+An alternative way to assess causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
+In response to novel social dilemmas, social norms and cooperative behaviors emerge as people discover the benefits and costs of different behaviors (35), enabling researchers to take multiple snapshots over time and establish temporal patterns.
+However, despite the promise of this approach, previous research has not studied the longitudinal interplay between social norms and cooperative behavior.
+
+To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, in the current work we focus on mask wearing during the COVID-19 pandemic.
 Before the pandemic, mask wearing was not a common behavior in the United States.
-In April 2020, two months into the pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a cooperative behavior that people should adopt to minimize the spread of COVID-19.
-But mask wearing posed a social dilemma to individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., ``flattening the curve'' to protect at-risk individuals).
-Thus, the evolution of mask wearing in the United States throughout the COVID-19 pandemic allows us to study, on a short timescale within a single population, the emergence of novel injunctive and descriptive social norms and the causal effects of these norms on cooperative behavior.
-
-Recent research has found positive relationships between perceptions of social norms and protective COVID-19 behaviors.
-In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (29), and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions (30).
-In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (31).
-These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time (32), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the entire pandemic.
-These studies also do not adequately distinguish between the effects of social norms and the effects of other non-social beliefs that are also predicted to shape behavior, such as factual beliefs and personal normative beliefs (27).
+In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a protective behavior that people should adopt to minimize the spread of the disease (36).
+Mask wearing has individual benefits, but the CDC also emphasized the collective benefits of the behavior in reducing the spread of the disease throughout the community (37).
+Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., ``flattening the curve'' to protect at-risk individuals).
+Thus, the development of mask wearing in the United States during the COVID-19 pandemic allows us to study the emergence of novel injunctive and descriptive social norms and their causal effects on cooperative behavior over a short timescale within a single population.
+
+Recent research has found positive relationships between individuals' perceptions of social norms and their protective COVID-19 behaviors.
+In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (38), and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions (39).
+In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (40).
+These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time (41), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the pandemic.
+Furthermore, these studies do not distinguish between the effects of social norms and the effects of other non-social beliefs that are predicted to shape behavior, such as factual beliefs and personal normative beliefs (33).
+Several of the studies also do not control for political ideology, which is important to account for since mask wearing was highly politicized in the United States during the COVID-19 pandemic (42).
 
@@
 Between September 2020 and October 2022, we asked participants to report their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
-We used these data to answer two main research questions.
-First, how do descriptive and injunctive mask wearing norms emerge and evolve over time in the United States population?
-Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing behavior over time?
+Guided by a causal model of social norms and mask wearing (Supplementary Figure \ref{fig:plotDAG}), we also asked participants about their factual beliefs (i.e., whether they believe mask wearing is effective), personal normative beliefs (i.e., whether they believed mask wearing is the right thing to do), and political ideology, and controlled for these factors.
+We used our longitudinal data to answer our two main research questions in a specific real-world context.
+First, how do descriptive and injunctive mask wearing norms emerge and develop over time in the United States population during the pandemic?
+Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing over time?
 
@@
 To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration.
-Figure \ref{fig:plotTimeline} plots self-reported mask wearing behavior and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum's COVID-19 Timeline (33) and are included here based on their direct relevance to protective behaviors (e.g., changing CDC guidance on mask wearing, vaccination access, etc.).
+Figure \ref{fig:plotTimeline} plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
+These events were obtained from the CDC Museum's COVID-19 Timeline (36).
 
@@
 \includegraphics{manuscript_files/figure-latex/plotTimeline-1.pdf}
-\caption{\label{fig:plotTimeline}\emph{Timeline of self-reported mask wearing behavior and perceived social norms in the United States throughout the COVID-19 pandemic.} (a) Points and line ranges indicate mean averages and standard errors for the self-reported mask wearing item.
+\caption{\label{fig:plotTimeline}\emph{Timeline of self-reported mask wearing and perceived social norms in the United States during the COVID-19 pandemic.} (a) Points and line ranges indicate means ± two standard errors for the self-reported mask wearing item.
 This item was measured across all eighteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
-(b) Points and line ranges indicate mean averages and standard errors for perceived injunctive mask wearing norms (green) and perceived descriptive mask wearing norms (blue).
+(b) Points and line ranges indicate means ± two standard errors for perceived injunctive mask wearing norms (green) and perceived descriptive mask wearing norms (blue).
 These items were measured across eleven time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
-(c) Smoothed data for new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; \url{https://ourworldindata.org/}).
+(c) Smoothed data for daily new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; \url{https://ourworldindata.org/}).
 Across all panels, gray dashed lines represent significant pandemic-related events in the United States, such as vaccine approval from the Food and Drug Administration (FDA) and public health recommendations from the Centers for Disease Control and Prevention (CDC).}
@@
 First, social norms and behavior were tightly coupled over time.
-Although social norms are measured on fewer occasions than mask wearing behavior, we can see that as mask wearing behavior decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
-Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing behavior and perceived social norms, before declining again in 2022.
-In line with these patterns, multilevel regression models revealed positive correlations between mask wearing behavior and perceived descriptive mask wearing norms (\emph{b} = 0.29, 95\% confidence interval {[}0.23 0.35{]}, \emph{p} \textless{} .001) and between mask wearing behavior and perceived injunctive mask wearing norms (\emph{b} = 0.26, 95\% CI {[}0.22 0.30{]}, \emph{p} \textless{} .001) across individuals and time points (Supplementary Figure \ref{fig:plotCorBehNorm}; Supplementary Table \ref{tab:modelSummaryTable1}).
-
-Second, fluctuations in mask wearing behavior and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States.
+Although social norms are measured on fewer occasions than mask wearing, we can see that as mask wearing decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
+Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing and perceived social norms, before declining again in 2022.
+In line with these patterns, multilevel regression models revealed positive correlations between mask wearing and perceived descriptive mask wearing norms (\emph{b} = 0.29, 95\% confidence interval {[}0.23 0.35{]}, \emph{p} \textless{} .001) and between mask wearing and perceived injunctive mask wearing norms (\emph{b} = 0.26, 95\% CI {[}0.22 0.30{]}, \emph{p} \textless{} .001) across individuals and time points (Supplementary Figure \ref{fig:plotCorBehNorm}; Supplementary Table \ref{tab:modelSummaryTable1}).
+
+Second, fluctuations in mask wearing and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States.
 We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020.
-Perceived social norms and mask wearing behavior subsequently declined after the CDC rescinded their mask wearing recommendation in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high risk areas in August 2021.
+Perceived social norms and mask wearing subsequently declined after the CDC rescinded their mask wearing recommendation following widespread vaccine availability in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high-risk areas in August 2021.
 Finally, perceived social norms and mask wearing declined again after the CDC eased mask wearing guidelines in March 2022.
-Impressively, shifts in mask wearing behavior and norms happened on the timescale of weeks.
 These trends were confirmed by a series of multilevel regression models with change points aligning with changes in CDC mask wearing recommendations (Supplementary Figure \ref{fig:plotCDCSens}; Supplementary Table \ref{tab:changePointsTable}).
 
-Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing behavior and perceived social norms over time.
-To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing behavior, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from trait levels for our main variables (self-reported mask wearing behavior, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
+Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing and perceived social norms over time.
+To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
+This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
 In line with our proposed causal model (Supplementary Figure \ref{fig:plotDAG}), we also included political orientation as an exogenous time-invariant control.
 According to established fit statistics, this model fitted the data well (root mean square error of approximation = 0.030, 95\% CI {[}0.028 0.033{]}; standardized root mean squared residual = 0.087; comparative fit index = 0.957).
-Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing behavior, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \ref{tab:lavaanTable} for full list of estimated autoregressive and cross-lagged effects).
-
-Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing behavior and perceived social norms.
+Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \ref{tab:lavaanTable} for full list of estimated autoregressive and cross-lagged effects).
+
+Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
 On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (\emph{r} = 0.19, 95\% CI {[}0.04 0.33{]}, \emph{p} = .019) and stronger injunctive mask wearing norms (\emph{r} = 0.27, 95\% CI {[}0.14 0.40{]}, \emph{p} \textless{} .001).
@@
 
-Regarding within-person dynamics over time, Figure \ref{fig:plotRICLPM} displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing behavior across the study duration, controlling for non-social beliefs and political orientation.
+Regarding within-person dynamics over time, Figure \ref{fig:plotRICLPM} displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing across the study duration, controlling for non-social beliefs and political orientation.
 In random intercept cross-lagged panel models, autoregressive effects represent ``persistence'' or ``inertia'' in within-person fluctuations from stable trait levels.
-In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the ``stability'' of the measure over time, which is captured by the random intercepts in our model).
+In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the ``stable trait level'' over time, which is captured by the random intercepts in our model).
+For example, an autoregressive effect from mask wearing in February 2021 to future mask wearing in June 2021 would suggest that wearing masks more than average in February predicts wearing masks more than average in June.
 By contrast, and most relevant for the current study, cross-lagged effects represent the effect of a within-person fluctuation in one measure on future within-person fluctuations in other measures.
 In other words, a positive cross-lagged effect indicates that being higher than average on one measure predicts being higher than average on \emph{another} measure in the following time point.
+For example, a cross-lagged effect from descriptive norms in February 2021 to future mask wearing in June 2021 would suggest that perceiving descriptive norms as stronger than average in February predicts wearing masks more than average in June.
 Cross-lagged effects are thus used to infer within-person causal influences over time.
@@
 \includegraphics{manuscript_files/figure-latex/plotRICLPM-1.pdf}
-\caption{\label{fig:plotRICLPM}\emph{Results of ten-wave unconstrained random-intercept cross-lagged panel model.} Arrows represent within-person autoregressive and cross-lagged effects for mask wearing behavior and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
+\caption{\label{fig:plotRICLPM}\emph{Results of ten-wave unconstrained random-intercept cross-lagged panel model.} Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
 Bolded arrows indicate significantly positive parameters, \emph{p} \textless{} 0.05.
-Gray arrows indicate non-significant parameters.}
+Gray arrows indicate non-significant parameters.
+There are no significant direct paths from injunctive norms to future mask wearing, showing that people's beliefs about what they should be doing did not have any direct influences on future mask wearing.
+On the other hand, there are significant paths from descriptive norms to future mask wearing, meaning that people's beliefs about what others are doing influenced their future mask wearing.}
 \end{figure}
 
-In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing behavior.
-On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing behavior.
-According to recent effect size guidelines for cross-lagged panel models (34), the standardized beta coefficients for these cross-lagged effects were large (Time 2, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; Time 5, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; Time 11, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; Time 13, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
-In 2022, these cross-lagged effects from descriptive norms to mask wearing behavior diminished.
-We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing behavior predicted future within-person increases in perceived descriptive norms.
+In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing.
+On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
+According to recent effect size guidelines for cross-lagged panel models (43), the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
+In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
+We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive norms.
 Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.
 
-However, the model reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did not predict future within-person changes in mask wearing behavior across the entire pandemic.
-All cross-lagged effects from perceived injunctive norms to mask wearing behavior are non-significant.
-Any causal effect that perceived injunctive norms might have had on future mask wearing behavior appears to be fully mediated by perceived descriptive norms.
-For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing behavior.
-But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing behavior over time within individuals.
+However, the model also reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did \emph{not} predict future within-person changes in mask wearing across our study duration.
+All cross-lagged effects from perceived injunctive norms to mask wearing are non-significant.
+Any causal effect that perceived injunctive norms might have had on future mask wearing appears to be fully mediated by perceived descriptive norms.
+This means that believing that others think that mask wearing is the right thing to do influences one's perception of what others are actually doing, which then influences future behavior.
+For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing.
+But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing over time within individuals.
 
@@
 
-Using longitudinal data from the United States across the entire COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma.
-The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior are tightly coupled, and both change quickly in response to recommendations from public health authorities.
-The results of our structural equation model also indicate that descriptive norms caused future increases in mask wearing behavior at the beginning of the pandemic.
-By contrast, injunctive norms were not directly causally related to future mask wearing behavior over the course of the pandemic.
-
-Our finding that social norms and mask wearing behavior are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (22).
-Moreover, the role of authorities, like the CDC, in shaping social norms and behavior supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (35).
-The CDC explicitly releases guidelines for the benefit of public health, one of many situations in which the function of authorities is to coordinate behavior on large scales.
-However, even when institutions do not intend to coordinate behavior, they still may introduce influential social norms into the population (e.g., through business practices or products produced).
-
-In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future increases in mask wearing across multiple time points.
+Using longitudinal data from the United States across two years of the COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma.
+The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior were tightly coupled and both changed dynamically in response to recommendations from public health authorities.
+Moreover, the results of our cross-lagged panel model indicate that descriptive norms caused future increases in mask wearing in the first year and a half of the pandemic.
+By contrast, injunctive norms were not directly causally related to future mask wearing over the entire two years of data collection.
+
+Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (28).
+Moreover, the fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (44).
+Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms (45).
+While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our descriptive trends have shown, can unfold over measurement intervals as short as four to six weeks.
+
+In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing across multiple time points.
 These cross-lagged effects were independent of the effects of non-social beliefs and political orientation.
-In line with this finding, descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) throughout the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects (31).
-Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (30).
-Why this specific effect of descriptive norms on mask wearing?
-Descriptive norms are particularly useful for fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (36).
-During times of uncertainty, people look to others to quickly coordinate their behavior, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms (37, 38).
-Supporting this situational-uncertainty explanation, our model revealed that descriptive norms predicted future mask wearing at the beginning of the pandemic, when uncertainty was likely at its peak.
-Unsure of how to respond to a novel threatening situation, people began to look to their neighbors and adapted their mask wearing behavior accordingly.
-However, as the uncertainty surrounding the pandemic lessened in 2022, reliance on descriptive norms appeared to decrease.
-
-Our finding that injunctive norms do not predict future mask wearing behavior is at odds with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults (29).
-One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous correlations.
-Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing behavior were made more salient than injunctive norms throughout the pandemic, and therefore had a greater influence on behavior (2).
+In line with this finding, similar descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) during the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects as we have done here (40).
+Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (i.e., a descriptive norm) (39).
+
+Why have descriptive norms had these positive effects on protective COVID-19 behaviors like mask wearing?
+Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (46).
+During times of uncertainty, people look to others to get information quickly about how they should behave, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms (47, 48).
+Moreover, humans tend to be conditional cooperators who adapt their levels of cooperation depending on the degree of cooperative behavior in the population (49).
+Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
+
+Our longitudinal model revealed that, at the between-person level, people who perceived stronger injunctive norms on average also reported more frequent mask wearing on average.
+This correlation is in line with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults (38).
+However, at the within-person level, we found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
+This finding is at odds with previous evidence that experimentally-induced injunctive social norms can cause increases in cooperative behavior (22).
+
+One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous experimental results.
+Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were made more salient than injunctive norms during the pandemic.
+According to the focus theory of normative conduct (2), this difference in norm salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people's behaviors, increasing the relative salience of injunctive norms.
@@
 
+While perceived injunctive norms did not directly predict future mask wearing, on one occasion they did indirectly predict future mask wearing \emph{through} perceived descriptive norms.
+In other words, perceptions of what others think people ought to be doing predicted future perceptions of what people were actually doing, which then predicted future mask wearing.
+This suggests that people may use injunctive norm information as a signal that they should check what others are doing, which could then aid in interpreting or perceiving other people's behavior as a descriptive norm.
+Future studies could manipulate injunctive norm presence and determine the effect of these norms on how people attend to and interpret behavioral information.
+
 We are limited in generalizing these findings due to the constraints of our sample and the variables considered.
 While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \ref{fig:plotAttrition}).
-This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, different ethnicities, or those with different political ideologies.
-Since injunctive norm strength varies based on which group is seen as the source of the norm (39), it would be interesting to learn whether different groups have different patterns of norm emergence over time.
-In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors (40).
-Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large (41).
+This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
+Since injunctive norm strength varies based on which group is seen as the source of the norm (50), it would be interesting to learn whether different groups have different patterns of norm emergence over time.
+In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors (42).
+Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large (51).
 Our results also might not generalize to all social norms, behaviors, and social dilemmas.
-Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more abstract and remote than the COVID-19 pandemic.
+Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more remote than the COVID-19 pandemic.
 For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.
 
-Regardless, in the case of mask wearing in the United States over the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and responded to both recommendations from authorities and current levels of cooperative behavior.
+For the case of mask wearing in the United States during the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and tracked ongoing changes in both recommendations from authorities and current levels of cooperative behavior.
 Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
-Importantly, this key finding slices both ways.
+Importantly, this key finding slices two ways.
 Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that \emph{low} local levels of mask wearing \emph{discouraged} future personal mask use.
-This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community (42).
+This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community (52).
 Our work thus underscores the importance of consistent, visible community adherence for encouraging personal protective behaviors in response to global pandemics like COVID-19.
@@
 From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
-This resulted in 18 unique time points of data collection throughout the pandemic.
+This resulted in 18 unique time points of data collection during the pandemic.
 The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
-634 of the initial 915 participants returned to complete the survey at Time 2, while 347 participants continued through to Time 18 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
+Of the initial 915 participants, 634 returned to complete the survey at Time 2, while 347 participants continued through to Time 18 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
 
@@
 
-\hypertarget{self-reported-mask-wearing-behavior}{%
-\subsubsection{Self-reported mask wearing behavior}\label{self-reported-mask-wearing-behavior}}
+\hypertarget{self-reported-mask-wearing}{%
+\subsubsection{Self-reported mask wearing}\label{self-reported-mask-wearing}}
 
 At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing behavior by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+Following this question, participants self-reported their mask wearing by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
 
@@
 Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
-We measured perceived descriptive social norms as the mean average of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
-
-Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules.
-We measured perceived injunctive social norms as the mean average of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
-
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the social norm items.
+We measured perceived descriptive social norms as the mean of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
+
+Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
+We measured perceived injunctive social norms as the mean of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
+
+To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
 We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
@@
 To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
-In this causal model, we included two kinds of non-social beliefs highlighted by previous research (27): factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the morally right thing to do).
-These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing behavior.
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research (33): factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
+These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
 In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States throughout the pandemic (40).
+This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic (42).
 Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
@@
 Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
-Factual beliefs were measured as the mean average of the following two items: ``\emph{I wear a face mask when going out in public to keep myself from getting sick}'' and ``\emph{I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet}''.
+Factual beliefs were measured as the mean of the following two items: ``\emph{I wear a face mask when going out in public to keep myself from getting sick}'' and ``\emph{I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet}''.
 Personal normative beliefs were measured with a single item: ``\emph{Wearing a face mask when going out in public is the right thing to do}''.
@@
 Political orientation was measured in the first time point only.
-We measured political orientation as the mean average of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
+We measured political orientation as the mean of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
 These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
@@
 
-To analyze average trends in self-reported mask wearing behavior and perceived social norms, we fitted several multilevel regression models.
-First, to determine whether mask wearing behavior and social norms were coupled over time, we regressed mask wearing behavior on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing behavior and perceived social norms onto a continuous time predictor.
+To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
+First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
+Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
 These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these multilevel regression models using the \emph{lme4} R package (43) and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data (32).
+We estimated these multilevel regression models using the \emph{lme4} R package (53) and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data (41).
 This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time\textsubscript{t-1} predicts being above average in another variable at time\textsubscript{t}.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets (32, 44).
-
-We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package (45).
-In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing behavior, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time \emph{t-1} predicts being above average in another variable at time \emph{t}.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets (41, 54).
+
+We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package (55).
+In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
 For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
-We modelled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
 We restricted the analysis to the ten time points with available data for all five variables.
@@
 
-All analyses were conducted in R v4.1.1 (46).
-Visualizations were generated using the \emph{cowplot} (47) and \emph{ggplot2} (48) packages.
-The manuscript was reproducibly generated using the \emph{targets} (49) and \emph{papaja} (50) packages.
+All analyses were conducted in R v4.1.1 (56).
+Visualizations were generated using the \emph{cowplot} (57) and \emph{ggplot2} (58) packages.
+The manuscript was reproducibly generated using the \emph{targets} (59) and \emph{papaja} (60) packages.
 
@@
 This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
+We would also like to thank the Language, Culture, and Cognition lab at the University of Auckland for providing feedback on a previous version of this manuscript.
 
@@
 
-SLH, ERH, and PMT conceptualized the study.
-SLH and SC oversaw the data curation, investigation, and methodology of the study, and wrote the first draft of the paper.
-SC conducted the formal analysis and created all visualisations.
-ERH and PMT provided funding and supervision for the study.
+SLH, JDA, DGB, AV, ERH, AA, and PMT conceptualized the study.
+SLH, SC, JDA, DGB, and AV oversaw the data curation, investigation, and methodology of the study.
+SLH and SC wrote the first draft of the paper.
+SC conducted the formal analysis and created all visualizations.
+ERH, AA, and PMT provided funding and supervision for the study.
 All authors reviewed and edited the final draft of the paper.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Legros2020}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
 \CSLLeftMargin{4. }%
-\CSLRightInline{S.
-Legros, B.
-Cislaghi, \href{https://doi.org/10.1177/1745691619866455}{Mapping the social-norms literature: An overview of reviews}.
-\emph{Perspectives on Psychological Science} \textbf{15}, 62--80 (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
-\CSLLeftMargin{5. }%
 \CSLRightInline{G.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Legros2020}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{S.
+Legros, B.
+Cislaghi, \href{https://doi.org/10.1177/1745691619866455}{Mapping the social-norms literature: An overview of reviews}.
+\emph{Perspectives on Psychological Science} \textbf{15}, 62--80 (2020).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Asch1956}{}}%
@@
 A minority of one against a unanimous majority}.
-\emph{Psychological Monographs: General and Applied} \textbf{70}, 1 (1956).}
+\emph{Psychological Monographs: General and Applied} \textbf{70}, 1--70 (1956).}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Riedl2012}{}}%
+\CSLLeftMargin{8. }%
+\CSLRightInline{K.
+Riedl, K.
+Jensen, J.
+Call, M.
+Tomasello, \href{https://doi.org/10.1073/pnas.1203179109}{No third-party punishment in chimpanzees}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{109}, 14824--14829 (2012).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Jensen2014}{}}%
-\CSLLeftMargin{8. }%
+\CSLLeftMargin{9. }%
 \CSLRightInline{K.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Rakoczy2013}{}}%
-\CSLLeftMargin{9. }%
+\CSLLeftMargin{10. }%
 \CSLRightInline{H. Rakoczy, M. F. H. Schmidt, \href{https://doi.org/10.1111/cdep.12010}{The early ontogeny of social norms}. \emph{Child Development Perspectives} \textbf{7}, 17--21 (2013).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Schmidt2012}{}}%
-\CSLLeftMargin{10. }%
+\CSLLeftMargin{11. }%
 \CSLRightInline{M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Vaish2011}{}}%
-\CSLLeftMargin{11. }%
+\CSLLeftMargin{12. }%
 \CSLRightInline{A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Schaumberg2022}{}}%
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{13. }%
 \CSLRightInline{R.
@@
 Skowronek, \href{https://doi.org/10.1177/09567976221075303}{Shame broadcasts social norms: The positive social effects of shame on norm acquisition and normative behavior}.
-\emph{Psychological Science}, 09567976221075303 (2022).}
+\emph{Psychological Science} \textbf{33}, 1257--1277 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fehr2004}{}}%
+\CSLLeftMargin{14. }%
+\CSLRightInline{E. Fehr, U. Fischbacher, \href{https://doi.org/10.1016/S1090-5138(04)00005-4}{Third-party punishment and social norms}. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2006}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{C.
+Bicchieri, \emph{The grammar of society: The nature and dynamics of social norms} (Cambridge University Press, 2005) https:/doi.org/\href{https://doi.org/10.1017/CBO9780511616037}{10.1017/CBO9780511616037}.}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Gelfand2011}{}}%
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{16. }%
 \CSLRightInline{M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
-\CSLLeftMargin{14. }%
-\CSLRightInline{J. Henrich, \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter} (Princeton University Press, 2015).}
+\CSLLeftMargin{17. }%
+\CSLRightInline{J.
+Henrich, \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter} (Princeton University Press, 2015) https:/doi.org/\href{https://doi.org/10.2307/j.ctvc77f0d}{10.2307/j.ctvc77f0d}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Roos2015}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{P.
+Roos, M.
+Gelfand, D.
+Nau, J.
+Lun, \href{https://doi.org/10.1016/j.obhdp.2015.01.003}{Societal threat and cultural variation in the strength of social norms: An evolutionary basis}.
+\emph{Organizational Behavior and Human Decision Processes} \textbf{129}, 14--23 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Aktipis2011}{}}%
+\CSLLeftMargin{19. }%
+\CSLRightInline{A.
+Aktipis, L.
+Cronk, R.
+de Aguiar, \href{https://doi.org/10.1007/s10745-010-9364-9}{Risk-pooling and herd survival: An agent-based model of a {M}aasai gift-giving system}.
+\emph{Human Ecology} \textbf{39}, 131--140 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Chudek2011}{}}%
+\CSLLeftMargin{20. }%
+\CSLRightInline{M.
+Chudek, J.
+Henrich, \href{https://doi.org/10.1016/j.tics.2011.03.003}{Culture--gene coevolution, norm-psychology and the emergence of human prosociality}.
+\emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Schultz2007}{}}%
-\CSLLeftMargin{16. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Cialdini2007}{}}%
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Cialdini2021}{}}%
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Boyd1985}{}}%
-\CSLLeftMargin{19. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{R. Boyd, P. J. Richerson, \emph{Culture and the evolutionary process} (University of Chicago Press, 1985).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Kushnick2014}{}}%
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Jordan2009}{}}%
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{F.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{K.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Fehr2018b}{}}%
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{D.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{L.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Ostrom1990}{}}%
+\CSLLeftMargin{35. }%
+\CSLRightInline{E.
+Ostrom, \emph{Governing the commons: The evolution of institutions for collective action} (Cambridge University Press, 1990) https:/doi.org/\href{https://doi.org/10.1017/CBO9780511807763}{10.1017/CBO9780511807763}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CDC2022}{}}%
+\CSLLeftMargin{36. }%
+\CSLRightInline{Centers for Disease Control and Prevention, \href{https://www.cdc.gov/museum/timeline/covid19.html}{{CDC Museum COVID-19 Timeline}} (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CDC2020}{}}%
+\CSLLeftMargin{37. }%
+\CSLRightInline{Centers for Disease Control and Prevention, \href{https://www.cdc.gov/media/releases/2020/p0714-americans-to-wear-masks.html}{{CDC calls on Americans to wear masks to prevent COVID-19 spread}} (2020).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{S.
@@
 Janke, \href{https://doi.org/10.1177/13684302211023562}{Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic}.
-\emph{Group Processes \& Intergroup Relations}, 13684302211023562 (2021).}
+\emph{Group Processes \& Intergroup Relations} \textbf{25}, 1819--1835 (2021).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{E.
@@
 Grasman, \href{https://doi.org/10.1037/a0038889}{A critique of the cross-lagged panel model}.
-\emph{Psychological Methods} \textbf{20}, 102 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CDC2022b}{}}%
-\CSLLeftMargin{33. }%
-\CSLRightInline{Centers for Disease Control and Prevention, \href{https://www.cdc.gov/museum/timeline/covid19.html}{{CDC Museum COVID-19 Timeline}} (2022).}
+\emph{Psychological Methods} \textbf{20}, 102--116 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{R.
+Baxter-King, J.
+R.
+Brown, R.
+D.
+Enos, A.
+Naeim, L.
+Vavreck, \href{https://doi.org/10.1073/pnas.2116311119}{How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{119}, e2116311119 (2022).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{U.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{H.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
+\CSLLeftMargin{45. }%
+\CSLRightInline{R.
+Amato, L.
+Lacasa, A.
+Díaz-Guilera, A.
+Baronchelli, \href{https://doi.org/10.1073/pnas.1721059115}{The dynamics of norm change in the cultural evolution of language}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{115}, 8260--8265 (2018).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Hogg2008}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wellen1998}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{J.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
+\CSLLeftMargin{49. }%
+\CSLRightInline{A.
+Chaudhuri, \href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
+\emph{Experimental Economics} \textbf{14}, 47--83 (2011).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Neighbors2008}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{C.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
-\CSLLeftMargin{40. }%
-\CSLRightInline{R.
-Baxter-King, J.
-R.
-Brown, R.
-D.
-Enos, A.
-Naeim, L.
-Vavreck, \href{https://doi.org/10.1073/pnas.2116311119}{How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}}.
-\emph{Proceedings of the National Academy of Sciences} \textbf{119}, e2116311119 (2022).}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Remedios2022}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{J. D. Remedios, \href{https://doi.org/10.1038/s44159-022-00024-4}{Psychology must grapple with whiteness}. \emph{Nature Reviews Psychology} \textbf{1}, 125--126 (2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{H.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{D.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Y. Rosseel, \href{https://doi.org/10.18637/jss.v048.i02}{{lavaan}: An {R} package for structural equation modeling}. \emph{Journal of Statistical Software} \textbf{48}, 1--36 (2012).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{R Core Team, \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}} (R Foundation for Statistical Computing, 2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{C. O. Wilke, \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}} (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{H. Wickham, \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}} (Springer-Verlag New York, 2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{F. Aust, M. Barth, \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}} (2022).}
@@
 
-\renewcommand{\appendixname}{\textbf{Supplementary Materials}}
+\renewcommand{\appendixname}{\textbf{Supplementary Material}}
 \renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
@@
 \centering
+\includegraphics{manuscript_files/figure-latex/plotDAG-1.pdf}
+\caption{\label{fig:plotDAG}\emph{Directed acyclic graph reflecting causal assumptions.} In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing (Mask).
+Perceptions of descriptive and injunctive norms also indirectly cause mask wearing through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
+Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
+Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
 \includegraphics{manuscript_files/figure-latex/plotCorBehNorm-1.pdf}
-\caption{\label{fig:plotCorBehNorm}\emph{Predictions from multilevel models with self-reported mask wearing behavior as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables.} Models contain random intercepts for participants and time points.
+\caption{\label{fig:plotCorBehNorm}\emph{Predictions from multilevel models with self-reported mask wearing as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables.} Models contain random intercepts for participants and time points.
 Lines are fixed effect regression lines from multilevel models, shaded areas are 95\% confidence intervals.}
@@
 \includegraphics{manuscript_files/figure-latex/plotCDCSens-1.pdf}
-\caption{\label{fig:plotCDCSens}\emph{Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations.} These models track temporal changes in (a) self-reported mask wearing behavior, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
+\caption{\label{fig:plotCDCSens}\emph{Predictions from multilevel models with change points in line with changes in CDC mask wearing recommendations.} These models track temporal changes in (a) self-reported mask wearing, (b) perceived strength of descriptive norms, and (c) perceived strength of injunctive norms.
 Time is included as a continuous linear predictor, scaled between 0 and 1, with three forced change points (dashed lines).
@@
 Bolded lines and shaded areas represent fixed effect regression lines from multilevel models and 95\% confidence intervals, respectively.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotDAG-1.pdf}
-\caption{\label{fig:plotDAG}\emph{Directed acyclic graph reflecting causal assumptions.} In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing behavior (Mask).
-Perceptions of descriptive and injunctive norms also indirectly cause mask wearing behavior through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
-Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
-Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
-Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
 \end{figure}
@@
 \begin{ThreePartTable}
+
+\small{
 
@@
 \endhead
-Intercept & 4.13, 95\% CI [4.06 4.21] & 5.00, 95\% CI [4.89 5.10] & 5.64, 95\% CI [5.53 5.75]\\
-Slope1 & 0.99, 95\% CI [0.56 1.43] & 1.96, 95\% CI [1.23 2.72] & 0.78, 95\% CI [0.07 1.50]\\
-Slope2 & -5.24, 95\% CI [-5.78 -4.70] & -10.37, 95\% CI [-11.08 -9.67] & -8.34, 95\% CI [-9.02 -7.66]\\
-Slope3 & 1.81, 95\% CI [1.31 2.28] & 1.82, 95\% CI [1.36 2.27] & 1.14, 95\% CI [0.69 1.59]\\
-Slope4 & -4.17, 95\% CI [-4.62 -3.70] & -5.41, 95\% CI [-5.79 -5.01] & -4.66, 95\% CI [-5.05 -4.26]\\
+Intercept & 4.13, 95\% CI [ 4.05\ \ 4.21] & \ \ 5.00, 95\% CI [\ \ 4.90\ \ 5.10] & 5.64, 95\% CI [ 5.53\ \ 5.74]\\
+Slope1 & 0.99, 95\% CI [ 0.56\ \ 1.42] & \ \ 1.98, 95\% CI [\ \ 1.24\ \ 2.72] & 0.78, 95\% CI [ 0.03\ \ 1.52]\\
+Slope2 & -5.23, 95\% CI [-5.73 -4.71] & -10.38, 95\% CI [-11.07 -9.67] & -8.36, 95\% CI [-9.05 -7.64]\\
+Slope3 & 1.80, 95\% CI [ 1.33\ \ 2.33] & \ \ 1.82, 95\% CI [\ \ 1.35\ \ 2.25] & 1.15, 95\% CI [ 0.68\ \ 1.59]\\
+Slope4 & -4.16, 95\% CI [-4.65 -3.68] & -5.40, 95\% CI [ -5.77 -4.99] & -4.66, 95\% CI [-5.03 -4.25]\\
 N & 8505 & 4851 & 4861\\
@@
 \end{longtable}
+
+}
 

### 2022-12-19 R/functions.R

@@ -625,7 +625,7 @@ plotTimeline <- function(d, owid) {
     geom_line() +
     # axes and theme
     scale_x_date(name = NULL, date_labels = "%b\n%Y", date_breaks = "2 month", limits = c(ymd("2020-09-20"), ymd("2022-09-05"))) +
-    scale_y_log10(name = "New\nCOVID-19 cases\nin United States", limits = c(1e+04, 1e+06), labels = trans_format("log10", math_format(10^.x))) +
+    scale_y_log10(name = "Daily new\nCOVID-19 cases\nin United States", limits = c(1e+04, 1e+06), labels = trans_format("log10", math_format(10^.x))) +
     theme_classic()
   # put together
   out <- plot_grid(pA, pB, pC, nrow = 3, align = "v", labels = c("a","b","c"))
@@ -772,7 +772,8 @@ plotRICLPM <- function(model) {
     tibble(
       date = as.Date(rep(dates[c(2,5,9,11,13:18)], times = 3)),
       var = factor(rep(c(var1, var2, var3), each = 10), levels = c(var1, var3, var2)),
-      col = factor(rep(c(col1, col2, col3), each = 10), levels = c(col1, col3, col2))
+      col = factor(rep(c(col1, col2, col3), each = 10), levels = c(col1, col3, col2)),
+      varNum = rep(c(3,1,2), each = 10)
     )
   # standardised coefficients and significance
   sc <- 
@@ -787,33 +788,61 @@ plotRICLPM <- function(model) {
     mutate(
       x      = as.Date(dates[as.numeric(substr(rhs, 4, 5))]),
       xend   = as.Date(dates[as.numeric(substr(lhs, 4, 5))]),
-      y      = c(var1, var2, var3)[as.numeric(substr(rhs, 2, 2))],
-      yend   = c(var1, var2, var3)[as.numeric(substr(lhs, 2, 2))],
+      y      = c(3, 1, 2)[as.numeric(substr(rhs, 2, 2))],
+      yend   = c(3, 1, 2)[as.numeric(substr(lhs, 2, 2))],
       size   = ifelse(est.std >= 0, est.std, 0), # for visualisation, min effect = 0
       colour = ifelse(pvalue.y < 0.05, "black", "lightgrey")
-    )
+    ) %>%
+    arrange(desc(colour))
+  # number to multiply y-axis by
+  i <- 150
   # plot
   p <-
-    # draw plot
[…]

### 2022-12-19 _targets.R

@@ -3,11 +3,12 @@ library(targets)
 library(tarchetypes)
 source("R/functions.R")
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("arm", "broom.mixed", "cowplot", "dagitty", "ggdag", 
-                            "ggeffects", "ggraph", "huxtable", "jtools", "knitr",
-                            "kableExtra", "lavaan", "lme4", "lmerTest", "lubridate", 
-                            "MuMIn", "papaja", "rnaturalearth", "rnaturalearthdata", 
-                            "scales", "semTools", "sf", "zipcodeR"))
+tar_option_set(packages = c("arm", "broom.mixed", "cowplot", "dagitty", "ggarchery",
+                            "ggdag", "ggeffects", "ggraph", "huxtable", "jtools", 
+                            "knitr","kableExtra", "lavaan", "lme4", "lmerTest", 
+                            "lubridate", "MuMIn", "papaja", "rnaturalearth", 
+                            "rnaturalearthdata", "scales", "semTools", "sf", 
+                            "zipcodeR"))
 # workflow
 list(
   # files
@@ -34,6 +35,10 @@ list(
   tar_target(conf2.1, confint(m2.1, method = "Wald")),
   tar_target(conf2.2, confint(m2.2, method = "Wald")),
   tar_target(conf2.3, confint(m2.3, method = "Wald")),
+  # change point parameters
+  tar_target(pars2.1, getChangePointPars(m2.1, outcome = "Mask wearing")),
+  tar_target(pars2.2, getChangePointPars(m2.2, outcome = "Descriptive norms")),
+  tar_target(pars2.3, getChangePointPars(m2.3, outcome = "Injunctive norms")),
   # random-intercept cross-lagged panel models
   tar_target(riclpm, fitRICLPM(d)),
   # fit measures
@@ -48,7 +53,8 @@ list(
   tar_target(plot7, plotRICLPM(riclpm)),
   # tables
   tar_target(itemTable, makeItemTable()),
-  tar_target(changePointsTable, makeChangePointsTable(m2.1, m2.2, m2.3)),
+  tar_target(changePointsTable, makeChangePointsTable(m2.1, m2.2, m2.3,
+                                                      pars2.1, pars2.2, pars2.3)),
   tar_target(lavaanTable, makeLavaanTable(riclpm)),
   # manuscript
   tar_render(manuscript, "manuscript.Rmd"),

### 2023-01-25 manuscript.Rmd
changes before resubmission

@@
     affiliation   : "3"
-  - name          : "Diego Guevara Beltran"
+  - name          : "Diego Guevara Beltrán"
     affiliation   : "4"
@@
 abstract: |
- Human sociality is governed by two types of social norms: injunctive social norms, which prescribe what people _ought_ to do, and descriptive social norms, which reflect what people _actually_ do.
-These norms enable people to cooperate in the face of group-wide challenges.
-While previous experimental work has shown that people’s behavior is influenced by social norms, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
-To understand how cooperative behavior emerges and is shaped by changing social norms in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
-Mask wearing has individual benefits, but it is also a cooperative behavior that provides collective benefits of reduced disease transmission in the community.
-Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = `r nrow(tar_read(d))`), we tracked people’s reported mask wearing and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
-Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior were tightly coupled and both changed quickly in response to recommendations from public health authorities.
-In addition, a random-intercept cross-lagged panel model revealed that perceived descriptive norms, but not perceived injunctive norms, caused future within-person increases in individuals' mask wearing.
-These findings show that, during uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
+ Human sociality is governed by two types of social norms: injunctive norms, which prescribe what people _ought_ to do, and descriptive norms, which reflect what people _actually_ do.
+The process by which these norms emerge and their causal influences on cooperative behavior over time are not well understood.
+Here, we study these questions through social norms influencing mask wearing during the COVID-19 pandemic.
+Leveraging two years of data from the United States (18 time points; _n_ = `r nrow(tar_read(d))`), we tracked mask wearing and perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
+Longitudinal trends suggested that norms and behavior were tightly coupled, changing quickly in response to public health recommendations.
+In addition, longitudinal modelling revealed that descriptive norms, not injunctive norms, caused future increases in mask wearing.
+During uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
 
@@
 keywords          : "social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
-wordcount         : "5276 words"
+wordcount         : "4521 words"
 bibliography      : ["r-references.bib"]
@@
 linenumbers       : yes
-mask              : no
+mask              : yes
 draft             : no
@@
 appendix          : "appendix.Rmd"
-csl               : pnas.csl
+csl               : apa.csl
 ---
@@
 
-# \normalfont Significance Statement
+# \normalfont Statement of Relevance
 
@@
 what people _ought_ to be doing).
+These findings show that visible community adherence is an important driver of mask wearing, suggesting potential avenues for behavior change in future global pandemics.
 
@@
 
-Social norms are a key aspect of human sociality [@Bicchieri2009; @Cialdini1991; @Fehr2018a; @VanKleef2019].
+Social norms are a key aspect of human sociality [@Bicchieri2009; @Cialdini1991; @Fehr2018a].
 Broadly, social norms are defined as commonly known behavioral guidelines enforced by groups of people [@Legros2020].
-Since Asch's early studies of normative conformity [@Asch1956], evidence has accrued that humans, unlike other species, are highly attuned to social norms [@Tomasello2014; @Riedl2012].
-From a young age, children begin to adhere to and enforce group-wide social norms [@Jensen2014; @Rakoczy2013; @Schmidt2012].
-Both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms [@Vaish2011; @Schaumberg2022].
-People also readily punish third-party norm violators [@Fehr2004].
-Emerging from this uniquely normative psychology, social norms specify the rules and practices that govern human social life, forming the "grammar of society" [@Bicchieri2006].
+By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as resource scarcity, natural disasters, and infectious diseases [@Roos2015].
+Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans [@Henrich2015].
 
-By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as inter-group conflict, resource scarcity, high population density, natural disasters, and infectious diseases [@Gelfand2011; @Henrich2015; @Roos2015].
-In the Maasai people of East Africa, for example, social norms governing transfers of livestock enable individuals to collectively manage the risks associated with nomadic pastoralism (e.g., famine, droughts) [@Aktipis2011].
-In countries with high population densities, social norms provide the rules and regularities that allow large groups of people to effectively align their behavior and avoid chaos and conflict [@Gelfand2011].
-Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans [@Henrich2015; @Chudek2011].
+Previous research has distinguished between two types of social norms: injunctive norms and descriptive norms [@Bicchieri2009; @Cialdini1991; @Deutsch1955].
+Injunctive norms indicate what others tend to approve or disapprove of and often involve social sanctions if violated.
+By contrast, descriptive norms simply describe what most people are doing in a given situation, but carry no prescriptive information _per se_.
+According to the focus theory of normative conduct [@Cialdini1991], these two kinds of social norms often align, but they can also be in conflict with one another and differentially affect behavior depending on which norm is more salient.
+For example, there may be an injunctive norm that cleaning up litter at a picnic site is the right thing to do: one _ought_ to behave this way.
+However, if an individual observes that most people are leaving their litter behind at the site, they may instead follow the descriptive norm and litter themselves.
 
-Previous research has distinguished between two primary types of social norms: injunctive norms and descriptive norms [@Bicchieri2009; @Cialdini1991; @Deutsch1955].
-Injunctive social norms indicate what others in the group tend to approve or disapprove of and often involve social sanctions if violated.
-By contrast, descriptive norms simply describe what most people are doing in a given situation.
-Though these two kinds of social norms tend to align, they can also be in conflict with one another.
-For example, there may be an injunctive norm that cleaning up litter at a picnic site is the right thing to do: one _ought_ to behave this way.
-However, if an individual observes that most people are leaving their litter behind at the site, the descriptive norm is to not clean up.
-It is thus possible for injunctive and descriptive norms to differentially affect behavior [@Schultz2007].
-
-Despite decades of research on the causes and consequences of injunctive and descriptive norms [@Bicchieri2009; @Cialdini1991; @Cialdini2007; @Cialdini2021; @Schultz2007], open questions remain regarding the emergence and causal influence of social norms [@Fehr2018a; @Legros2020; @VanKleef2019].
+Despite decades of research on injunctive and descriptive norms [@Cialdini1991; @Cialdini2021; @Schultz2007], open questions remain regarding the emergence and causal influence of social norms [@Legros2020; @VanKleef2019].
 First, how do injunctive and descriptive norms emerge over time within a population?
@@
 
-Research in cultural evolution and behavioral economics has investigated how social norms emerge in a population over time.
-In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring human populations [@Boyd1985; @Henrich2015].
-For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership [@Kushnick2014] and post-marital residence [@Jordan2009].
+Research has investigated how social norms emerge in a population over time.
+In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring populations [@Henrich2015].
 However, less is known about how social norms arise endogenously within populations in the short term.
-Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019] and require peer enforcement to become stable [@Fehr2018b].
+Recent work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019].
 But it remains unclear whether these findings generalize beyond the laboratory to real human populations.
 
-With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the impact of social norms on behavior.
-For example, field experiments in the United Kingdom and United States have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling [@Nigbur2010], paying taxes [@Larkin2018], and sustainably reusing towels in hotels [@Goldstein2008].
-Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals’ attention on opposing injunctive norms (e.g., seeing that the litter is swept into piles, showcasing that it should be cleaned up) [@Cialdini1991; @Schultz2007].
+With regards to normative influences on behavior, studies have demonstrated positive effects of descriptive norms on a variety of cooperative behaviors, including recycling [@Nigbur2010], paying taxes [@Larkin2018], and sustainably reusing towels in hotels [@Goldstein2008].
+However, these studies have two main limitations.
+First, studies have not accounted for other potential non-social explanations for behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
+These non-social beliefs, labeled "factual beliefs" and "personal normative beliefs" [@Bicchieri2014], often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
+For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common behavior).
+It is thus important to control for non-social beliefs in studies of social norms.
+Second, studies have tended to follow cross-sectional experimental designs.
+However, social norms are not static: they change dynamically over time through processes of deliberation and interaction [@Smith2015].
+An alternative but underutilized method of assessing causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
 
-However, these cross-sectional studies have two main limitations.
-First, studies have not adequately controlled for other potential non-social influences on behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
-Previous research has labeled these non-social beliefs as "factual beliefs" and "personal normative beliefs" respectively and emphasized that they should be distinguished from injunctive and descriptive social norms [@Bicchieri2014].
-For example, willingness to sustainably reuse towels might be driven by perceptions that towel reuse has a positive impact on the environment and/or personal beliefs that towel reuse is the right thing to do, rather than by social norms.
-Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point.
-But social norms are not static: they change dynamically over time through processes of deliberation and interaction [@Smith2015].
-An alternative way to assess causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
-In response to novel social dilemmas, social norms and cooperative behaviors emerge as people discover the benefits and costs of different behaviors [@Ostrom1990], enabling researchers to take multiple snapshots over time and establish temporal patterns.
-However, despite the promise of this approach, previous research has not studied the longitudinal interplay between social norms and cooperative behavior.
+To understand how social norms emerge over time and shape cooperative behavior in a non-experimental setting, we focus on mask wearing in the United States during the COVID-19 pandemic.
+In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease [@CDC2022].
+Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread [@CDC2020].
+Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the community (e.g., "flattening the curve" to protect at-risk individuals).
+Thus, the development of mask wearing during the COVID-19 pandemic enables us to study the emergence of social norms and their causal effects on cooperative behavior over a short timescale within a single population.
 
-To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, in the current work we focus on mask wearing during the COVID-19 pandemic.
-Before the pandemic, mask wearing was not a common behavior in the United States.
-In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a protective behavior that people should adopt to minimize the spread of the disease [@CDC2022].
-Mask wearing has individual benefits, but the CDC also emphasized the collective benefits of the behavior in reducing the spread of the disease throughout the community [@CDC2020].
-Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., "flattening the curve" to protect at-risk individuals).
-Thus, the development of mask wearing in the United States during the COVID-19 pandemic allows us to study the emergence of novel injunctive and descriptive social norms and their causal effects on cooperative behavior over a short timescale within a single population.
+Recent research has found positive relationships between social norms and protective COVID-19 behaviors.
+In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure [@Macy2021], and another study found that experimentally-induced descriptive norms increased mask wearing intentions [@Bokemper2021].
+In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing [@Rudert2021].
+These studies are telling, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
+Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States [@BaxterKing2022].
 
-Recent research has found positive relationships between individuals' perceptions of social norms and their protective COVID-19 behaviors.
-In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure [@Macy2021], and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions [@Bokemper2021].
-In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing [@Rudert2021].
-These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time [@Hamaker2015], nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the pandemic.
-Furthermore, these studies do not distinguish between the effects of social norms and the effects of other non-social beliefs that are predicted to shape behavior, such as factual beliefs and personal normative beliefs [@Bicchieri2014].
-Several of the studies also do not control for political ideology, which is important to account for since mask wearing was highly politicized in the United States during the COVID-19 pandemic [@BaxterKing2022].
-
-Here, we use two years of longitudinal data from a representative sample of adults in the United States (18 time points; _n_ = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
-Between September 2020 and October 2022, we asked participants to report their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
-Guided by a causal model of social norms and mask wearing (Supplementary Figure \@ref(fig:plotDAG)), we also asked participants about their factual beliefs (i.e., whether they believe mask wearing is effective), personal normative beliefs (i.e., whether they believed mask wearing is the right thing to do), and political ideology, and controlled for these factors.
-We used our longitudinal data to answer our two main research questions in a specific real-world context.
-First, how do descriptive and injunctive mask wearing norms emerge and develop over time in the United States population during the pandemic?
-Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing over time?
-
-# Results
-
-To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration.
-Figure \@ref(fig:plotTimeline) plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum’s COVID-19 Timeline [@CDC2022].
-
-(ref:plotTimelineCaption) _Timeline of self-reported mask wearing and perceived social norms in the United States during the COVID-19 pandemic._ (a) Points and line ranges indicate means ± two standard errors for the self-reported mask wearing item.
-This item was measured across all eighteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
-(b) Points and line ranges indicate means ± two standard errors for perceived injunctive mask wearing norms (green) and perceived descriptive mask wearing norms (blue).
-These items were measured across eleven time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
-(c) Smoothed data for daily new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; https://ourworldindata.org/).
-Across all panels, gray dashed lines represent significant pandemic-related events in the United States, such as vaccine approval from the Food and Drug Administration (FDA) and public health recommendations from the Centers for Disease Control and Prevention (CDC).
-
-```{r plotTimeline, echo=F, fig.width=6, fig.height=6, fig.cap="(ref:plotTimelineCaption)"}
-plot3
-```
-
-Two main observations can be made about the emergence and stability of social norms from these visualizations.
-First, social norms and behavior were tightly coupled over time.
-Although social norms are measured on fewer occasions than mask wearing, we can see that as mask wearing decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
-Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing and perceived social norms, before declining again in 2022.
-In line with these patterns, multilevel regression models revealed positive correlations between mask wearing and perceived descriptive mask wearing norms (_b_ = `r printnum(fixef(m1.1)["DesNorms"])`, 95% confidence interval [`r printnum(conf1.1["DesNorms","2.5 %"])` `r printnum(conf1.1["DesNorms","97.5 %"])`], _p_ `r printp(summary(m1.1)$coefficients["DesNorms","Pr(>|t|)"])`) and between mask wearing and perceived injunctive mask wearing norms (_b_ = `r printnum(fixef(m1.2)["InjNorms"])`, 95% CI [`r printnum(conf1.2["InjNorms","2.5 %"])` `r printnum(conf1.2["InjNorms","97.5 %"])`], _p_ `r printp(summary(m1.2)$coefficients["InjNorms","Pr(>|t|)"])`) across individuals and time points (Supplementary Figure \@ref(fig:plotCorBehNorm); Supplementary Table \@ref(tab:modelSummaryTable1)).
-
-Second, fluctuations in mask wearing and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States.
-We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020.
-Perceived social norms and mask wearing subsequently declined after the CDC rescinded their mask wearing recommendation following widespread vaccine availability in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high-risk areas in August 2021.
-Finally, perceived social norms and mask wearing declined again after the CDC eased mask wearing guidelines in March 2022.
-These trends were confirmed by a series of multilevel regression models with change points aligning with changes in CDC mask wearing recommendations (Supplementary Figure \@ref(fig:plotCDCSens); Supplementary Table \@ref(tab:changePointsTable)).
-
-Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing and perceived social norms over time.
-To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
-In line with our proposed causal model (Supplementary Figure \@ref(fig:plotDAG)), we also included political orientation as an exogenous time-invariant control.
-According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).
-Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
-
-```{r, echo = FALSE}
-sc <- standardizedSolution(riclpm)
-pe <- parameterEstimates(riclpm)
-```
-
-Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
-On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`) and stronger injunctive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri2","pvalue"])`).
-Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (_r_ = `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri2" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`).
-
-Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing across the study duration, controlling for non-social beliefs and political orientation.
-In random intercept cross-lagged panel models, autoregressive effects represent "persistence" or "inertia" in within-person fluctuations from stable trait levels.
-In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the "stable trait level" over time, which is captured by the random intercepts in our model).
-For example, an autoregressive effect from mask wearing in February 2021 to future mask wearing in June 2021 would suggest that wearing masks more than average in February predicts wearing masks more than average in June.
-By contrast, and most relevant for the current study, cross-lagged effects represent the effect of a within-person fluctuation in one measure on future within-person fluctuations in other measures.
-In other words, a positive cross-lagged effect indicates that being higher than average on one measure predicts being higher than average on _another_ measure in the following time point.
-For example, a cross-lagged effect from descriptive norms in February 2021 to future mask wearing in June 2021 would suggest that perceiving descriptive norms as stronger than average in February predicts wearing masks more than average in June.
-Cross-lagged effects are thus used to infer within-person causal influences over time.
-
-(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Circles represent data collection time points.
-Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
-Arrow thickness is scaled according to standardized effect size.
-Bolded arrows indicate significantly positive parameters, _p_ < 0.05.
-Gray arrows indicate non-significant parameters.
-There are no significant direct paths from injunctive norms to future mask wearing, showing that people’s beliefs about what they should be doing did not have any direct influences on future mask wearing.
-On the other hand, there are significant paths from descriptive norms to future mask wearing, meaning that people’s beliefs about what others are doing influenced their future mask wearing.
-
-```{r plotRICLPM, echo=F, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotRICLPMCaption)"}
-plot7
-```
-
-In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing.
-On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
-According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (first wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.lower"])` `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_05" & pe$op == "~" & pe$rhs == "w3_02","pvalue"])`; second wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.lower"])` `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_09" & pe$op == "~" & pe$rhs == "w3_05","pvalue"])`; fourth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.lower"])` `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_13" & pe$op == "~" & pe$rhs == "w3_11","pvalue"])`; fifth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.lower"])` `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_14" & pe$op == "~" & pe$rhs == "w3_13","pvalue"])`).
-In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
-We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive norms.
-Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.
-
-However, the model also reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did _not_ predict future within-person changes in mask wearing across our study duration.
-All cross-lagged effects from perceived injunctive norms to mask wearing are non-significant.
-Any causal effect that perceived injunctive norms might have had on future mask wearing appears to be fully mediated by perceived descriptive norms.
-This means that believing that others think that mask wearing is the right thing to do influences one’s perception of what others are actually doing, which then influences future behavior.
-For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing.
-But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing over time within individuals.
-
-# Discussion
-
-Using longitudinal data from the United States across two years of the COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma.
-The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior were tightly coupled and both changed dynamically in response to recommendations from public health authorities.
-Moreover, the results of our cross-lagged panel model indicate that descriptive norms caused future increases in mask wearing in the first year and a half of the pandemic.
-By contrast, injunctive norms were not directly causally related to future mask wearing over the entire two years of data collection.
-
-Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction [@Titlestad2019].
-Moreover, the fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one’s own behaviors are mutually constructed [@Markus2010].
-Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms [@Amato2018].
-While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our descriptive trends have shown, can unfold over measurement intervals as short as four to six weeks.
-
-In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing across multiple time points.
-These cross-lagged effects were independent of the effects of non-social beliefs and political orientation.
-In line with this finding, similar descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) during the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects as we have done here [@Rudert2021].
-Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (i.e., a descriptive norm) [@Bokemper2021].
-
-Why have descriptive norms had these positive effects on protective COVID-19 behaviors like mask wearing?
-Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic [@Gelfand2015].
-During times of uncertainty, people look to others to get information quickly about how they should behave, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms [@Hogg2008; @Wellen1998].
-Moreover, humans tend to be conditional cooperators who adapt their levels of cooperation depending on the degree of cooperative behavior in the population [@Chaudhuri2011].
-Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
-
-Our longitudinal model revealed that, at the between-person level, people who perceived stronger injunctive norms on average also reported more frequent mask wearing on average.
-This correlation is in line with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults [@Macy2021].
-However, at the within-person level, we found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
-This finding is at odds with previous evidence that experimentally-induced injunctive social norms can cause increases in cooperative behavior [@Schultz2007].
-
-One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous experimental results.
-Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were made more salient than injunctive norms during the pandemic.
-According to the focus theory of normative conduct [@Cialdini1991], this difference in norm salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
-By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people’s behaviors, increasing the relative salience of injunctive norms.
-To test this idea, future research should expand our longitudinal cross-lagged approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
-
-While perceived injunctive norms did not directly predict future mask wearing, on one occasion they did indirectly predict future mask wearing _through_ perceived descriptive norms.
-In other words, perceptions of what others think people ought to be doing predicted future perceptions of what people were actually doing, which then predicted future mask wearing.
-This suggests that people may use injunctive norm information as a signal that they should check what others are doing, which could then aid in interpreting or perceiving other people's behavior as a descriptive norm.
-Future studies could manipulate injunctive norm presence and determine the effect of these norms on how people attend to and interpret behavioral information.
-
-We are limited in generalizing these findings due to the constraints of our sample and the variables considered.
-While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \@ref(fig:plotAttrition)).
-This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
-Since injunctive norm strength varies based on which group is seen as the source of the norm [@Neighbors2008], it would be interesting to learn whether different groups have different patterns of norm emergence over time.
-In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors [@BaxterKing2022].
-Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large [@Remedios2022].
-Our results also might not generalize to all social norms, behaviors, and social dilemmas.
-Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more remote than the COVID-19 pandemic.
-For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.
-
-For the case of mask wearing in the United States during the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and tracked ongoing changes in both recommendations from authorities and current levels of cooperative behavior.
-Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
-Importantly, this key finding slices two ways.
-Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that _low_ local levels of mask wearing _discouraged_ future personal mask use.
-This echoes recent reports of people in the United States not wanting to be "singled out" by being the only one wearing a mask in their community [@Natanson2022].
-Our work thus underscores the importance of consistent, visible community adherence for encouraging personal protective behaviors in response to global pandemics like COVID-19.
+Here, we use two years of data from a representative sample of adults in the United States (18 time points; _n_ = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
+Participants reported their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
+We also asked participants about their non-social mask wearing beliefs and political ideology, and controlled for these factors.
+We used these longitudinal data to answer two main research questions in a specific real-world context.
+First, how do descriptive and injunctive mask wearing norms emerge over time?
+Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing?
 
@@
 The manuscript was reproducibly generated using the _targets_ [@Landau2021] and _papaja_ [@Aust2022] packages.
+All code and data are publicly available on GitHub [@Heiman2023].
+
+# Results
+
+To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration.
+Figure \@ref(fig:plotTimeline) plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
+These events were obtained from the CDC Museum’s COVID-19 Timeline [@CDC2022].
+
+(ref:plotTimelineCaption) _Timeline of self-reported mask wearing and perceived social norms in the United States during the COVID-19 pandemic._ (a) Points and line ranges indicate means ± two standard errors for the self-reported mask wearing item.
+This item was measured across all eighteen time points on a 5-point Likert scale, with higher values indicating increased frequency of personal mask wearing during in-person interactions.
+(b) Points and line ranges indicate means ± two standard errors for perceived injunctive mask wearing norms (green) and perceived descriptive mask wearing norms (blue).
+These items were measured across eleven time points on a 7-point Likert scale, with higher values indicating stronger perceived social norms.
+(c) Smoothed data for daily new COVID-19 cases in the United States, displayed on the log scale (data retrieved from Our World in Data; https://ourworldindata.org/).
+Across all panels, gray dashed lines represent significant pandemic-related events in the United States, such as vaccine approval from the Food and Drug Administration (FDA) and public health recommendations from the Centers for Disease Control and Prevention (CDC).
+
+```{r plotTimeline, echo=F, fig.width=6, fig.height=6, fig.cap="(ref:plotTimelineCaption)"}
+plot3
+```
+
+Two main observations can be made about the emergence and stability of social norms from these visualizations.
+First, social norms and behavior were tightly coupled over time.
+Although social norms are measured on fewer occasions than mask wearing, we can see that as mask wearing decreased in the summer of 2021, so too did perceived descriptive and injunctive mask wearing norms.
+Subsequently, the steep rise in COVID-19 case numbers in the fall of 2021 saw concomitant increases in both mask wearing and perceived social norms, before declining again in 2022.
+In line with these patterns, multilevel regression models revealed positive correlations between mask wearing and perceived descriptive mask wearing norms (_b_ = `r printnum(fixef(m1.1)["DesNorms"])`, 95% confidence interval [`r printnum(conf1.1["DesNorms","2.5 %"])` `r printnum(conf1.1["DesNorms","97.5 %"])`], _p_ `r printp(summary(m1.1)$coefficients["DesNorms","Pr(>|t|)"])`) and between mask wearing and perceived injunctive mask wearing norms (_b_ = `r printnum(fixef(m1.2)["InjNorms"])`, 95% CI [`r printnum(conf1.2["InjNorms","2.5 %"])` `r printnum(conf1.2["InjNorms","97.5 %"])`], _p_ `r printp(summary(m1.2)$coefficients["InjNorms","Pr(>|t|)"])`) across individuals and time points (Supplementary Figure \@ref(fig:plotCorBehNorm); Supplementary Table \@ref(tab:modelSummaryTable1)).
+
+Second, fluctuations in mask wearing and perceived social norms are in line with recommendations broadcasted by the CDC, an important institution governing public health in the United States.
+We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020.
+Perceived social norms and mask wearing subsequently declined after the CDC rescinded their mask wearing recommendation following widespread vaccine availability in March 2021, and then increased again after the CDC updated their guidelines for indoor mask use in high-risk areas in August 2021.
+Finally, perceived social norms and mask wearing declined again after the CDC eased mask wearing guidelines in March 2022.
+These trends were confirmed by a series of multilevel regression models with change points aligning with changes in CDC mask wearing recommendations (Supplementary Figure \@ref(fig:plotCDCSens); Supplementary Table \@ref(tab:changePointsTable)).
+
+Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing and perceived social norms over time.
+To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
+This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
+In line with our proposed causal model (Supplementary Figure \@ref(fig:plotDAG)), we also included political orientation as an exogenous time-invariant control.
+According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).
+Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
+
+```{r, echo = FALSE}
+sc <- standardizedSolution(riclpm)
+pe <- parameterEstimates(riclpm)
+```
+
+Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
+On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`) and stronger injunctive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri2","pvalue"])`).
+Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (_r_ = `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri2" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`).
+
+Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing across the study duration, controlling for non-social beliefs and political orientation.
+In random intercept cross-lagged panel models, autoregressive effects represent "persistence" or "inertia" in within-person fluctuations from stable trait levels.
+In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the "stable trait level" over time, which is captured by the random intercepts in our model).
+For example, an autoregressive effect from mask wearing in February 2021 to future mask wearing in June 2021 would suggest that wearing masks more than average in February predicts wearing masks more than average in June.
+By contrast, and most relevant for the current study, cross-lagged effects represent the effect of a within-person fluctuation in one measure on future within-person fluctuations in other measures.
+In other words, a positive cross-lagged effect indicates that being higher than average on one measure predicts being higher than average on _another_ measure in the following time point.
+For example, a cross-lagged effect from descriptive norms in February 2021 to future mask wearing in June 2021 would suggest that perceiving descriptive norms as stronger than average in February predicts wearing masks more than average in June.
+Cross-lagged effects are thus used to infer within-person causal influences over time.
+
+(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
+Arrow thickness is scaled according to standardized effect size.
+Bolded arrows indicate significantly positive parameters, _p_ < 0.05.
+Gray arrows indicate non-significant parameters.
+There are no significant direct paths from injunctive norms to future mask wearing, showing that people’s beliefs about what they should be doing did not have any direct influences on future mask wearing.
+On the other hand, there are significant paths from descriptive norms to future mask wearing, meaning that people’s beliefs about what others are doing influenced their future mask wearing.
+
+```{r plotRICLPM, echo=F, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotRICLPMCaption)"}
+plot7
+```
+
+In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing.
+On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
+According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (first wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.lower"])` `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_05" & pe$op == "~" & pe$rhs == "w3_02","pvalue"])`; second wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.lower"])` `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_09" & pe$op == "~" & pe$rhs == "w3_05","pvalue"])`; fourth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.lower"])` `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_13" & pe$op == "~" & pe$rhs == "w3_11","pvalue"])`; fifth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.lower"])` `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_14" & pe$op == "~" & pe$rhs == "w3_13","pvalue"])`).
+In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
+We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive norms.
+Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.
+
+However, the model also reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did _not_ predict future within-person changes in mask wearing across our study duration.
+All cross-lagged effects from perceived injunctive norms to mask wearing are non-significant.
+Any causal effect that perceived injunctive norms might have had on future mask wearing appears to be fully mediated by perceived descriptive norms.
+This means that believing that others think that mask wearing is the right thing to do influences one’s perception of what others are actually doing, which then influences future behavior.
+For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing.
+But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing over time within individuals.
+
+# Discussion
+
+Using longitudinal data from the United States across two years of the COVID-19 pandemic, we aimed to understand how descriptive and injunctive mask wearing norms emerge and influence behavior in response to a naturally unfolding social dilemma.
+The trends of norm perceptions and self-reported mask wearing over time suggest that norms and behavior were tightly coupled and both changed dynamically in response to recommendations from public health authorities.
+Moreover, the results of our cross-lagged panel model indicate that descriptive norms caused future increases in mask wearing in the first year and a half of the pandemic.
+By contrast, injunctive norms were not directly causally related to future mask wearing over the entire two years of data collection.
+
+Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction [@Titlestad2019].
+The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one’s own behaviors are mutually constructed [@Markus2010].
+Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms [@Amato2018].
+While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our trends have shown, can unfold over measurement intervals as short as four to six weeks.
+
+We found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing, independent of the effects of non-social beliefs and political orientation.
+This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors [@Bokemper2021; @Rudert2021].
+There are several explanations for why descriptive norms have had these positive effects on protective COVID-19 behaviors like mask wearing.
+First, people may have followed descriptive norms to quickly coordinate their behavior with others during the pandemic.
+Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic [@Gelfand2015].
+Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population [@Chaudhuri2011].
+Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
+Third, the increased frequency of mask wearing in the population might have created a bandwagon effect [@SchmittBeck2015], encouraging conformist copying.
+Under this view, people wear masks not to coordinate or cooperate, but simply to fit in with the crowd.
+Future research will be required to determine the motivations underlying adherence to descriptive norms during uncertain times.
+
+We found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
+One possible explanation for this result is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were more salient than injunctive norms during the pandemic.
+According to focus theory [@Cialdini1991], this difference in salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
+By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people’s behaviors, increasing the relative salience of injunctive norms.
+To test this idea, future research should expand our longitudinal approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
+
+We are limited in generalizing these findings due to the constraints of our sample and the variables considered.
+While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \@ref(fig:plotAttrition)).
+This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
+Our results also might not generalize to all social norms, behaviors, and social dilemmas.
+Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more remote than the COVID-19 pandemic.
+For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.
+
+For the case of mask wearing in the United States during the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and tracked ongoing changes in both recommendations from authorities and current levels of cooperative behavior.
+Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
+Our work thus underscores the importance of consistent, visible community adherence for encouraging protective behaviors in response to global pandemics like COVID-19.
 
@@
 
-This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
-We would also like to thank the Language, Culture, and Cognition lab at the University of Auckland for providing feedback on a previous version of this manuscript.
+[REDACTED]
 
@@
 
-SLH, JDA, DGB, AV, ERH, AA, and PMT conceptualized the study.
-SLH, SC, JDA, DGB, and AV oversaw the data curation, investigation, and methodology of the study.
-SLH and SC wrote the first draft of the paper.
-SC conducted the formal analysis and created all visualizations.
-ERH, AA, and PMT provided funding and supervision for the study.
-All authors reviewed and edited the final draft of the paper.
+[REDACTED]
 
@@
 
-# Research Transparency and Reproducibility
+# Open Practices Statement
 
-All data and code to reproduce the statistical analyses in this manuscript can be found on GitHub: https://github.com/ScottClaessens/covidMaskWearing
+All data and code to reproduce the statistical analyses in this manuscript are publicly available on GitHub: https://github.com/ScottClaessens/covidMaskWearing This study was not preregistered.
 

### 2023-01-25 manuscript.tex
changes before resubmission

@@
 \documentclass[
-  man, donotrepeattitle,floatsintext]{apa6}
+  man, donotrepeattitle,mask,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 {}{\PatchFailed}
-\keywords{social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 5276 words}
+\keywords{social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 4521 words}
 \usepackage{lineno}
@@
   pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic},
-  pdfauthor={Samantha L. Heiman*,1, Scott Claessens*,2, Jessica D. Ayers3, Diego Guevara Beltran4, Andrew Van Horn5,6, Edward R. Hirt1, Athena Aktipis†,4, \& Peter M. Todd†,1,7},
   pdflang={en-EN},
@@
 Heiman\textsuperscript{*,1}, Scott Claessens\textsuperscript{*,2}, Jessica D.
-Ayers\textsuperscript{3}, Diego Guevara Beltran\textsuperscript{4}, Andrew Van Horn\textsuperscript{5,6}, Edward R.
+Ayers\textsuperscript{3}, Diego Guevara Beltrán\textsuperscript{4}, Andrew Van Horn\textsuperscript{5,6}, Edward R.
 Hirt\textsuperscript{1}, Athena Aktipis\textsuperscript{†,4}, \& Peter M.
@@
 \abstract{%
-Human sociality is governed by two types of social norms: injunctive social norms, which prescribe what people \emph{ought} to do, and descriptive social norms, which reflect what people \emph{actually} do.
-These norms enable people to cooperate in the face of group-wide challenges.
-While previous experimental work has shown that people's behavior is influenced by social norms, several open questions remain about the natural emergence of injunctive and descriptive social norms within populations and their causal influences on cooperative behavior over time.
-To understand how cooperative behavior emerges and is shaped by changing social norms in a non-experimental setting, we studied mask wearing during the COVID-19 pandemic.
-Mask wearing has individual benefits, but it is also a cooperative behavior that provides collective benefits of reduced disease transmission in the community.
-Leveraging two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 915), we tracked people's reported mask wearing and their perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
-Longitudinal trends of norm perceptions and self-reported mask wearing suggested that norms and behavior were tightly coupled and both changed quickly in response to recommendations from public health authorities.
-In addition, a random-intercept cross-lagged panel model revealed that perceived descriptive norms, but not perceived injunctive norms, caused future within-person increases in individuals' mask wearing.
-These findings show that, during uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
+Human sociality is governed by two types of social norms: injunctive norms, which prescribe what people \emph{ought} to do, and descriptive norms, which reflect what people \emph{actually} do.
+The process by which these norms emerge and their causal influences on cooperative behavior over time are not well understood.
+Here, we study these questions through social norms influencing mask wearing during the COVID-19 pandemic.
+Leveraging two years of data from the United States (18 time points; \emph{n} = 915), we tracked mask wearing and perceived injunctive and descriptive mask wearing norms as the pandemic unfolded.
+Longitudinal trends suggested that norms and behavior were tightly coupled, changing quickly in response to public health recommendations.
+In addition, longitudinal modelling revealed that descriptive norms, not injunctive norms, caused future increases in mask wearing.
+During uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
 }
@@
 
-\hypertarget{significance-statement}{%
-\section{\texorpdfstring{\normalfont Significance Statement}{Significance Statement}}\label{significance-statement}}
+\hypertarget{statement-of-relevance}{%
+\section{\texorpdfstring{\normalfont Statement of Relevance}{Statement of Relevance}}\label{statement-of-relevance}}
 
@@
 Moreover, longitudinal modeling suggests that mask wearing is causally preceded by perceptions of descriptive norms (i.e.~what people \emph{are} doing) but not injunctive norms (i.e.~what people \emph{ought} to be doing).
+These findings show that visible community adherence is an important driver of mask wearing, suggesting potential avenues for behavior change in future global pandemics.
 
@@
 
-Social norms are a key aspect of human sociality (1--4).
-Broadly, social norms are defined as commonly known behavioral guidelines enforced by groups of people (5).
-Since Asch's early studies of normative conformity (6), evidence has accrued that humans, unlike other species, are highly attuned to social norms (7, 8).
-From a young age, children begin to adhere to and enforce group-wide social norms (9--11).
-Both children and adults rely on normative emotions, such as shame and guilt, to determine when they or others have violated social norms (12, 13).
-People also readily punish third-party norm violators (14).
-Emerging from this uniquely normative psychology, social norms specify the rules and practices that govern human social life, forming the ``grammar of society'' (15).
-
-By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as inter-group conflict, resource scarcity, high population density, natural disasters, and infectious diseases (16--18).
-In the Maasai people of East Africa, for example, social norms governing transfers of livestock enable individuals to collectively manage the risks associated with nomadic pastoralism (e.g., famine, droughts) (19).
-In countries with high population densities, social norms provide the rules and regularities that allow large groups of people to effectively align their behavior and avoid chaos and conflict (16).
-Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans (17, 20).
-
-Previous research has distinguished between two primary types of social norms: injunctive norms and descriptive norms (1, 2, 21).
-Injunctive social norms indicate what others in the group tend to approve or disapprove of and often involve social sanctions if violated.
-By contrast, descriptive norms simply describe what most people are doing in a given situation.
-Though these two kinds of social norms tend to align, they can also be in conflict with one another.
+Social norms are a key aspect of human sociality (Bicchieri \& Xiao, 2009; Cialdini et al., 1991; Fehr \& Schurtenberger, 2018).
+Broadly, social norms are defined as commonly known behavioral guidelines enforced by groups of people (Legros \& Cislaghi, 2020).
+By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as resource scarcity, natural disasters, and infectious diseases (Roos et al., 2015).
+Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans (Henrich, 2015).
+
+Previous research has distinguished between two types of social norms: injunctive norms and descriptive norms (Bicchieri \& Xiao, 2009; Cialdini et al., 1991; Deutsch \& Gerard, 1955).
+Injunctive norms indicate what others tend to approve or disapprove of and often involve social sanctions if violated.
+By contrast, descriptive norms simply describe what most people are doing in a given situation, but carry no prescriptive information \emph{per se}.
+According to the focus theory of normative conduct (Cialdini et al., 1991), these two kinds of social norms often align, but they can also be in conflict with one another and differentially affect behavior depending on which norm is more salient.
 For example, there may be an injunctive norm that cleaning up litter at a picnic site is the right thing to do: one \emph{ought} to behave this way.
-However, if an individual observes that most people are leaving their litter behind at the site, the descriptive norm is to not clean up.
-It is thus possible for injunctive and descriptive norms to differentially affect behavior (22).
-
-Despite decades of research on the causes and consequences of injunctive and descriptive norms (1, 2, 22--24), open questions remain regarding the emergence and causal influence of social norms (3--5).
+However, if an individual observes that most people are leaving their litter behind at the site, they may instead follow the descriptive norm and litter themselves.
+
+Despite decades of research on injunctive and descriptive norms (Cialdini et al., 1991; Cialdini \& Jacobson, 2021; Schultz et al., 2007), open questions remain regarding the emergence and causal influence of social norms (Legros \& Cislaghi, 2020; van Kleef et al., 2019).
 First, how do injunctive and descriptive norms emerge over time within a population?
@@
 
-Research in cultural evolution and behavioral economics has investigated how social norms emerge in a population over time.
-In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring human populations (17, 25).
-For example, cultural phylogenetic studies have revealed patterns of vertical cultural inheritance across societies for a variety of injunctive social norms, such as norms governing land ownership (26) and post-marital residence (27).
+Research has investigated how social norms emerge in a population over time.
+In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring populations (Henrich, 2015).
 However, less is known about how social norms arise endogenously within populations in the short term.
-Recent experimental work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (28) and require peer enforcement to become stable (29).
+Recent work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (Titlestad et al., 2019).
 But it remains unclear whether these findings generalize beyond the laboratory to real human populations.
 
-With regards to normative influences on behavior, there is a wealth of cross-sectional evidence demonstrating the impact of social norms on behavior.
-For example, field experiments in the United Kingdom and United States have demonstrated the positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (30), paying taxes (31), and sustainably reusing towels in hotels (32).
-Evidence also suggests that any potentially deleterious effects of descriptive social norms (e.g., choosing to litter at a picnic site that already contains visible signs of littering) can be counteracted by instead focusing individuals' attention on opposing injunctive norms (e.g., seeing that the litter is swept into piles, showcasing that it should be cleaned up) (2, 22).
-
-However, these cross-sectional studies have two main limitations.
-First, studies have not adequately controlled for other potential non-social influences on behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
-Previous research has labeled these non-social beliefs as ``factual beliefs'' and ``personal normative beliefs'' respectively and emphasized that they should be distinguished from injunctive and descriptive social norms (33).
-For example, willingness to sustainably reuse towels might be driven by perceptions that towel reuse has a positive impact on the environment and/or personal beliefs that towel reuse is the right thing to do, rather than by social norms.
-Second, cross-sectional studies have tended to follow experimental designs in which perceptions of social norms are manipulated by the researchers at a single time point.
-But social norms are not static: they change dynamically over time through processes of deliberation and interaction (34).
-An alternative way to assess causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
-In response to novel social dilemmas, social norms and cooperative behaviors emerge as people discover the benefits and costs of different behaviors (35), enabling researchers to take multiple snapshots over time and establish temporal patterns.
-However, despite the promise of this approach, previous research has not studied the longitudinal interplay between social norms and cooperative behavior.
-
-To understand how novel injunctive and descriptive social norms emerge over time and shape cooperative behavior in a non-experimental setting, in the current work we focus on mask wearing during the COVID-19 pandemic.
-Before the pandemic, mask wearing was not a common behavior in the United States.
-In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) as a protective behavior that people should adopt to minimize the spread of the disease (36).
-Mask wearing has individual benefits, but the CDC also emphasized the collective benefits of the behavior in reducing the spread of the disease throughout the community (37).
-Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the wider community (e.g., ``flattening the curve'' to protect at-risk individuals).
-Thus, the development of mask wearing in the United States during the COVID-19 pandemic allows us to study the emergence of novel injunctive and descriptive social norms and their causal effects on cooperative behavior over a short timescale within a single population.
-
-Recent research has found positive relationships between individuals' perceptions of social norms and their protective COVID-19 behaviors.
-In the United States, one cross-sectional study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (38), and another vignette study found that experimentally-manipulated descriptive norms increased personal mask wearing intentions (39).
-In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (40).
-These studies are telling, but since they are experimental, cross-sectional, or only minimally longitudinal, they are unable to distinguish between within-person and between-person change over time (41), nor do they have the temporal granularity to capture fluctuating changes in norm strength and norm adherence across the pandemic.
-Furthermore, these studies do not distinguish between the effects of social norms and the effects of other non-social beliefs that are predicted to shape behavior, such as factual beliefs and personal normative beliefs (33).
-Several of the studies also do not control for political ideology, which is important to account for since mask wearing was highly politicized in the United States during the COVID-19 pandemic (42).
-
-Here, we use two years of longitudinal data from a representative sample of adults in the United States (18 time points; \emph{n} = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
-Between September 2020 and October 2022, we asked participants to report their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
-Guided by a causal model of social norms and mask wearing (Supplementary Figure \ref{fig:plotDAG}), we also asked participants about their factual beliefs (i.e., whether they believe mask wearing is effective), personal normative beliefs (i.e., whether they believed mask wearing is the right thing to do), and political ideology, and controlled for these factors.
-We used our longitudinal data to answer our two main research questions in a specific real-world context.
-First, how do descriptive and injunctive mask wearing norms emerge and develop over time in the United States population during the pandemic?
-Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing over time?
+With regards to normative influences on behavior, studies have demonstrated positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (Nigbur et al., 2010), paying taxes (Larkin et al., 2018), and sustainably reusing towels in hotels (Goldstein et al., 2008).
+However, these studies have two main limitations.
+First, studies have not accounted for other potential non-social explanations for behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
+These non-social beliefs, labeled ``factual beliefs'' and ``personal normative beliefs'' (Bicchieri et al., 2014), often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
+For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common behavior).
+It is thus important to control for non-social beliefs in studies of social norms.
+Second, studies have tended to follow cross-sectional experimental designs.
+However, social norms are not static: they change dynamically over time through processes of deliberation and interaction (Smith et al., 2015).
+An alternative but underutilized method of assessing causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
+
+To understand how social norms emerge over time and shape cooperative behavior in a non-experimental setting, we focus on mask wearing in the United States during the COVID-19 pandemic.
+In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease (Centers for Disease Control and Prevention, 2022).
+Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread (Centers for Disease Control and Prevention, 2020).
+Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the community (e.g., ``flattening the curve'' to protect at-risk individuals).
+Thus, the development of mask wearing during the COVID-19 pandemic enables us to study the emergence of social norms and their causal effects on cooperative behavior over a short timescale within a single population.
+
+Recent research has found positive relationships between social norms and protective COVID-19 behaviors.
+In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (Macy et al., 2021), and another study found that experimentally-induced descriptive norms increased mask wearing intentions (Bokemper et al., 2021).
+In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (Rudert \& Janke, 2021).
+These studies are telling, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
+Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States (Baxter-King et al., 2022).
+
+Here, we use two years of data from a representative sample of adults in the United States (18 time points; \emph{n} = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
+Participants reported their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
+We also asked participants about their non-social mask wearing beliefs and political ideology, and controlled for these factors.
+We used these longitudinal data to answer two main research questions in a specific real-world context.
+First, how do descriptive and injunctive mask wearing norms emerge over time?
+Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing?
+
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
+
+\hypertarget{ethical-approval}{%
+\subsection{Ethical approval}\label{ethical-approval}}
+
+This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
+
+\hypertarget{participants-and-sampling}{%
+\subsection{Participants and sampling}\label{participants-and-sampling}}
+
+Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution).
+From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
+This resulted in 18 unique time points of data collection during the pandemic.
+The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
+Of the initial 915 participants, 634 returned to complete the survey at Time 2, while 347 participants continued through to Time 18 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
+
+\hypertarget{measures}{%
+\subsection{Measures}\label{measures}}
+
+\hypertarget{self-reported-mask-wearing}{%
+\subsubsection{Self-reported mask wearing}\label{self-reported-mask-wearing}}
+
+At every time point, participants were asked about the number of in-person interactions they had in the last week.
+Following this question, participants self-reported their mask wearing by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+
+\hypertarget{perceived-descriptive-and-injunctive-social-norms}{%
+\subsubsection{Perceived descriptive and injunctive social norms}\label{perceived-descriptive-and-injunctive-social-norms}}
+
+In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
+
+Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
+We measured perceived descriptive social norms as the mean of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
+
+Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
+We measured perceived injunctive social norms as the mean of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
+
+To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
+We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
+Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).
+
+\hypertarget{additional-control-variables}{%
+\subsubsection{Additional control variables}\label{additional-control-variables}}
+
+To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research (Bicchieri et al., 2014): factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
+These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
+In addition, we also included political orientation as a common cause of all other variables.
+This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic (Baxter-King et al., 2022).
+Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
+
+Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
+Factual beliefs were measured as the mean of the following two items: ``\emph{I wear a face mask when going out in public to keep myself from getting sick}'' and ``\emph{I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet}''.
+Personal normative beliefs were measured with a single item: ``\emph{Wearing a face mask when going out in public is the right thing to do}''.
+These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
+
+Political orientation was measured in the first time point only.
+We measured political orientation as the mean of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
+These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
+
+\hypertarget{statistical-analysis}{%
+\subsection{Statistical analysis}\label{statistical-analysis}}
+
+To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
+First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
+Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
+These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
+We estimated these multilevel regression models using the \emph{lme4} R package (Bates et al., 2015) and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data (Hamaker et al., 2015).
+This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time \emph{t-1} predicts being above average in another variable at time \emph{t}.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets (Granger, 1969; Hamaker et al., 2015).
+
+We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package (Rosseel, 2012).
+In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
+For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
+We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We restricted the analysis to the ten time points with available data for all five variables.
+Full information maximum likelihood estimation was used to deal with missing data.
+
+All analyses were conducted in R v4.1.1 (R Core Team, 2022).
+Visualizations were generated using the \emph{cowplot} (Wilke, 2020) and \emph{ggplot2} (Wickham, 2016) packages.
+The manuscript was reproducibly generated using the \emph{targets} (Landau, 2021) and \emph{papaja} (Aust \& Barth, 2022) packages.
+All code and data are publicly available on GitHub (Heiman et al., 2023).
 
@@
 Figure \ref{fig:plotTimeline} plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum's COVID-19 Timeline (36).
+These events were obtained from the CDC Museum's COVID-19 Timeline (Centers for Disease Control and Prevention, 2022).
 
@@
 
-Second, fluctuations in mask wearing and perceived social norms are in line with recommendations broadcasted by the CDC, the main national public health agency of the United States.
+Second, fluctuations in mask wearing and perceived social norms are in line with recommendations broadcasted by the CDC, an important institution governing public health in the United States.
 We do not have data for the very start of the pandemic in early 2020, but the high levels of mask wearing and strong perceived social norms at the start of our observation window likely emerged after the initial mask wearing recommendation from the CDC in April 2020.
@@
 On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
-According to recent effect size guidelines for cross-lagged panel models (43), the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
+According to recent effect size guidelines for cross-lagged panel models (Orth et al., 2022), the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
 In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
@@
 
-Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (28).
-Moreover, the fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (44).
-Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms (45).
-While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our descriptive trends have shown, can unfold over measurement intervals as short as four to six weeks.
-
-In our longitudinal analysis, we found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing across multiple time points.
-These cross-lagged effects were independent of the effects of non-social beliefs and political orientation.
-In line with this finding, similar descriptive norms have also been shown to predict future increases in physical distancing and prosocial behaviors (e.g., neighborhood help, charitable donations) during the COVID-19 pandemic, though this previous work did not adequately disentangle between-person and within-person effects as we have done here (40).
-Similarly, experimental work has revealed that people in the United States are more likely to report intentions to wear a mask if they are told that others are wearing masks (i.e., a descriptive norm) (39).
-
-Why have descriptive norms had these positive effects on protective COVID-19 behaviors like mask wearing?
-Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (46).
-During times of uncertainty, people look to others to get information quickly about how they should behave, and attempt to alleviate uncertainty-related stress by identifying with their group and its social norms (47, 48).
-Moreover, humans tend to be conditional cooperators who adapt their levels of cooperation depending on the degree of cooperative behavior in the population (49).
+Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (Titlestad et al., 2019).
+The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (Markus \& Kitayama, 2010).
+Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms (Amato et al., 2018).
+While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our trends have shown, can unfold over measurement intervals as short as four to six weeks.
+
+We found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing, independent of the effects of non-social beliefs and political orientation.
+This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors (Bokemper et al., 2021; Rudert \& Janke, 2021).
+There are several explanations for why descriptive norms have had these positive effects on protective COVID-19 behaviors like mask wearing.
+First, people may have followed descriptive norms to quickly coordinate their behavior with others during the pandemic.
+Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (Gelfand \& Harrington, 2015).
+Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population (Chaudhuri, 2011).
 Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
-
-Our longitudinal model revealed that, at the between-person level, people who perceived stronger injunctive norms on average also reported more frequent mask wearing on average.
-This correlation is in line with cross-sectional evidence showing that perceived injunctive norms are positively correlated with intentions to stay indoors during the pandemic among older adults (38).
-However, at the within-person level, we found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
-This finding is at odds with previous evidence that experimentally-induced injunctive social norms can cause increases in cooperative behavior (22).
-
-One possible explanation for these conflicting findings is that, unlike previous work, our model systematically controlled for non-social beliefs, such as factual beliefs and personal normative beliefs, which could potentially have driven previous experimental results.
-Another possible explanation is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were made more salient than injunctive norms during the pandemic.
-According to the focus theory of normative conduct (2), this difference in norm salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
+Third, the increased frequency of mask wearing in the population might have created a bandwagon effect (Schmitt-Beck, 2015), encouraging conformist copying.
+Under this view, people wear masks not to coordinate or cooperate, but simply to fit in with the crowd.
+Future research will be required to determine the motivations underlying adherence to descriptive norms during uncertain times.
+
+We found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
+One possible explanation for this result is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were more salient than injunctive norms during the pandemic.
+According to focus theory (Cialdini et al., 1991), this difference in salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people's behaviors, increasing the relative salience of injunctive norms.
-To test this idea, future research should expand our longitudinal cross-lagged approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
-
-While perceived injunctive norms did not directly predict future mask wearing, on one occasion they did indirectly predict future mask wearing \emph{through} perceived descriptive norms.
-In other words, perceptions of what others think people ought to be doing predicted future perceptions of what people were actually doing, which then predicted future mask wearing.
-This suggests that people may use injunctive norm information as a signal that they should check what others are doing, which could then aid in interpreting or perceiving other people's behavior as a descriptive norm.
-Future studies could manipulate injunctive norm presence and determine the effect of these norms on how people attend to and interpret behavioral information.
+To test this idea, future research should expand our longitudinal approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
 
@@
 This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
-Since injunctive norm strength varies based on which group is seen as the source of the norm (50), it would be interesting to learn whether different groups have different patterns of norm emergence over time.
-In particular, future analyses with larger samples should consider political ideology as a group identity, due to the political polarization of COVID-19 protective behaviors (42).
-Our sample was predominantly White, and so future larger samples should be intentionally more diverse, answering calls to avoid generalizing White samples as representative of human behavior at large (51).
 Our results also might not generalize to all social norms, behaviors, and social dilemmas.
@@
 Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
-Importantly, this key finding slices two ways.
-Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that \emph{low} local levels of mask wearing \emph{discouraged} future personal mask use.
-This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community (52).
-Our work thus underscores the importance of consistent, visible community adherence for encouraging personal protective behaviors in response to global pandemics like COVID-19.
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{ethical-approval}{%
-\subsection{Ethical approval}\label{ethical-approval}}
-
-This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
-
-\hypertarget{participants-and-sampling}{%
-\subsection{Participants and sampling}\label{participants-and-sampling}}
-
-Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution).
-From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
-This resulted in 18 unique time points of data collection during the pandemic.
-The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
-Of the initial 915 participants, 634 returned to complete the survey at Time 2, while 347 participants continued through to Time 18 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
-
-\hypertarget{measures}{%
-\subsection{Measures}\label{measures}}
-
-\hypertarget{self-reported-mask-wearing}{%
-\subsubsection{Self-reported mask wearing}\label{self-reported-mask-wearing}}
-
-At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
-
-\hypertarget{perceived-descriptive-and-injunctive-social-norms}{%
-\subsubsection{Perceived descriptive and injunctive social norms}\label{perceived-descriptive-and-injunctive-social-norms}}
-
-In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
-
-Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
-We measured perceived descriptive social norms as the mean of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
-
-Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
-We measured perceived injunctive social norms as the mean of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
-
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
-We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
-Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).
-
-\hypertarget{additional-control-variables}{%
-\subsubsection{Additional control variables}\label{additional-control-variables}}
-
-To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
-In this causal model, we included two kinds of non-social beliefs highlighted by previous research (33): factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
-These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
-In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic (42).
-Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
-
-Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
-Factual beliefs were measured as the mean of the following two items: ``\emph{I wear a face mask when going out in public to keep myself from getting sick}'' and ``\emph{I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet}''.
-Personal normative beliefs were measured with a single item: ``\emph{Wearing a face mask when going out in public is the right thing to do}''.
-These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
-
-Political orientation was measured in the first time point only.
-We measured political orientation as the mean of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
-These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
-
-\hypertarget{statistical-analysis}{%
-\subsection{Statistical analysis}\label{statistical-analysis}}
-
-To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
-First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
-These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these multilevel regression models using the \emph{lme4} R package (53) and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data (41).
-This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time \emph{t-1} predicts being above average in another variable at time \emph{t}.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets (41, 54).
-
-We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package (55).
-In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
-For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
-We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
-We restricted the analysis to the ten time points with available data for all five variables.
-Full information maximum likelihood estimation was used to deal with missing data.
-
-All analyses were conducted in R v4.1.1 (56).
-Visualizations were generated using the \emph{cowplot} (57) and \emph{ggplot2} (58) packages.
-The manuscript was reproducibly generated using the \emph{targets} (59) and \emph{papaja} (60) packages.
+Our work thus underscores the importance of consistent, visible community adherence for encouraging protective behaviors in response to global pandemics like COVID-19.
 
@@
 
-This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
-We would also like to thank the Language, Culture, and Cognition lab at the University of Auckland for providing feedback on a previous version of this manuscript.
+{[}REDACTED{]}
 
@@
 
-SLH, JDA, DGB, AV, ERH, AA, and PMT conceptualized the study.
-SLH, SC, JDA, DGB, and AV oversaw the data curation, investigation, and methodology of the study.
-SLH and SC wrote the first draft of the paper.
-SC conducted the formal analysis and created all visualizations.
-ERH, AA, and PMT provided funding and supervision for the study.
-All authors reviewed and edited the final draft of the paper.
+{[}REDACTED{]}
 
@@
 
-\hypertarget{research-transparency-and-reproducibility}{%
-\section{Research Transparency and Reproducibility}\label{research-transparency-and-reproducibility}}
-
-All data and code to reproduce the statistical analyses in this manuscript can be found on GitHub: \url{https://github.com/ScottClaessens/covidMaskWearing}
+\hypertarget{open-practices-statement}{%
+\section{Open Practices Statement}\label{open-practices-statement}}
+
+All data and code to reproduce the statistical analyses in this manuscript are publicly available on GitHub: \url{https://github.com/ScottClaessens/covidMaskWearing} This study was not preregistered.
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
+Amato, R., Lacasa, L., Díaz-Guilera, A., \& Baronchelli, A.
+(2018).
+The dynamics of norm change in the cultural evolution of language.
+\emph{Proceedings of the National Academy of Sciences}, \emph{115}(33), 8260--8265.
+\url{https://doi.org/10.1073/pnas.1721059115}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
+Aust, F., \& Barth, M. (2022). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://github.com/crsh/papaja}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
+Bates, D., Mächler, M., Bolker, B., \& Walker, S.
+(2015).
+Fitting linear mixed-effects models using lme4.
+\emph{Journal of Statistical Software}, \emph{67}(1), 1--48.
+\url{https://doi.org/10.18637/jss.v067.i01}
+
+\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
+Baxter-King, R., Brown, J.
+R., Enos, R.
+D., Naeim, A., \& Vavreck, L.
+(2022).
+How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}.
+\emph{Proceedings of the National Academy of Sciences}, \emph{119}(21), e2116311119.
+\url{https://doi.org/10.1073/pnas.2116311119}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
+Bicchieri, C., Lindemans, J.
+W., \& Jiang, T.
+(2014).
+A structured approach to a diagnostic of collective practices.
+\emph{Frontiers in Psychology}, \emph{5}, 1418.
+\url{https://doi.org/10.3389/fpsyg.2014.01418}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2009}{}}%
-\CSLLeftMargin{1. }%
-\CSLRightInline{C. Bicchieri, E. Xiao, \href{https://doi.org/10.1002/bdm.621}{Do the right thing: But only if others do so}. \emph{Journal of Behavioral Decision Making} \textbf{22}, 191--208 (2009).}
+Bicchieri, C., \& Xiao, E. (2009). Do the right thing: But only if others do so. \emph{Journal of Behavioral Decision Making}, \emph{22}(2), 191--208. \url{https://doi.org/10.1002/bdm.621}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
+Bokemper, S.
+E., Cucciniello, M., Rotesi, T., Pin, P., Malik, A.
+A., Willebrand, K., Paintsil, E.
+E., Omer, S.
+B., Huber, G.
+A., \& Melegaro, A.
+(2021).
+Experimental evidence that changing beliefs about mask efficacy and social norms increase mask wearing for {COVID-19} risk reduction: Results from the {United States} and {Italy}.
+\emph{PLOS ONE}, \emph{16}(10), e0258282.
+\url{https://doi.org/10.1371/journal.pone.0258282}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CDC2020}{}}%
+Centers for Disease Control and Prevention.
+(2020).
+\emph{{CDC calls on Americans to wear masks to prevent COVID-19 spread}}.
+\url{https://www.cdc.gov/media/releases/2020/p0714-americans-to-wear-masks.html}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CDC2022}{}}%
+Centers for Disease Control and Prevention. (2022). \emph{{CDC Museum COVID-19 Timeline}}. \url{https://www.cdc.gov/museum/timeline/covid19.html}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
+Chaudhuri, A.
+(2011).
+Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature.
+\emph{Experimental Economics}, \emph{14}(1), 47--83.
+\url{https://doi.org/10.1007/s10683-010-9257-1}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2021}{}}%
+Cialdini, R.
+B., \& Jacobson, R.
+P.
+(2021).
+Influences of social norms on climate change-related behaviors.
+\emph{Current Opinion in Behavioral Sciences}, \emph{42}, 1--8.
+\url{https://doi.org/10.1016/j.cobeha.2021.01.005}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Cialdini1991}{}}%
-\CSLLeftMargin{2. }%
-\CSLRightInline{R.
+Cialdini, R.
+B., Kallgren, C.
+A., \& Reno, R.
+R.
+(1991).
+\emph{A focus theory of normative conduct: A theoretical refinement and reevaluation of the role of norms in human behavior} (M.
+P.
+Zanna, Ed.; Vol.
+24, pp.
+201--234).
+Academic Press.
+\url{https://doi.org/10.1016/S0065-2601(08)60330-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
+Deutsch, M., \& Gerard, H.
 B.
-Cialdini, C.
+(1955).
+A study of normative and informational social influences upon individual judgment.
+\emph{The Journal of Abnormal and Social Psychology}, \emph{51}(3), 629--636.
+\url{https://doi.org/10.1037/h0046408}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018a}{}}%
+Fehr, E., \& Schurtenberger, I. (2018). Normative foundations of human cooperation. \emph{Nature Human Behaviour}, \emph{2}(7), 458--468. \url{https://doi.org/10.1038/s41562-018-0385-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
+Gelfand, M.
+J., \& Harrington, J.
+R.
+(2015).
+The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?
+\emph{Journal of Cross-Cultural Psychology}, \emph{46}(10), 1273--1278.
+\url{https://doi.org/10.1177/0022022115600796}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
+Goldstein, N.
+J., Cialdini, R.
+B., \& Griskevicius, V.
+(2008).
+A room with a viewpoint: Using social norms to motivate environmental conservation in hotels.
+\emph{Journal of Consumer Research}, \emph{35}(3), 472--482.
+\url{https://doi.org/10.1086/586910}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
+Granger, C. W. J. (1969). Investigating causal relations by econometric models and cross-spectral methods. \emph{Econometrica}, \emph{37}(3), 424--438. \url{https://doi.org/10.2307/1912791}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
+Hamaker, E. L., Kuiper, R. M., \& Grasman, R. P. P. P. (2015). A critique of the cross-lagged panel model. \emph{Psychological Methods}, \emph{20}(1), 102--116. \url{https://doi.org/10.1037/a0038889}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Heiman2023}{}}%
+Heiman, S.
+L., Claessens, S., Ayers, J.
+D., Beltrán, D.
+G., Van Horn, A., Hirt, E.
+R., Aktipis, A., \& Todd, P.
+M.
+(2023).
+\emph{Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic}.
+GitHub repository.
+\url{https://github.com/ScottClaessens/covidMaskWearing}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
+Henrich, J.
+(2015).
+\emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}.
+Princeton University Press.
+\url{https://doi.org/10.2307/j.ctvc77f0d}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
+Landau, W.
+M.
+(2021).
+The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing.
+\emph{Journal of Open Source Software}, \emph{6}(57), 2959.
+\url{https://doi.org/10.21105/joss.02959}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
+Larkin, C., Sanders, M., Andresen, I., \& Algate, F.
+(2018).
+\emph{Testing local descriptive norms and salience of enforcement action: A field experiment to increase tax collection} (SSRN Scholarly Paper No.
+3167575).
+\url{https://doi.org/10.2139/ssrn.3167575}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Legros2020}{}}%
+Legros, S., \& Cislaghi, B.
+(2020).
+Mapping the social-norms literature: An overview of reviews.
+\emph{Perspectives on Psychological Science}, \emph{15}(1), 62--80.
+\url{https://doi.org/10.1177/1745691619866455}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
+Macy, J.
+T., Owens, C., Mullis, K., \& Middlestadt, S.
+E.
+(2021).
+The role of self-efficacy and injunctive norms in helping older adults decide to stay home during the {COVID-19} pandemic.
+\emph{Frontiers in Public Health}, \emph{9}, 660813.
+\url{https://doi.org/10.3389/fpubh.2021.660813}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
+Markus, H.
+R., \& Kitayama, S.
+(2010).
+Cultures and selves: A cycle of mutual constitution.
+\emph{Perspectives on Psychological Science}, \emph{5}(4), 420--430.
+\url{https://doi.org/10.1177/1745691610375557}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
+Nigbur, D., Lyons, E., \& Uzzell, D.
+(2010).
+Attitudes, norms, identity and environmental behaviour: Using an expanded theory of planned behaviour to predict participation in a kerbside recycling programme.
+\emph{British Journal of Social Psychology}, \emph{49}(2), 259--284.
+\url{https://doi.org/10.1348/014466609X449395}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
+Orth, U., Meier, L.
+L., Bühler, J.
+L., Dapp, L.
+C., Krauss, S., Messerli, D., \& Robins, R.
+W.
+(2022).
+Effect size guidelines for cross-lagged effects.
+\emph{Psychological Methods}.
+\url{https://doi.org/10.1037/met0000499}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
+R Core Team. (2022). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Roos2015}{}}%
+Roos, P., Gelfand, M., Nau, D., \& Lun, J.
+(2015).
+Societal threat and cultural variation in the strength of social norms: An evolutionary basis.
+\emph{Organizational Behavior and Human Decision Processes}, \emph{129}, 14--23.
+\url{https://doi.org/10.1016/j.obhdp.2015.01.003}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
+Rosseel, Y. (2012). {lavaan}: An {R} package for structural equation modeling. \emph{Journal of Statistical Software}, \emph{48}(2), 1--36. \url{https://doi.org/10.18637/jss.v048.i02}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
+Rudert, S.
+C., \& Janke, S.
+(2021).
+Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic.
+\emph{Group Processes \& Intergroup Relations}, \emph{25}, 1819--1835.
+\url{https://doi.org/10.1177/13684302211023562}
+
+\leavevmode\vadjust pre{\hypertarget{ref-SchmittBeck2015}{}}%
+Schmitt-Beck, R.
+(2015).
+Bandwagon effect.
+In \emph{The international encyclopedia of political communication} (pp.
+1--5).
+John Wiley \& Sons, Ltd.
+\url{https://doi.org/10.1002/9781118541555.wbiepc015}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schultz2007}{}}%
+Schultz, P.
+W., Nolan, J.
+M., Cialdini, R.
+B., Goldstein, N.
+J., \& Griskevicius, V.
+(2007).
+The constructive, destructive, and reconstructive power of social norms.
+\emph{Psychological Science}, \emph{18}(5), 429--434.
+\url{https://doi.org/10.1111/j.1467-9280.2007.01917.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
+Smith, L.
+G.
+E., Thomas, E.
+F., \& McGarty, C.
+(2015).
+"We must be the change we want to see in the world": Integrating norms and identities through social interaction.
+\emph{Political Psychology}, \emph{36}(5), 543--557.
+\url{https://doi.org/10.1111/pops.12180}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
+Titlestad, K., Snijders, T.
 A.
-Kallgren, R.
-R.
-Reno, {``\href{https://doi.org/10.1016/S0065-2601(08)60330-5}{A focus theory of normative conduct: A theoretical refinement and reevaluation of the role of norms in human behavior}''} in Advances in experimental social psychology., M.
-P.
-Zanna, Ed.
-(Academic Press, 1991), pp.
-201--234.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018a}{}}%
-\CSLLeftMargin{3. }%
-\CSLRightInline{E. Fehr, I. Schurtenberger, \href{https://doi.org/10.1038/s41562-018-0385-5}{Normative foundations of human cooperation}. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).}
+B., Durrheim, K., Quayle, M., \& Postmes, T.
+(2019).
+The dynamic emergence of cooperative norms in a social dilemma.
+\emph{Journal of Experimental Social Psychology}, \emph{84}, 103799.
+\url{https://doi.org/10.1016/j.jesp.2019.03.010}
 
 \leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
-\CSLLeftMargin{4. }%
-\CSLRightInline{G.
-A.
-van Kleef, M.
-J.
-Gelfand, J.
-Jetten, \href{https://doi.org/10.1016/j.jesp.2019.05.002}{The dynamic nature of social norms: New perspectives on norm development, impact, violation, and enforcement}.
-\emph{Journal of Experimental Social Psychology} \textbf{84}, 103814 (2019).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Legros2020}{}}%
-\CSLLeftMargin{5. }%
-\CSLRightInline{S.
-Legros, B.
-Cislaghi, \href{https://doi.org/10.1177/1745691619866455}{Mapping the social-norms literature: An overview of reviews}.
-\emph{Perspectives on Psychological Science} \textbf{15}, 62--80 (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Asch1956}{}}%
-\CSLLeftMargin{6. }%
-\CSLRightInline{S.
-E.
-Asch, \href{https://doi.org/10.1037/h0093718}{Studies of independence and conformity: I.
-A minority of one against a unanimous majority}.
-\emph{Psychological Monographs: General and Applied} \textbf{70}, 1--70 (1956).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Tomasello2014}{}}%
-\CSLLeftMargin{7. }%
-\CSLRightInline{M. Tomasello, \href{https://doi.org/10.1002/ejsp.2015}{The ultra-social animal}. \emph{European Journal of Social Psychology} \textbf{44}, 187--194 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Riedl2012}{}}%
-\CSLLeftMargin{8. }%
-\CSLRightInline{K.
-Riedl, K.
-Jensen, J.
-Call, M.
-Tomasello, \href{https://doi.org/10.1073/pnas.1203179109}{No third-party punishment in chimpanzees}.
-\emph{Proceedings of the National Academy of Sciences} \textbf{109}, 14824--14829 (2012).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Jensen2014}{}}%
-\CSLLeftMargin{9. }%
-\CSLRightInline{K.
-Jensen, A.
-Vaish, M.
-F.
-H.
-Schmidt, \href{https://doi.org/10.3389/fpsyg.2014.00822}{The emergence of human prosociality: Aligning with others through feelings, concerns, and norms}.
-\emph{Frontiers in Psychology} \textbf{5}, 822 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rakoczy2013}{}}%
-\CSLLeftMargin{10. }%
-\CSLRightInline{H. Rakoczy, M. F. H. Schmidt, \href{https://doi.org/10.1111/cdep.12010}{The early ontogeny of social norms}. \emph{Child Development Perspectives} \textbf{7}, 17--21 (2013).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Schmidt2012}{}}%
-\CSLLeftMargin{11. }%
-\CSLRightInline{M.
-F.
-H.
-Schmidt, M.
-Tomasello, \href{https://doi.org/10.1177/0963721412448659}{Young children enforce social norms}.
-\emph{Current Directions in Psychological Science} \textbf{21}, 232--236 (2012).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Vaish2011}{}}%
-\CSLLeftMargin{12. }%
-\CSLRightInline{A.
-Vaish, M.
-Carpenter, M.
-Tomasello, \href{https://doi.org/10.1037/a0024462}{Young children's responses to guilt displays}.
-\emph{Developmental Psychology} \textbf{47}, 1248--1262 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Schaumberg2022}{}}%
-\CSLLeftMargin{13. }%
-\CSLRightInline{R.
-L.
-Schaumberg, S.
-E.
-Skowronek, \href{https://doi.org/10.1177/09567976221075303}{Shame broadcasts social norms: The positive social effects of shame on norm acquisition and normative behavior}.
-\emph{Psychological Science} \textbf{33}, 1257--1277 (2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2004}{}}%
-\CSLLeftMargin{14. }%
-\CSLRightInline{E. Fehr, U. Fischbacher, \href{https://doi.org/10.1016/S1090-5138(04)00005-4}{Third-party punishment and social norms}. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2006}{}}%
-\CSLLeftMargin{15. }%
-\CSLRightInline{C.
-Bicchieri, \emph{The grammar of society: The nature and dynamics of social norms} (Cambridge University Press, 2005) https:/doi.org/\href{https://doi.org/10.1017/CBO9780511616037}{10.1017/CBO9780511616037}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2011}{}}%
-\CSLLeftMargin{16. }%
-\CSLRightInline{M.
-J.
-Gelfand, \emph{et al.}, \href{https://doi.org/10.1126/science.1197754}{Differences between tight and loose cultures: A 33-nation study}.
-\emph{Science} \textbf{332}, 1100--1104 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
-\CSLLeftMargin{17. }%
-\CSLRightInline{J.
-Henrich, \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter} (Princeton University Press, 2015) https:/doi.org/\href{https://doi.org/10.2307/j.ctvc77f0d}{10.2307/j.ctvc77f0d}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Roos2015}{}}%
-\CSLLeftMargin{18. }%
-\CSLRightInline{P.
-Roos, M.
-Gelfand, D.
-Nau, J.
-Lun, \href{https://doi.org/10.1016/j.obhdp.2015.01.003}{Societal threat and cultural variation in the strength of social norms: An evolutionary basis}.
-\emph{Organizational Behavior and Human Decision Processes} \textbf{129}, 14--23 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Aktipis2011}{}}%
-\CSLLeftMargin{19. }%
-\CSLRightInline{A.
-Aktipis, L.
-Cronk, R.
-de Aguiar, \href{https://doi.org/10.1007/s10745-010-9364-9}{Risk-pooling and herd survival: An agent-based model of a {M}aasai gift-giving system}.
-\emph{Human Ecology} \textbf{39}, 131--140 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Chudek2011}{}}%
-\CSLLeftMargin{20. }%
-\CSLRightInline{M.
-Chudek, J.
-Henrich, \href{https://doi.org/10.1016/j.tics.2011.03.003}{Culture--gene coevolution, norm-psychology and the emergence of human prosociality}.
-\emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
-\CSLLeftMargin{21. }%
-\CSLRightInline{M.
-Deutsch, H.
-B.
-Gerard, \href{https://doi.org/10.1037/h0046408}{A study of normative and informational social influences upon individual judgment}.
-\emph{The Journal of Abnormal and Social Psychology} \textbf{51}, 629--636 (1955).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Schultz2007}{}}%
-\CSLLeftMargin{22. }%
-\CSLRightInline{P.
-W.
-Schultz, J.
-M.
-Nolan, R.
-B.
-Cialdini, N.
-J.
-Goldstein, V.
-Griskevicius, \href{https://doi.org/10.1111/j.1467-9280.2007.01917.x}{The constructive, destructive, and reconstructive power of social norms}.
-\emph{Psychological Science} \textbf{18}, 429--434 (2007).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2007}{}}%
-\CSLLeftMargin{23. }%
-\CSLRightInline{R.
-B.
-Cialdini, \href{https://doi.org/10.1007/s11336-006-1560-6}{Descriptive social norms as underappreciated sources of social control}.
-\emph{Psychometrika} \textbf{72}, 263--268 (2007).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cialdini2021}{}}%
-\CSLLeftMargin{24. }%
-\CSLRightInline{R.
-B.
-Cialdini, R.
-P.
-Jacobson, \href{https://doi.org/10.1016/j.cobeha.2021.01.005}{Influences of social norms on climate change-related behaviors}.
-\emph{Current Opinion in Behavioral Sciences} \textbf{42}, 1--8 (2021).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Boyd1985}{}}%
-\CSLLeftMargin{25. }%
-\CSLRightInline{R. Boyd, P. J. Richerson, \emph{Culture and the evolutionary process} (University of Chicago Press, 1985).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Kushnick2014}{}}%
-\CSLLeftMargin{26. }%
-\CSLRightInline{G.
-Kushnick, R.
-D.
-Gray, F.
-M.
-Jordan, \href{https://doi.org/10.1016/j.evolhumbehav.2014.03.001}{The sequential evolution of land tenure norms}.
-\emph{Evolution and Human Behavior} \textbf{35}, 309--318 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Jordan2009}{}}%
-\CSLLeftMargin{27. }%
-\CSLRightInline{F.
-M.
-Jordan, R.
-D.
-Gray, S.
-J.
-Greenhill, R.
-Mace, \href{https://doi.org/10.1098/rspb.2009.0088}{Matrilocal residence is ancestral in {A}ustronesian societies}.
-\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{276}, 1957--1964 (2009).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
-\CSLLeftMargin{28. }%
-\CSLRightInline{K.
-Titlestad, T.
-A.
-B.
-Snijders, K.
-Durrheim, M.
-Quayle, T.
-Postmes, \href{https://doi.org/10.1016/j.jesp.2019.03.010}{The dynamic emergence of cooperative norms in a social dilemma}.
-\emph{Journal of Experimental Social Psychology} \textbf{84}, 103799 (2019).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018b}{}}%
-\CSLLeftMargin{29. }%
-\CSLRightInline{E.
-Fehr, I.
-Schurtenberger, {``The dynamics of norm formation and norm decay''} (Department of Economics 278, University of Zurich, 2018) https:/doi.org/\href{https://doi.org/10.5167/uzh-147925}{10.5167/uzh-147925}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
-\CSLLeftMargin{30. }%
-\CSLRightInline{D.
-Nigbur, E.
-Lyons, D.
-Uzzell, \href{https://doi.org/10.1348/014466609X449395}{Attitudes, norms, identity and environmental behaviour: Using an expanded theory of planned behaviour to predict participation in a kerbside recycling programme}.
-\emph{British Journal of Social Psychology} \textbf{49}, 259--284 (2010).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
-\CSLLeftMargin{31. }%
-\CSLRightInline{C.
-Larkin, M.
-Sanders, I.
-Andresen, F.
-Algate, {``Testing local descriptive norms and salience of enforcement action: A field experiment to increase tax collection''} (2018) https:/doi.org/\href{https://doi.org/10.2139/ssrn.3167575}{10.2139/ssrn.3167575}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
-\CSLLeftMargin{32. }%
-\CSLRightInline{N.
-J.
-Goldstein, R.
-B.
-Cialdini, V.
-Griskevicius, \href{https://doi.org/10.1086/586910}{A room with a viewpoint: Using social norms to motivate environmental conservation in hotels}.
-\emph{Journal of Consumer Research} \textbf{35}, 472--482 (2008).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
-\CSLLeftMargin{33. }%
-\CSLRightInline{C.
-Bicchieri, J.
-W.
-Lindemans, T.
-Jiang, \href{https://doi.org/10.3389/fpsyg.2014.01418}{A structured approach to a diagnostic of collective practices}.
-\emph{Frontiers in Psychology} \textbf{5}, 1418 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
-\CSLLeftMargin{34. }%
-\CSLRightInline{L.
-G.
-E.
-Smith, E.
-F.
-Thomas, C.
-McGarty, \href{https://doi.org/10.1111/pops.12180}{"We must be the change we want to see in the world": Integrating norms and identities through social interaction}.
-\emph{Political Psychology} \textbf{36}, 543--557 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ostrom1990}{}}%
-\CSLLeftMargin{35. }%
-\CSLRightInline{E.
-Ostrom, \emph{Governing the commons: The evolution of institutions for collective action} (Cambridge University Press, 1990) https:/doi.org/\href{https://doi.org/10.1017/CBO9780511807763}{10.1017/CBO9780511807763}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CDC2022}{}}%
-\CSLLeftMargin{36. }%
-\CSLRightInline{Centers for Disease Control and Prevention, \href{https://www.cdc.gov/museum/timeline/covid19.html}{{CDC Museum COVID-19 Timeline}} (2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CDC2020}{}}%
-\CSLLeftMargin{37. }%
-\CSLRightInline{Centers for Disease Control and Prevention, \href{https://www.cdc.gov/media/releases/2020/p0714-americans-to-wear-masks.html}{{CDC calls on Americans to wear masks to prevent COVID-19 spread}} (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
-\CSLLeftMargin{38. }%
-\CSLRightInline{J.
-T.
-Macy, C.
-Owens, K.
-Mullis, S.
-E.
-Middlestadt, \href{https://doi.org/10.3389/fpubh.2021.660813}{The role of self-efficacy and injunctive norms in helping older adults decide to stay home during the {COVID-19} pandemic}.
-\emph{Frontiers in Public Health} \textbf{9}, 660813 (2021).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
-\CSLLeftMargin{39. }%
-\CSLRightInline{S.
-E.
-Bokemper, \emph{et al.}, \href{https://doi.org/10.1371/journal.pone.0258282}{Experimental evidence that changing beliefs about mask efficacy and social norms increase mask wearing for {COVID-19} risk reduction: Results from the {United States} and {Italy}}.
-\emph{PLOS ONE} \textbf{16}, e0258282 (2021).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
-\CSLLeftMargin{40. }%
-\CSLRightInline{S.
-C.
-Rudert, S.
-Janke, \href{https://doi.org/10.1177/13684302211023562}{Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic}.
-\emph{Group Processes \& Intergroup Relations} \textbf{25}, 1819--1835 (2021).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{E.
-L.
-Hamaker, R.
-M.
-Kuiper, R.
-P.
-P.
-P.
-Grasman, \href{https://doi.org/10.1037/a0038889}{A critique of the cross-lagged panel model}.
-\emph{Psychological Methods} \textbf{20}, 102--116 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
-\CSLLeftMargin{42. }%
-\CSLRightInline{R.
-Baxter-King, J.
-R.
-Brown, R.
-D.
-Enos, A.
-Naeim, L.
-Vavreck, \href{https://doi.org/10.1073/pnas.2116311119}{How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}}.
-\emph{Proceedings of the National Academy of Sciences} \textbf{119}, e2116311119 (2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
-\CSLLeftMargin{43. }%
-\CSLRightInline{U.
-Orth, \emph{et al.}, Effect size guidelines for cross-lagged effects.
-\emph{Psychological Methods} (2022) https:/doi.org/\href{https://doi.org/10.1037/met0000499}{10.1037/met0000499}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
-\CSLLeftMargin{44. }%
-\CSLRightInline{H.
-R.
-Markus, S.
-Kitayama, \href{https://doi.org/10.1177/1745691610375557}{Cultures and selves: A cycle of mutual constitution}.
-\emph{Perspectives on Psychological Science} \textbf{5}, 420--430 (2010).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
-\CSLLeftMargin{45. }%
-\CSLRightInline{R.
-Amato, L.
-Lacasa, A.
-Díaz-Guilera, A.
-Baronchelli, \href{https://doi.org/10.1073/pnas.1721059115}{The dynamics of norm change in the cultural evolution of language}.
-\emph{Proceedings of the National Academy of Sciences} \textbf{115}, 8260--8265 (2018).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
-\CSLLeftMargin{46. }%
-\CSLRightInline{M.
-J.
-Gelfand, J.
-R.
-Harrington, \href{https://doi.org/10.1177/0022022115600796}{The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?} \emph{Journal of Cross-Cultural Psychology} \textbf{46}, 1273--1278 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hogg2008}{}}%
-\CSLLeftMargin{47. }%
-\CSLRightInline{M.
-A.
-Hogg, Z.
-P.
-Hohman, J.
-E.
-Rivera, \href{https://doi.org/10.1111/j.1751-9004.2008.00099.x}{Why do people join groups?
-Three motivational accounts from social psychology}.
-\emph{Social and Personality Psychology Compass} \textbf{2}, 1269--1280 (2008).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wellen1998}{}}%
-\CSLLeftMargin{48. }%
-\CSLRightInline{J.
-M.
-Wellen, M.
-A.
-Hogg, D.
-J.
-Terry, \href{https://doi.org/10.1037/1089-2699.2.1.48}{Group norms and attitude-behavior consistency: The role of group salience and mood}.
-\emph{Group Dynamics: Theory, Research, and Practice} \textbf{2}, 48--56 (1998).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
-\CSLLeftMargin{49. }%
-\CSLRightInline{A.
-Chaudhuri, \href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
-\emph{Experimental Economics} \textbf{14}, 47--83 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Neighbors2008}{}}%
-\CSLLeftMargin{50. }%
-\CSLRightInline{C.
-Neighbors, \emph{et al.}, \href{https://doi.org/10.1037/a0013043}{The relative impact of injunctive norms on college student drinking: The role of reference group.} \emph{Psychology of Addictive Behaviors} \textbf{22}, 576--581 (2008).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Remedios2022}{}}%
-\CSLLeftMargin{51. }%
-\CSLRightInline{J. D. Remedios, \href{https://doi.org/10.1038/s44159-022-00024-4}{Psychology must grapple with whiteness}. \emph{Nature Reviews Psychology} \textbf{1}, 125--126 (2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
-\CSLLeftMargin{52. }%
-\CSLRightInline{H.
-Natanson, \href{https://www.washingtonpost.com/education/2022/02/25/peer-pressure-mask-optional-schools/}{Peer pressure is ending mask usage in schools}.
-\emph{The Washingston Post} (2022) (August 18, 2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
-\CSLLeftMargin{53. }%
-\CSLRightInline{D.
-Bates, M.
-Mächler, B.
-Bolker, S.
-Walker, \href{https://doi.org/10.18637/jss.v067.i01}{Fitting linear mixed-effects models using lme4}.
-\emph{Journal of Statistical Software} \textbf{67}, 1--48 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
-\CSLLeftMargin{54. }%
-\CSLRightInline{C.
-W.
-J.
-Granger, \href{https://doi.org/10.2307/1912791}{Investigating causal relations by econometric models and cross-spectral methods}.
-\emph{Econometrica} \textbf{37}, 424--438 (1969).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
-\CSLLeftMargin{55. }%
-\CSLRightInline{Y. Rosseel, \href{https://doi.org/10.18637/jss.v048.i02}{{lavaan}: An {R} package for structural equation modeling}. \emph{Journal of Statistical Software} \textbf{48}, 1--36 (2012).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-\CSLLeftMargin{56. }%
-\CSLRightInline{R Core Team, \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}} (R Foundation for Statistical Computing, 2022).}
+van Kleef, G.
+A., Gelfand, M.
+J., \& Jetten, J.
+(2019).
+The dynamic nature of social norms: New perspectives on norm development, impact, violation, and enforcement.
+\emph{Journal of Experimental Social Psychology}, \emph{84}, 103814.
+\url{https://doi.org/10.1016/j.jesp.2019.05.002}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
+Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-\CSLLeftMargin{57. }%
-\CSLRightInline{C. O. Wilke, \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}} (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-\CSLLeftMargin{58. }%
-\CSLRightInline{H. Wickham, \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}} (Springer-Verlag New York, 2016).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{59. }%
-\CSLRightInline{W.
-M.
-Landau, \href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
-\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-\CSLLeftMargin{60. }%
-\CSLRightInline{F. Aust, M. Barth, \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}} (2022).}
+Wilke, C. O. (2020). \emph{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}. \url{https://CRAN.R-project.org/package=cowplot}
 
@@
 \subsection{Supplementary Figures}\label{supplementary-figures}}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotUSMap-1.pdf}
+\caption{\label{fig:plotUSMap}\emph{Map of the United States with participant zip code locations.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAttrition-1.pdf}
+\caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
+\end{figure}
+
+\newpage
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotAttrition-1.pdf}
-\caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotUSMap-1.pdf}
-\caption{\label{fig:plotUSMap}\emph{Map of the United States with participant zip code locations.}}
-\end{figure}
-
-\newpage
-
 \hypertarget{supplementary-tables}{%
 \subsection{Supplementary Tables}\label{supplementary-tables}}
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{longtable}{m{2.5cm}m{2.5cm}m{9cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:itemTable}List of norm interpretation questions asked at Time 7.
+\emph{These questions were preceded by the following text:} ``There may or may not be a difference between what people around you are doing and what they should be doing.
+You can learn about what people are doing and what they should be doing in different ways.
+For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both''.
+\emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
+\toprule
+Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:itemTable} continued}}\\
+\toprule
+Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
+\midrule
+\endhead
+Provides descriptive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone is doing?\\
+ &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone is doing?\\
+ & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone is doing?\\
+ &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone is doing?\\
+Provides injunctive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone should be doing?\\
+ &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone should be doing?\\
+ & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone should be doing?\\
+ &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone should be doing?\\
+\bottomrule
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
 
@@
 
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{longtable}{m{2.5cm}m{2.5cm}m{9cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:itemTable}List of norm interpretation questions asked at Time 7.
-\emph{These questions were preceded by the following text:} ``There may or may not be a difference between what people around you are doing and what they should be doing.
-You can learn about what people are doing and what they should be doing in different ways.
-For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both''.
-\emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
-\toprule
-Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:itemTable} continued}}\\
-\toprule
-Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
-\midrule
-\endhead
-Provides descriptive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone is doing?\\
- &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone is doing?\\
- & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone is doing?\\
- &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone is doing?\\
-Provides injunctive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone should be doing?\\
- &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone should be doing?\\
- & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone should be doing?\\
- &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone should be doing?\\
-\bottomrule
-\end{longtable}
-
-\end{ThreePartTable}
-\end{center}
-
-\newpage
-
 \hypertarget{supplementary-references}{%

### 2023-03-14 manuscript.Rmd
resubmission

@@
 keywords          : "social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
-wordcount         : "4521 words"
+wordcount         : "4381 words"
 bibliography      : ["r-references.bib"]
@@
 linenumbers       : yes
-mask              : yes
+mask              : no
 draft             : no
@@
 appendix          : "appendix.Rmd"
-csl               : apa.csl
+csl               : nature.csl
 ---
@@
            plot1, plot2, plot3, plot4, plot5, plot6, plot7,
-           itemTable, changePointsTable, lavaanTable))
+           plot8, itemTable, changePointsTable, lavaanTable))
 ```
-
-# \normalfont Statement of Relevance
-
-Social norms have been identified as important drivers of human cooperation, but the emergence of social norms within populations and their subsequent effects on cooperative behavior are not well understood.
-Here, we use mask wearing during the COVID-19 pandemic as a real-world setting in which to study the emergence of social norms.
-Over two years in the United States, we find that social norms and mask wearing are tightly coupled and change quickly in response to public health recommendations.
-Moreover, longitudinal modeling suggests that mask wearing is causally preceded by perceptions of descriptive norms (i.e.
-what people _are_ doing) but not injunctive norms (i.e.
-what people _ought_ to be doing).
-These findings show that visible community adherence is an important driver of mask wearing, suggesting potential avenues for behavior change in future global pandemics.
-
-\newpage
-
-\begin{center}Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic\end{center}
 
@@
 In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring populations [@Henrich2015].
-However, less is known about how social norms arise endogenously within populations in the short term.
+However, less is known about how social norms arise endogenously within populations in the short term, including how descriptive norms arise and interact with injunctive norms to shape behavior.
 Recent work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019].
@@
 With regards to normative influences on behavior, studies have demonstrated positive effects of descriptive norms on a variety of cooperative behaviors, including recycling [@Nigbur2010], paying taxes [@Larkin2018], and sustainably reusing towels in hotels [@Goldstein2008].
-However, these studies have two main limitations.
+However, these studies have two key aspects that limit their ability to assess the causal impact of norms, both of which we address in our current work.
 First, studies have not accounted for other potential non-social explanations for behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
 These non-social beliefs, labeled "factual beliefs" and "personal normative beliefs" [@Bicchieri2014], often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
-For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common behavior).
+For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common or socially approved behavior).
 It is thus important to control for non-social beliefs in studies of social norms.
-Second, studies have tended to follow cross-sectional experimental designs.
+Second, studies have tended to follow cross-sectional experimental designs in which social norm perceptions are manipulated by the researchers.
 However, social norms are not static: they change dynamically over time through processes of deliberation and interaction [@Smith2015].
@@
 Second, how do descriptive and injunctive mask wearing norms causally influence mask wearing?
-
-# Materials and Methods
-
-## Ethical approval
-
-This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
-
-## Participants and sampling
-
-Using the platform Prolific (https://www.prolific.co/), we distributed surveys to a representative sample of adults from the United States (_n_ = `r nrow(d)`, _M_~age~ = `r round(mean(d$Age.1, na.rm = TRUE))` years, `r round(mean(d$Ethnicity.1 == "White or Caucasian", na.rm = TRUE)*100)`% White, `r round(mean(d$Sex.1 == 2, na.rm = TRUE)*100)`% Women; see Supplementary Figure \@ref(fig:plotUSMap) for geographic distribution).
-From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
-This resulted in 18 unique time points of data collection during the pandemic.
-The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
-Of the initial `r nrow(d)` participants, `r sum(!is.na(d$StartDate.2))` returned to complete the survey at Time 2, while `r sum(!is.na(d$StartDate.18))` participants continued through to Time 18 (see Supplementary Figure \@ref(fig:plotAttrition) for attrition rates across all time points).
-
-## Measures
-
-### Self-reported mask wearing
-
-At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
-
-### Perceived descriptive and injunctive social norms
-
-In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
-
-Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
-We measured perceived descriptive social norms as the mean of the following two items: "_What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?_" and "_What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?_" These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
-
-Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
-We measured perceived injunctive social norms as the mean of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
-
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
-We asked participants whether each of the four items informed them about what people _are_ doing or what people _should_ be doing (i.e., giving descriptive or injunctive information).
-Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \@ref(tab:itemTable)).
-
-### Additional control variables
-
-To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \@ref(fig:plotDAG)).
-In this causal model, we included two kinds of non-social beliefs highlighted by previous research [@Bicchieri2014]: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
-These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
-In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic [@BaxterKing2022].
-Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
-
-Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
-Factual beliefs were measured as the mean of the following two items: "_I wear a face mask when going out in public to keep myself from getting sick_" and "_I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet_".
-Personal normative beliefs were measured with a single item: "_Wearing a face mask when going out in public is the right thing to do_".
-These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
-
-Political orientation was measured in the first time point only.
-We measured political orientation as the mean of the following two items: "_How would you describe your political orientation with regard to social issues?_" and "_How would you describe your political orientation with regard to economic issues?_".
-These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
-
-## Statistical analysis
-
-To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
-First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
-These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these multilevel regression models using the _lme4_ R package [@Bates2015] and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data [@Hamaker2015].
-This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time _t-1_ predicts being above average in another variable at time _t_.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015; @Granger1969].
-
-We estimated the random-intercept cross-lagged panel model using the _lavaan_ R package [@Rosseel2012].
-In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
-For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
-We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
-We restricted the analysis to the ten time points with available data for all five variables.
-Full information maximum likelihood estimation was used to deal with missing data.
-
-All analyses were conducted in R v4.1.1 [@RCoreTeam].
-Visualizations were generated using the _cowplot_ [@Wilke2020] and _ggplot2_ [@Wickham2016] packages.
-The manuscript was reproducibly generated using the _targets_ [@Landau2021] and _papaja_ [@Aust2022] packages.
-All code and data are publicly available on GitHub [@Heiman2023].
 
@@
 We are limited in generalizing these findings due to the constraints of our sample and the variables considered.
-While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \@ref(fig:plotAttrition)).
-This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
+While the attrition in our study did not substantially affect the representativeness of our sample (Supplementary Figure \@ref(fig:plotAttritionBreakdown)),
+the reduction in our sample size did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
 Our results also might not generalize to all social norms, behaviors, and social dilemmas.
@@
 Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
+Importantly, this key finding slices two ways.
+Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that _low_ local levels of mask wearing _discouraged_ future personal mask use.
+This echoes recent reports of people in the United States not wanting to be "singled out" by being the only one wearing a mask in their community[@Natanson2022].
 Our work thus underscores the importance of consistent, visible community adherence for encouraging protective behaviors in response to global pandemics like COVID-19.
+
+# Materials and Methods
+
+## Ethical approval
+
+This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
+
+## Participants and sampling
+
+Using the platform Prolific (https://www.prolific.co/), we distributed surveys to a representative sample of adults from the United States (_n_ = `r nrow(d)`, _M_~age~ = `r round(mean(d$Age.1, na.rm = TRUE))` years, `r round(mean(d$Ethnicity.1 == "White or Caucasian", na.rm = TRUE)*100)`% White, `r round(mean(d$Sex.1 == 2, na.rm = TRUE)*100)`% Women; see Supplementary Figure \@ref(fig:plotUSMap) for geographic distribution).
+From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
+This resulted in 18 unique time points of data collection during the pandemic.
+The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
+Of the initial `r nrow(d)` participants, `r sum(!is.na(d$StartDate.2))` returned to complete the survey at Time 2, while `r sum(!is.na(d$StartDate.18))` participants continued through to Time 18 (see Supplementary Figure \@ref(fig:plotAttrition) for attrition rates across all time points).
+However, this attrition did not substantially affect the demographic makeup of the sample through time (Supplementary Figure \@ref(fig:plotAttritionBreakdown)).
+
+## Measures
+
+### Self-reported mask wearing
+
+At every time point, participants were asked about the number of in-person interactions they had in the last week.
+Following this question, participants self-reported their mask wearing by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+
+### Perceived descriptive and injunctive social norms
+
+In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
+
+Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
+We measured perceived descriptive social norms as the mean of the following two items: "_What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?_" and "_What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?_" These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
+
+Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
+We measured perceived injunctive social norms as the mean of the following two items: "_In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?_" and "_How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?_" These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
+
+To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
+We asked participants whether each of the four items informed them about what people _are_ doing or what people _should_ be doing (i.e., giving descriptive or injunctive information).
+Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \@ref(tab:itemTable)).
+
+### Additional control variables
+
+To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \@ref(fig:plotDAG)).
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research [@Bicchieri2014]: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
+These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
+In addition, we also included political orientation as a common cause of all other variables.
+This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic [@BaxterKing2022].
+Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
+
+Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
+Factual beliefs were measured as the mean of the following two items: "_I wear a face mask when going out in public to keep myself from getting sick_" and "_I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet_".
+Personal normative beliefs were measured with a single item: "_Wearing a face mask when going out in public is the right thing to do_".
+These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
+
+Political orientation was measured in the first time point only.
+We measured political orientation as the mean of the following two items: "_How would you describe your political orientation with regard to social issues?_" and "_How would you describe your political orientation with regard to economic issues?_".
+These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
+
+## Statistical analysis
+
+To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
+First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
+Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
+These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
+We estimated these multilevel regression models using the _lme4_ R package [@Bates2015] and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data [@Hamaker2015].
+This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time _t-1_ predicts being above average in another variable at time _t_.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015; @Granger1969].
+
+We estimated the random-intercept cross-lagged panel model using the _lavaan_ R package [@Rosseel2012].
+In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
+For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
+We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We restricted the analysis to the ten time points with available data for all five variables.
+Full information maximum likelihood estimation was used to deal with missing data.
+
+All analyses were conducted in R v4.1.1 [@RCoreTeam].
+Visualizations were generated using the _cowplot_ [@Wilke2020] and _ggplot2_ [@Wickham2016] packages.
+The manuscript was reproducibly generated using the _targets_ [@Landau2021] and _papaja_ [@Aust2022] packages.
+All code and data are publicly available on GitHub [@Heiman2023].
 
@@
 
-[REDACTED]
+This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
+We would also like to thank the Language, Culture, and Cognition lab at the University of Auckland for providing feedback on a previous version of this manuscript.
 
@@
 
-[REDACTED]
+SLH, JDA, DGB, AV, ERH, AA, and PMT conceptualized the study.
+SLH, SC, JDA, DGB, and AV oversaw the data curation, investigation, and methodology of the study.
+SLH and SC wrote the first draft of the paper.
+SC conducted the formal analysis and created all visualizations.
+ERH, AA, and PMT provided funding and supervision for the study.
+All authors reviewed and edited the final draft of the paper.
 

### 2023-03-14 manuscript.tex
resubmission

@@
 \documentclass[
-  man, donotrepeattitle,mask,floatsintext]{apa6}
+  man, donotrepeattitle,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 {}{\PatchFailed}
-\keywords{social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 4521 words}
+\keywords{social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 4381 words}
 \usepackage{lineno}
@@
   pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic},
+  pdfauthor={Samantha L. Heiman*,1, Scott Claessens*,2, Jessica D. Ayers3, Diego Guevara Beltrán4, Andrew Van Horn5,6, Edward R. Hirt1, Athena Aktipis†,4, \& Peter M. Todd†,1,7},
   pdflang={en-EN},
@@
 
-\hypertarget{statement-of-relevance}{%
-\section{\texorpdfstring{\normalfont Statement of Relevance}{Statement of Relevance}}\label{statement-of-relevance}}
-
-Social norms have been identified as important drivers of human cooperation, but the emergence of social norms within populations and their subsequent effects on cooperative behavior are not well understood.
-Here, we use mask wearing during the COVID-19 pandemic as a real-world setting in which to study the emergence of social norms.
-Over two years in the United States, we find that social norms and mask wearing are tightly coupled and change quickly in response to public health recommendations.
-Moreover, longitudinal modeling suggests that mask wearing is causally preceded by perceptions of descriptive norms (i.e.~what people \emph{are} doing) but not injunctive norms (i.e.~what people \emph{ought} to be doing).
-These findings show that visible community adherence is an important driver of mask wearing, suggesting potential avenues for behavior change in future global pandemics.
-
-\newpage
-
-\begin{center}Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic\end{center}
-
-Social norms are a key aspect of human sociality (Bicchieri \& Xiao, 2009; Cialdini et al., 1991; Fehr \& Schurtenberger, 2018).
-Broadly, social norms are defined as commonly known behavioral guidelines enforced by groups of people (Legros \& Cislaghi, 2020).
-By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as resource scarcity, natural disasters, and infectious diseases (Roos et al., 2015).
-Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans (Henrich, 2015).
-
-Previous research has distinguished between two types of social norms: injunctive norms and descriptive norms (Bicchieri \& Xiao, 2009; Cialdini et al., 1991; Deutsch \& Gerard, 1955).
+Social norms are a key aspect of human sociality\textsuperscript{1--3}.
+Broadly, social norms are defined as commonly known behavioral guidelines enforced by groups of people\textsuperscript{4}.
+By coordinating the behavior of many individuals, social norms enable human groups to cooperate in the face of group-wide challenges and threats, such as resource scarcity, natural disasters, and infectious diseases\textsuperscript{5}.
+Social norms are thus hypothesized to have played a key role in the evolution of large-scale cooperation in humans\textsuperscript{6}.
+
+Previous research has distinguished between two types of social norms: injunctive norms and descriptive norms\textsuperscript{1,2,7}.
 Injunctive norms indicate what others tend to approve or disapprove of and often involve social sanctions if violated.
 By contrast, descriptive norms simply describe what most people are doing in a given situation, but carry no prescriptive information \emph{per se}.
-According to the focus theory of normative conduct (Cialdini et al., 1991), these two kinds of social norms often align, but they can also be in conflict with one another and differentially affect behavior depending on which norm is more salient.
+According to the focus theory of normative conduct\textsuperscript{2}, these two kinds of social norms often align, but they can also be in conflict with one another and differentially affect behavior depending on which norm is more salient.
 For example, there may be an injunctive norm that cleaning up litter at a picnic site is the right thing to do: one \emph{ought} to behave this way.
@@
 
-Despite decades of research on injunctive and descriptive norms (Cialdini et al., 1991; Cialdini \& Jacobson, 2021; Schultz et al., 2007), open questions remain regarding the emergence and causal influence of social norms (Legros \& Cislaghi, 2020; van Kleef et al., 2019).
+Despite decades of research on injunctive and descriptive norms\textsuperscript{2,8,9}, open questions remain regarding the emergence and causal influence of social norms\textsuperscript{4,10}.
 First, how do injunctive and descriptive norms emerge over time within a population?
@@
 Research has investigated how social norms emerge in a population over time.
-In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring populations (Henrich, 2015).
-However, less is known about how social norms arise endogenously within populations in the short term.
-Recent work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions (Titlestad et al., 2019).
+In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring populations\textsuperscript{6}.
+However, less is known about how social norms arise endogenously within populations in the short term, including how descriptive norms arise and interact with injunctive norms to shape behavior.
+Recent work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions\textsuperscript{11}.
 But it remains unclear whether these findings generalize beyond the laboratory to real human populations.
 
-With regards to normative influences on behavior, studies have demonstrated positive effects of descriptive norms on a variety of cooperative behaviors, including recycling (Nigbur et al., 2010), paying taxes (Larkin et al., 2018), and sustainably reusing towels in hotels (Goldstein et al., 2008).
-However, these studies have two main limitations.
+With regards to normative influences on behavior, studies have demonstrated positive effects of descriptive norms on a variety of cooperative behaviors, including recycling\textsuperscript{12}, paying taxes\textsuperscript{13}, and sustainably reusing towels in hotels\textsuperscript{14}.
+However, these studies have two key aspects that limit their ability to assess the causal impact of norms, both of which we address in our current work.
 First, studies have not accounted for other potential non-social explanations for behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
-These non-social beliefs, labeled ``factual beliefs'' and ``personal normative beliefs'' (Bicchieri et al., 2014), often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
-For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common behavior).
+These non-social beliefs, labeled ``factual beliefs'' and ``personal normative beliefs''\textsuperscript{15}, often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
+For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common or socially approved behavior).
 It is thus important to control for non-social beliefs in studies of social norms.
-Second, studies have tended to follow cross-sectional experimental designs.
-However, social norms are not static: they change dynamically over time through processes of deliberation and interaction (Smith et al., 2015).
+Second, studies have tended to follow cross-sectional experimental designs in which social norm perceptions are manipulated by the researchers.
+However, social norms are not static: they change dynamically over time through processes of deliberation and interaction\textsuperscript{16}.
 An alternative but underutilized method of assessing causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
@@
 To understand how social norms emerge over time and shape cooperative behavior in a non-experimental setting, we focus on mask wearing in the United States during the COVID-19 pandemic.
-In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease (Centers for Disease Control and Prevention, 2022).
-Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread (Centers for Disease Control and Prevention, 2020).
+In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease\textsuperscript{17}.
+Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread\textsuperscript{18}.
 Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the community (e.g., ``flattening the curve'' to protect at-risk individuals).
@@
 Recent research has found positive relationships between social norms and protective COVID-19 behaviors.
-In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure (Macy et al., 2021), and another study found that experimentally-induced descriptive norms increased mask wearing intentions (Bokemper et al., 2021).
-In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing (Rudert \& Janke, 2021).
+In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure\textsuperscript{19}, and another study found that experimentally-induced descriptive norms increased mask wearing intentions\textsuperscript{20}.
+In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing\textsuperscript{21}.
 These studies are telling, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
-Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States (Baxter-King et al., 2022).
+Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States\textsuperscript{22}.
 
@@
 
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{ethical-approval}{%
-\subsection{Ethical approval}\label{ethical-approval}}
-
-This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
-
-\hypertarget{participants-and-sampling}{%
-\subsection{Participants and sampling}\label{participants-and-sampling}}
-
-Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution).
-From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
-This resulted in 18 unique time points of data collection during the pandemic.
-The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
-Of the initial 915 participants, 634 returned to complete the survey at Time 2, while 347 participants continued through to Time 18 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
-
-\hypertarget{measures}{%
-\subsection{Measures}\label{measures}}
-
-\hypertarget{self-reported-mask-wearing}{%
-\subsubsection{Self-reported mask wearing}\label{self-reported-mask-wearing}}
-
-At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
-
-\hypertarget{perceived-descriptive-and-injunctive-social-norms}{%
-\subsubsection{Perceived descriptive and injunctive social norms}\label{perceived-descriptive-and-injunctive-social-norms}}
-
-In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
-
-Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
-We measured perceived descriptive social norms as the mean of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
-
-Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
-We measured perceived injunctive social norms as the mean of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
-
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
-We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
-Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).
-
-\hypertarget{additional-control-variables}{%
-\subsubsection{Additional control variables}\label{additional-control-variables}}
-
-To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
-In this causal model, we included two kinds of non-social beliefs highlighted by previous research (Bicchieri et al., 2014): factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
-These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
-In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic (Baxter-King et al., 2022).
-Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
-
-Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
-Factual beliefs were measured as the mean of the following two items: ``\emph{I wear a face mask when going out in public to keep myself from getting sick}'' and ``\emph{I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet}''.
-Personal normative beliefs were measured with a single item: ``\emph{Wearing a face mask when going out in public is the right thing to do}''.
-These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
-
-Political orientation was measured in the first time point only.
-We measured political orientation as the mean of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
-These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
-
-\hypertarget{statistical-analysis}{%
-\subsection{Statistical analysis}\label{statistical-analysis}}
-
-To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
-First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
-Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
-These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these multilevel regression models using the \emph{lme4} R package (Bates et al., 2015) and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data (Hamaker et al., 2015).
-This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time \emph{t-1} predicts being above average in another variable at time \emph{t}.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets (Granger, 1969; Hamaker et al., 2015).
-
-We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package (Rosseel, 2012).
-In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
-For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
-We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
-We restricted the analysis to the ten time points with available data for all five variables.
-Full information maximum likelihood estimation was used to deal with missing data.
-
-All analyses were conducted in R v4.1.1 (R Core Team, 2022).
-Visualizations were generated using the \emph{cowplot} (Wilke, 2020) and \emph{ggplot2} (Wickham, 2016) packages.
-The manuscript was reproducibly generated using the \emph{targets} (Landau, 2021) and \emph{papaja} (Aust \& Barth, 2022) packages.
-All code and data are publicly available on GitHub (Heiman et al., 2023).
-
 \hypertarget{results}{%
@@
 Figure \ref{fig:plotTimeline} plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum's COVID-19 Timeline (Centers for Disease Control and Prevention, 2022).
+These events were obtained from the CDC Museum's COVID-19 Timeline\textsuperscript{17}.
 
@@
 On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
-According to recent effect size guidelines for cross-lagged panel models (Orth et al., 2022), the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
+According to recent effect size guidelines for cross-lagged panel models\textsuperscript{23}, the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
 In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
@@
 
-Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction (Titlestad et al., 2019).
-The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed (Markus \& Kitayama, 2010).
-Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms (Amato et al., 2018).
+Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction\textsuperscript{11}.
+The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed\textsuperscript{24}.
+Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms\textsuperscript{25}.
 While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our trends have shown, can unfold over measurement intervals as short as four to six weeks.
@@
 We found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing, independent of the effects of non-social beliefs and political orientation.
-This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors (Bokemper et al., 2021; Rudert \& Janke, 2021).
+This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors\textsuperscript{20,21}.
 There are several explanations for why descriptive norms have had these positive effects on protective COVID-19 behaviors like mask wearing.
 First, people may have followed descriptive norms to quickly coordinate their behavior with others during the pandemic.
-Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic (Gelfand \& Harrington, 2015).
-Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population (Chaudhuri, 2011).
+Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic\textsuperscript{26}.
+Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population\textsuperscript{27}.
 Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
-Third, the increased frequency of mask wearing in the population might have created a bandwagon effect (Schmitt-Beck, 2015), encouraging conformist copying.
+Third, the increased frequency of mask wearing in the population might have created a bandwagon effect\textsuperscript{28}, encouraging conformist copying.
 Under this view, people wear masks not to coordinate or cooperate, but simply to fit in with the crowd.
@@
 One possible explanation for this result is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were more salient than injunctive norms during the pandemic.
-According to focus theory (Cialdini et al., 1991), this difference in salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
+According to focus theory\textsuperscript{2}, this difference in salience would produce behavior in line with descriptive norms and potentially suppress the effects of injunctive norms.
 By contrast, for more private behaviors like remaining indoors, it would have been less possible to observe other people's behaviors, increasing the relative salience of injunctive norms.
@@
 We are limited in generalizing these findings due to the constraints of our sample and the variables considered.
-While our sample began as representative of the United States, there was significant attrition over the course of the study (Supplementary Figure \ref{fig:plotAttrition}).
-This attrition did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
+While the attrition in our study did not substantially affect the representativeness of our sample (Supplementary Figure \ref{fig:plotAttritionBreakdown}),
+the reduction in our sample size did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
 Our results also might not generalize to all social norms, behaviors, and social dilemmas.
@@
 Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
+Importantly, this key finding slices two ways.
+Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that \emph{low} local levels of mask wearing \emph{discouraged} future personal mask use.
+This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community\textsuperscript{29}.
 Our work thus underscores the importance of consistent, visible community adherence for encouraging protective behaviors in response to global pandemics like COVID-19.
+
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
+
+\hypertarget{ethical-approval}{%
+\subsection{Ethical approval}\label{ethical-approval}}
+
+This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
+
+\hypertarget{participants-and-sampling}{%
+\subsection{Participants and sampling}\label{participants-and-sampling}}
+
+Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution).
+From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
+This resulted in 18 unique time points of data collection during the pandemic.
+The first 12 time points were distributed monthly, while the remaining six time points were distributed every two months.
+Of the initial 915 participants, 634 returned to complete the survey at Time 2, while 347 participants continued through to Time 18 (see Supplementary Figure \ref{fig:plotAttrition} for attrition rates across all time points).
+However, this attrition did not substantially affect the demographic makeup of the sample through time (Supplementary Figure \ref{fig:plotAttritionBreakdown}).
+
+\hypertarget{measures}{%
+\subsection{Measures}\label{measures}}
+
+\hypertarget{self-reported-mask-wearing}{%
+\subsubsection{Self-reported mask wearing}\label{self-reported-mask-wearing}}
+
+At every time point, participants were asked about the number of in-person interactions they had in the last week.
+Following this question, participants self-reported their mask wearing by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+
+\hypertarget{perceived-descriptive-and-injunctive-social-norms}{%
+\subsubsection{Perceived descriptive and injunctive social norms}\label{perceived-descriptive-and-injunctive-social-norms}}
+
+In 11 of the 18 time points (Time 2, 3, 5, 9, 11, 13, 14, 15, 16, 17, and 18), we asked questions about perceived descriptive and injunctive mask wearing norms.
+
+Descriptive social norms were operationalized as the proportion of individuals in participants' local areas wearing masks in routine and recreational settings.
+We measured perceived descriptive social norms as the mean of the following two items: ``\emph{What proportion of people in your area wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work)?}'' and ``\emph{What proportion of people in your area wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party)?}'' These perceived descriptive social norm items were measured on 7-point Likert scales, from None (1) to All (7).
+
+Injunctive social norms were operationalized as respected individuals wearing masks and community encouragement of mask wearing rules to emphasize the perceived social approval of the behavior from group leaders and the community at large.
+We measured perceived injunctive social norms as the mean of the following two items: ``\emph{In general, how often do you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.)?}'' and ``\emph{How much are mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.)?}'' These perceived injunctive social norm items were measured on 7-point Likert scales, from Never/Rarely (1) to Very Often (7) for the first item, and from Strongly Discouraged (1) to Strongly Encouraged (7) for the second item.
+
+To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
+We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
+Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).
+
+\hypertarget{additional-control-variables}{%
+\subsubsection{Additional control variables}\label{additional-control-variables}}
+
+To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research\textsuperscript{15}: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
+These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
+In addition, we also included political orientation as a common cause of all other variables.
+This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic\textsuperscript{22}.
+Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
+
+Non-social beliefs were measured in 12 of the 18 time points (Time 2, 4, 5, 7, 9, 11, 13, 14, 15, 16, 17, and 18).
+Factual beliefs were measured as the mean of the following two items: ``\emph{I wear a face mask when going out in public to keep myself from getting sick}'' and ``\emph{I wear a face mask when going out in public to prevent others from getting sick in case I may be infected but don't know it yet}''.
+Personal normative beliefs were measured with a single item: ``\emph{Wearing a face mask when going out in public is the right thing to do}''.
+These non-social belief items were measured on 7-point Likert scales, from Strongly Disagree (1) to Strongly Agree (7).
+
+Political orientation was measured in the first time point only.
+We measured political orientation as the mean of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
+These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
+
+\hypertarget{statistical-analysis}{%
+\subsection{Statistical analysis}\label{statistical-analysis}}
+
+To analyze average trends in self-reported mask wearing and perceived social norms, we fitted several multilevel regression models.
+First, to determine whether mask wearing and social norms were coupled over time, we regressed mask wearing on perceived descriptive and injunctive norms separately, including random intercepts and slopes for participants and time points.
+Second, to analyze whether changes over time were related to recommendations from the CDC, we regressed mask wearing and perceived social norms onto a continuous time predictor.
+These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
+We estimated these multilevel regression models using the \emph{lme4} R package\textsuperscript{30} and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data\textsuperscript{31}.
+This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time \emph{t-1} predicts being above average in another variable at time \emph{t}.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets\textsuperscript{31,32}.
+
+We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package\textsuperscript{33}.
+In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
+For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
+We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We restricted the analysis to the ten time points with available data for all five variables.
+Full information maximum likelihood estimation was used to deal with missing data.
+
+All analyses were conducted in R v4.1.1\textsuperscript{34}.
+Visualizations were generated using the \emph{cowplot}\textsuperscript{35} and \emph{ggplot2}\textsuperscript{36} packages.
+The manuscript was reproducibly generated using the \emph{targets}\textsuperscript{37} and \emph{papaja}\textsuperscript{38} packages.
+All code and data are publicly available on GitHub\textsuperscript{39}.
 
@@
 
-{[}REDACTED{]}
+This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
+We would also like to thank the Language, Culture, and Cognition lab at the University of Auckland for providing feedback on a previous version of this manuscript.
 
@@
 
-{[}REDACTED{]}
+SLH, JDA, DGB, AV, ERH, AA, and PMT conceptualized the study.
+SLH, SC, JDA, DGB, and AV oversaw the data curation, investigation, and methodology of the study.
+SLH and SC wrote the first draft of the paper.
+SC conducted the formal analysis and created all visualizations.
+ERH, AA, and PMT provided funding and supervision for the study.
+All authors reviewed and edited the final draft of the paper.
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
-Amato, R., Lacasa, L., Díaz-Guilera, A., \& Baronchelli, A.
-(2018).
-The dynamics of norm change in the cultural evolution of language.
-\emph{Proceedings of the National Academy of Sciences}, \emph{115}(33), 8260--8265.
-\url{https://doi.org/10.1073/pnas.1721059115}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-Aust, F., \& Barth, M. (2022). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. \url{https://github.com/crsh/papaja}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
-Bates, D., Mächler, M., Bolker, B., \& Walker, S.
-(2015).
-Fitting linear mixed-effects models using lme4.
-\emph{Journal of Statistical Software}, \emph{67}(1), 1--48.
-\url{https://doi.org/10.18637/jss.v067.i01}
-
-\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
-Baxter-King, R., Brown, J.
-R., Enos, R.
-D., Naeim, A., \& Vavreck, L.
-(2022).
-How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}.
-\emph{Proceedings of the National Academy of Sciences}, \emph{119}(21), e2116311119.
-\url{https://doi.org/10.1073/pnas.2116311119}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
-Bicchieri, C., Lindemans, J.
-W., \& Jiang, T.
-(2014).
-A structured approach to a diagnostic of collective practices.
-\emph{Frontiers in Psychology}, \emph{5}, 1418.
-\url{https://doi.org/10.3389/fpsyg.2014.01418}
-
+\begin{CSLReferences}{0}{0}
 \leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2009}{}}%
-Bicchieri, C., \& Xiao, E. (2009). Do the right thing: But only if others do so. \emph{Journal of Behavioral Decision Making}, \emph{22}(2), 191--208. \url{https://doi.org/10.1002/bdm.621}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
-Bokemper, S.
-E., Cucciniello, M., Rotesi, T., Pin, P., Malik, A.
-A., Willebrand, K., Paintsil, E.
-E., Omer, S.
-B., Huber, G.
-A., \& Melegaro, A.
-(2021).
-Experimental evidence that changing beliefs about mask efficacy and social norms increase mask wearing for {COVID-19} risk reduction: Results from the {United States} and {Italy}.
-\emph{PLOS ONE}, \emph{16}(10), e0258282.
-\url{https://doi.org/10.1371/journal.pone.0258282}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CDC2020}{}}%
-Centers for Disease Control and Prevention.
-(2020).
-\emph{{CDC calls on Americans to wear masks to prevent COVID-19 spread}}.
-\url{https://www.cdc.gov/media/releases/2020/p0714-americans-to-wear-masks.html}
-
-\leavevmode\vadjust pre{\hypertarget{ref-CDC2022}{}}%
-Centers for Disease Control and Prevention. (2022). \emph{{CDC Museum COVID-19 Timeline}}. \url{https://www.cdc.gov/museum/timeline/covid19.html}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
-Chaudhuri, A.
-(2011).
-Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature.
-\emph{Experimental Economics}, \emph{14}(1), 47--83.
-\url{https://doi.org/10.1007/s10683-010-9257-1}
+\CSLLeftMargin{1. }%
+\CSLRightInline{Bicchieri, C.
+\& Xiao, E.
+\href{https://doi.org/10.1002/bdm.621}{Do the right thing: But only if others do so}.
+\emph{Journal of Behavioral Decision Making} \textbf{22}, 191--208 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Cialdini1991}{}}%
+\CSLLeftMargin{2. }%
+\CSLRightInline{Cialdini, R.
+B., Kallgren, C.
+A.
+\& Reno, R.
+R.
+\href{https://doi.org/10.1016/S0065-2601(08)60330-5}{A focus theory of normative conduct: A theoretical refinement and reevaluation of the role of norms in human behavior}.
+in (ed.
+Zanna, M.
+P.) vol.
+24 201--234 (Academic Press, 1991).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018a}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{Fehr, E. \& Schurtenberger, I. \href{https://doi.org/10.1038/s41562-018-0385-5}{Normative foundations of human cooperation}. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Legros2020}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{Legros, S.
+\& Cislaghi, B.
+\href{https://doi.org/10.1177/1745691619866455}{Mapping the social-norms literature: An overview of reviews}.
+\emph{Perspectives on Psychological Science} \textbf{15}, 62--80 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Roos2015}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{Roos, P., Gelfand, M., Nau, D.
+\& Lun, J.
+\href{https://doi.org/10.1016/j.obhdp.2015.01.003}{Societal threat and cultural variation in the strength of social norms: An evolutionary basis}.
+\emph{Organizational Behavior and Human Decision Processes} \textbf{129}, 14--23 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
+\CSLLeftMargin{6. }%
+\CSLRightInline{Henrich, J.
+\emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}.
+(Princeton University Press, 2015).
+doi:\href{https://doi.org/10.2307/j.ctvc77f0d}{10.2307/j.ctvc77f0d}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
+\CSLLeftMargin{7. }%
+\CSLRightInline{Deutsch, M.
+\& Gerard, H.
+B.
+\href{https://doi.org/10.1037/h0046408}{A study of normative and informational social influences upon individual judgment}.
+\emph{The Journal of Abnormal and Social Psychology} \textbf{51}, 629--636 (1955).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Cialdini2021}{}}%
-Cialdini, R.
-B., \& Jacobson, R.
+\CSLLeftMargin{8. }%
+\CSLRightInline{Cialdini, R.
+B.
+\& Jacobson, R.
 P.
-(2021).
-Influences of social norms on climate change-related behaviors.
-\emph{Current Opinion in Behavioral Sciences}, \emph{42}, 1--8.
-\url{https://doi.org/10.1016/j.cobeha.2021.01.005}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Cialdini1991}{}}%
-Cialdini, R.
-B., Kallgren, C.
-A., \& Reno, R.
-R.
-(1991).
-\emph{A focus theory of normative conduct: A theoretical refinement and reevaluation of the role of norms in human behavior} (M.
-P.
-Zanna, Ed.; Vol.
-24, pp.
-201--234).
-Academic Press.
-\url{https://doi.org/10.1016/S0065-2601(08)60330-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Deutsch1955}{}}%
-Deutsch, M., \& Gerard, H.
-B.
-(1955).
-A study of normative and informational social influences upon individual judgment.
-\emph{The Journal of Abnormal and Social Psychology}, \emph{51}(3), 629--636.
-\url{https://doi.org/10.1037/h0046408}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018a}{}}%
-Fehr, E., \& Schurtenberger, I. (2018). Normative foundations of human cooperation. \emph{Nature Human Behaviour}, \emph{2}(7), 458--468. \url{https://doi.org/10.1038/s41562-018-0385-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
-Gelfand, M.
-J., \& Harrington, J.
-R.
-(2015).
-The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?
-\emph{Journal of Cross-Cultural Psychology}, \emph{46}(10), 1273--1278.
-\url{https://doi.org/10.1177/0022022115600796}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
-Goldstein, N.
-J., Cialdini, R.
-B., \& Griskevicius, V.
-(2008).
-A room with a viewpoint: Using social norms to motivate environmental conservation in hotels.
-\emph{Journal of Consumer Research}, \emph{35}(3), 472--482.
-\url{https://doi.org/10.1086/586910}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
-Granger, C. W. J. (1969). Investigating causal relations by econometric models and cross-spectral methods. \emph{Econometrica}, \emph{37}(3), 424--438. \url{https://doi.org/10.2307/1912791}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
-Hamaker, E. L., Kuiper, R. M., \& Grasman, R. P. P. P. (2015). A critique of the cross-lagged panel model. \emph{Psychological Methods}, \emph{20}(1), 102--116. \url{https://doi.org/10.1037/a0038889}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Heiman2023}{}}%
-Heiman, S.
-L., Claessens, S., Ayers, J.
-D., Beltrán, D.
-G., Van Horn, A., Hirt, E.
-R., Aktipis, A., \& Todd, P.
-M.
-(2023).
-\emph{Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic}.
-GitHub repository.
-\url{https://github.com/ScottClaessens/covidMaskWearing}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Henrich2015}{}}%
-Henrich, J.
-(2015).
-\emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}.
-Princeton University Press.
-\url{https://doi.org/10.2307/j.ctvc77f0d}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-Landau, W.
-M.
-(2021).
-The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing.
-\emph{Journal of Open Source Software}, \emph{6}(57), 2959.
-\url{https://doi.org/10.21105/joss.02959}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
-Larkin, C., Sanders, M., Andresen, I., \& Algate, F.
-(2018).
-\emph{Testing local descriptive norms and salience of enforcement action: A field experiment to increase tax collection} (SSRN Scholarly Paper No.
-3167575).
-\url{https://doi.org/10.2139/ssrn.3167575}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Legros2020}{}}%
-Legros, S., \& Cislaghi, B.
-(2020).
-Mapping the social-norms literature: An overview of reviews.
-\emph{Perspectives on Psychological Science}, \emph{15}(1), 62--80.
-\url{https://doi.org/10.1177/1745691619866455}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
-Macy, J.
-T., Owens, C., Mullis, K., \& Middlestadt, S.
-E.
-(2021).
-The role of self-efficacy and injunctive norms in helping older adults decide to stay home during the {COVID-19} pandemic.
-\emph{Frontiers in Public Health}, \emph{9}, 660813.
-\url{https://doi.org/10.3389/fpubh.2021.660813}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
-Markus, H.
-R., \& Kitayama, S.
-(2010).
-Cultures and selves: A cycle of mutual constitution.
-\emph{Perspectives on Psychological Science}, \emph{5}(4), 420--430.
-\url{https://doi.org/10.1177/1745691610375557}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
-Nigbur, D., Lyons, E., \& Uzzell, D.
-(2010).
-Attitudes, norms, identity and environmental behaviour: Using an expanded theory of planned behaviour to predict participation in a kerbside recycling programme.
-\emph{British Journal of Social Psychology}, \emph{49}(2), 259--284.
-\url{https://doi.org/10.1348/014466609X449395}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
-Orth, U., Meier, L.
-L., Bühler, J.
-L., Dapp, L.
-C., Krauss, S., Messerli, D., \& Robins, R.
-W.
-(2022).
-Effect size guidelines for cross-lagged effects.
-\emph{Psychological Methods}.
-\url{https://doi.org/10.1037/met0000499}
-
-\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-R Core Team. (2022). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Roos2015}{}}%
-Roos, P., Gelfand, M., Nau, D., \& Lun, J.
-(2015).
-Societal threat and cultural variation in the strength of social norms: An evolutionary basis.
-\emph{Organizational Behavior and Human Decision Processes}, \emph{129}, 14--23.
-\url{https://doi.org/10.1016/j.obhdp.2015.01.003}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
-Rosseel, Y. (2012). {lavaan}: An {R} package for structural equation modeling. \emph{Journal of Statistical Software}, \emph{48}(2), 1--36. \url{https://doi.org/10.18637/jss.v048.i02}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
-Rudert, S.
-C., \& Janke, S.
-(2021).
-Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic.
-\emph{Group Processes \& Intergroup Relations}, \emph{25}, 1819--1835.
-\url{https://doi.org/10.1177/13684302211023562}
-
-\leavevmode\vadjust pre{\hypertarget{ref-SchmittBeck2015}{}}%
-Schmitt-Beck, R.
-(2015).
-Bandwagon effect.
-In \emph{The international encyclopedia of political communication} (pp.
-1--5).
-John Wiley \& Sons, Ltd.
-\url{https://doi.org/10.1002/9781118541555.wbiepc015}
+\href{https://doi.org/10.1016/j.cobeha.2021.01.005}{Influences of social norms on climate change-related behaviors}.
+\emph{Current Opinion in Behavioral Sciences} \textbf{42}, 1--8 (2021).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Schultz2007}{}}%
-Schultz, P.
+\CSLLeftMargin{9. }%
+\CSLRightInline{Schultz, P.
 W., Nolan, J.
@@
 B., Goldstein, N.
-J., \& Griskevicius, V.
-(2007).
-The constructive, destructive, and reconstructive power of social norms.
-\emph{Psychological Science}, \emph{18}(5), 429--434.
-\url{https://doi.org/10.1111/j.1467-9280.2007.01917.x}
+J.
+\& Griskevicius, V.
+\href{https://doi.org/10.1111/j.1467-9280.2007.01917.x}{The constructive, destructive, and reconstructive power of social norms}.
+\emph{Psychological Science} \textbf{18}, 429--434 (2007).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
+\CSLLeftMargin{10. }%
+\CSLRightInline{van Kleef, G.
+A., Gelfand, M.
+J.
+\& Jetten, J.
+\href{https://doi.org/10.1016/j.jesp.2019.05.002}{The dynamic nature of social norms: New perspectives on norm development, impact, violation, and enforcement}.
+\emph{Journal of Experimental Social Psychology} \textbf{84}, 103814 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
+\CSLLeftMargin{11. }%
+\CSLRightInline{Titlestad, K., Snijders, T.
+A.
+B., Durrheim, K., Quayle, M.
+\& Postmes, T.
+\href{https://doi.org/10.1016/j.jesp.2019.03.010}{The dynamic emergence of cooperative norms in a social dilemma}.
+\emph{Journal of Experimental Social Psychology} \textbf{84}, 103799 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{Nigbur, D., Lyons, E.
+\& Uzzell, D.
+\href{https://doi.org/10.1348/014466609X449395}{Attitudes, norms, identity and environmental behaviour: Using an expanded theory of planned behaviour to predict participation in a kerbside recycling programme}.
+\emph{British Journal of Social Psychology} \textbf{49}, 259--284 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
+\CSLLeftMargin{13. }%
+\CSLRightInline{Larkin, C., Sanders, M., Andresen, I.
+\& Algate, F.
+\emph{Testing local descriptive norms and salience of enforcement action: A field experiment to increase tax collection}.
+(2018) doi:\href{https://doi.org/10.2139/ssrn.3167575}{10.2139/ssrn.3167575}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
+\CSLLeftMargin{14. }%
+\CSLRightInline{Goldstein, N.
+J., Cialdini, R.
+B.
+\& Griskevicius, V.
+\href{https://doi.org/10.1086/586910}{A room with a viewpoint: Using social norms to motivate environmental conservation in hotels}.
+\emph{Journal of Consumer Research} \textbf{35}, 472--482 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{Bicchieri, C., Lindemans, J.
+W.
+\& Jiang, T.
+\href{https://doi.org/10.3389/fpsyg.2014.01418}{A structured approach to a diagnostic of collective practices}.
+\emph{Frontiers in Psychology} \textbf{5}, 1418 (2014).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
-Smith, L.
+\CSLLeftMargin{16. }%
+\CSLRightInline{Smith, L.
 G.
 E., Thomas, E.
-F., \& McGarty, C.
-(2015).
-"We must be the change we want to see in the world": Integrating norms and identities through social interaction.
-\emph{Political Psychology}, \emph{36}(5), 543--557.
-\url{https://doi.org/10.1111/pops.12180}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
-Titlestad, K., Snijders, T.
-A.
-B., Durrheim, K., Quayle, M., \& Postmes, T.
-(2019).
-The dynamic emergence of cooperative norms in a social dilemma.
-\emph{Journal of Experimental Social Psychology}, \emph{84}, 103799.
-\url{https://doi.org/10.1016/j.jesp.2019.03.010}
-
-\leavevmode\vadjust pre{\hypertarget{ref-VanKleef2019}{}}%
-van Kleef, G.
-A., Gelfand, M.
-J., \& Jetten, J.
-(2019).
-The dynamic nature of social norms: New perspectives on norm development, impact, violation, and enforcement.
-\emph{Journal of Experimental Social Psychology}, \emph{84}, 103814.
-\url{https://doi.org/10.1016/j.jesp.2019.05.002}
+F.
+\& McGarty, C.
+\href{https://doi.org/10.1111/pops.12180}{"We must be the change we want to see in the world": Integrating norms and identities through social interaction}.
+\emph{Political Psychology} \textbf{36}, 543--557 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CDC2022}{}}%
+\CSLLeftMargin{17. }%
+\CSLRightInline{Centers for Disease Control and Prevention. \href{https://www.cdc.gov/museum/timeline/covid19.html}{{CDC Museum COVID-19 Timeline}}. (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-CDC2020}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{Centers for Disease Control and Prevention.
+\href{https://www.cdc.gov/media/releases/2020/p0714-americans-to-wear-masks.html}{{CDC calls on Americans to wear masks to prevent COVID-19 spread}}.
+(2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
+\CSLLeftMargin{19. }%
+\CSLRightInline{Macy, J.
+T., Owens, C., Mullis, K.
+\& Middlestadt, S.
+E.
+\href{https://doi.org/10.3389/fpubh.2021.660813}{The role of self-efficacy and injunctive norms in helping older adults decide to stay home during the {COVID-19} pandemic}.
+\emph{Frontiers in Public Health} \textbf{9}, 660813 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
+\CSLLeftMargin{20. }%
+\CSLRightInline{Bokemper, S.
+E.
+\emph{et al.} \href{https://doi.org/10.1371/journal.pone.0258282}{Experimental evidence that changing beliefs about mask efficacy and social norms increase mask wearing for {COVID-19} risk reduction: Results from the {United States} and {Italy}}.
+\emph{PLOS ONE} \textbf{16}, e0258282 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
+\CSLLeftMargin{21. }%
+\CSLRightInline{Rudert, S.
+C.
+\& Janke, S.
+\href{https://doi.org/10.1177/13684302211023562}{Following the crowd in times of crisis: Descriptive norms predict physical distancing, stockpiling, and prosocial behavior during the {COVID-19} pandemic}.
+\emph{Group Processes \& Intergroup Relations} \textbf{25}, 1819--1835 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
+\CSLLeftMargin{22. }%
+\CSLRightInline{Baxter-King, R., Brown, J.
+R., Enos, R.
+D., Naeim, A.
+\& Vavreck, L.
+\href{https://doi.org/10.1073/pnas.2116311119}{How local partisan context conditions prosocial behaviors: Mask wearing during {COVID-19}}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{119}, e2116311119 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
+\CSLLeftMargin{23. }%
+\CSLRightInline{Orth, U. \emph{et al.} Effect size guidelines for cross-lagged effects. \emph{Psychological Methods} (2022) doi:\href{https://doi.org/10.1037/met0000499}{10.1037/met0000499}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
+\CSLLeftMargin{24. }%
+\CSLRightInline{Markus, H.
+R.
+\& Kitayama, S.
+\href{https://doi.org/10.1177/1745691610375557}{Cultures and selves: A cycle of mutual constitution}.
+\emph{Perspectives on Psychological Science} \textbf{5}, 420--430 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
+\CSLLeftMargin{25. }%
+\CSLRightInline{Amato, R., Lacasa, L., Díaz-Guilera, A.
+\& Baronchelli, A.
+\href{https://doi.org/10.1073/pnas.1721059115}{The dynamics of norm change in the cultural evolution of language}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{115}, 8260--8265 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
+\CSLLeftMargin{26. }%
+\CSLRightInline{Gelfand, M.
+J.
+\& Harrington, J.
+R.
+\href{https://doi.org/10.1177/0022022115600796}{The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?} \emph{Journal of Cross-Cultural Psychology} \textbf{46}, 1273--1278 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
+\CSLLeftMargin{27. }%
+\CSLRightInline{Chaudhuri, A.
+\href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
+\emph{Experimental Economics} \textbf{14}, 47--83 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-SchmittBeck2015}{}}%
+\CSLLeftMargin{28. }%
+\CSLRightInline{Schmitt-Beck, R.
+Bandwagon effect.
+in \emph{The international encyclopedia of political communication} 1--5 (John Wiley \& Sons, Ltd, 2015).
+doi:\href{https://doi.org/10.1002/9781118541555.wbiepc015}{10.1002/9781118541555.wbiepc015}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
+\CSLLeftMargin{29. }%
+\CSLRightInline{Natanson, H.
+\href{https://www.washingtonpost.com/education/2022/02/25/peer-pressure-mask-optional-schools/}{Peer pressure is ending mask usage in schools}.
+\emph{The Washingston Post} (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{Bates, D., Mächler, M., Bolker, B.
+\& Walker, S.
+\href{https://doi.org/10.18637/jss.v067.i01}{Fitting linear mixed-effects models using lme4}.
+\emph{Journal of Statistical Software} \textbf{67}, 1--48 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
+\CSLLeftMargin{31. }%
+\CSLRightInline{Hamaker, E.
+L., Kuiper, R.
+M.
+\& Grasman, R.
+P.
+P.
+P.
+\href{https://doi.org/10.1037/a0038889}{A critique of the cross-lagged panel model}.
+\emph{Psychological Methods} \textbf{20}, 102--116 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
+\CSLLeftMargin{32. }%
+\CSLRightInline{Granger, C.
+W.
+J.
+\href{https://doi.org/10.2307/1912791}{Investigating causal relations by econometric models and cross-spectral methods}.
+\emph{Econometrica} \textbf{37}, 424--438 (1969).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
+\CSLLeftMargin{33. }%
+\CSLRightInline{Rosseel, Y. \href{https://doi.org/10.18637/jss.v048.i02}{{lavaan}: An {R} package for structural equation modeling}. \emph{Journal of Statistical Software} \textbf{48}, 1--36 (2012).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
+\CSLLeftMargin{34. }%
+\CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
+\CSLLeftMargin{35. }%
+\CSLRightInline{Wilke, C. O. \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}}. (2020).}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-Wilke, C. O. (2020). \emph{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}. \url{https://CRAN.R-project.org/package=cowplot}
+\CSLLeftMargin{36. }%
+\CSLRightInline{Wickham, H. \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}}. (Springer-Verlag New York, 2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
+\CSLLeftMargin{37. }%
+\CSLRightInline{Landau, W.
+M.
+\href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
+\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
+\CSLLeftMargin{38. }%
+\CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}}. (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Heiman2023}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{Heiman, S.
+L.
+\emph{et al.} \href{https://github.com/ScottClaessens/covidMaskWearing}{Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic}.
+(2023).}
 
@@
 \caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAttritionBreakdown-1.pdf}
+\caption{\label{fig:plotAttritionBreakdown}\emph{Demographics of the sample across the course of the study.}}
 \end{figure}

### 2023-03-14 R/functions.R

@@ -496,7 +496,7 @@ plotAttrition <- function(d) {
     transmute(time = as.numeric(time), value) %>%
     # convert time to date
     mutate(time = ymd(dates[time])) %>%
-    # calculate mean and standard errors
+    # calculate attrition
     group_by(time) %>%
     summarise(count = sum(!is.na(value)), .groups = "drop") %>%
     # plot
@@ -509,6 +509,89 @@ plotAttrition <- function(d) {
   # save
   ggsave(out, filename = "figures/attrition.pdf", width = 5.5, height = 3)
   ggsave(out, filename = "figures/attrition.png", width = 5.5, height = 3)
+  return(out)
+}
+
+# plot sample breakdown over time
+plotAttritionBreakdown <- function(d) {
+  # vector of dates for plot
+  dates <- c("2020-09-27", "2020-10-27", "2020-11-28",
+             "2020-12-28", "2021-01-27", "2021-02-26",
+             "2021-03-28", "2021-04-27", "2021-05-27",
+             "2021-06-26", "2021-07-26", "2021-08-26",
+             "2021-10-25", "2021-12-16", "2022-02-25",
+             "2022-04-26", "2022-06-25", "2022-08-29")
+  # data with demographic breakdown over time
+  dDem <-
+    d %>%
+    # pivot into long format
+    rename(StartDate.5 = StartDate_A, 
+           StartDate.7 = StartDate...1230,
+           StartDate.16 = StartDate...2919) %>%
+    pivot_longer((starts_with("StartDate"))) %>%
+    separate(name, c("var", "time")) %>%
+    # keep only relevant vars
+    transmute(sex = Sex.1, age = Age.1, ethnicity = Ethnicity.1,
+              time = as.numeric(time), value) %>%
+    # convert time to date, and empty demographic vars if no response from participant
+    mutate(
+      time = ymd(dates[time]),
+      sex = ifelse(is.na(value), NA, sex),
+      age = ifelse(is.na(value), NA, age),
+      ethnicity = ifelse(is.na(value), NA, ethnicity),
+      ) %>%
+    # calculate attrition
+    group_by(time) %>%
+    summarise(
+      # sex
+      propMale = mean(sex == 1, na.rm = TRUE),
+      # age
+      meanAge = mean(age, na.rm = TRUE),
+      # ethnicity
[…]

### 2023-03-14 _targets.R

@@ -51,6 +51,7 @@ list(
   tar_target(plot5, plotCorBehNorm(m1.1, m1.2)),
   tar_target(plot6, plotCDCSens(m2.1, m2.2, m2.3)),
   tar_target(plot7, plotRICLPM(riclpm)),
+  tar_target(plot8, plotAttritionBreakdown(d)),
   # tables
   tar_target(itemTable, makeItemTable()),
   tar_target(changePointsTable, makeChangePointsTable(m2.1, m2.2, m2.3,

### 2023-03-15 manuscript.Rmd
update manuscript

@@
 In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring populations [@Henrich2015].
-However, less is known about how social norms arise endogenously within populations in the short term, including how descriptive norms arise and interact with injunctive norms to shape behavior.
-Recent work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019].
+However, less is known about how social norms arise endogenously within populations in the short term.
+While researchers have simulated the emergence of descriptive norms [@Muldoon2014a;@Muldoon2014b], this modelling work does not capture how descriptive norms develop alongside injunctive norms in real-world settings.
+Recent work in behavioral economics has also suggested that injunctive norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019].
 But it remains unclear whether these findings generalize beyond the laboratory to real human populations.

### 2023-03-15 manuscript.tex
update manuscript

@@
 In the long term, cultural evolutionary models show that injunctive social norms can be vertically transmitted across generations by imitation or teaching, or horizontally diffused from neighboring populations\textsuperscript{6}.
-However, less is known about how social norms arise endogenously within populations in the short term, including how descriptive norms arise and interact with injunctive norms to shape behavior.
-Recent work in behavioral economics suggests that social norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions\textsuperscript{11}.
+However, less is known about how social norms arise endogenously within populations in the short term.
+While researchers have simulated the emergence of descriptive norms\textsuperscript{11,12}, this modelling work does not capture how descriptive norms develop alongside injunctive norms in real-world settings.
+Recent work in behavioral economics has also suggested that injunctive norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions\textsuperscript{13}.
 But it remains unclear whether these findings generalize beyond the laboratory to real human populations.
 
-With regards to normative influences on behavior, studies have demonstrated positive effects of descriptive norms on a variety of cooperative behaviors, including recycling\textsuperscript{12}, paying taxes\textsuperscript{13}, and sustainably reusing towels in hotels\textsuperscript{14}.
+With regards to normative influences on behavior, studies have demonstrated positive effects of descriptive norms on a variety of cooperative behaviors, including recycling\textsuperscript{14}, paying taxes\textsuperscript{15}, and sustainably reusing towels in hotels\textsuperscript{16}.
 However, these studies have two key aspects that limit their ability to assess the causal impact of norms, both of which we address in our current work.
 First, studies have not accounted for other potential non-social explanations for behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
-These non-social beliefs, labeled ``factual beliefs'' and ``personal normative beliefs''\textsuperscript{15}, often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
+These non-social beliefs, labeled ``factual beliefs'' and ``personal normative beliefs''\textsuperscript{17}, often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
 For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common or socially approved behavior).
@@
 Second, studies have tended to follow cross-sectional experimental designs in which social norm perceptions are manipulated by the researchers.
-However, social norms are not static: they change dynamically over time through processes of deliberation and interaction\textsuperscript{16}.
+However, social norms are not static: they change dynamically over time through processes of deliberation and interaction\textsuperscript{18}.
 An alternative but underutilized method of assessing causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
@@
 To understand how social norms emerge over time and shape cooperative behavior in a non-experimental setting, we focus on mask wearing in the United States during the COVID-19 pandemic.
-In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease\textsuperscript{17}.
-Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread\textsuperscript{18}.
+In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease\textsuperscript{19}.
+Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread\textsuperscript{20}.
 Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the community (e.g., ``flattening the curve'' to protect at-risk individuals).
@@
 Recent research has found positive relationships between social norms and protective COVID-19 behaviors.
-In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure\textsuperscript{19}, and another study found that experimentally-induced descriptive norms increased mask wearing intentions\textsuperscript{20}.
-In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing\textsuperscript{21}.
+In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure\textsuperscript{21}, and another study found that experimentally-induced descriptive norms increased mask wearing intentions\textsuperscript{22}.
+In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing\textsuperscript{23}.
 These studies are telling, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
-Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States\textsuperscript{22}.
+Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States\textsuperscript{24}.
 
@@
 Figure \ref{fig:plotTimeline} plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum's COVID-19 Timeline\textsuperscript{17}.
+These events were obtained from the CDC Museum's COVID-19 Timeline\textsuperscript{19}.
 
@@
 On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
-According to recent effect size guidelines for cross-lagged panel models\textsuperscript{23}, the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
+According to recent effect size guidelines for cross-lagged panel models\textsuperscript{25}, the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
 In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
@@
 
-Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction\textsuperscript{11}.
-The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed\textsuperscript{24}.
-Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms\textsuperscript{25}.
+Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction\textsuperscript{13}.
+The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed\textsuperscript{26}.
+Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms\textsuperscript{27}.
 While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our trends have shown, can unfold over measurement intervals as short as four to six weeks.
@@
 We found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing, independent of the effects of non-social beliefs and political orientation.
-This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors\textsuperscript{20,21}.
+This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors\textsuperscript{22,23}.
 There are several explanations for why descriptive norms have had these positive effects on protective COVID-19 behaviors like mask wearing.
 First, people may have followed descriptive norms to quickly coordinate their behavior with others during the pandemic.
-Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic\textsuperscript{26}.
-Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population\textsuperscript{27}.
+Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic\textsuperscript{28}.
+Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population\textsuperscript{29}.
 Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
-Third, the increased frequency of mask wearing in the population might have created a bandwagon effect\textsuperscript{28}, encouraging conformist copying.
+Third, the increased frequency of mask wearing in the population might have created a bandwagon effect\textsuperscript{30}, encouraging conformist copying.
 Under this view, people wear masks not to coordinate or cooperate, but simply to fit in with the crowd.
@@
 Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that \emph{low} local levels of mask wearing \emph{discouraged} future personal mask use.
-This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community\textsuperscript{29}.
+This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community\textsuperscript{31}.
 Our work thus underscores the importance of consistent, visible community adherence for encouraging protective behaviors in response to global pandemics like COVID-19.
@@
 To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
-In this causal model, we included two kinds of non-social beliefs highlighted by previous research\textsuperscript{15}: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research\textsuperscript{17}: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
 These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
 In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic\textsuperscript{22}.
+This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic\textsuperscript{24}.
 Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
@@
 These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these multilevel regression models using the \emph{lme4} R package\textsuperscript{30} and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data\textsuperscript{31}.
+We estimated these multilevel regression models using the \emph{lme4} R package\textsuperscript{32} and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data\textsuperscript{33}.
 This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
 Positive cross-lagged effects from this model indicate that being above average on one variable at time \emph{t-1} predicts being above average in another variable at time \emph{t}.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets\textsuperscript{31,32}.
-
-We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package\textsuperscript{33}.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets\textsuperscript{33,34}.
+
+We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package\textsuperscript{35}.
 In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
@@
 
-All analyses were conducted in R v4.1.1\textsuperscript{34}.
-Visualizations were generated using the \emph{cowplot}\textsuperscript{35} and \emph{ggplot2}\textsuperscript{36} packages.
-The manuscript was reproducibly generated using the \emph{targets}\textsuperscript{37} and \emph{papaja}\textsuperscript{38} packages.
-All code and data are publicly available on GitHub\textsuperscript{39}.
+All analyses were conducted in R v4.1.1\textsuperscript{36}.
+Visualizations were generated using the \emph{cowplot}\textsuperscript{37} and \emph{ggplot2}\textsuperscript{38} packages.
+The manuscript was reproducibly generated using the \emph{targets}\textsuperscript{39} and \emph{papaja}\textsuperscript{40} packages.
+All code and data are publicly available on GitHub\textsuperscript{41}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Muldoon2014a}{}}%
+\CSLLeftMargin{11. }%
+\CSLRightInline{Muldoon, R., Lisciandra, C. \& Hartmann, S. Why are there descriptive norms? Because we looked for them. \emph{Synthese} \textbf{191}, 4409--4429 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Muldoon2014b}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{Muldoon, R., Lisciandra, C., Bicchieri, C., Hartmann, S. \& Sprenger, J. On the emergence of descriptive norms. \emph{Politics, Philosophy \& Economics} \textbf{13}, 3--22 (2014).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Titlestad2019}{}}%
-\CSLLeftMargin{11. }%
+\CSLLeftMargin{13. }%
 \CSLRightInline{Titlestad, K., Snijders, T.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nigbur2010}{}}%
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{14. }%
 \CSLRightInline{Nigbur, D., Lyons, E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Larkin2018}{}}%
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{15. }%
 \CSLRightInline{Larkin, C., Sanders, M., Andresen, I.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goldstein2008}{}}%
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{16. }%
 \CSLRightInline{Goldstein, N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{17. }%
 \CSLRightInline{Bicchieri, C., Lindemans, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
-\CSLLeftMargin{16. }%
+\CSLLeftMargin{18. }%
 \CSLRightInline{Smith, L.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CDC2022}{}}%
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{19. }%
 \CSLRightInline{Centers for Disease Control and Prevention. \href{https://www.cdc.gov/museum/timeline/covid19.html}{{CDC Museum COVID-19 Timeline}}. (2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CDC2020}{}}%
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{Centers for Disease Control and Prevention.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
-\CSLLeftMargin{19. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{Macy, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{Bokemper, S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{Rudert, S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{Baxter-King, R., Brown, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{Orth, U. \emph{et al.} Effect size guidelines for cross-lagged effects. \emph{Psychological Methods} (2022) doi:\href{https://doi.org/10.1037/met0000499}{10.1037/met0000499}.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Markus, H.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{Amato, R., Lacasa, L., Díaz-Guilera, A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{Gelfand, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{Chaudhuri, A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-SchmittBeck2015}{}}%
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{Schmitt-Beck, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{Natanson, H.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{Bates, D., Mächler, M., Bolker, B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{Hamaker, E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Granger, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Rosseel, Y. \href{https://doi.org/10.18637/jss.v048.i02}{{lavaan}: An {R} package for structural equation modeling}. \emph{Journal of Statistical Software} \textbf{48}, 1--36 (2012).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Wilke, C. O. \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}}. (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Wickham, H. \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}}. (Springer-Verlag New York, 2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Landau, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}}. (2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Heiman2023}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Heiman, S.

### 2023-03-27 manuscript.Rmd
update ethics statement and README

@@
 
-This project was granted exemption from the Institutional Review Board of Arizona State University (STUDY00011678). All participants in this study provided informed consent.
+All experimental protocols were approved by the Institutional Review Board of Arizona State University (STUDY00011678).
+All methods were carried out in accordance with relevant guidelines and regulations.
+All participants in this study provided informed consent.
 

### 2023-03-30 manuscript.Rmd
change data availability statement

@@
 
-# Open Practices Statement
+# Data Availability Statement
 
-All data and code to reproduce the statistical analyses in this manuscript are publicly available on GitHub: https://github.com/ScottClaessens/covidMaskWearing This study was not preregistered.
+All data and code to reproduce the statistical analyses in this manuscript are publicly available on GitHub: https://github.com/ScottClaessens/covidMaskWearing
 

### 2023-03-30 manuscript.tex
suppress supps page and line numbering
change data availability statement

@@
 
-\hypertarget{open-practices-statement}{%
-\section{Open Practices Statement}\label{open-practices-statement}}
-
-All data and code to reproduce the statistical analyses in this manuscript are publicly available on GitHub: \url{https://github.com/ScottClaessens/covidMaskWearing} This study was not preregistered.
+\hypertarget{data-availability-statement}{%
+\section{Data Availability Statement}\label{data-availability-statement}}
+
+All data and code to reproduce the statistical analyses in this manuscript are publicly available on GitHub: \url{https://github.com/ScottClaessens/covidMaskWearing}
 
@@
 \renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}
+\pagenumbering{gobble}
+\nolinenumbers
 

### 2023-07-06 appendix.Rmd
changes before resubmission
resubmission
suppress supps page and line numbering
second round of reviews

@@
----
-output:
-  pdf_document: default
-  html_document: default
----
 \renewcommand{\appendixname}{\textbf{Supplementary Material}}
@@
 \renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
-\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}
+\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
+\pagenumbering{gobble}
 
@@
 
-Results showed that participants did differentiate the perceived norm items as expected.
-Participants rated the perceived descriptive norm items as providing more descriptive information than the perceived injunctive norm items, _t_(`r printnum(constructValDes$parameter, digits = 0)`) = `r printnum(constructValDes$statistic)`, _p_ `r printp(constructValDes$p.value)` (mean descriptive items = `r printnum(mean((d$DescriptiveLearning3.7 + d$DescriptiveLearning4.7) / 2, na.rm = T))`; mean injunctive items = `r printnum(mean((d$DescriptiveLearning1.7 + d$DescriptiveLearning2.7) / 2, na.rm = T))`).
-By contrast, participants rated the perceived injunctive norm items as providing more injunctive information than the perceived descriptive norm items, _t_(`r printnum(constructValInj$parameter, digits = 0)`) = `r printnum(constructValInj$statistic)`, _p_ `r printp(constructValInj$p.value)` (mean descriptive items = `r printnum(mean((d$InjunctiveLearning3.7 + d$InjunctiveLearning4.7) / 2, na.rm = T))`; mean injunctive items = `r printnum(mean((d$InjunctiveLearning1.7 + d$InjunctiveLearning2.7) / 2, na.rm = T))`).
+We fitted a multilevel model to determine whether participants rated the descriptive norm items as providing more descriptive information than the injunctive norm items, and vice versa.
+Pairwise comparisons from this model revealed that participants did differentiate the perceived norm items as expected (see Supplementary Table \@ref(tab:constructValTable) for pairwise comparisons).
+Both descriptive norm items were seen as providing more descriptive information than both injunctive norm items.
+Conversely, both injunctive norm items were seen as providing more injunctive information than both descriptive norm items.
+
+### Between-person results from random-intercept cross-lagged panel model
+
+The random-intercept cross-lagged panel model partitions the data into two components: (1) between-person trait-like individual differences that persist over time, and (2) within-person fluctuations from trait levels at individual time points.
+In the main text, we focused on the results from the within-person component of the model, specifically the autoregressive and cross-lagged parameters, as we were interested in within-person causal effects.
+However, the model also estimates the covariances between the random intercepts in the model, which can be interpreted as the correlations between between-person trait levels for each variable (i.e., whether people who are generally high on one variable tend to be generally high on another variable).
+
+```{r echo = FALSE}
+# standardized estimates from constrained model
+sc_con1 <- standardizedSolution(riclpm1_con)
+```
+
+The covariances between the random intercepts in the full time-invariant model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
+On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(sc_con1[sc_con1$lhs == "ri1" & sc_con1$op == "~~" & sc_con1$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc_con1[sc_con1$lhs == "ri1" & sc_con1$op == "~~" & sc_con1$rhs == "ri3","ci.lower"])` `r printnum(sc_con1[sc_con1$lhs == "ri1" & sc_con1$op == "~~" & sc_con1$rhs == "ri3","ci.upper"])`]) and stronger injunctive mask wearing norms (_r_ = `r printnum(sc_con1[sc_con1$lhs == "ri1" & sc_con1$op == "~~" & sc_con1$rhs == "ri2","est.std"])`, 95% CI [`r printnum(sc_con1[sc_con1$lhs == "ri1" & sc_con1$op == "~~" & sc_con1$rhs == "ri2","ci.lower"])` `r printnum(sc_con1[sc_con1$lhs == "ri1" & sc_con1$op == "~~" & sc_con1$rhs == "ri2","ci.upper"])`]).
+Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (_r_ = `r printnum(sc_con1[sc_con1$lhs == "ri2" & sc_con1$op == "~~" & sc_con1$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc_con1[sc_con1$lhs == "ri2" & sc_con1$op == "~~" & sc_con1$rhs == "ri3","ci.lower"])` `r printnum(sc_con1[sc_con1$lhs == "ri2" & sc_con1$op == "~~" & sc_con1$rhs == "ri3","ci.upper"])`]).
+This general pattern of correlations at the between-person level was unchanged when removing all covariates and when unconstraining the within-person effects over time.
+
+### Multi-group random-intercept cross-lagged panel model
+
+To test whether the causal effects of descriptive and injunctive norms varied across regions of the United States with different mask wearing policies, we fitted a multi-group random-intercept cross-lagged panel model.
+The model was fitted separately to two groups: majority-Democrat states and majority-Republican states.
+Participants were linked to US states via self-reported zip codes.
+We operationalised state-level partisanship using the state-level results for the 2020 Presidential election.
+For ease of interpretation, we constrained the within-person autoregressive effects, cross-lagged effects, covariances, and variances to equality over time.
+This multi-group model fitted the data well according to one fit statistic (RMSEA = `r printnum(fitMeasures4_con['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures4_con['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures4_con['rmsea.ci.upper'], digits = 3)`]) but not according to others (SRMR = `r printnum(fitMeasures4_con['srmr'], digits = 3)`; CFI = `r printnum(fitMeasures4_con['cfi'], digits = 3)`).
+
+```{r echo=FALSE}
+# parameter estimates
+pe_con4 <- parameterEstimates(riclpm4_con)
+# cross-lagged effects for democrat states
+crossInj_con4_dem <- pe_con4[pe_con4$lhs == "w1_05" & pe_con4$rhs == "w2_02" & pe_con4$group == 1,]
+crossDes_con4_dem <- pe_con4[pe_con4$lhs == "w1_05" & pe_con4$rhs == "w3_02" & pe_con4$group == 1,]
+# cross-lagged effects for republican states
+crossInj_con4_rep <- pe_con4[pe_con4$lhs == "w1_05" & pe_con4$rhs == "w2_02" & pe_con4$group == 2,]
+crossDes_con4_rep <- pe_con4[pe_con4$lhs == "w1_05" & pe_con4$rhs == "w3_02" & pe_con4$group == 2,]
+```
+
+In majority-Republican states, the main results are unchanged from the full model fitted to the whole sample: perceived descriptive norms predicted future mask wearing (_b_ = `r printnum(crossDes_con4_rep[,"est"])`, 95% CI [`r printnum(crossDes_con4_rep[,"ci.lower"])` `r printnum(crossDes_con4_rep[,"ci.upper"])`], _p_ = `r printp(crossDes_con4_rep[,"pvalue"])`) while perceived injunctive norms did not (_b_ = `r printnum(crossInj_con4_rep[,"est"])`, 95% CI [`r printnum(crossInj_con4_rep[,"ci.lower"])` `r printnum(crossInj_con4_rep[,"ci.upper"])`], _p_ = `r printp(crossInj_con4_rep[,"pvalue"])`).
+In majority-Democrat states, perceived descriptive norms have a slightly weaker effect on future mask wearing (_b_ = `r printnum(crossDes_con4_dem[,"est"])`, 95% CI [`r printnum(crossDes_con4_dem[,"ci.lower"])` `r printnum(crossDes_con4_dem[,"ci.upper"])`], _p_ = `r printp(crossDes_con4_dem[,"pvalue"])`) but the effect of perceived injunctive norms remains non-significant (_b_ = `r printnum(crossInj_con4_dem[,"est"])`, 95% CI [`r printnum(crossInj_con4_dem[,"ci.lower"])` `r printnum(crossInj_con4_dem[,"ci.upper"])`], _p_ = `r printp(crossInj_con4_dem[,"pvalue"])`).
+For full list of autoregressive and cross-lagged effects from the multi-group model, see Supplementary Table \@ref(tab:constrainedTable4).
 
@@
 
-(ref:plotDAGCaption) _Directed acyclic graph reflecting causal assumptions._ In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing (Mask).
-Perceptions of descriptive and injunctive norms also indirectly cause mask wearing through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
-Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
-Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
-Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.
-
-```{r plotDAG, echo=F, fig.width=5.6, fig.height=4.5, fig.cap="(ref:plotDAGCaption)"}
-plot4
-```
-
-\newpage
-
 (ref:plotCorBehNormCaption) _Predictions from multilevel models with self-reported mask wearing as the outcome variable and (a) perceived strength of descriptive norms and (b) perceived strength of injunctive norms as independent predictor variables._ Models contain random intercepts for participants and time points.
@@
 ```{r plotCorBehNorm, echo=F, fig.width=6.5, fig.height=3.8, fig.cap="(ref:plotCorBehNormCaption)"}
-plot5
+plot05
 ```
@@
 ```{r plotCDCSens, echo=F, fig.width=4, fig.height=6, fig.cap="(ref:plotCDCSensCaption)"}
-plot6
+plot06
+```
+
+\newpage
+
+(ref:plotDAGCaption) _Directed acyclic graph reflecting causal assumptions._ In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing (Mask).
+Perceptions of descriptive and injunctive norms also indirectly cause mask wearing through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
+Finally, demographic variables such as gender, age, ethnicity, socioeconomic status, and political orientation (summarised as Demographics) are exogenous variables that are common causes of all other variables.
+Using the backdoor criterion[@Pearl1995], this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived injunctive norms on mask wearing.
+
+```{r plotDAG, echo=F, fig.width=8, fig.height=7, fig.cap="(ref:plotDAGCaption)"}
+plot04
+```
+
+\newpage
+
+(ref:plotRICLPMPath2Caption) _Path diagram of ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed)._ Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for demographics and political orientation.
+Arrow thickness is scaled according to standardized effect size.
+Bolded arrows indicate significantly positive parameters, _p_ < 0.05.
+Gray arrows indicate non-significant parameters.
+
+```{r plotRICLPMPath2, echo=F, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotRICLPMPath2Caption)"}
+plot08
+```
+
+\newpage
+
+(ref:plotRICLPMPath3Caption) _Path diagram of ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed)._ Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences.
+Arrow thickness is scaled according to standardized effect size.
+Bolded arrows indicate significantly positive parameters, _p_ < 0.05.
+Gray arrows indicate non-significant parameters.
+
+```{r plotRICLPMPath3, echo=F, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotRICLPMPath3Caption)"}
+plot09
+```
+
+\newpage
+
+(ref:plotRICLPMCoef2Caption) _Standardized cross-lagged coefficients for descriptive norms and injunctive norms predicting future mask wearing in the ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed)._ Points are standardized estimates, lines are 95% confidence intervals.
+
+```{r plotRICLPMCoef2, echo=F, fig.width=6.5, fig.height=3, fig.cap="(ref:plotRICLPMCoef2Caption)"}
+plot11
+```
+
+\newpage
+
+(ref:plotRICLPMCoef3Caption) _Standardized cross-lagged coefficients for descriptive norms and injunctive norms predicting future mask wearing in the ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed)._ Points are standardized estimates, lines are 95% confidence intervals.
+
+```{r plotRICLPMCoef3, echo=F, fig.width=6.5, fig.height=3, fig.cap="(ref:plotRICLPMCoef3Caption)"}
+plot12
 ```
@@
 ```{r plotAttrition, echo=F, fig.width=5.5, fig.height=3, fig.cap="(ref:plotAttritionCaption)"}
-plot2
+plot02
+```
+
+\newpage
+
+(ref:plotAttritionBreakdownCaption) _Demographics of the sample across the course of the study._
+
+```{r plotAttritionBreakdown, echo=F, fig.width=7, fig.height=6, fig.cap="(ref:plotAttritionBreakdownCaption)"}
+plot13
 ```
@@
 ```{r plotUSMap, echo=F, fig.width=5, fig.height=3, fig.cap="(ref:plotUSMapCaption)"}
-plot1
+plot01
+```
+
+\newpage
+
+(ref:plotIntFreqCaption) _The relationship between self-reported mask wearing and the logged number of in-person interactions in the last 24 hours (a) and in the last 7 days (b)._ Lines and shaded areas are predictions and 95% confidence intervals from multilevel models with random intercepts for timepoint and participant.
+The first multilevel model revealed a negative relationship between mask wearing and the log number of in-person interactions in the last 24 hours (b = `r printnum(fixef(intFreq24hrs)[2])`, SE = `r printnum(summary(intFreq24hrs)$coefficients["logContacts24hrs","Std.
+Error"])`, p = `r printp(summary(intFreq24hrs)$coefficients["logContacts24hrs","Pr(>|t|)"])`).
+The second multilevel model revealed a negative relationship between mask wearing and the log number of in-person interactions in the last 7 days (b = `r printnum(fixef(intFreq7days)[2])`, SE = `r printnum(summary(intFreq7days)$coefficients["logContacts7days","Std.
+Error"])`, p `r printp(summary(intFreq7days)$coefficients["logContacts7days","Pr(>|t|)"])`).
+
+```{r plotIntFreq, echo=F, fig.width=8, fig.height=4, fig.cap="(ref:plotIntFreqCaption)"}
+plot14
 ```
@@
 
-(ref:lavaanTableCaption) Standardised autoregressive and cross-lagged parameters from random-intercept cross-lagged panel model.
-_Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms, Fact = factual beliefs, Pers = personal normative beliefs.
-Variable name suffixes indicate time points.
+(ref:constrainedTable2Caption) Unstandardized autoregressive and cross-lagged parameters from time-invariant random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed).
+_Arrows indicate the direction of prediction._
+
+```{r constrainedTable2, echo=F, warning=F, message=F}
+apa_table(
+  constrainedTable2,
+  booktabs = TRUE,
+  longtable = TRUE,
+  align = c("l", rep("c", 5)),
+  caption = "(ref:constrainedTable2Caption)"
+)
+```
+
+\newpage
+
+(ref:constrainedTable3Caption) Unstandardized autoregressive and cross-lagged parameters from time-invariant random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed).
+_Arrows indicate the direction of prediction._
+
+```{r constrainedTable3, echo=F, warning=F, message=F}
+apa_table(
+  constrainedTable3,
+  booktabs = TRUE,
+  longtable = TRUE,
+  align = c("l", rep("c", 5)),
+  caption = "(ref:constrainedTable3Caption)"
+)
+```
+
+\newpage
+
+(ref:unconstrainedTable1Caption) Unstandardized autoregressive and cross-lagged parameters from time-varying random-intercept cross-lagged panel model (full model).
+_Variable name suffixes indicate time points.
 Arrows indicate the direction of prediction._
 
-(ref:rightArrow) →
-
-```{r lavaanTable, echo=F, warning=F, message=F}
+```{r unconstrainedTable1, echo=F, warning=F, message=F}
 apa_table(
-  lavaanTable,
+  unconstrainedTable1,
   booktabs = TRUE,
@@
   align = c("l", rep("c", 5)),
-  caption = "(ref:lavaanTableCaption)"
+  caption = "(ref:unconstrainedTable1Caption)"
+)
+```
+
+\newpage
+
+(ref:unconstrainedTable2Caption) Unstandardized autoregressive and cross-lagged parameters from time-varying random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed).
+_Variable name suffixes indicate time points.
+Arrows indicate the direction of prediction._
+
+```{r unconstrainedTable2, echo=F, warning=F, message=F}
+apa_table(
+  unconstrainedTable2,
+  booktabs = TRUE,
+  longtable = TRUE,
+  font_size = "footnotesize",
+  align = c("l", rep("c", 5)),
+  caption = "(ref:unconstrainedTable2Caption)"
+)
+```
+
+\newpage
+
+(ref:unconstrainedTable3Caption) Unstandardized autoregressive and cross-lagged parameters from time-varying random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed).
+_Variable name suffixes indicate time points.
+Arrows indicate the direction of prediction._
+
+```{r unconstrainedTable3, echo=F, warning=F, message=F}
+apa_table(
+  unconstrainedTable3,
+  booktabs = TRUE,
+  longtable = TRUE,
+  font_size = "footnotesize",
+  align = c("l", rep("c", 5)),
+  caption = "(ref:unconstrainedTable3Caption)"
 )
@@
   booktabs = TRUE,
-  longtable = TRUE,
-  align = c("m{2.5cm}", "m{2.5cm}", "m{9cm}"),
+  landscape = TRUE,
+  align = c("m{2.5cm}", "m{2cm}", "m{12.5cm}", "m{1cm}", "m{1cm}"),
   caption = "(ref:itemTableCaption)"
@@
 
-## Supplementary References
+(ref:constructValTableCaption) Pairwise comparisons from a multilevel model testing the construct validity of the four self-reported social norm items.
+_Data were analysed in long format, with the self-report social norm item and the type of normative information provided (descriptive or injunctive) as predictors and a random intercept for participant.
+All pairwise comparisons account for multiple comparisons using Tukey adjustment._
 
-Pearl, J. (1995). Causal diagrams for empirical research. _Biometrika_, _82_(4), 669–688, https://doi.org/10.1093/biomet/82.4.669
+```{r constructValTable, echo=F, warning=F, message=F}
+apa_table(
+  constructValTable,
+  booktabs = TRUE,
+  longtable = TRUE,
+  caption = "(ref:constructValTableCaption)"
+)
+```
+
+\newpage
+
+(ref:constrainedTable4Caption) Unstandardized autoregressive and cross-lagged parameters from multi-group time-invariant random-intercept cross-lagged panel model, split by majority-Democrat and majority-Republican states.
+_The model contains all covariates.
+Arrows indicate the direction of prediction._
+
+```{r constrainedTable4, echo=F, warning=F, message=F}
+apa_table(
+  constrainedTable4,
+  booktabs = TRUE,
+  longtable = TRUE,
+  align = c("l", "l", rep("c", 5)),
+  caption = "(ref:constrainedTable4Caption)"
+)
+```

### 2023-07-06 manuscript.Rmd
second round of reviews

@@
 ---
-title             : "Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic"
+title             : "Descriptive norms caused increases in mask wearing during the COVID-19 pandemic"
 shorttitle        : "Norms and mask wearing"
@@
 Longitudinal trends suggested that norms and behavior were tightly coupled, changing quickly in response to public health recommendations.
-In addition, longitudinal modelling revealed that descriptive norms, not injunctive norms, caused future increases in mask wearing.
-During uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
+In addition, longitudinal modeling revealed that descriptive norms caused future increases in mask wearing across multiple waves of data collection.
+These cross-lagged causal effects of descriptive norms were large, even after controlling for non-social beliefs and demographic variables.
+Injunctive norms, by contrast, had less frequent and generally weaker causal effects on future mask wearing.
+During uncertain times, cooperative behavior is more strongly driven by what others are actually doing, rather than what others think ought to be done.
 
-note              : "\\footnotesize \\raggedright \\mbox{* indicates shared first authorship, † indicates shared senior authorship} \\par Correspondence concerning this article should be addressed to Samantha L.
-Heiman, 1101 E 10th St, Bloomington, IN 47405, United States.
-E-mail: slheiman@iu.edu \\par This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts & Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
-\\par This working paper has not yet been peer-reviewed."
 keywords          : "social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation"
-wordcount         : "4381 words"
+wordcount         : "Main text (excluding Methods): 4553 words"
 bibliography      : ["r-references.bib"]
@@
   - \captionsetup[figure]{font={stretch=1,small}}
+  - \note{\raggedright * indicates shared first authorship, † indicates shared senior authorship \par Correspondence concerning this article should be addressed to Samantha L.
+Heiman, 1101 E 10th St, Bloomington, IN 47405, United States.
+\text{E-mail:} \href{mailto:slheiman@iu.edu}{\nolinkurl{slheiman@iu.edu}} \par This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts and Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.}
 appendix          : "appendix.Rmd"
@@
 knitr::opts_chunk$set(warning = FALSE, message = FALSE)
-tar_load(c(fileData, fileOwid, d, owid,
-           constructValDes, constructValInj,
-           m1.1, m1.2, m2.1, m2.2, m2.3,
-           conf1.1, conf1.2, conf2.1, conf2.2, conf2.3,
-           pars2.1, pars2.2, pars2.3, riclpm, fitMeasures,
-           plot1, plot2, plot3, plot4, plot5, plot6, plot7,
-           plot8, itemTable, changePointsTable, lavaanTable))
+tar_load(
+  c(
+    d, owid, constructVal, intFreq24hrs, intFreq7days,
+    m1.1, m1.2, m2.1, m2.2, m2.3, conf1.1, conf1.2, 
+    conf2.1, conf2.2, conf2.3, pars2.1, pars2.2, pars2.3,
+    riclpm1_uncon, riclpm2_uncon, riclpm3_uncon,
+    riclpm1_con, riclpm2_con, riclpm3_con, riclpm4_con,
+    compare1, compare2, compare3, fitMeasures1_uncon, 
+    fitMeasures2_uncon, fitMeasures3_uncon,
+    fitMeasures1_con, fitMeasures2_con, fitMeasures3_con, 
+    fitMeasures4_con, plot01, plot02, plot03, plot04, plot05, 
+    plot06, plot07, plot08, plot09, plot10, plot11, plot12, plot13, plot14,
+    constrainedTable1, constrainedTable2, constrainedTable3,
+    constrainedTable4, itemTable, changePointsTable, unconstrainedTable1,
+    unconstrainedTable2, unconstrainedTable3, constructValTable
+  )
+)
 ```
@@
 However, less is known about how social norms arise endogenously within populations in the short term.
-While researchers have simulated the emergence of descriptive norms [@Muldoon2014a;@Muldoon2014b], this modelling work does not capture how descriptive norms develop alongside injunctive norms in real-world settings.
+While researchers have simulated the emergence of descriptive norms [@Muldoon2014a;@Muldoon2014b], this modeling work does not capture how descriptive norms develop alongside injunctive norms in real-world settings.
 Recent work in behavioral economics has also suggested that injunctive norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions [@Titlestad2019].
@@
 First, studies have not accounted for other potential non-social explanations for behavior, such as perceptions of the effectiveness of the behavior and personal beliefs that one should behave in a certain way.
-These non-social beliefs, labeled "factual beliefs" and "personal normative beliefs" [@Bicchieri2014], often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
+These non-social beliefs, labeled "factual beliefs" and "personal normative beliefs" [@Bicchieri2014a], often correlate with descriptive and injunctive norms, but they are fundamentally different because they can cause behavior separately from social expectations about what others do or think should be done.
 For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common or socially approved behavior).
-It is thus important to control for non-social beliefs in studies of social norms.
+It is thus important to control for non-social beliefs in studies of social norms, especially considering that personal norms have previously been shown to influence prosocial behavior[@Bicchieri2010;@Capraro2021].
 Second, studies have tended to follow cross-sectional experimental designs in which social norm perceptions are manipulated by the researchers.
@@
 
-Recent research has found positive relationships between social norms and protective COVID-19 behaviors.
+Recent research has studied the relationships between social norms and protective COVID-19 behaviors.
 In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure [@Macy2021], and another study found that experimentally-induced descriptive norms increased mask wearing intentions [@Bokemper2021].
+In Italy, an experimental study found no effect of messages highlighting descriptive norms, injunctive norms, or personal norms on time spent reading information about COVID-19 governmental rules [@Bilancini2020].
 In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing [@Rudert2021].
-These studies are telling, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
-Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States [@BaxterKing2022].
+These studies are informative, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
+Furthermore, several of the studies do not control for potential confounding variables, such as demographics and political ideology.
+These variables are important to account for as they have previously been shown to be related to COVID-19 attitudes and behaviors [@BaxterKing2022;@Heiman2022].
 
-Here, we use two years of data from a representative sample of adults in the United States (18 time points; _n_ = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
+Here, we use two years of data from a representative sample of adults in the United States (18 time points; _n_ = `r nrow(tar_read(d))`) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
 Participants reported their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
-We also asked participants about their non-social mask wearing beliefs and political ideology, and controlled for these factors.
+We also asked participants about their non-social mask wearing beliefs, demographics, and political ideology, and controlled for these factors.
 We used these longitudinal data to answer two main research questions in a specific real-world context.
@@
 
-To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration.
+To answer our first research question about the emergence of mask wearing norms, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration, showing how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic.
 Figure \@ref(fig:plotTimeline) plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
@@
 ```{r plotTimeline, echo=F, fig.width=6, fig.height=6, fig.cap="(ref:plotTimelineCaption)"}
-plot3
+plot03
 ```
@@
 
-Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing and perceived social norms over time.
-To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
-In line with our proposed causal model (Supplementary Figure \@ref(fig:plotDAG)), we also included political orientation as an exogenous time-invariant control.
-According to established fit statistics, this model fitted the data well (root mean square error of approximation = `r printnum(fitMeasures['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual = `r printnum(fitMeasures['srmr'], digits = 3)`; comparative fit index = `r printnum(fitMeasures['cfi'], digits = 3)`).
-Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \@ref(tab:lavaanTable) for full list of estimated autoregressive and cross-lagged effects).
+Sample averages can provide informative trends, but they do not allow us to determine whether within-person changes in social norms caused future within-person changes in mask wearing over time.
+To answer our second research question about causal effects, we fitted a ten-wave random-intercept cross-lagged panel model[@Hamaker2015;@Mulder2021] to the longitudinal data.
+This model separately estimates stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and control variables (factual beliefs and personal normative beliefs).
+In line with our proposed causal model (Supplementary Figure \@ref(fig:plotDAG)), we also control for potential confounding in this model by including demographics (gender, age, ethnicity, socioeconomic status) and political orientation as exogenous controls.
 
-```{r, echo = FALSE}
-sc <- standardizedSolution(riclpm)
-pe <- parameterEstimates(riclpm)
-```
-
-Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
-On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`) and stronger injunctive mask wearing norms (_r_ = `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.lower"])` `r printnum(sc[sc$lhs == "ri1" & sc$op == "~~" & sc$rhs == "ri2","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri1" & pe$op == "~~" & pe$rhs == "ri2","pvalue"])`).
-Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (_r_ = `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.lower"])` `r printnum(sc[sc$lhs == "ri2" & sc$op == "~~" & sc$rhs == "ri3","ci.upper"])`], _p_ `r printp(pe[pe$lhs == "ri2" & pe$op == "~~" & pe$rhs == "ri3","pvalue"])`).
-
-Regarding within-person dynamics over time, Figure \@ref(fig:plotRICLPM) displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing across the study duration, controlling for non-social beliefs and political orientation.
-In random intercept cross-lagged panel models, autoregressive effects represent "persistence" or "inertia" in within-person fluctuations from stable trait levels.
+Random-intercept cross-lagged panel models capture within-person changes over time with autoregressive and cross-lagged effects.
+Autoregressive effects represent "persistence" or "inertia" in within-person fluctuations from stable trait levels.
 In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the "stable trait level" over time, which is captured by the random intercepts in our model).
@@
 Cross-lagged effects are thus used to infer within-person causal influences over time.
+In what follows, we focus on the within-person autoregressive and cross-lagged effects for mask wearing and perceived social norms (see Supplementary Results for between-person results).
 
-(ref:plotRICLPMCaption) _Results of ten-wave unconstrained random-intercept cross-lagged panel model._ Circles represent data collection time points.
-Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
+```{r echo=FALSE}
+# parameter estimates
+pe_con <- parameterEstimates(riclpm1_con)
+# cross-lagged effects from constrained model
+crossInj_con <- pe_con[pe_con$lhs == "w1_05" & pe_con$rhs == "w2_02",]
+crossDes_con <- pe_con[pe_con$lhs == "w1_05" & pe_con$rhs == "w3_02",]
+```
+
+We first fitted a time-invariant model, which constrained the autoregressive effects, cross-lagged effects, covariances, and variances to equality over time.
+This model assumes that the relationships between variables are identical across all time points, giving us a sense of the average causal effects of social norms on mask wearing over time.
+According to established fit statistics, the time-invariant model fitted the data reasonably well (root mean square error of approximation (RMSEA) = `r printnum(fitMeasures1_con['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures1_con['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures1_con['rmsea.ci.upper'], digits = 3)`]; standardized root mean squared residual (SRMR) = `r printnum(fitMeasures1_con['srmr'], digits = 3)`; comparative fit index (CFI) = `r printnum(fitMeasures1_con['cfi'], digits = 3)`).
+Table \@ref(tab:constrainedTable1) summarizes the autoregressive and cross-lagged parameters from the time-invariant model.
+Autoregressive effects were significantly positive for mask wearing, perceived descriptive norms, and perceived injunctive norms, indicating that being higher than average on these variables at time $t$ generally predicted being higher than average on the same variables at time $t + 1$.
+Crucially, cross-lagged effects from the time-invariant model revealed that perceived descriptive norms predicted future mask wearing (unstandardized _b_ = `r printnum(crossDes_con[,"est"])`, 95% CI [`r printnum(crossDes_con[,"ci.lower"])` `r printnum(crossDes_con[,"ci.upper"])`], _p_ `r printp(crossDes_con[,"pvalue"])`) while perceived injunctive norms did not (_b_ = `r printnum(crossInj_con[,"est"])`, 95% CI [`r printnum(crossInj_con[,"ci.lower"])` `r printnum(crossInj_con[,"ci.upper"])`], _p_ = `r printp(crossInj_con[,"pvalue"])`).
+In other words, perceiving descriptive norms as stronger than average at time $t$ generally predicted wearing masks more frequently than average at time $t + 1$, while no such effect existed for injunctive norms.
+All other cross-lagged effects between mask wearing, perceived descriptive norms, and perceived injunctive norms were significantly positive in the time-invariant model.
+This general pattern of results was unchanged when removing factual beliefs and personal normative beliefs from the time-invariant model (Supplementary Table \@ref(tab:constrainedTable2)) and when removing factual beliefs, personal normative beliefs, and all exogenous covariates from the time-invariant model (Supplementary Table \@ref(tab:constrainedTable3)).
+
+(ref:rightArrow) →
+
+(ref:constrainedTable1Caption) Unstandardized autoregressive and cross-lagged parameters from time-invariant random-intercept cross-lagged panel model.
+_Arrows indicate the direction of prediction.
+Note that the effects of factual beliefs and personal normative beliefs are omitted from this table for clarity.
+SE = standard error._
+
+```{r constrainedTable1, echo=F, warning=F, message=F}
+apa_table(
+  constrainedTable1,
+  booktabs = TRUE,
+  align = c("l", rep("c", 5)),
+  caption = "(ref:constrainedTable1Caption)"
+)
+```
+
+Given that the strength of perceived social norms varied throughout our data collection window (Figure \@ref(fig:plotTimeline)), it is plausible that the causal effects of social norms on mask wearing may have changed over time as well, rather than being identical at each time point.
+To test whether the equality constraints over time were tenable[@Mulder2021], we compared the time-invariant model to an alternative time-varying model that freely estimated the autoregressive effects, cross-lagged effects, covariances, and variances at each time point.
+This model assumes that the relationships between variables are different across different time points, allowing the causal effects of social norms on future mask wearing to vary over time.
+The time-varying model fitted the data better than the time-invariant model (RMSEA = `r printnum(fitMeasures1_uncon['rmsea'], digits = 3)`, 95% CI [`r printnum(fitMeasures1_uncon['rmsea.ci.lower'], digits = 3)` `r printnum(fitMeasures1_uncon['rmsea.ci.upper'], digits = 3)`]; SRMR = `r printnum(fitMeasures1_uncon['srmr'], digits = 3)`; CFI = `r printnum(fitMeasures1_uncon['cfi'], digits = 3)`).
+Model comparison revealed that freely estimating the parameters over time resulted in improved model fit ($\Delta$AIC = `r printnum(compare1[1,2] - compare1[2,2])`, $\Delta\chi^2$(`r round(compare1[2,6], digits = 0)`) = `r printnum(-compare1[2,5], big.mark = "")`, _p_ `r printp(compare1[2,7])`) suggesting that there was substantial variability in the relationships between mask wearing and social norms across the course of the pandemic.
+Accordingly, we now turn to the results from the time-varying model.
+
+```{r echo=FALSE}
+# parameter estimates and standardized parameters
+pe_uncon <- parameterEstimates(riclpm1_uncon)
+sc_uncon <- standardizedSolution(riclpm1_uncon)
+# unstandardized cross-lagged effects from unconstrained model
+crossDes05_unstd_uncon <- pe_uncon[pe_uncon$lhs == "w1_05" & pe_uncon$rhs == "w3_02",]
+crossDes09_unstd_uncon <- pe_uncon[pe_uncon$lhs == "w1_09" & pe_uncon$rhs == "w3_05",]
+crossDes13_unstd_uncon <- pe_uncon[pe_uncon$lhs == "w1_14" & pe_uncon$rhs == "w3_13",]
+crossInj15_unstd_uncon <- pe_uncon[pe_uncon$lhs == "w1_16" & pe_uncon$rhs == "w2_15",]
+# standardized cross-lagged effects from unconstrained model
+crossDes05_std_uncon <- sc_uncon[sc_uncon$lhs == "w1_05" & sc_uncon$rhs == "w3_02",]
+crossDes09_std_uncon <- sc_uncon[sc_uncon$lhs == "w1_09" & sc_uncon$rhs == "w3_05",]
+crossDes13_std_uncon <- sc_uncon[sc_uncon$lhs == "w1_14" & sc_uncon$rhs == "w3_13",]
+crossInj15_std_uncon <- sc_uncon[sc_uncon$lhs == "w1_16" & sc_uncon$rhs == "w2_15",]
+```
+
+Figure \@ref(fig:plotRICLPMPath1) displays the autoregressive and cross-lagged effects for social norms and mask wearing from the time-varying model (see Supplementary Table \@ref(tab:unconstrainedTable1) for full list of estimated autoregressive and cross-lagged effects).
+In late 2020 and 2021, we find three occasions where within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
+According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized coefficients for these cross-lagged effects were large (first wave, standardized $\beta$ = `r printnum(crossDes05_std_uncon$est.std)`, _b_ = `r printnum(crossDes05_unstd_uncon$est)`, 95% CI [`r printnum(crossDes05_unstd_uncon$ci.lower)` `r printnum(crossDes05_unstd_uncon$ci.upper)`], _p_ = `r printp(crossDes05_unstd_uncon$pvalue)`; second wave, $\beta$ = `r printnum(crossDes09_std_uncon$est.std)`, _b_ = `r printnum(crossDes09_unstd_uncon$est)`, 95% CI [`r printnum(crossDes09_unstd_uncon$ci.lower)` `r printnum(crossDes09_unstd_uncon$ci.upper)`], _p_ = `r printp(crossDes09_unstd_uncon$pvalue)`; fifth wave, $\beta$ = `r printnum(crossDes13_std_uncon$est.std)`, _b_ = `r printnum(crossDes13_unstd_uncon$est)`, 95% CI [`r printnum(crossDes13_unstd_uncon$ci.lower)` `r printnum(crossDes13_unstd_uncon$ci.upper)`], _p_ = `r printp(crossDes13_unstd_uncon$pvalue)`).
+At other time points, the cross-lagged effects for descriptive norms tended to be estimated in a positive direction, though these estimates were not statistically significant and decreased in magnitude in 2022 (Figure \@ref(fig:plotRICLPMCoef1)).
+Time-varying models without covariates revealed an additional significant cross-lagged effect of descriptive norms on future mask wearing at the fourth wave, but otherwise the general pattern of results for descriptive norms was unchanged (Supplementary Figures \@ref(fig:plotRICLPMPath2) - \@ref(fig:plotRICLPMCoef3); Supplementary Tables \@ref(tab:unconstrainedTable2) and \@ref(tab:unconstrainedTable3)).
+
+(ref:plotRICLPMPath1Caption) _Path diagram of ten-wave time-varying random-intercept cross-lagged panel model._ Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, demographics, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
@@
 Gray arrows indicate non-significant parameters.
-There are no significant direct paths from injunctive norms to future mask wearing, showing that people’s beliefs about what they should be doing did not have any direct influences on future mask wearing.
-On the other hand, there are significant paths from descriptive norms to future mask wearing, meaning that people’s beliefs about what others are doing influenced their future mask wearing.
 
-```{r plotRICLPM, echo=F, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotRICLPMCaption)"}
-plot7
+```{r plotRICLPMPath1, echo=F, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotRICLPMPath1Caption)"}
+plot07
 ```
 
-In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing.
-On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
-According to recent effect size guidelines for cross-lagged panel models [@Orth2022], the standardized beta coefficients for these cross-lagged effects were large (first wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.lower"])` `r printnum(sc[sc$lhs == "w1_05" & sc$op == "~" & sc$rhs == "w3_02","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_05" & pe$op == "~" & pe$rhs == "w3_02","pvalue"])`; second wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.lower"])` `r printnum(sc[sc$lhs == "w1_09" & sc$op == "~" & sc$rhs == "w3_05","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_09" & pe$op == "~" & pe$rhs == "w3_05","pvalue"])`; fourth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.lower"])` `r printnum(sc[sc$lhs == "w1_13" & sc$op == "~" & sc$rhs == "w3_11","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_13" & pe$op == "~" & pe$rhs == "w3_11","pvalue"])`; fifth wave, $\beta$ = `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","est.std"])`, 95% CI [`r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.lower"])` `r printnum(sc[sc$lhs == "w1_14" & sc$op == "~" & sc$rhs == "w3_13","ci.upper"])`], _p_ = `r printp(pe[pe$lhs == "w1_14" & pe$op == "~" & pe$rhs == "w3_13","pvalue"])`).
-In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
-We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive norms.
-Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.
+(ref:plotRICLPMCoef1Caption) _Standardized cross-lagged coefficients for descriptive norms and injunctive norms predicting future mask wearing in the ten-wave time-varying random-intercept cross-lagged panel model._ Points are standardized estimates, lines are 95% confidence intervals.
 
-However, the model also reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did _not_ predict future within-person changes in mask wearing across our study duration.
-All cross-lagged effects from perceived injunctive norms to mask wearing are non-significant.
-Any causal effect that perceived injunctive norms might have had on future mask wearing appears to be fully mediated by perceived descriptive norms.
-This means that believing that others think that mask wearing is the right thing to do influences one’s perception of what others are actually doing, which then influences future behavior.
-For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing.
-But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing over time within individuals.
+```{r plotRICLPMCoef1, echo=F, fig.width=6.5, fig.height=3, fig.cap="(ref:plotRICLPMCoef1Caption)"}
+plot10
+```
+
+In contrast, within-person increases in perceived injunctive norms only predicted future within-person increases in mask wearing at a single time point in March 2022 (seventh wave, $\beta$ = `r printnum(crossInj15_std_uncon$est.std)`, _b_ = `r printnum(crossInj15_unstd_uncon$est)`, 95% CI [`r printnum(crossInj15_unstd_uncon$ci.lower)` `r printnum(crossInj15_unstd_uncon$ci.upper)`], _p_ = `r printp(crossInj15_unstd_uncon$pvalue)`).
+This cross-lagged effect was robust to the removal of covariates (Supplementary Figures \@ref(fig:plotRICLPMPath2) - \@ref(fig:plotRICLPMCoef3); Supplementary Tables \@ref(tab:unconstrainedTable2) and \@ref(tab:unconstrainedTable3)).
+At all other time points, the cross-lagged effects of injunctive norms on future mask wearing were non-significant and estimated to be generally weaker than the cross-lagged effects of descriptive norms (Figure \@ref(fig:plotRICLPMCoef1)), explaining the lack of effect of injunctive norms in the time-invariant model.
+
+One possible explanation for the generally weaker causal influence of injunctive norms on future mask wearing is that injunctive norms might only have been sufficiently salient in regions of the United States with stronger governmental enforcement of mask wearing.
+Previous work has shown that, during the pandemic, states with Democratic leadership tended to have more stringent mask wearing policies than states with Republican leadership[@Adolph2022].
+It is thus plausible that injunctive norms had a stronger causal effect on future mask wearing in majority-Democrat states compared to majority-Republican states.
+To test this explanation, we fitted a multi-group time-invariant random-intercept cross-lagged panel model to the data, with separate groups for majority-Democrat and majority-Republican states.
+However, contrary to the salience account, perceived injunctive norms failed to predict future mask wearing in both Democrat states and Republican states (see Supplementary Results).
+
+Beyond the effects of descriptive and injunctive norms on future mask wearing, our main time-varying model revealed other interesting patterns (Figure \@ref(fig:plotRICLPMPath1)).
+On a few occasions, we find evidence for a reciprocal relationship between social norms and mask wearing, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive and injunctive norms.
+Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person influences between these variables as the pandemic unfolded.
 
@@
 Moreover, the results of our cross-lagged panel model indicate that descriptive norms caused future increases in mask wearing in the first year and a half of the pandemic.
-By contrast, injunctive norms were not directly causally related to future mask wearing over the entire two years of data collection.
+By contrast, injunctive norms were less frequently related to future mask wearing during the pandemic, with generally weaker effects than descriptive norms.
 
@@
 The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one’s own behaviors are mutually constructed [@Markus2010].
-Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms [@Amato2018].
+Indeed, previous work has shown that formal institutions are critical for the emergence and rapid adoption of novel social norms [@Amato2018].
 While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our trends have shown, can unfold over measurement intervals as short as four to six weeks.
 
-We found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing, independent of the effects of non-social beliefs and political orientation.
+We found that descriptive norms predicted future within-person increases in mask wearing, independent of the effects of injunctive norms, non-social beliefs, and demographic variables.
 This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors [@Bokemper2021; @Rudert2021].
@@
 Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic [@Gelfand2015].
+This may help to explain why the effects of descriptive norms were more prevalent in earlier waves of our data, when uncertainty was highest.
 Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population [@Chaudhuri2011].
@@
 Third, the increased frequency of mask wearing in the population might have created a bandwagon effect [@SchmittBeck2015], encouraging conformist copying.
-Under this view, people wear masks not to coordinate or cooperate, but simply to fit in with the crowd.
+Under this view, people wear masks not to coordinate or cooperate, but simply because they see a majority of others engaging in the behavior.
 Future research will be required to determine the motivations underlying adherence to descriptive norms during uncertain times.
 
-We found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
+On the whole, we found that perceived injunctive norms tended not to predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing were not strongly causally related during our data collection window.
 One possible explanation for this result is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were more salient than injunctive norms during the pandemic.
@@
 To test this idea, future research should expand our longitudinal approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
+Nevertheless, it is worth noting that injunctive norms continued to have no overall causal effect on mask wearing when we focused specifically on majority-Democrat US states that tended to employ state-wide mask mandates[@Adolph2022], suggesting that a lack of saliency cannot entirely explain why injunctive norms did not have a consistent effect on mask wearing during the pandemic.
 
-We are limited in generalizing these findings due to the constraints of our sample and the variables considered.
-While the attrition in our study did not substantially affect the representativeness of our sample (Supplementary Figure \@ref(fig:plotAttritionBreakdown)),
-the reduction in our sample size did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
-Our results also might not generalize to all social norms, behaviors, and social dilemmas.
+Regarding injunctive mask wearing norms, several open questions remain.
+First, given that we see one time point in the later stages of the pandemic in which injunctive norms directly predicted future mask wearing, it is possible that there were also earlier time points where injunctive norms had effects on behavior.
+We do not have data from the very beginning of the pandemic (March-September 2020) to directly test this, but future work could test this by examining the effects of injunctive norms on behaviour directly after the onset of a crisis.
+Second, it remains unclear how the source of the injunctive norm influences its efficacy.
+Our operationalization of injunctive norms referred in part to the behavior of respected others, and so the influence of the perceived approval of prestigious political and nationwide organizational leaders may be an important lever to test.
+In addition, given that we operationalized injunctive norms in part by local-level encouragement, this might explain why we see the only significant effect of injunctive norms in February 2022, as by this point the CDC was easing their nationwide mask guidelines.
+At this stage, people may have begun to look more to injunctive information from their local areas for encouragement on what was appropriate.
+Third, our longitudinal modelling showed that changes in perceived descriptive norms consistently influenced future changes in perceived injunctive norms throughout our study period, while the reverse was only true in the latter half of our study period.
+This suggests that initially changing empirical expectations of others' behavior could potentially motivate compliance and shift later perceptions of injunctive norms [@Bicchieri2014b].
+Although our model was not specifically designed to address this question, future work should follow up on the temporal relationship between descriptive and injunctive norms with more fine-grained longitudinal data, perhaps experimentally intervening on perceptions of descriptive or injunctive norms to infer directions of causation.
+
+It is unclear the extent to which our pattern of results might generalize to other cultural contexts.
+The United States is relatively close to the global average on cultural tightness-looseness, a cross-cultural dimension that measures nationwide strength of social norms[@Gelfand2011;@Gelfand2021].
+We might therefore expect to see similar effects of social norms on mask wearing in countries with similar scores on this dimension.
+However, for countries with higher levels of cultural tightness than the United States, we might expect to see stronger effects of descriptive norms on mask wearing, additional effects of injunctive norms on mask wearing, and different dynamics between descriptive and injunctive norms.
+If borne out by the data, such patterns might explain why countries with higher levels of cultural tightness had fewer COVID-19 cases and deaths compared to culturally looser countries[@Gelfand2021].
+To test this, future research could expand our longitudinal approach to other cultural contexts.
+
+Our results might not generalize to all social norms, behaviors, and social dilemmas.
 Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more remote than the COVID-19 pandemic.
 For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.
+Mask wearing is also a unique cooperative behavior in that it is not "purely" cooperative (i.e., imposing costs on the actor while providing benefits to targets).
+Much of past research on normative influences on prosocial behavior has focused on these purely unselfish cooperative behaviors[@Capraro2021].
+Mask wearing is different as it does have individual benefits (e.g., reduced likelihood of contracting the disease).
+However, mask wearing also often imposes non-trivial costs on individuals (e.g., difficulty breathing, discomfort, disrupted social interaction) for the benefit of the wider group, meaning that it is still useful to conceptualise it as a prosocial or cooperative behavior.
+With these considerations, our results may be most pertinent for easily observable behaviors in response to an immediate or short-term threat that offer both benefits and costs to the individual in service to the wider group (e.g., other protective health behaviors, joining a protest or strike).
 
-For the case of mask wearing in the United States during the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and tracked ongoing changes in both recommendations from authorities and current levels of cooperative behavior.
+There are some limitations associated with our longitudinal survey design.
+First, we asked participants to self-report their frequency of mask wearing.
+However, self-report measures may be biased by anchoring effects and participants’ imperfect recall of their own behavior, resulting in measurement error[@Hansen2022].
+Such measures were necessary to implement our two-year longitudinal study, but future field studies of mask wearing norms could avoid this pitfall by directly measuring mask wearing through naturalistic observation.
+Second, as is common in longitudinal designs, there was substantial attrition over the course of the study (Supplementary Figure \@ref(fig:plotAttrition)).
+While this attrition did not substantially affect the demographic representativeness of our sample (Supplementary Figure \@ref(fig:plotAttritionBreakdown)), the sample may have been self-selected in other unknown ways that introduced confounds.
+Future longitudinal work could encourage higher retention rates with more frequent reminders and fewer survey questions to maintain attention and interest.
+
+Despite these caveats, we have shown that mask wearing norms developed rapidly in the United States population during the COVID-19 pandemic and tracked ongoing changes in both recommendations from authorities and current levels of mask wearing behavior.
 Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
@@
 This echoes recent reports of people in the United States not wanting to be "singled out" by being the only one wearing a mask in their community[@Natanson2022].
-Our work thus underscores the importance of consistent, visible community adherence for encouraging protective behaviors in response to global pandemics like COVID-19.
+Organizations interested in combating such backfire effects might consider making masks easily and cheaply available in local settings (e.g., within smaller institutions or at events) and running media campaigns to show that mask wearing is common among "everyday people" doing "everyday things".
+Once tested in target settings[@IJzerman2020], such interventions could leverage the power of consistent, visible community adherence to encourage protective behaviors in response to global pandemics like COVID-19.
 
@@
 Using the platform Prolific (https://www.prolific.co/), we distributed surveys to a representative sample of adults from the United States (_n_ = `r nrow(d)`, _M_~age~ = `r round(mean(d$Age.1, na.rm = TRUE))` years, `r round(mean(d$Ethnicity.1 == "White or Caucasian", na.rm = TRUE)*100)`% White, `r round(mean(d$Sex.1 == 2, na.rm = TRUE)*100)`% Women; see Supplementary Figure \@ref(fig:plotUSMap) for geographic distribution).
-From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
+From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors (for a visualization of all questions asked of this sample, see https://navigateobscurity.com/cooperation-conflict-lab/us-variable-lookup).
 This resulted in 18 unique time points of data collection during the pandemic.
@@
 However, this attrition did not substantially affect the demographic makeup of the sample through time (Supplementary Figure \@ref(fig:plotAttritionBreakdown)).
+On average, participants were paid approximately $8 USD per hour for completing the surveys.
 
@@
 
-At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+At every time point, participants were asked about the number of in-person interactions they had in the last 7 days and the last 24 hours.
+These two separate items specified that the interactions could have been either recreational or routine: "_How many people (outside of your household) have you had in-person interactions with during the last [7 days / 24 hours] (e.g., visiting with friends, buying food or supplies from a cashier)?_" Following these questions, participants self-reported their mask wearing by answering: "_During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?_" This question was asked regardless of how many in-person interactions the participants reported in the two interaction questions.
+Participants responded to the mask wearing question on a 5-point Likert scale, from Never (1) to Always (5).
+Responses to this question were weakly negatively associated with the number of in-person interactions that participants reported in the last 7 days and the last 24 hours, since wearing masks and social distancing are both COVID-19 protective behaviors (Supplementary Figure \@ref(fig:plotIntFreq)).
 
@@
 
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
+There is potential overlap between these operationalizations of descriptive and injunctive norms.
+For example, the item asking how often participants see people that they respect and trust wearing masks could be capturing descriptive norms as well as injunctive norms.
+To address this concern about the construct validity of the four social norm items, at time point 7 we asked participants about their interpretations of the items.
 We asked participants whether each of the four items informed them about what people _are_ doing or what people _should_ be doing (i.e., giving descriptive or injunctive information).
-Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \@ref(tab:itemTable)).
+We found that participants rated the two descriptive norm questions as providing more descriptive information than the two injunctive norm questions, and vice versa, suggesting that the items are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Tables \@ref(tab:itemTable) and \@ref(tab:constructValTable)).
 
@@
 
-To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \@ref(fig:plotDAG)).
-In this causal model, we included two kinds of non-social beliefs highlighted by previous research [@Bicchieri2014]: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
+To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (Supplementary Figure \@ref(fig:plotDAG)).
+In this causal model, we included two kinds of non-social beliefs highlighted by previous research [@Bicchieri2014a]: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
 These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
-In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic [@BaxterKing2022].
-Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
+In addition, we also included demographics (gender, age, ethnicity, socioeconomic status, and political orientation) as common causes of all other variables.
+This is justified by evidence showing that these variables are associated with COVID-19 attitudes and behaviors[@BaxterKing2022;@Heiman2022].
+Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and all demographic variables in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
 
@@
 
-Political orientation was measured in the first time point only.
-We measured political orientation as the mean of the following two items: "_How would you describe your political orientation with regard to social issues?_" and "_How would you describe your political orientation with regard to economic issues?_".
+All demographic variables were measured at the first time point only.
+Gender was measured using a question that asked about participants' biological sex ("_What is your sex?_") with three possible categories (Male, Female, or Other).
+We are assuming within these analyses that participants identify as the gender associated with their biological sex.
+Age was measured numerically ("_What is your age?_").
+Ethnicity was measured with a single item ("_What is your ethnicity?_") with six possible categories (Asian or Pacific Islander, Hispanic or Latino/a, White or Caucasian, Black or African American, Native American, or Other).
+Socioeconomic status was operationalized through a composite measure averaging income, education level, and subjective socioeconomic status.
+Income was captured by a single item measured on a 9-point scale of increasing income amounts ("_What was your combined household income in the previous year before taxes?_"), education was captured by a single item measured on a 7-point scale of increasing education levels ("_What is the highest level of education that you have completed?_"), and subjective socioeconomic status was represented by a ladder with ten rungs, from those who are "worst off" at the bottom to those who are "best off" at the top.
+Political orientation was measured as the mean of the following two items: "_How would you describe your political orientation with regard to social issues?_" and "_How would you describe your political orientation with regard to economic issues?_".
 These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
@@
 
-To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data [@Hamaker2015].
+To quantify the within-person relationships between our variables over time, we fitted random-intercept cross-lagged panel models to our longitudinal data [@Hamaker2015;@Mulder2021].
 This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time _t-1_ predicts being above average in another variable at time _t_.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015; @Granger1969].
+Positive cross-lagged effects from this model indicate that being above average on one variable at time $t$ predicts being above average in another variable at time $t + 1$.
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets [@Hamaker2015;@Granger1969;@Mulder2021].
 
-We estimated the random-intercept cross-lagged panel model using the _lavaan_ R package [@Rosseel2012].
-In line with our directed acyclic graph (see Supplementary Figure \@ref(fig:plotDAG)), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
+We estimated the random-intercept cross-lagged panel models using the _lavaan_ R package [@Rosseel2012].
+In line with our directed acyclic graph (Supplementary Figure \@ref(fig:plotDAG)), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two control variables (factual beliefs and personal normative beliefs) in the model.
 For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
-We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We modeled autoregressive and cross-lagged effects between all five variables, and included gender, age, ethnicity, socioeconomic status, and political ideology as exogenous covariates.
 We restricted the analysis to the ten time points with available data for all five variables.
-Full information maximum likelihood estimation was used to deal with missing data.
+We fitted both a time-invariant model (i.e., a model that constrained the within-person autoregressive effects, cross-lagged effects, covariances, and variance to equality over time) and a time-varying model that freely estimated all parameters.
+In both cases, full information maximum likelihood estimation was used to deal with missing data.
 
+In addition to the full random-intercept cross-lagged panel model containing all covariates, we additionally fitted (1) a model with factual beliefs and personal normative beliefs removed and (2) a model with all covariates removed.
+We fitted both time-invariant and time-varying versions of these models.
+Finally, we fitted a time-invariant multi-group model that estimated parameters separately for majority-Democrat states and majority-Republican states, operationalised using the state-level results from the 2020 United States Presidential election.
+Results for these additional models are reported in the Supplementary Material.
+
+Analyses were conducted retrospectively after data collection during the COVID-19 pandemic, meaning that no analyses were preregistered and all analyses should thus be considered exploratory.
 All analyses were conducted in R v4.1.1 [@RCoreTeam].
@@
 The manuscript was reproducibly generated using the _targets_ [@Landau2021] and _papaja_ [@Aust2022] packages.
-All code and data are publicly available on GitHub [@Heiman2023].
+All code and data are publicly available on GitHub: https://github.com/ScottClaessens/covidMaskWearing
 
 \newpage
+\nolinenumbers
 
@@
 SLH and SC wrote the first draft of the paper.
-SC conducted the formal analysis and created all visualizations.
+SC conducted the analyses and created all visualizations.
 ERH, AA, and PMT provided funding and supervision for the study.

### 2023-07-06 manuscript.tex
second round of reviews

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 
-% \usepackage{etoolbox}
 \makeatletter
-\patchcmd{\HyOrg@maketitle}
+\usepackage{etoolbox}
+\patchcmd{\maketitle}
   {\section{\normalfont\normalsize\abstractname}}
@@
   {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
+\patchcmd{\maketitle}
   {\section{\protect\normalfont{\@title}}}
@@
 {}{\PatchFailed}
-\keywords{social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: 4381 words}
+\keywords{social norms; descriptive norms; injunctive norms; longitudinal; COVID-19; mask wearing; cooperation\newline\indent Word count: Main text (excluding Methods): 4553 words}
 \usepackage{lineno}
@@
 \captionsetup[figure]{font={stretch=1,small}}
+\note{\raggedright * indicates shared first authorship, † indicates shared senior authorship \par Correspondence concerning this article should be addressed to Samantha L.
+Heiman, 1101 E 10th St, Bloomington, IN 47405, United States.
+\text{E-mail:} \href{mailto:slheiman@iu.edu}{\nolinkurl{slheiman@iu.edu}} \par This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts and Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.}
 \ifLuaTeX
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
-  pdftitle={Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic},
+  pdftitle={Descriptive norms caused increases in mask wearing during the COVID-19 pandemic},
   pdfauthor={Samantha L. Heiman*,1, Scott Claessens*,2, Jessica D. Ayers3, Diego Guevara Beltrán4, Andrew Van Horn5,6, Edward R. Hirt1, Athena Aktipis†,4, \& Peter M. Todd†,1,7},
@@
 
-\title{Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic}
+\title{Descriptive norms caused increases in mask wearing during the COVID-19 pandemic}
 \author{Samantha L.
@@
 
-\note{
-
-\footnotesize
-
-\raggedright
-
-\mbox{* indicates shared first authorship, † indicates shared senior authorship}
-
-\par
-
-Correspondence concerning this article should be addressed to Samantha L.
-Heiman, 1101 E 10th St, Bloomington, IN 47405, United States.
-E-mail: \href{mailto:slheiman@iu.edu}{\nolinkurl{slheiman@iu.edu}}
-
-\par
-
-This study was funded by the Interdisciplinary Cooperation Initiative, ASU President's Office, the Cooperation Science Network, the Institute for Mental Health Research, the University of New Mexico, the Indiana University College of Arts \& Sciences, the Rutgers University Center for Human Evolutionary Studies, the Charles Koch Foundation, and the John Templeton Foundation.
-
-\par
-
-This working paper has not yet been peer-reviewed.
-
-}
-
 \abstract{%
@@
 Longitudinal trends suggested that norms and behavior were tightly coupled, changing quickly in response to public health recommendations.
-In addition, longitudinal modelling revealed that descriptive norms, not injunctive norms, caused future increases in mask wearing.
-During uncertain times, cooperative behavior is driven by what others are actually doing, rather than what others think ought to be done.
+In addition, longitudinal modeling revealed that descriptive norms caused future increases in mask wearing across multiple waves of data collection.
+These cross-lagged causal effects of descriptive norms were large, even after controlling for non-social beliefs and demographic variables.
+Injunctive norms, by contrast, had less frequent and generally weaker causal effects on future mask wearing.
+During uncertain times, cooperative behavior is more strongly driven by what others are actually doing, rather than what others think ought to be done.
 }
@@
 However, less is known about how social norms arise endogenously within populations in the short term.
-While researchers have simulated the emergence of descriptive norms\textsuperscript{11,12}, this modelling work does not capture how descriptive norms develop alongside injunctive norms in real-world settings.
+While researchers have simulated the emergence of descriptive norms\textsuperscript{11,12}, this modeling work does not capture how descriptive norms develop alongside injunctive norms in real-world settings.
 Recent work in behavioral economics has also suggested that injunctive norms of public good provisioning develop in tandem with cooperative behavior through repeated interactions\textsuperscript{13}.
@@
 For example, willingness to recycle might be driven by perceptions that recycling has a positive impact on the environment and/or personal beliefs that recycling is the right thing to do, even if social norms actively discourage recycling (e.g., recycling is not a common or socially approved behavior).
-It is thus important to control for non-social beliefs in studies of social norms.
+It is thus important to control for non-social beliefs in studies of social norms, especially considering that personal norms have previously been shown to influence prosocial behavior\textsuperscript{18,19}.
 Second, studies have tended to follow cross-sectional experimental designs in which social norm perceptions are manipulated by the researchers.
-However, social norms are not static: they change dynamically over time through processes of deliberation and interaction\textsuperscript{18}.
+However, social norms are not static: they change dynamically over time through processes of deliberation and interaction\textsuperscript{20}.
 An alternative but underutilized method of assessing causality between social norms and cooperative behavior, while retaining ecological validity, is to follow these variables over time amidst a real, unfolding social dilemma.
@@
 To understand how social norms emerge over time and shape cooperative behavior in a non-experimental setting, we focus on mask wearing in the United States during the COVID-19 pandemic.
-In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease\textsuperscript{19}.
-Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread\textsuperscript{20}.
+In April 2020, one month after the World Health Organization declared COVID-19 a global pandemic, mask wearing was officially recommended by the Centers for Disease Control and Prevention (CDC) to minimize the spread of the disease\textsuperscript{21}.
+Mask wearing has individual benefits, but the CDC also emphasized the collective benefits in reducing disease spread\textsuperscript{22}.
 Indeed, mask wearing posed a social dilemma to many individuals, in that it imposed personal costs (e.g., difficulty breathing, disrupted social interaction) for the benefit of the community (e.g., ``flattening the curve'' to protect at-risk individuals).
@@
 
-Recent research has found positive relationships between social norms and protective COVID-19 behaviors.
-In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure\textsuperscript{21}, and another study found that experimentally-induced descriptive norms increased mask wearing intentions\textsuperscript{22}.
-In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing\textsuperscript{23}.
-These studies are telling, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
-Furthermore, several of the studies do not control for political ideology, which is important to account for since COVID-19 was highly politicized in the United States\textsuperscript{24}.
-
-Here, we use two years of data from a representative sample of adults in the United States (18 time points; \emph{n} = 916) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
+Recent research has studied the relationships between social norms and protective COVID-19 behaviors.
+In the United States, one study found that perceptions of injunctive norms positively predicted intentions to stay at home to minimize exposure\textsuperscript{23}, and another study found that experimentally-induced descriptive norms increased mask wearing intentions\textsuperscript{24}.
+In Italy, an experimental study found no effect of messages highlighting descriptive norms, injunctive norms, or personal norms on time spent reading information about COVID-19 governmental rules\textsuperscript{25}.
+In Germany, a two-wave study found that perceptions of descriptive norms positively predicted future protective behaviors, such as physical distancing\textsuperscript{26}.
+These studies are informative, but since they are cross-sectional or only minimally longitudinal, they do not have the temporal granularity to capture fluctuating changes in norm strength and adherence across the pandemic.
+Furthermore, several of the studies do not control for potential confounding variables, such as demographics and political ideology.
+These variables are important to account for as they have previously been shown to be related to COVID-19 attitudes and behaviors\textsuperscript{27,28}.
+
+Here, we use two years of data from a representative sample of adults in the United States (18 time points; \emph{n} = 915) to track the development of descriptive and injunctive mask wearing norms and mask wearing behavior over the course of the COVID-19 pandemic.
 Participants reported their frequency of mask wearing during in-person interactions, as well as their perceptions of descriptive and injunctive mask wearing norms.
-We also asked participants about their non-social mask wearing beliefs and political ideology, and controlled for these factors.
+We also asked participants about their non-social mask wearing beliefs, demographics, and political ideology, and controlled for these factors.
 We used these longitudinal data to answer two main research questions in a specific real-world context.
@@
 
-To understand how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration.
+To answer our first research question about the emergence of mask wearing norms, we first visualized the average descriptive trends of self-reported norm perceptions across the entire study duration, showing how mask wearing social norms emerged and fluctuated over the course of the COVID-19 pandemic.
 Figure \ref{fig:plotTimeline} plots self-reported mask wearing and perceptions of descriptive and injunctive mask wearing norms alongside relevant pandemic-related events in the United States, such as CDC public health recommendations and COVID-19 case numbers.
-These events were obtained from the CDC Museum's COVID-19 Timeline\textsuperscript{19}.
+These events were obtained from the CDC Museum's COVID-19 Timeline\textsuperscript{21}.
 
@@
 
-Sample averages can provide informative trends, but they do not allow us to estimate within-person changes in mask wearing and perceived social norms over time.
-To determine whether within-person changes in social norms temporally preceded within-person changes in mask wearing, we fitted a ten-wave unconstrained random-intercept cross-lagged panel model to the longitudinal data.
-This structural equation model separately estimated stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and time-variant control variables (factual beliefs and personal normative beliefs).
-In line with our proposed causal model (Supplementary Figure \ref{fig:plotDAG}), we also included political orientation as an exogenous time-invariant control.
-According to established fit statistics, this model fitted the data well (root mean square error of approximation = 0.030, 95\% CI {[}0.028 0.033{]}; standardized root mean squared residual = 0.087; comparative fit index = 0.957).
-Since we are primarily interested in the causal effects of social norms on behavior, in what follows we focus on the results for mask wearing, perceived descriptive norms, and perceived injunctive norms (but see Supplementary Table \ref{tab:lavaanTable} for full list of estimated autoregressive and cross-lagged effects).
-
-Regarding between-person individual differences, the covariances between the random intercepts in the model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
-On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (\emph{r} = 0.19, 95\% CI {[}0.04 0.33{]}, \emph{p} = .019) and stronger injunctive mask wearing norms (\emph{r} = 0.27, 95\% CI {[}0.14 0.40{]}, \emph{p} \textless{} .001).
-Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (\emph{r} = 0.71, 95\% CI {[}0.65 0.78{]}, \emph{p} \textless{} .001).
-
-Regarding within-person dynamics over time, Figure \ref{fig:plotRICLPM} displays autoregressive and cross-lagged effects for perceived descriptive norms, perceived injunctive norms, and mask wearing across the study duration, controlling for non-social beliefs and political orientation.
-In random intercept cross-lagged panel models, autoregressive effects represent ``persistence'' or ``inertia'' in within-person fluctuations from stable trait levels.
+Sample averages can provide informative trends, but they do not allow us to determine whether within-person changes in social norms caused future within-person changes in mask wearing over time.
+To answer our second research question about causal effects, we fitted a ten-wave random-intercept cross-lagged panel model\textsuperscript{29,30} to the longitudinal data.
+This model separately estimates stable trait-like between-person individual differences and within-person fluctuations from those trait levels for our main variables (self-reported mask wearing, perceived descriptive mask wearing norms, and perceived injunctive mask wearing norms) and control variables (factual beliefs and personal normative beliefs).
+In line with our proposed causal model (Supplementary Figure \ref{fig:plotDAG}), we also control for potential confounding in this model by including demographics (gender, age, ethnicity, socioeconomic status) and political orientation as exogenous controls.
+
+Random-intercept cross-lagged panel models capture within-person changes over time with autoregressive and cross-lagged effects.
+Autoregressive effects represent ``persistence'' or ``inertia'' in within-person fluctuations from stable trait levels.
 In other words, a positive autoregressive effect indicates that being higher than average on one measure predicts being higher than average on that same measure in the following time point (this is not to be confused with the ``stable trait level'' over time, which is captured by the random intercepts in our model).
@@
 Cross-lagged effects are thus used to infer within-person causal influences over time.
+In what follows, we focus on the within-person autoregressive and cross-lagged effects for mask wearing and perceived social norms (see Supplementary Results for between-person results).
+
+We first fitted a time-invariant model, which constrained the autoregressive effects, cross-lagged effects, covariances, and variances to equality over time.
+This model assumes that the relationships between variables are identical across all time points, giving us a sense of the average causal effects of social norms on mask wearing over time.
+According to established fit statistics, the time-invariant model fitted the data reasonably well (root mean square error of approximation (RMSEA) = 0.038, 95\% CI {[}0.036 0.040{]}; standardized root mean squared residual (SRMR) = 0.093; comparative fit index (CFI) = 0.906).
+Table \ref{tab:constrainedTable1} summarizes the autoregressive and cross-lagged parameters from the time-invariant model.
+Autoregressive effects were significantly positive for mask wearing, perceived descriptive norms, and perceived injunctive norms, indicating that being higher than average on these variables at time \(t\) generally predicted being higher than average on the same variables at time \(t + 1\).
+Crucially, cross-lagged effects from the time-invariant model revealed that perceived descriptive norms predicted future mask wearing (unstandardized \emph{b} = 0.12, 95\% CI {[}0.06 0.18{]}, \emph{p} \textless{} .001) while perceived injunctive norms did not (\emph{b} = 0.00, 95\% CI {[}-0.05 0.06{]}, \emph{p} = .893).
+In other words, perceiving descriptive norms as stronger than average at time \(t\) generally predicted wearing masks more frequently than average at time \(t + 1\), while no such effect existed for injunctive norms.
+All other cross-lagged effects between mask wearing, perceived descriptive norms, and perceived injunctive norms were significantly positive in the time-invariant model.
+This general pattern of results was unchanged when removing factual beliefs and personal normative beliefs from the time-invariant model (Supplementary Table \ref{tab:constrainedTable2}) and when removing factual beliefs, personal normative beliefs, and all exogenous covariates from the time-invariant model (Supplementary Table \ref{tab:constrainedTable3}).
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:constrainedTable1}Unstandardized autoregressive and cross-lagged parameters from time-invariant random-intercept cross-lagged panel model.
+\emph{Arrows indicate the direction of prediction.
+Note that the effects of factual beliefs and personal normative beliefs are omitted from this table for clarity.
+SE = standard error.}}
+
+\begin{tabular}{lccccc}
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+Mask wearing → Mask wearing & 0.21 & 0.03 & 0.16 & 0.26 & 0.00\\
+Mask wearing → Injunctive norms & 0.04 & 0.02 & 0.00 & 0.07 & 0.04\\
+Mask wearing → Descriptive norms & 0.05 & 0.02 & 0.01 & 0.08 & 0.01\\
+Injunctive norms → Mask wearing & 0.00 & 0.03 & -0.05 & 0.06 & 0.89\\
+Injunctive norms → Injunctive norms & 0.27 & 0.03 & 0.21 & 0.32 & 0.00\\
+Injunctive norms → Descriptive norms & 0.12 & 0.02 & 0.07 & 0.17 & 0.00\\
+Descriptive norms → Mask wearing & 0.12 & 0.03 & 0.06 & 0.18 & 0.00\\
+Descriptive norms → Injunctive norms & 0.17 & 0.02 & 0.13 & 0.22 & 0.00\\
+Descriptive norms → Descriptive norms & 0.32 & 0.03 & 0.27 & 0.37 & 0.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+Given that the strength of perceived social norms varied throughout our data collection window (Figure \ref{fig:plotTimeline}), it is plausible that the causal effects of social norms on mask wearing may have changed over time as well, rather than being identical at each time point.
+To test whether the equality constraints over time were tenable\textsuperscript{30}, we compared the time-invariant model to an alternative time-varying model that freely estimated the autoregressive effects, cross-lagged effects, covariances, and variances at each time point.
+This model assumes that the relationships between variables are different across different time points, allowing the causal effects of social norms on future mask wearing to vary over time.
+The time-varying model fitted the data better than the time-invariant model (RMSEA = 0.031, 95\% CI {[}0.029 0.033{]}; SRMR = 0.073; CFI = 0.955).
+Model comparison revealed that freely estimating the parameters over time resulted in improved model fit (\(\Delta\)AIC = -483.09, \(\Delta\chi^2\)(320) = -1123.09, \emph{p} \textless{} .001) suggesting that there was substantial variability in the relationships between mask wearing and social norms across the course of the pandemic.
+Accordingly, we now turn to the results from the time-varying model.
+
+Figure \ref{fig:plotRICLPMPath1} displays the autoregressive and cross-lagged effects for social norms and mask wearing from the time-varying model (see Supplementary Table \ref{tab:unconstrainedTable1} for full list of estimated autoregressive and cross-lagged effects).
+In late 2020 and 2021, we find three occasions where within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
+According to recent effect size guidelines for cross-lagged panel models\textsuperscript{31}, the standardized coefficients for these cross-lagged effects were large (first wave, standardized \(\beta\) = 0.16, \emph{b} = 0.20, 95\% CI {[}0.06 0.34{]}, \emph{p} = .006; second wave, \(\beta\) = 0.21, \emph{b} = 0.26, 95\% CI {[}0.10 0.42{]}, \emph{p} = .002; fifth wave, \(\beta\) = 0.15, \emph{b} = 0.15, 95\% CI {[}0.01 0.29{]}, \emph{p} = .033).
+At other time points, the cross-lagged effects for descriptive norms tended to be estimated in a positive direction, though these estimates were not statistically significant and decreased in magnitude in 2022 (Figure \ref{fig:plotRICLPMCoef1}).
+Time-varying models without covariates revealed an additional significant cross-lagged effect of descriptive norms on future mask wearing at the fourth wave, but otherwise the general pattern of results for descriptive norms was unchanged (Supplementary Figures \ref{fig:plotRICLPMPath2} - \ref{fig:plotRICLPMCoef3}; Supplementary Tables \ref{tab:unconstrainedTable2} and \ref{tab:unconstrainedTable3}).
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/plotRICLPM-1.pdf}
-\caption{\label{fig:plotRICLPM}\emph{Results of ten-wave unconstrained random-intercept cross-lagged panel model.} Circles represent data collection time points.
-Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, and political orientation.
+\includegraphics{manuscript_files/figure-latex/plotRICLPMPath1-1.pdf}
+\caption{\label{fig:plotRICLPMPath1}\emph{Path diagram of ten-wave time-varying random-intercept cross-lagged panel model.} Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for factual beliefs, personal normative beliefs, demographics, and political orientation.
 Arrow thickness is scaled according to standardized effect size.
 Bolded arrows indicate significantly positive parameters, \emph{p} \textless{} 0.05.
-Gray arrows indicate non-significant parameters.
-There are no significant direct paths from injunctive norms to future mask wearing, showing that people's beliefs about what they should be doing did not have any direct influences on future mask wearing.
-On the other hand, there are significant paths from descriptive norms to future mask wearing, meaning that people's beliefs about what others are doing influenced their future mask wearing.}
+Gray arrows indicate non-significant parameters.}
 \end{figure}
 
-In late 2020 and throughout 2021, we see several cross-lagged effects from perceived descriptive norms to future mask wearing.
-On four occasions, within-person increases in perceived descriptive norms predicted future within-person increases in mask wearing.
-According to recent effect size guidelines for cross-lagged panel models\textsuperscript{25}, the standardized beta coefficients for these cross-lagged effects were large (first wave, \(\beta\) = 0.17, 95\% CI {[}0.06 0.28{]}, \emph{p} = .002; second wave, \(\beta\) = 0.21, 95\% CI {[}0.08 0.34{]}, \emph{p} = .001; fourth wave, \(\beta\) = 0.15, 95\% CI {[}0.01 0.30{]}, \emph{p} = .041; fifth wave, \(\beta\) = 0.16, 95\% CI {[}0.02 0.29{]}, \emph{p} = .023).
-In 2022, these cross-lagged effects from descriptive norms to mask wearing diminished.
-We also find some evidence for a reciprocal effect, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive norms.
-Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person causal effects between these variables.
-
-However, the model also reveals that, after controlling for perceived descriptive norms, non-social beliefs, and political orientation, within-person changes in perceived injunctive norms did \emph{not} predict future within-person changes in mask wearing across our study duration.
-All cross-lagged effects from perceived injunctive norms to mask wearing are non-significant.
-Any causal effect that perceived injunctive norms might have had on future mask wearing appears to be fully mediated by perceived descriptive norms.
-This means that believing that others think that mask wearing is the right thing to do influences one's perception of what others are actually doing, which then influences future behavior.
-For example, between August 2021 and December 2021, perceived injunctive norms predicted future perceived descriptive norms, which themselves predicted future mask wearing.
-But aside from these indirect effects, perceived injunctive norms did not have a direct causal effect on mask wearing over time within individuals.
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotRICLPMCoef1-1.pdf}
+\caption{\label{fig:plotRICLPMCoef1}\emph{Standardized cross-lagged coefficients for descriptive norms and injunctive norms predicting future mask wearing in the ten-wave time-varying random-intercept cross-lagged panel model.} Points are standardized estimates, lines are 95\% confidence intervals.}
+\end{figure}
+
+In contrast, within-person increases in perceived injunctive norms only predicted future within-person increases in mask wearing at a single time point in March 2022 (seventh wave, \(\beta\) = 0.17, \emph{b} = 0.20, 95\% CI {[}0.03 0.37{]}, \emph{p} = .018).
+This cross-lagged effect was robust to the removal of covariates (Supplementary Figures \ref{fig:plotRICLPMPath2} - \ref{fig:plotRICLPMCoef3}; Supplementary Tables \ref{tab:unconstrainedTable2} and \ref{tab:unconstrainedTable3}).
+At all other time points, the cross-lagged effects of injunctive norms on future mask wearing were non-significant and estimated to be generally weaker than the cross-lagged effects of descriptive norms (Figure \ref{fig:plotRICLPMCoef1}), explaining the lack of effect of injunctive norms in the time-invariant model.
+
+One possible explanation for the generally weaker causal influence of injunctive norms on future mask wearing is that injunctive norms might only have been sufficiently salient in regions of the United States with stronger governmental enforcement of mask wearing.
+Previous work has shown that, during the pandemic, states with Democratic leadership tended to have more stringent mask wearing policies than states with Republican leadership\textsuperscript{32}.
+It is thus plausible that injunctive norms had a stronger causal effect on future mask wearing in majority-Democrat states compared to majority-Republican states.
+To test this explanation, we fitted a multi-group time-invariant random-intercept cross-lagged panel model to the data, with separate groups for majority-Democrat and majority-Republican states.
+However, contrary to the salience account, perceived injunctive norms failed to predict future mask wearing in both Democrat states and Republican states (see Supplementary Results).
+
+Beyond the effects of descriptive and injunctive norms on future mask wearing, our main time-varying model revealed other interesting patterns (Figure \ref{fig:plotRICLPMPath1}).
+On a few occasions, we find evidence for a reciprocal relationship between social norms and mask wearing, whereby within-person increases in mask wearing predicted future within-person increases in perceived descriptive and injunctive norms.
+Moreover, several cross-lagged effects emerged between perceived descriptive and injunctive norms, demonstrating reciprocal within-person influences between these variables as the pandemic unfolded.
 
@@
 Moreover, the results of our cross-lagged panel model indicate that descriptive norms caused future increases in mask wearing in the first year and a half of the pandemic.
-By contrast, injunctive norms were not directly causally related to future mask wearing over the entire two years of data collection.
+By contrast, injunctive norms were less frequently related to future mask wearing during the pandemic, with generally weaker effects than descriptive norms.
 
 Our finding that social norms and mask wearing are tightly coupled over time provides real-world support for experimental evidence that social norms and cooperative behavior develop synergistically within groups via processes of social interaction\textsuperscript{13}.
-The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed\textsuperscript{26}.
-Indeed, empirical work in cultural evolution suggests that formal institutions are critical for the emergence and rapid adoption of novel social norms\textsuperscript{27}.
+The fact that these changes closely tracked the release of guidelines by the CDC supports the idea that institutions are part of the process by which culture and one's own behaviors are mutually constructed\textsuperscript{33}.
+Indeed, previous work has shown that formal institutions are critical for the emergence and rapid adoption of novel social norms\textsuperscript{34}.
 While new norms can and do emerge spontaneously in populations, the process is slow compared to institution-driven norm change, which, as our trends have shown, can unfold over measurement intervals as short as four to six weeks.
 
-We found that descriptive norms, not injunctive norms, predicted future within-person increases in mask wearing, independent of the effects of non-social beliefs and political orientation.
-This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors\textsuperscript{22,23}.
+We found that descriptive norms predicted future within-person increases in mask wearing, independent of the effects of injunctive norms, non-social beliefs, and demographic variables.
+This finding is in line with previous evidence showing that perceptions of descriptive norms were positively related to other protective COVID-19 behaviors\textsuperscript{24,26}.
 There are several explanations for why descriptive norms have had these positive effects on protective COVID-19 behaviors like mask wearing.
 First, people may have followed descriptive norms to quickly coordinate their behavior with others during the pandemic.
-Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic\textsuperscript{28}.
-Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population\textsuperscript{29}.
+Descriptive norms are particularly useful for coordinating behavior during fast changing, threatening situations with a high degree of uncertainty, such as the COVID-19 pandemic\textsuperscript{35}.
+This may help to explain why the effects of descriptive norms were more prevalent in earlier waves of our data, when uncertainty was highest.
+Second, people might have engaged in conditional cooperation, adapting their cooperation levels to the degree of cooperation in the population\textsuperscript{36}.
 Descriptive mask wearing norms provide evidence that others are cooperating, increasing the likelihood that individuals will themselves contribute to the public good by wearing masks.
-Third, the increased frequency of mask wearing in the population might have created a bandwagon effect\textsuperscript{30}, encouraging conformist copying.
-Under this view, people wear masks not to coordinate or cooperate, but simply to fit in with the crowd.
+Third, the increased frequency of mask wearing in the population might have created a bandwagon effect\textsuperscript{37}, encouraging conformist copying.
+Under this view, people wear masks not to coordinate or cooperate, but simply because they see a majority of others engaging in the behavior.
 Future research will be required to determine the motivations underlying adherence to descriptive norms during uncertain times.
 
-We found that perceived injunctive norms did not directly predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing are not directly causally related.
+On the whole, we found that perceived injunctive norms tended not to predict future within-person increases in mask wearing, suggesting that injunctive norms and mask wearing were not strongly causally related during our data collection window.
 One possible explanation for this result is that, due to the increased opportunities to observe mask wearing in public, descriptive norms of mask wearing were more salient than injunctive norms during the pandemic.
@@
 To test this idea, future research should expand our longitudinal approach to protective behaviors beyond mask wearing, including both public behaviors (e.g., physical distancing) and private behaviors (e.g., hand washing and home isolation).
-
-We are limited in generalizing these findings due to the constraints of our sample and the variables considered.
-While the attrition in our study did not substantially affect the representativeness of our sample (Supplementary Figure \ref{fig:plotAttritionBreakdown}),
-the reduction in our sample size did not leave us with enough data to test the robustness of our results within different identity groups, such as different genders, ethnicities, or political ideologies.
-Our results also might not generalize to all social norms, behaviors, and social dilemmas.
+Nevertheless, it is worth noting that injunctive norms continued to have no overall causal effect on mask wearing when we focused specifically on majority-Democrat US states that tended to employ state-wide mask mandates\textsuperscript{32}, suggesting that a lack of saliency cannot entirely explain why injunctive norms did not have a consistent effect on mask wearing during the pandemic.
+
+Regarding injunctive mask wearing norms, several open questions remain.
+First, given that we see one time point in the later stages of the pandemic in which injunctive norms directly predicted future mask wearing, it is possible that there were also earlier time points where injunctive norms had effects on behavior.
+We do not have data from the very beginning of the pandemic (March-September 2020) to directly test this, but future work could test this by examining the effects of injunctive norms on behaviour directly after the onset of a crisis.
+Second, it remains unclear how the source of the injunctive norm influences its efficacy.
+Our operationalization of injunctive norms referred in part to the behavior of respected others, and so the influence of the perceived approval of prestigious political and nationwide organizational leaders may be an important lever to test.
+In addition, given that we operationalized injunctive norms in part by local-level encouragement, this might explain why we see the only significant effect of injunctive norms in February 2022, as by this point the CDC was easing their nationwide mask guidelines.
+At this stage, people may have begun to look more to injunctive information from their local areas for encouragement on what was appropriate.
+Third, our longitudinal modelling showed that changes in perceived descriptive norms consistently influenced future changes in perceived injunctive norms throughout our study period, while the reverse was only true in the latter half of our study period.
+This suggests that initially changing empirical expectations of others' behavior could potentially motivate compliance and shift later perceptions of injunctive norms\textsuperscript{38}.
+Although our model was not specifically designed to address this question, future work should follow up on the temporal relationship between descriptive and injunctive norms with more fine-grained longitudinal data, perhaps experimentally intervening on perceptions of descriptive or injunctive norms to infer directions of causation.
+
+It is unclear the extent to which our pattern of results might generalize to other cultural contexts.
+The United States is relatively close to the global average on cultural tightness-looseness, a cross-cultural dimension that measures nationwide strength of social norms\textsuperscript{39,40}.
+We might therefore expect to see similar effects of social norms on mask wearing in countries with similar scores on this dimension.
+However, for countries with higher levels of cultural tightness than the United States, we might expect to see stronger effects of descriptive norms on mask wearing, additional effects of injunctive norms on mask wearing, and different dynamics between descriptive and injunctive norms.
+If borne out by the data, such patterns might explain why countries with higher levels of cultural tightness had fewer COVID-19 cases and deaths compared to culturally looser countries\textsuperscript{40}.
+To test this, future research could expand our longitudinal approach to other cultural contexts.
+
+Our results might not generalize to all social norms, behaviors, and social dilemmas.
 Norms governing sustainability in response to climate change, for example, might take longer to emerge, since the threat of climate change is more remote than the COVID-19 pandemic.
 For more distant social dilemmas that do not cause immediate day-to-day uncertainty, descriptive social norms may not necessarily drive cooperative behavior.
-
-For the case of mask wearing in the United States during the COVID-19 pandemic, we have shown that social norms developed rapidly in the population and tracked ongoing changes in both recommendations from authorities and current levels of cooperative behavior.
+Mask wearing is also a unique cooperative behavior in that it is not ``purely'' cooperative (i.e., imposing costs on the actor while providing benefits to targets).
+Much of past research on normative influences on prosocial behavior has focused on these purely unselfish cooperative behaviors\textsuperscript{19}.
+Mask wearing is different as it does have individual benefits (e.g., reduced likelihood of contracting the disease).
+However, mask wearing also often imposes non-trivial costs on individuals (e.g., difficulty breathing, discomfort, disrupted social interaction) for the benefit of the wider group, meaning that it is still useful to conceptualise it as a prosocial or cooperative behavior.
+With these considerations, our results may be most pertinent for easily observable behaviors in response to an immediate or short-term threat that offer both benefits and costs to the individual in service to the wider group (e.g., other protective health behaviors, joining a protest or strike).
+
+There are some limitations associated with our longitudinal survey design.
+First, we asked participants to self-report their frequency of mask wearing.
+However, self-report measures may be biased by anchoring effects and participants' imperfect recall of their own behavior, resulting in measurement error\textsuperscript{41}.
+Such measures were necessary to implement our two-year longitudinal study, but future field studies of mask wearing norms could avoid this pitfall by directly measuring mask wearing through naturalistic observation.
+Second, as is common in longitudinal designs, there was substantial attrition over the course of the study (Supplementary Figure \ref{fig:plotAttrition}).
+While this attrition did not substantially affect the demographic representativeness of our sample (Supplementary Figure \ref{fig:plotAttritionBreakdown}), the sample may have been self-selected in other unknown ways that introduced confounds.
+Future longitudinal work could encourage higher retention rates with more frequent reminders and fewer survey questions to maintain attention and interest.
+
+Despite these caveats, we have shown that mask wearing norms developed rapidly in the United States population during the COVID-19 pandemic and tracked ongoing changes in both recommendations from authorities and current levels of mask wearing behavior.
 Moreover, we found that descriptive norms, rather than injunctive norms, were the main driver for future mask wearing.
@@
 Not only does it imply that high local levels of mask wearing encouraged future personal mask use, but it also implies that \emph{low} local levels of mask wearing \emph{discouraged} future personal mask use.
-This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community\textsuperscript{31}.
-Our work thus underscores the importance of consistent, visible community adherence for encouraging protective behaviors in response to global pandemics like COVID-19.
+This echoes recent reports of people in the United States not wanting to be ``singled out'' by being the only one wearing a mask in their community\textsuperscript{42}.
+Organizations interested in combating such backfire effects might consider making masks easily and cheaply available in local settings (e.g., within smaller institutions or at events) and running media campaigns to show that mask wearing is common among ``everyday people'' doing ``everyday things''.
+Once tested in target settings\textsuperscript{43}, such interventions could leverage the power of consistent, visible community adherence to encourage protective behaviors in response to global pandemics like COVID-19.
 
@@
 Using the platform Prolific (\url{https://www.prolific.co/}), we distributed surveys to a representative sample of adults from the United States (\emph{n} = 915, \emph{M}\textsubscript{age} = 46 years, 75\% White, 52\% Women; see Supplementary Figure \ref{fig:plotUSMap} for geographic distribution).
-From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors.
+From September 2020 to October 2022, we asked participants to complete regular surveys of COVID-19 related attitudes and behaviors (for a visualization of all questions asked of this sample, see \url{https://navigateobscurity.com/cooperation-conflict-lab/us-variable-lookup}).
 This resulted in 18 unique time points of data collection during the pandemic.
@@
 However, this attrition did not substantially affect the demographic makeup of the sample through time (Supplementary Figure \ref{fig:plotAttritionBreakdown}).
+On average, participants were paid approximately \$8 USD per hour for completing the surveys.
 
@@
 
-At every time point, participants were asked about the number of in-person interactions they had in the last week.
-Following this question, participants self-reported their mask wearing by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' Participants responded on a 5-point Likert scale, from Never (1) to Always (5).
+At every time point, participants were asked about the number of in-person interactions they had in the last 7 days and the last 24 hours.
+These two separate items specified that the interactions could have been either recreational or routine: ``\emph{How many people (outside of your household) have you had in-person interactions with during the last {[}7 days / 24 hours{]} (e.g., visiting with friends, buying food or supplies from a cashier)?}'' Following these questions, participants self-reported their mask wearing by answering: ``\emph{During these in-person interactions, if you were closer than 6 feet (2 meters) from the person(s) did you wear a face mask?}'' This question was asked regardless of how many in-person interactions the participants reported in the two interaction questions.
+Participants responded to the mask wearing question on a 5-point Likert scale, from Never (1) to Always (5).
+Responses to this question were weakly negatively associated with the number of in-person interactions that participants reported in the last 7 days and the last 24 hours, since wearing masks and social distancing are both COVID-19 protective behaviors (Supplementary Figure \ref{fig:plotIntFreq}).
 
@@
 
-To check the construct validity of these measures, at time point 7 we asked participants about their interpretations of the four social norm items above.
+There is potential overlap between these operationalizations of descriptive and injunctive norms.
+For example, the item asking how often participants see people that they respect and trust wearing masks could be capturing descriptive norms as well as injunctive norms.
+To address this concern about the construct validity of the four social norm items, at time point 7 we asked participants about their interpretations of the items.
 We asked participants whether each of the four items informed them about what people \emph{are} doing or what people \emph{should} be doing (i.e., giving descriptive or injunctive information).
-Participants were able to correctly distinguish between the two sets of items, suggesting that they are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Table \ref{tab:itemTable}).
+We found that participants rated the two descriptive norm questions as providing more descriptive information than the two injunctive norm questions, and vice versa, suggesting that the items are valid measures of perceived descriptive and injunctive social norms (see Supplementary Results and Supplementary Tables \ref{tab:itemTable} and \ref{tab:constructValTable}).
 
@@
 
-To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (see Supplementary Figure \ref{fig:plotDAG}).
+To identify direct causal effects in our longitudinal analysis, we constructed a directed acyclic causal graph outlining the expected causal relationships between our variables (Supplementary Figure \ref{fig:plotDAG}).
 In this causal model, we included two kinds of non-social beliefs highlighted by previous research\textsuperscript{17}: factual beliefs (i.e., beliefs about the effectiveness or consequences of mask wearing) and personal normative beliefs (i.e., personal beliefs about whether mask wearing is the right thing to do).
 These variables were included as potential mediators of the effects of descriptive and injunctive social norms on mask wearing.
-In addition, we also included political orientation as a common cause of all other variables.
-This is justified by evidence showing that mask wearing was heavily politicized in the United States during the pandemic\textsuperscript{24}.
-Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and political orientation in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
+In addition, we also included demographics (gender, age, ethnicity, socioeconomic status, and political orientation) as common causes of all other variables.
+This is justified by evidence showing that these variables are associated with COVID-19 attitudes and behaviors\textsuperscript{27,28}.
+Given this causal graph, it is necessary to control for factual beliefs, personal normative beliefs, and all demographic variables in order to estimate the direct causal effects of descriptive and injunctive norms on mask wearing behavior over time.
 
@@
 
-Political orientation was measured in the first time point only.
-We measured political orientation as the mean of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
+All demographic variables were measured at the first time point only.
+Gender was measured using a question that asked about participants' biological sex (``\emph{What is your sex?}'') with three possible categories (Male, Female, or Other).
+We are assuming within these analyses that participants identify as the gender associated with their biological sex.
+Age was measured numerically (``\emph{What is your age?}'').
+Ethnicity was measured with a single item (``\emph{What is your ethnicity?}'') with six possible categories (Asian or Pacific Islander, Hispanic or Latino/a, White or Caucasian, Black or African American, Native American, or Other).
+Socioeconomic status was operationalized through a composite measure averaging income, education level, and subjective socioeconomic status.
+Income was captured by a single item measured on a 9-point scale of increasing income amounts (``\emph{What was your combined household income in the previous year before taxes?}''), education was captured by a single item measured on a 7-point scale of increasing education levels (``\emph{What is the highest level of education that you have completed?}''), and subjective socioeconomic status was represented by a ladder with ten rungs, from those who are ``worst off'' at the bottom to those who are ``best off'' at the top.
+Political orientation was measured as the mean of the following two items: ``\emph{How would you describe your political orientation with regard to social issues?}'' and ``\emph{How would you describe your political orientation with regard to economic issues?}''.
 These items were measured on 7-point Likert scales, from Very Liberal (1) to Very Conservative (7).
@@
 These models included random intercepts and slopes for participants, as well as change points aligning with changes in CDC mask wearing recommendations.
-We estimated these multilevel regression models using the \emph{lme4} R package\textsuperscript{32} and dealt with missing data via listwise deletion.
-
-To quantify the within-person relationships between our variables over time, we fitted a random-intercept cross-lagged panel model to our longitudinal data\textsuperscript{33}.
+We estimated these multilevel regression models using the \emph{lme4} R package\textsuperscript{44} and dealt with missing data via listwise deletion.
+
+To quantify the within-person relationships between our variables over time, we fitted random-intercept cross-lagged panel models to our longitudinal data\textsuperscript{29,30}.
 This structural equation model distinguishes between stable between-person trait levels and within-person fluctuations from trait levels.
-Positive cross-lagged effects from this model indicate that being above average on one variable at time \emph{t-1} predicts being above average in another variable at time \emph{t}.
-These models are considered the gold standard for identifying Granger causality in longitudinal datasets\textsuperscript{33,34}.
-
-We estimated the random-intercept cross-lagged panel model using the \emph{lavaan} R package\textsuperscript{35}.
-In line with our directed acyclic graph (see Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two time-variant control variables (factual beliefs and personal normative beliefs) in the model.
+Positive cross-lagged effects from this model indicate that being above average on one variable at time \(t\) predicts being above average in another variable at time \(t + 1\).
+These models are considered the gold standard for identifying Granger causality in longitudinal datasets\textsuperscript{29,30,45}.
+
+We estimated the random-intercept cross-lagged panel models using the \emph{lavaan} R package\textsuperscript{46}.
+In line with our directed acyclic graph (Supplementary Figure \ref{fig:plotDAG}), we included three main variables (self-reported mask wearing, perceived descriptive norms, and perceived injunctive norms) and two control variables (factual beliefs and personal normative beliefs) in the model.
 For each of these variables, the model estimated a stable between-person trait level (random intercept) and time-specific within-person fluctuations from this trait level.
-We modeled autoregressive and cross-lagged effects between all five variables, and included political orientation as a time-invariant covariate.
+We modeled autoregressive and cross-lagged effects between all five variables, and included gender, age, ethnicity, socioeconomic status, and political ideology as exogenous covariates.
 We restricted the analysis to the ten time points with available data for all five variables.
-Full information maximum likelihood estimation was used to deal with missing data.
-
-All analyses were conducted in R v4.1.1\textsuperscript{36}.
-Visualizations were generated using the \emph{cowplot}\textsuperscript{37} and \emph{ggplot2}\textsuperscript{38} packages.
-The manuscript was reproducibly generated using the \emph{targets}\textsuperscript{39} and \emph{papaja}\textsuperscript{40} packages.
-All code and data are publicly available on GitHub\textsuperscript{41}.
-
-\newpage
+We fitted both a time-invariant model (i.e., a model that constrained the within-person autoregressive effects, cross-lagged effects, covariances, and variance to equality over time) and a time-varying model that freely estimated all parameters.
+In both cases, full information maximum likelihood estimation was used to deal with missing data.
+
+In addition to the full random-intercept cross-lagged panel model containing all covariates, we additionally fitted (1) a model with factual beliefs and personal normative beliefs removed and (2) a model with all covariates removed.
+We fitted both time-invariant and time-varying versions of these models.
+Finally, we fitted a time-invariant multi-group model that estimated parameters separately for majority-Democrat states and majority-Republican states, operationalised using the state-level results from the 2020 United States Presidential election.
+Results for these additional models are reported in the Supplementary Material.
+
+Analyses were conducted retrospectively after data collection during the COVID-19 pandemic, meaning that no analyses were preregistered and all analyses should thus be considered exploratory.
+All analyses were conducted in R v4.1.1\textsuperscript{47}.
+Visualizations were generated using the \emph{cowplot}\textsuperscript{48} and \emph{ggplot2}\textsuperscript{49} packages.
+The manuscript was reproducibly generated using the \emph{targets}\textsuperscript{50} and \emph{papaja}\textsuperscript{51} packages.
+All code and data are publicly available on GitHub: \url{https://github.com/ScottClaessens/covidMaskWearing}
+
+\newpage
+\nolinenumbers
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014a}{}}%
 \CSLLeftMargin{17. }%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2010}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{Bicchieri, C.
+\& Chavez, A.
+\href{https://doi.org/10.1002/bdm.648}{Behaving as expected: Public information and fairness norms}.
+\emph{Journal of Behavioral Decision Making} \textbf{23}, 161--178 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Capraro2021}{}}%
+\CSLLeftMargin{19. }%
+\CSLRightInline{Capraro, V.
+\& Perc, M.
+\href{https://doi.org/10.1098/rsif.2020.0880}{Mathematical foundations of moral preferences}.
+\emph{Journal of the Royal Society Interface} \textbf{18}, 20200880 (2021).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Smith2015}{}}%
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{Smith, L.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CDC2022}{}}%
-\CSLLeftMargin{19. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{Centers for Disease Control and Prevention. \href{https://www.cdc.gov/museum/timeline/covid19.html}{{CDC Museum COVID-19 Timeline}}. (2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-CDC2020}{}}%
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{Centers for Disease Control and Prevention.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Macy2021}{}}%
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{Macy, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bokemper2021}{}}%
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{Bokemper, S.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Bilancini2020}{}}%
+\CSLLeftMargin{25. }%
+\CSLRightInline{Bilancini, E., Boncinelli, L., Capraro, V.
+\& Paolo, R.
+D.
+The effect of norm-based messages on reading and understanding COVID-19 pandemic response governmental rules.
+\emph{Journal of Behavioral Economics for Policy} \textbf{4}, 45--55 (2020).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Rudert2021}{}}%
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Rudert, S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-BaxterKing2022}{}}%
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{Baxter-King, R., Brown, J.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
-\CSLLeftMargin{25. }%
-\CSLRightInline{Orth, U. \emph{et al.} Effect size guidelines for cross-lagged effects. \emph{Psychological Methods} (2022) doi:\href{https://doi.org/10.1037/met0000499}{10.1037/met0000499}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
-\CSLLeftMargin{26. }%
-\CSLRightInline{Markus, H.
-R.
-\& Kitayama, S.
-\href{https://doi.org/10.1177/1745691610375557}{Cultures and selves: A cycle of mutual constitution}.
-\emph{Perspectives on Psychological Science} \textbf{5}, 420--430 (2010).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
-\CSLLeftMargin{27. }%
-\CSLRightInline{Amato, R., Lacasa, L., Díaz-Guilera, A.
-\& Baronchelli, A.
-\href{https://doi.org/10.1073/pnas.1721059115}{The dynamics of norm change in the cultural evolution of language}.
-\emph{Proceedings of the National Academy of Sciences} \textbf{115}, 8260--8265 (2018).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Heiman2022}{}}%
 \CSLLeftMargin{28. }%
-\CSLRightInline{Gelfand, M.
-J.
-\& Harrington, J.
-R.
-\href{https://doi.org/10.1177/0022022115600796}{The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?} \emph{Journal of Cross-Cultural Psychology} \textbf{46}, 1273--1278 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
+\CSLRightInline{Heiman, S.
+L.
+\emph{et al.} Identities as predictors of vaccine hesitancy during the COVID-19 pandemic.
+\emph{Journal of Social Issues} 1--22 (2022) doi:\href{https://doi.org/10.1111/josi.12569}{10.1111/josi.12569}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
 \CSLLeftMargin{29. }%
-\CSLRightInline{Chaudhuri, A.
-\href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
-\emph{Experimental Economics} \textbf{14}, 47--83 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-SchmittBeck2015}{}}%
-\CSLLeftMargin{30. }%
-\CSLRightInline{Schmitt-Beck, R.
-Bandwagon effect.
-in \emph{The international encyclopedia of political communication} 1--5 (John Wiley \& Sons, Ltd, 2015).
-doi:\href{https://doi.org/10.1002/9781118541555.wbiepc015}{10.1002/9781118541555.wbiepc015}.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
-\CSLLeftMargin{31. }%
-\CSLRightInline{Natanson, H.
-\href{https://www.washingtonpost.com/education/2022/02/25/peer-pressure-mask-optional-schools/}{Peer pressure is ending mask usage in schools}.
-\emph{The Washingston Post} (2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
-\CSLLeftMargin{32. }%
-\CSLRightInline{Bates, D., Mächler, M., Bolker, B.
-\& Walker, S.
-\href{https://doi.org/10.18637/jss.v067.i01}{Fitting linear mixed-effects models using lme4}.
-\emph{Journal of Statistical Software} \textbf{67}, 1--48 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hamaker2015}{}}%
-\CSLLeftMargin{33. }%
 \CSLRightInline{Hamaker, E.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Mulder2021}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{Mulder, J.
+D.
+\& Hamaker, E.
+L.
+\href{https://doi.org/10.1080/10705511.2020.1784738}{Three extensions of the random intercept cross-lagged panel model}.
+\emph{Structural Equation Modeling: A Multidisciplinary Journal} \textbf{28}, 638--648 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Orth2022}{}}%
+\CSLLeftMargin{31. }%
+\CSLRightInline{Orth, U. \emph{et al.} Effect size guidelines for cross-lagged effects. \emph{Psychological Methods} (2022) doi:\href{https://doi.org/10.1037/met0000499}{10.1037/met0000499}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Adolph2022}{}}%
+\CSLLeftMargin{32. }%
+\CSLRightInline{Adolph, C.
+\emph{et al.} \href{https://doi.org/10.1017/spq.2021.22}{Governor partisanship explains the adoption of statewide mask mandates in response to COVID-19}.
+\emph{State Politics \& Policy Quarterly} \textbf{22}, 24--49 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Markus2010}{}}%
+\CSLLeftMargin{33. }%
+\CSLRightInline{Markus, H.
+R.
+\& Kitayama, S.
+\href{https://doi.org/10.1177/1745691610375557}{Cultures and selves: A cycle of mutual constitution}.
+\emph{Perspectives on Psychological Science} \textbf{5}, 420--430 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Amato2018}{}}%
+\CSLLeftMargin{34. }%
+\CSLRightInline{Amato, R., Lacasa, L., Díaz-Guilera, A.
+\& Baronchelli, A.
+\href{https://doi.org/10.1073/pnas.1721059115}{The dynamics of norm change in the cultural evolution of language}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{115}, 8260--8265 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2015}{}}%
+\CSLLeftMargin{35. }%
+\CSLRightInline{Gelfand, M.
+J.
+\& Harrington, J.
+R.
+\href{https://doi.org/10.1177/0022022115600796}{The motivational force of descriptive norms: For whom and when are descriptive norms most predictive of behavior?} \emph{Journal of Cross-Cultural Psychology} \textbf{46}, 1273--1278 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
+\CSLLeftMargin{36. }%
+\CSLRightInline{Chaudhuri, A.
+\href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
+\emph{Experimental Economics} \textbf{14}, 47--83 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-SchmittBeck2015}{}}%
+\CSLLeftMargin{37. }%
+\CSLRightInline{Schmitt-Beck, R.
+Bandwagon effect.
+in \emph{The international encyclopedia of political communication} 1--5 (John Wiley \& Sons, Ltd, 2015).
+doi:\href{https://doi.org/10.1002/9781118541555.wbiepc015}{10.1002/9781118541555.wbiepc015}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bicchieri2014b}{}}%
+\CSLLeftMargin{38. }%
+\CSLRightInline{Bicchieri, C.
+\& Mercier, H.
+Norms and beliefs: How change occurs.
+in \emph{The complexity of social norms} (eds.
+Xenitidou, M.
+\& Edmonds, B.) 37--54 (Springer International Publishing, 2014).
+doi:\href{https://doi.org/10.1007/978-3-319-05308-0_3}{10.1007/978-3-319-05308-0\_3}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2011}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{Gelfand, M.
+J.
+\emph{et al.} \href{https://doi.org/10.1126/science.1197754}{Differences between tight and loose cultures: A 33-nation study}.
+\emph{Science} \textbf{332}, 1100--1104 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2021}{}}%
+\CSLLeftMargin{40. }%
+\CSLRightInline{Gelfand, M.
+J.
+\emph{et al.} \href{https://doi.org/10.1016/S2542-5196(20)30301-6}{The relationship between cultural tightness-looseness and {COVID-19} cases and deaths: A global analysis}.
+\emph{The Lancet Planetary Health} \textbf{5}, e135--e144 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hansen2022}{}}%
+\CSLLeftMargin{41. }%
+\CSLRightInline{Hansen, P.
+G., Larsen, E.
+G.
+\& Gundersen, C.
+D.
+\href{https://doi.org/10.1017/bpp.2021.13}{Reporting on one's behavior: A survey experiment on the nonvalidity of self-reported {COVID-19} hygiene-relevant routine behaviors}.
+\emph{Behavioural Public Policy} \textbf{6}, 34--51 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Natanson2022}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{Natanson, H.
+\href{https://www.washingtonpost.com/education/2022/02/25/peer-pressure-mask-optional-schools/}{Peer pressure is ending mask usage in schools}.
+\emph{The Washingston Post} (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-IJzerman2020}{}}%
+\CSLLeftMargin{43. }%
+\CSLRightInline{IJzerman, H.
+\emph{et al.} \href{https://doi.org/10.1038/s41562-020-00990-w}{Use caution when applying behavioural science to policy}.
+\emph{Nature Human Behaviour} \textbf{4}, 1092--1094 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bates2015}{}}%
+\CSLLeftMargin{44. }%
+\CSLRightInline{Bates, D., Mächler, M., Bolker, B.
+\& Walker, S.
+\href{https://doi.org/10.18637/jss.v067.i01}{Fitting linear mixed-effects models using lme4}.
+\emph{Journal of Statistical Software} \textbf{67}, 1--48 (2015).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Granger1969}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Granger, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Rosseel2012}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Rosseel, Y. \href{https://doi.org/10.18637/jss.v048.i02}{{lavaan}: An {R} package for structural equation modeling}. \emph{Journal of Statistical Software} \textbf{48}, 1--36 (2012).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Wilke, C. O. \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}}. (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Wickham, H. \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}}. (Springer-Verlag New York, 2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Landau, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}}. (2022).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Heiman2023}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{Heiman, S.
-L.
-\emph{et al.} \href{https://github.com/ScottClaessens/covidMaskWearing}{Descriptive, not injunctive, social norms caused increases in mask wearing during the COVID-19 pandemic}.
-(2023).}
+\leavevmode\vadjust pre{\hypertarget{ref-Pearl1995}{}}%
+\CSLLeftMargin{52. }%
+\CSLRightInline{Pearl, J. \href{https://doi.org/10.2307/2337329}{Causal diagrams for empirical research}. \emph{Biometrika} \textbf{82}, 669--688 (1995).}
 
@@
 \renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
-\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}
+\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
 \pagenumbering{gobble}
-\nolinenumbers
 
@@
 
-Results showed that participants did differentiate the perceived norm items as expected.
-Participants rated the perceived descriptive norm items as providing more descriptive information than the perceived injunctive norm items, \emph{t}(442) = -7.28, \emph{p} \textless{} .001 (mean descriptive items = 4.75; mean injunctive items = 4.25).
-By contrast, participants rated the perceived injunctive norm items as providing more injunctive information than the perceived descriptive norm items, \emph{t}(444) = 7.15, \emph{p} \textless{} .001 (mean descriptive items = 5.11; mean injunctive items = 5.54).
+We fitted a multilevel model to determine whether participants rated the descriptive norm items as providing more descriptive information than the injunctive norm items, and vice versa.
+Pairwise comparisons from this model revealed that participants did differentiate the perceived norm items as expected (see Supplementary Table \ref{tab:constructValTable} for pairwise comparisons).
+Both descriptive norm items were seen as providing more descriptive information than both injunctive norm items.
+Conversely, both injunctive norm items were seen as providing more injunctive information than both descriptive norm items.
+
+\hypertarget{between-person-results-from-random-intercept-cross-lagged-panel-model}{%
+\subsubsection{Between-person results from random-intercept cross-lagged panel model}\label{between-person-results-from-random-intercept-cross-lagged-panel-model}}
+
+The random-intercept cross-lagged panel model partitions the data into two components: (1) between-person trait-like individual differences that persist over time, and (2) within-person fluctuations from trait levels at individual time points.
+In the main text, we focused on the results from the within-person component of the model, specifically the autoregressive and cross-lagged parameters, as we were interested in within-person causal effects.
+However, the model also estimates the covariances between the random intercepts in the model, which can be interpreted as the correlations between between-person trait levels for each variable (i.e., whether people who are generally high on one variable tend to be generally high on another variable).
+
+The covariances between the random intercepts in the full time-invariant model revealed positive correlations between stable trait levels of mask wearing and perceived social norms.
+On average across the whole study, participants who more frequently wore masks during in-person interactions also perceived stronger descriptive mask wearing norms (\emph{r} = 0.16, 95\% CI {[}0.01 0.30{]}) and stronger injunctive mask wearing norms (\emph{r} = 0.30, 95\% CI {[}0.17 0.43{]}).
+Stable trait perceptions of descriptive and injunctive mask wearing norms were also highly positively correlated (\emph{r} = 0.68, 95\% CI {[}0.61 0.75{]}).
+This general pattern of correlations at the between-person level was unchanged when removing all covariates and when unconstraining the within-person effects over time.
+
+\hypertarget{multi-group-random-intercept-cross-lagged-panel-model}{%
+\subsubsection{Multi-group random-intercept cross-lagged panel model}\label{multi-group-random-intercept-cross-lagged-panel-model}}
+
+To test whether the causal effects of descriptive and injunctive norms varied across regions of the United States with different mask wearing policies, we fitted a multi-group random-intercept cross-lagged panel model.
+The model was fitted separately to two groups: majority-Democrat states and majority-Republican states.
+Participants were linked to US states via self-reported zip codes.
+We operationalised state-level partisanship using the state-level results for the 2020 Presidential election.
+For ease of interpretation, we constrained the within-person autoregressive effects, cross-lagged effects, covariances, and variances to equality over time.
+This multi-group model fitted the data well according to one fit statistic (RMSEA = 0.049, 95\% CI {[}0.046 0.051{]}) but not according to others (SRMR = 0.100; CFI = 0.856).
+
+In majority-Republican states, the main results are unchanged from the full model fitted to the whole sample: perceived descriptive norms predicted future mask wearing (\emph{b} = 0.14, 95\% CI {[}0.05 0.24{]}, \emph{p} = .004) while perceived injunctive norms did not (\emph{b} = 0.00, 95\% CI {[}-0.08 0.09{]}, \emph{p} = .910).
+In majority-Democrat states, perceived descriptive norms have a slightly weaker effect on future mask wearing (\emph{b} = 0.08, 95\% CI {[}0.00 0.17{]}, \emph{p} = .059) but the effect of perceived injunctive norms remains non-significant (\emph{b} = -0.02, 95\% CI {[}-0.11 0.06{]}, \emph{p} = .593).
+For full list of autoregressive and cross-lagged effects from the multi-group model, see Supplementary Table \ref{tab:constrainedTable4}.
 
@@
 \subsection{Supplementary Figures}\label{supplementary-figures}}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotUSMap-1.pdf}
-\caption{\label{fig:plotUSMap}\emph{Map of the United States with participant zip code locations.}}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotAttrition-1.pdf}
-\caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotAttritionBreakdown-1.pdf}
-\caption{\label{fig:plotAttritionBreakdown}\emph{Demographics of the sample across the course of the study.}}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotDAG-1.pdf}
-\caption{\label{fig:plotDAG}\emph{Directed acyclic graph reflecting causal assumptions.} In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing (Mask).
-Perceptions of descriptive and injunctive norms also indirectly cause mask wearing through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
-Finally, political orientation (PolOri) is an exogenous variable that is a common cause of all other variables.
-Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived descriptive norms on mask wearing.
-Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and political orientation to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
-\end{figure}
-
-\newpage
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotDAG-1.pdf}
+\caption{\label{fig:plotDAG}\emph{Directed acyclic graph reflecting causal assumptions.} In this model, a general unobserved sensitivity to social norms (NormSens) causes perceptions of descriptive social norms (DesNorm) and perceptions of injunctive social norms (InjNorm), and perceptions of descriptive and injunctive norms directly cause mask wearing (Mask).
+Perceptions of descriptive and injunctive norms also indirectly cause mask wearing through non-social beliefs, specifically factual beliefs (Fact) and personal normative beliefs (PersNorm).
+Finally, demographic variables such as gender, age, ethnicity, socioeconomic status, and political orientation (summarised as Demographics) are exogenous variables that are common causes of all other variables.
+Using the backdoor criterion\textsuperscript{52}, this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotRICLPMPath2-1.pdf}
+\caption{\label{fig:plotRICLPMPath2}\emph{Path diagram of ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed).} Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences and controlling for demographics and political orientation.
+Arrow thickness is scaled according to standardized effect size.
+Bolded arrows indicate significantly positive parameters, \emph{p} \textless{} 0.05.
+Gray arrows indicate non-significant parameters.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotRICLPMPath3-1.pdf}
+\caption{\label{fig:plotRICLPMPath3}\emph{Path diagram of ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed).} Circles represent data collection time points.
+Arrows represent within-person autoregressive effects (on one horizontal level) and cross-lagged effects (across levels) for mask wearing and perceived descriptive and injunctive norms, partitioning out stable between-person individual differences.
+Arrow thickness is scaled according to standardized effect size.
+Bolded arrows indicate significantly positive parameters, \emph{p} \textless{} 0.05.
+Gray arrows indicate non-significant parameters.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotRICLPMCoef2-1.pdf}
+\caption{\label{fig:plotRICLPMCoef2}\emph{Standardized cross-lagged coefficients for descriptive norms and injunctive norms predicting future mask wearing in the ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed).} Points are standardized estimates, lines are 95\% confidence intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotRICLPMCoef3-1.pdf}
+\caption{\label{fig:plotRICLPMCoef3}\emph{Standardized cross-lagged coefficients for descriptive norms and injunctive norms predicting future mask wearing in the ten-wave time-varying random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed).} Points are standardized estimates, lines are 95\% confidence intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAttrition-1.pdf}
+\caption{\label{fig:plotAttrition}\emph{Attrition across the course of the study.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAttritionBreakdown-1.pdf}
+\caption{\label{fig:plotAttritionBreakdown}\emph{Demographics of the sample across the course of the study.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotUSMap-1.pdf}
+\caption{\label{fig:plotUSMap}\emph{Map of the United States with participant zip code locations.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotIntFreq-1.pdf}
+\caption{\label{fig:plotIntFreq}\emph{The relationship between self-reported mask wearing and the logged number of in-person interactions in the last 24 hours (a) and in the last 7 days (b).} Lines and shaded areas are predictions and 95\% confidence intervals from multilevel models with random intercepts for timepoint and participant.
+The first multilevel model revealed a negative relationship between mask wearing and the log number of in-person interactions in the last 24 hours (b = -0.04, SE = 0.02, p = .007).
+The second multilevel model revealed a negative relationship between mask wearing and the log number of in-person interactions in the last 7 days (b = -0.06, SE = 0.02, p \textless{} .001).}
+\end{figure}
+
+\newpage
+
 \hypertarget{supplementary-tables}{%
 \subsection{Supplementary Tables}\label{supplementary-tables}}
-
-\begin{center}
-\begin{ThreePartTable}
-
-\begin{longtable}{m{2.5cm}m{2.5cm}m{9cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:itemTable}List of norm interpretation questions asked at Time 7.
-\emph{These questions were preceded by the following text:} ``There may or may not be a difference between what people around you are doing and what they should be doing.
-You can learn about what people are doing and what they should be doing in different ways.
-For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both''.
-\emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
-\toprule
-Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:itemTable} continued}}\\
-\toprule
-Interpretation & \multicolumn{1}{c}{Perceived norm item} & \multicolumn{1}{c}{Question}\\
-\midrule
-\endhead
-Provides descriptive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone is doing?\\
- &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone is doing?\\
- & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone is doing?\\
- &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone is doing?\\
-Provides injunctive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone should be doing?\\
- &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone should be doing?\\
- & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone should be doing?\\
- &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone should be doing?\\
-\bottomrule
-\end{longtable}
-
-\end{ThreePartTable}
-\end{center}
-
-\newpage
 
@@
 
+\begin{longtable}{lccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:constrainedTable2}Unstandardized autoregressive and cross-lagged parameters from time-invariant random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed).
+\emph{Arrows indicate the direction of prediction.}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:constrainedTable2} continued}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endhead
+Mask wearing → Mask wearing & 0.23 & 0.02 & 0.18 & 0.28 & 0.00\\
+Mask wearing → Injunctive norms & 0.02 & 0.02 & -0.02 & 0.05 & 0.27\\
+Mask wearing → Descriptive norms & 0.04 & 0.02 & 0.01 & 0.08 & 0.02\\
+Injunctive norms → Mask wearing & 0.02 & 0.03 & -0.04 & 0.08 & 0.48\\
+Injunctive norms → Injunctive norms & 0.26 & 0.03 & 0.21 & 0.31 & 0.00\\
+Injunctive norms → Descriptive norms & 0.12 & 0.02 & 0.07 & 0.17 & 0.00\\
+Descriptive norms → Mask wearing & 0.15 & 0.03 & 0.09 & 0.21 & 0.00\\
+Descriptive norms → Injunctive norms & 0.17 & 0.02 & 0.12 & 0.21 & 0.00\\
+Descriptive norms → Descriptive norms & 0.32 & 0.03 & 0.27 & 0.38 & 0.00\\
+\bottomrule
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{longtable}{lccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:constrainedTable3}Unstandardized autoregressive and cross-lagged parameters from time-invariant random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed).
+\emph{Arrows indicate the direction of prediction.}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:constrainedTable3} continued}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endhead
+Mask wearing → Mask wearing & 0.25 & 0.02 & 0.20 & 0.30 & 0.00\\
+Mask wearing → Injunctive norms & 0.03 & 0.02 & 0.00 & 0.07 & 0.06\\
+Mask wearing → Descriptive norms & 0.07 & 0.02 & 0.03 & 0.10 & 0.00\\
+Injunctive norms → Mask wearing & 0.03 & 0.03 & -0.03 & 0.09 & 0.32\\
+Injunctive norms → Injunctive norms & 0.28 & 0.03 & 0.23 & 0.33 & 0.00\\
+Injunctive norms → Descriptive norms & 0.14 & 0.02 & 0.09 & 0.18 & 0.00\\
+Descriptive norms → Mask wearing & 0.18 & 0.03 & 0.12 & 0.24 & 0.00\\
+Descriptive norms → Injunctive norms & 0.19 & 0.02 & 0.14 & 0.24 & 0.00\\
+Descriptive norms → Descriptive norms & 0.34 & 0.03 & 0.29 & 0.40 & 0.00\\
+\bottomrule
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
+
+\begin{center}
+\begin{ThreePartTable}
+
 \footnotesize{
 
-\begin{longtable}{lcccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:lavaanTable}Standardised autoregressive and cross-lagged parameters from random-intercept cross-lagged panel model.
-\emph{Variable name prefixes: Mask = mask wearing, Des = perceived descriptive norms, Inj = perceived injunctive norms, Fact = factual beliefs, Pers = personal normative beliefs.
-Variable name suffixes indicate time points.
+\begin{longtable}{lccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:unconstrainedTable1}Unstandardized autoregressive and cross-lagged parameters from time-varying random-intercept cross-lagged panel model (full model).
+\emph{Variable name suffixes indicate time points.
 Arrows indicate the direction of prediction.}}\\
 \toprule
-Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%}\\
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
 \midrule
 \endfirsthead
-\caption*{\normalfont{Table \ref{tab:lavaanTable} continued}}\\
+\caption*{\normalfont{Table \ref{tab:unconstrainedTable1} continued}}\\
 \toprule
-Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%}\\
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
 \midrule
 \endhead
-Des\_02 → Mask\_05 & 0.17 & 0.05 & 0.06 & 0.28\\
-Des\_02 → Inj\_05 & 0.17 & 0.06 & 0.06 & 0.28\\
-Des\_02 → Des\_05 & 0.37 & 0.05 & 0.26 & 0.47\\
-Des\_02 → Fact\_05 & 0.09 & 0.06 & -0.04 & 0.21\\
-Des\_02 → Pers\_05 & 0.04 & 0.06 & -0.08 & 0.17\\
-Des\_05 → Mask\_09 & 0.21 & 0.06 & 0.08 & 0.34\\
-Des\_05 → Inj\_09 & 0.23 & 0.07 & 0.10 & 0.36\\
-Des\_05 → Des\_09 & 0.26 & 0.06 & 0.14 & 0.39\\
-Des\_05 → Fact\_09 & 0.16 & 0.07 & 0.02 & 0.30\\
-Des\_05 → Pers\_09 & 0.27 & 0.07 & 0.12 & 0.42\\
-Des\_09 → Mask\_11 & 0.04 & 0.07 & -0.09 & 0.18\\
-Des\_09 → Inj\_11 & 0.20 & 0.07 & 0.07 & 0.33\\
-Des\_09 → Des\_11 & 0.26 & 0.07 & 0.13 & 0.39\\
-Des\_09 → Fact\_11 & 0.03 & 0.06 & -0.09 & 0.16\\
-Des\_09 → Pers\_11 & 0.07 & 0.07 & -0.06 & 0.20\\
-Des\_11 → Mask\_13 & 0.15 & 0.07 & 0.01 & 0.30\\
-Des\_11 → Inj\_13 & 0.03 & 0.07 & -0.12 & 0.17\\
-Des\_11 → Des\_13 & 0.27 & 0.07 & 0.14 & 0.41\\
-Des\_11 → Fact\_13 & 0.07 & 0.07 & -0.07 & 0.21\\
-Des\_11 → Pers\_13 & 0.06 & 0.07 & -0.08 & 0.20\\
-Des\_13 → Mask\_14 & 0.16 & 0.07 & 0.02 & 0.29\\
-Des\_13 → Inj\_14 & 0.21 & 0.06 & 0.09 & 0.33\\
-Des\_13 → Des\_14 & 0.40 & 0.06 & 0.28 & 0.51\\
-Des\_13 → Fact\_14 & 0.03 & 0.06 & -0.09 & 0.14\\
-Des\_13 → Pers\_14 & 0.01 & 0.06 & -0.11 & 0.12\\
-Des\_14 → Mask\_15 & 0.05 & 0.08 & -0.09 & 0.20\\
-Des\_14 → Inj\_15 & -0.01 & 0.07 & -0.16 & 0.13\\
-Des\_14 → Des\_15 & 0.34 & 0.06 & 0.22 & 0.46\\
-Des\_14 → Fact\_15 & 0.12 & 0.07 & -0.01 & 0.25\\
-Des\_14 → Pers\_15 & 0.09 & 0.07 & -0.05 & 0.23\\
-Des\_15 → Mask\_16 & 0.03 & 0.07 & -0.11 & 0.18\\
-Des\_15 → Inj\_16 & 0.23 & 0.08 & 0.08 & 0.38\\
-Des\_15 → Des\_16 & 0.30 & 0.07 & 0.15 & 0.45\\
-Des\_15 → Fact\_16 & 0.13 & 0.07 & 0.00 & 0.26\\
-Des\_15 → Pers\_16 & 0.01 & 0.07 & -0.12 & 0.14\\
-Des\_16 → Mask\_17 & 0.06 & 0.08 & -0.10 & 0.21\\
-Des\_16 → Inj\_17 & 0.24 & 0.08 & 0.08 & 0.39\\
-Des\_16 → Des\_17 & 0.53 & 0.07 & 0.40 & 0.66\\
-Des\_16 → Fact\_17 & 0.06 & 0.07 & -0.08 & 0.20\\
-Des\_16 → Pers\_17 & 0.03 & 0.07 & -0.10 & 0.16\\
-Des\_17 → Mask\_18 & 0.08 & 0.07 & -0.06 & 0.21\\
-Des\_17 → Inj\_18 & 0.30 & 0.07 & 0.17 & 0.43\\
-Des\_17 → Des\_18 & 0.46 & 0.06 & 0.34 & 0.58\\
-Des\_17 → Fact\_18 & 0.12 & 0.06 & 0.00 & 0.24\\
-Des\_17 → Pers\_18 & 0.07 & 0.06 & -0.05 & 0.20\\
-Fact\_02 → Mask\_05 & 0.06 & 0.07 & -0.08 & 0.19\\
-Fact\_02 → Inj\_05 & -0.10 & 0.07 & -0.24 & 0.03\\
-Fact\_02 → Des\_05 & -0.02 & 0.07 & -0.15 & 0.12\\
-Fact\_02 → Fact\_05 & 0.22 & 0.08 & 0.07 & 0.38\\
-Fact\_02 → Pers\_05 & -0.08 & 0.08 & -0.23 & 0.08\\
-Fact\_05 → Mask\_09 & 0.15 & 0.08 & -0.01 & 0.31\\
-Fact\_05 → Inj\_09 & -0.07 & 0.08 & -0.23 & 0.09\\
-Fact\_05 → Des\_09 & -0.05 & 0.08 & -0.20 & 0.11\\
-Fact\_05 → Fact\_09 & 0.07 & 0.09 & -0.10 & 0.25\\
-Fact\_05 → Pers\_09 & -0.03 & 0.09 & -0.20 & 0.15\\
-Fact\_09 → Mask\_11 & 0.15 & 0.08 & -0.01 & 0.30\\
-Fact\_09 → Inj\_11 & 0.03 & 0.08 & -0.12 & 0.18\\
-Fact\_09 → Des\_11 & 0.10 & 0.08 & -0.05 & 0.24\\
-Fact\_09 → Fact\_11 & 0.26 & 0.07 & 0.12 & 0.40\\
-Fact\_09 → Pers\_11 & 0.14 & 0.07 & -0.01 & 0.28\\
-Fact\_11 → Mask\_13 & 0.18 & 0.09 & 0.00 & 0.35\\
-Fact\_11 → Inj\_13 & 0.05 & 0.09 & -0.13 & 0.22\\
-Fact\_11 → Des\_13 & -0.12 & 0.08 & -0.28 & 0.04\\
-Fact\_11 → Fact\_13 & 0.19 & 0.08 & 0.03 & 0.36\\
-Fact\_11 → Pers\_13 & 0.16 & 0.08 & 0.00 & 0.33\\
-Fact\_13 → Mask\_14 & 0.05 & 0.08 & -0.12 & 0.21\\
-Fact\_13 → Inj\_14 & 0.04 & 0.07 & -0.11 & 0.18\\
-Fact\_13 → Des\_14 & 0.01 & 0.08 & -0.14 & 0.16\\
-Fact\_13 → Fact\_14 & 0.25 & 0.07 & 0.11 & 0.39\\
-Fact\_13 → Pers\_14 & 0.19 & 0.07 & 0.06 & 0.33\\
-Fact\_14 → Mask\_15 & 0.32 & 0.08 & 0.16 & 0.48\\
-Fact\_14 → Inj\_15 & -0.06 & 0.08 & -0.22 & 0.10\\
-Fact\_14 → Des\_15 & 0.15 & 0.07 & 0.01 & 0.29\\
-Fact\_14 → Fact\_15 & 0.47 & 0.07 & 0.33 & 0.60\\
-Fact\_14 → Pers\_15 & 0.31 & 0.08 & 0.16 & 0.47\\
-Fact\_15 → Mask\_16 & 0.10 & 0.09 & -0.08 & 0.28\\
-Fact\_15 → Inj\_16 & 0.08 & 0.10 & -0.11 & 0.27\\
-Fact\_15 → Des\_16 & 0.10 & 0.10 & -0.09 & 0.29\\
-Fact\_15 → Fact\_16 & 0.39 & 0.08 & 0.23 & 0.55\\
-Fact\_15 → Pers\_16 & 0.10 & 0.08 & -0.06 & 0.27\\
-Fact\_16 → Mask\_17 & 0.21 & 0.09 & 0.03 & 0.39\\
-Fact\_16 → Inj\_17 & -0.01 & 0.09 & -0.19 & 0.18\\
-Fact\_16 → Des\_17 & -0.05 & 0.09 & -0.22 & 0.12\\
-Fact\_16 → Fact\_17 & 0.22 & 0.08 & 0.06 & 0.39\\
-Fact\_16 → Pers\_17 & 0.06 & 0.08 & -0.10 & 0.22\\
-Fact\_17 → Mask\_18 & 0.10 & 0.09 & -0.08 & 0.28\\
-Fact\_17 → Inj\_18 & -0.10 & 0.09 & -0.28 & 0.08\\
-Fact\_17 → Des\_18 & 0.08 & 0.09 & -0.10 & 0.25\\
-Fact\_17 → Fact\_18 & 0.37 & 0.08 & 0.21 & 0.53\\
-Fact\_17 → Pers\_18 & 0.48 & 0.08 & 0.32 & 0.64\\
-Inj\_02 → Mask\_05 & 0.01 & 0.05 & -0.10 & 0.11\\
-Inj\_02 → Inj\_05 & 0.28 & 0.05 & 0.17 & 0.38\\
-Inj\_02 → Des\_05 & 0.07 & 0.05 & -0.03 & 0.18\\
-Inj\_02 → Fact\_05 & 0.05 & 0.06 & -0.08 & 0.17\\
-Inj\_02 → Pers\_05 & -0.01 & 0.06 & -0.13 & 0.11\\
-Inj\_05 → Mask\_09 & -0.07 & 0.06 & -0.19 & 0.05\\
-Inj\_05 → Inj\_09 & 0.08 & 0.06 & -0.04 & 0.21\\
-Inj\_05 → Des\_09 & -0.02 & 0.06 & -0.14 & 0.11\\
-Inj\_05 → Fact\_09 & 0.02 & 0.07 & -0.11 & 0.16\\
-Inj\_05 → Pers\_09 & -0.04 & 0.07 & -0.18 & 0.10\\
-Inj\_09 → Mask\_11 & 0.08 & 0.08 & -0.07 & 0.23\\
-Inj\_09 → Inj\_11 & 0.11 & 0.07 & -0.03 & 0.26\\
-Inj\_09 → Des\_11 & -0.03 & 0.07 & -0.17 & 0.11\\
-Inj\_09 → Fact\_11 & 0.01 & 0.07 & -0.12 & 0.15\\
-Inj\_09 → Pers\_11 & 0.05 & 0.07 & -0.08 & 0.19\\
-Inj\_11 → Mask\_13 & -0.01 & 0.07 & -0.15 & 0.13\\
-Inj\_11 → Inj\_13 & 0.29 & 0.07 & 0.15 & 0.43\\
-Inj\_11 → Des\_13 & 0.21 & 0.07 & 0.08 & 0.34\\
-Inj\_11 → Fact\_13 & 0.12 & 0.07 & -0.01 & 0.26\\
-Inj\_11 → Pers\_13 & 0.09 & 0.07 & -0.04 & 0.23\\
-Inj\_13 → Mask\_14 & -0.05 & 0.07 & -0.19 & 0.08\\
-Inj\_13 → Inj\_14 & 0.40 & 0.06 & 0.28 & 0.52\\
-Inj\_13 → Des\_14 & 0.15 & 0.07 & 0.03 & 0.28\\
-Inj\_13 → Fact\_14 & 0.02 & 0.06 & -0.10 & 0.14\\
-Inj\_13 → Pers\_14 & 0.09 & 0.06 & -0.02 & 0.21\\
-Inj\_14 → Mask\_15 & 0.08 & 0.07 & -0.06 & 0.22\\
-Inj\_14 → Inj\_15 & 0.45 & 0.07 & 0.32 & 0.58\\
-Inj\_14 → Des\_15 & 0.29 & 0.06 & 0.16 & 0.41\\
-Inj\_14 → Fact\_15 & 0.10 & 0.06 & -0.02 & 0.22\\
-Inj\_14 → Pers\_15 & 0.06 & 0.07 & -0.07 & 0.20\\
-Inj\_15 → Mask\_16 & 0.14 & 0.07 & 0.00 & 0.28\\
-Inj\_15 → Inj\_16 & 0.21 & 0.07 & 0.06 & 0.35\\
-Inj\_15 → Des\_16 & 0.06 & 0.07 & -0.08 & 0.21\\
-Inj\_15 → Fact\_16 & 0.01 & 0.06 & -0.12 & 0.13\\
-Inj\_15 → Pers\_16 & 0.10 & 0.06 & -0.03 & 0.22\\
-Inj\_16 → Mask\_17 & -0.01 & 0.07 & -0.15 & 0.13\\
-Inj\_16 → Inj\_17 & 0.38 & 0.07 & 0.23 & 0.52\\
-Inj\_16 → Des\_17 & 0.13 & 0.07 & 0.00 & 0.27\\
-Inj\_16 → Fact\_17 & 0.00 & 0.07 & -0.14 & 0.13\\
-Inj\_16 → Pers\_17 & -0.03 & 0.06 & -0.16 & 0.09\\
-Inj\_17 → Mask\_18 & -0.02 & 0.07 & -0.15 & 0.11\\
-Inj\_17 → Inj\_18 & 0.45 & 0.06 & 0.33 & 0.57\\
-Inj\_17 → Des\_18 & 0.19 & 0.06 & 0.07 & 0.32\\
-Inj\_17 → Fact\_18 & 0.01 & 0.06 & -0.11 & 0.13\\
-Inj\_17 → Pers\_18 & 0.01 & 0.06 & -0.11 & 0.13\\
-Mask\_02 → Mask\_05 & 0.21 & 0.05 & 0.11 & 0.31\\
-Mask\_02 → Inj\_05 & 0.09 & 0.05 & -0.01 & 0.20\\
-Mask\_02 → Des\_05 & 0.04 & 0.05 & -0.07 & 0.14\\
-Mask\_02 → Fact\_05 & -0.05 & 0.06 & -0.17 & 0.07\\
-Mask\_02 → Pers\_05 & -0.05 & 0.06 & -0.17 & 0.06\\
-Mask\_05 → Mask\_09 & 0.19 & 0.06 & 0.07 & 0.30\\
-Mask\_05 → Inj\_09 & 0.13 & 0.06 & 0.01 & 0.26\\
-Mask\_05 → Des\_09 & 0.02 & 0.06 & -0.10 & 0.14\\
-Mask\_05 → Fact\_09 & 0.14 & 0.07 & 0.01 & 0.27\\
-Mask\_05 → Pers\_09 & 0.06 & 0.07 & -0.07 & 0.20\\
-Mask\_09 → Mask\_11 & -0.01 & 0.07 & -0.14 & 0.12\\
-Mask\_09 → Inj\_11 & 0.06 & 0.06 & -0.07 & 0.18\\
-Mask\_09 → Des\_11 & 0.16 & 0.06 & 0.04 & 0.28\\
-Mask\_09 → Fact\_11 & 0.19 & 0.06 & 0.08 & 0.31\\
-Mask\_09 → Pers\_11 & 0.16 & 0.06 & 0.05 & 0.28\\
-Mask\_11 → Mask\_13 & 0.07 & 0.07 & -0.06 & 0.21\\
-Mask\_11 → Inj\_13 & 0.06 & 0.07 & -0.07 & 0.19\\
-Mask\_11 → Des\_13 & 0.07 & 0.06 & -0.06 & 0.19\\
-Mask\_11 → Fact\_13 & 0.04 & 0.06 & -0.09 & 0.16\\
-Mask\_11 → Pers\_13 & 0.03 & 0.07 & -0.10 & 0.16\\
-Mask\_13 → Mask\_14 & 0.19 & 0.06 & 0.08 & 0.31\\
-Mask\_13 → Inj\_14 & 0.07 & 0.05 & -0.03 & 0.18\\
-Mask\_13 → Des\_14 & 0.12 & 0.06 & 0.01 & 0.23\\
-Mask\_13 → Fact\_14 & 0.07 & 0.05 & -0.03 & 0.17\\
-Mask\_13 → Pers\_14 & 0.01 & 0.05 & -0.09 & 0.11\\
-Mask\_14 → Mask\_15 & 0.21 & 0.06 & 0.09 & 0.33\\
-Mask\_14 → Inj\_15 & 0.06 & 0.06 & -0.06 & 0.18\\
-Mask\_14 → Des\_15 & 0.08 & 0.05 & -0.02 & 0.18\\
-Mask\_14 → Fact\_15 & 0.05 & 0.05 & -0.06 & 0.15\\
-Mask\_14 → Pers\_15 & -0.05 & 0.06 & -0.17 & 0.06\\
-Mask\_15 → Mask\_16 & 0.25 & 0.07 & 0.12 & 0.39\\
-Mask\_15 → Inj\_16 & 0.02 & 0.07 & -0.12 & 0.16\\
-Mask\_15 → Des\_16 & 0.01 & 0.07 & -0.13 & 0.15\\
-Mask\_15 → Fact\_16 & 0.10 & 0.06 & -0.03 & 0.22\\
-Mask\_15 → Pers\_16 & 0.09 & 0.06 & -0.03 & 0.22\\
-Mask\_16 → Mask\_17 & 0.33 & 0.07 & 0.20 & 0.46\\
-Mask\_16 → Inj\_17 & -0.04 & 0.07 & -0.18 & 0.10\\
-Mask\_16 → Des\_17 & 0.16 & 0.06 & 0.03 & 0.28\\
-Mask\_16 → Fact\_17 & 0.22 & 0.06 & 0.09 & 0.34\\
-Mask\_16 → Pers\_17 & 0.12 & 0.06 & 0.01 & 0.24\\
-Mask\_17 → Mask\_18 & 0.39 & 0.06 & 0.27 & 0.51\\
-Mask\_17 → Inj\_18 & -0.05 & 0.06 & -0.17 & 0.07\\
-Mask\_17 → Des\_18 & 0.02 & 0.06 & -0.10 & 0.13\\
-Mask\_17 → Fact\_18 & 0.13 & 0.06 & 0.02 & 0.24\\
-Mask\_17 → Pers\_18 & 0.04 & 0.06 & -0.07 & 0.16\\
-Pers\_02 → Mask\_05 & 0.05 & 0.07 & -0.09 & 0.18\\
-Pers\_02 → Inj\_05 & 0.09 & 0.07 & -0.05 & 0.22\\
-Pers\_02 → Des\_05 & 0.03 & 0.07 & -0.10 & 0.17\\
-Pers\_02 → Fact\_05 & 0.06 & 0.08 & -0.09 & 0.22\\
-Pers\_02 → Pers\_05 & 0.36 & 0.07 & 0.21 & 0.50\\
-Pers\_05 → Mask\_09 & -0.27 & 0.08 & -0.42 & -0.12\\
-Pers\_05 → Inj\_09 & -0.16 & 0.08 & -0.31 & 0.00\\
-Pers\_05 → Des\_09 & -0.06 & 0.08 & -0.21 & 0.10\\
-Pers\_05 → Fact\_09 & -0.21 & 0.08 & -0.37 & -0.05\\
-Pers\_05 → Pers\_09 & -0.21 & 0.09 & -0.38 & -0.04\\
-Pers\_09 → Mask\_11 & 0.04 & 0.08 & -0.11 & 0.20\\
-Pers\_09 → Inj\_11 & 0.08 & 0.08 & -0.07 & 0.23\\
-Pers\_09 → Des\_11 & 0.04 & 0.07 & -0.10 & 0.19\\
-Pers\_09 → Fact\_11 & 0.06 & 0.07 & -0.08 & 0.20\\
-Pers\_09 → Pers\_11 & 0.16 & 0.07 & 0.02 & 0.31\\
-Pers\_11 → Mask\_13 & 0.08 & 0.08 & -0.08 & 0.24\\
-Pers\_11 → Inj\_13 & 0.09 & 0.08 & -0.07 & 0.24\\
-Pers\_11 → Des\_13 & 0.12 & 0.08 & -0.03 & 0.27\\
-Pers\_11 → Fact\_13 & 0.18 & 0.08 & 0.03 & 0.33\\
-Pers\_11 → Pers\_13 & 0.20 & 0.08 & 0.05 & 0.35\\
-Pers\_13 → Mask\_14 & 0.24 & 0.08 & 0.08 & 0.40\\
-Pers\_13 → Inj\_14 & -0.07 & 0.07 & -0.21 & 0.07\\
-Pers\_13 → Des\_14 & -0.03 & 0.08 & -0.18 & 0.12\\
-Pers\_13 → Fact\_14 & 0.34 & 0.07 & 0.21 & 0.48\\
-Pers\_13 → Pers\_14 & 0.41 & 0.07 & 0.29 & 0.54\\
-Pers\_14 → Mask\_15 & -0.05 & 0.08 & -0.22 & 0.11\\
-Pers\_14 → Inj\_15 & 0.15 & 0.08 & -0.02 & 0.31\\
-Pers\_14 → Des\_15 & -0.07 & 0.07 & -0.21 & 0.07\\
-Pers\_14 → Fact\_15 & 0.02 & 0.07 & -0.13 & 0.16\\
-Pers\_14 → Pers\_15 & 0.14 & 0.08 & -0.02 & 0.30\\
-Pers\_15 → Mask\_16 & 0.11 & 0.09 & -0.05 & 0.28\\
-Pers\_15 → Inj\_16 & 0.08 & 0.09 & -0.10 & 0.25\\
-Pers\_15 → Des\_16 & 0.17 & 0.09 & 0.00 & 0.35\\
-Pers\_15 → Fact\_16 & 0.11 & 0.08 & -0.05 & 0.26\\
-Pers\_15 → Pers\_16 & 0.41 & 0.08 & 0.27 & 0.56\\
-Pers\_16 → Mask\_17 & 0.00 & 0.08 & -0.17 & 0.17\\
-Pers\_16 → Inj\_17 & 0.05 & 0.09 & -0.12 & 0.23\\
-Pers\_16 → Des\_17 & -0.02 & 0.08 & -0.18 & 0.14\\
-Pers\_16 → Fact\_17 & 0.26 & 0.08 & 0.11 & 0.41\\
-Pers\_16 → Pers\_17 & 0.56 & 0.07 & 0.42 & 0.69\\
-Pers\_17 → Mask\_18 & 0.09 & 0.09 & -0.08 & 0.26\\
-Pers\_17 → Inj\_18 & -0.01 & 0.08 & -0.17 & 0.15\\
-Pers\_17 → Des\_18 & -0.02 & 0.08 & -0.18 & 0.14\\
-Pers\_17 → Fact\_18 & 0.16 & 0.08 & 0.01 & 0.31\\
-Pers\_17 → Pers\_18 & 0.12 & 0.08 & -0.03 & 0.27\\
+MaskWearing\_02 → MaskWearing\_05 & 0.24 & 0.06 & 0.12 & 0.36 & 0.00\\
+MaskWearing\_02 → InjunctiveNorms\_05 & 0.04 & 0.05 & -0.05 & 0.14 & 0.39\\
+MaskWearing\_02 → DescriptiveNorms\_05 & 0.02 & 0.05 & -0.08 & 0.12 & 0.64\\
+MaskWearing\_02 → FactualBeliefs\_05 & 0.00 & 0.05 & -0.10 & 0.09 & 0.97\\
+MaskWearing\_02 → PersonalBeliefs\_05 & -0.01 & 0.04 & -0.10 & 0.07 & 0.74\\
+MaskWearing\_05 → MaskWearing\_09 & 0.18 & 0.06 & 0.06 & 0.31 & 0.00\\
+MaskWearing\_05 → InjunctiveNorms\_09 & 0.14 & 0.06 & 0.03 & 0.26 & 0.01\\
+MaskWearing\_05 → DescriptiveNorms\_09 & 0.01 & 0.06 & -0.12 & 0.13 & 0.92\\
+MaskWearing\_05 → FactualBeliefs\_09 & 0.15 & 0.07 & 0.01 & 0.29 & 0.03\\
+MaskWearing\_05 → PersonalBeliefs\_09 & 0.06 & 0.07 & -0.07 & 0.19 & 0.35\\
+MaskWearing\_09 → MaskWearing\_11 & -0.04 & 0.08 & -0.19 & 0.12 & 0.62\\
+MaskWearing\_09 → InjunctiveNorms\_11 & 0.04 & 0.06 & -0.08 & 0.16 & 0.51\\
+MaskWearing\_09 → DescriptiveNorms\_11 & 0.10 & 0.06 & -0.02 & 0.21 & 0.09\\
+MaskWearing\_09 → FactualBeliefs\_11 & 0.22 & 0.08 & 0.07 & 0.37 & 0.00\\
+MaskWearing\_09 → PersonalBeliefs\_11 & 0.15 & 0.06 & 0.02 & 0.27 & 0.02\\
+MaskWearing\_11 → MaskWearing\_13 & 0.05 & 0.06 & -0.06 & 0.17 & 0.36\\
+MaskWearing\_11 → InjunctiveNorms\_13 & 0.04 & 0.05 & -0.06 & 0.14 & 0.47\\
+MaskWearing\_11 → DescriptiveNorms\_13 & 0.05 & 0.06 & -0.06 & 0.16 & 0.38\\
+MaskWearing\_11 → FactualBeliefs\_13 & 0.03 & 0.06 & -0.08 & 0.15 & 0.55\\
+MaskWearing\_11 → PersonalBeliefs\_13 & 0.03 & 0.05 & -0.08 & 0.14 & 0.58\\
+MaskWearing\_13 → MaskWearing\_14 & 0.19 & 0.06 & 0.06 & 0.31 & 0.00\\
+MaskWearing\_13 → InjunctiveNorms\_14 & 0.08 & 0.05 & -0.03 & 0.18 & 0.14\\
+MaskWearing\_13 → DescriptiveNorms\_14 & 0.09 & 0.05 & -0.01 & 0.20 & 0.08\\
+MaskWearing\_13 → FactualBeliefs\_14 & 0.09 & 0.06 & -0.02 & 0.21 & 0.12\\
+MaskWearing\_13 → PersonalBeliefs\_14 & 0.01 & 0.05 & -0.09 & 0.11 & 0.82\\
+MaskWearing\_14 → MaskWearing\_15 & 0.22 & 0.06 & 0.10 & 0.35 & 0.00\\
+MaskWearing\_14 → InjunctiveNorms\_15 & 0.06 & 0.06 & -0.05 & 0.17 & 0.31\\
+MaskWearing\_14 → DescriptiveNorms\_15 & 0.07 & 0.05 & -0.03 & 0.17 & 0.18\\
+MaskWearing\_14 → FactualBeliefs\_15 & 0.05 & 0.06 & -0.07 & 0.18 & 0.42\\
+MaskWearing\_14 → PersonalBeliefs\_15 & -0.07 & 0.06 & -0.20 & 0.05 & 0.27\\
+MaskWearing\_15 → MaskWearing\_16 & 0.29 & 0.08 & 0.14 & 0.44 & 0.00\\
+MaskWearing\_15 → InjunctiveNorms\_16 & 0.00 & 0.06 & -0.12 & 0.13 & 0.94\\
+MaskWearing\_15 → DescriptiveNorms\_16 & -0.01 & 0.06 & -0.12 & 0.10 & 0.91\\
+MaskWearing\_15 → FactualBeliefs\_16 & 0.09 & 0.08 & -0.06 & 0.25 & 0.24\\
+MaskWearing\_15 → PersonalBeliefs\_16 & 0.08 & 0.07 & -0.06 & 0.22 & 0.24\\
+MaskWearing\_16 → MaskWearing\_17 & 0.37 & 0.07 & 0.23 & 0.51 & 0.00\\
+MaskWearing\_16 → InjunctiveNorms\_17 & 0.01 & 0.05 & -0.08 & 0.10 & 0.83\\
+MaskWearing\_16 → DescriptiveNorms\_17 & 0.10 & 0.04 & 0.01 & 0.18 & 0.02\\
+MaskWearing\_16 → FactualBeliefs\_17 & 0.27 & 0.08 & 0.12 & 0.42 & 0.00\\
+MaskWearing\_16 → PersonalBeliefs\_17 & 0.15 & 0.07 & 0.01 & 0.28 & 0.03\\
+MaskWearing\_17 → MaskWearing\_18 & 0.38 & 0.06 & 0.26 & 0.50 & 0.00\\
+MaskWearing\_17 → InjunctiveNorms\_18 & -0.04 & 0.04 & -0.11 & 0.04 & 0.32\\
+MaskWearing\_17 → DescriptiveNorms\_18 & 0.02 & 0.03 & -0.04 & 0.09 & 0.49\\
+MaskWearing\_17 → FactualBeliefs\_18 & 0.13 & 0.07 & 0.00 & 0.26 & 0.05\\
+MaskWearing\_17 → PersonalBeliefs\_18 & 0.06 & 0.06 & -0.06 & 0.18 & 0.34\\
+InjunctiveNorms\_02 → MaskWearing\_05 & 0.03 & 0.07 & -0.11 & 0.17 & 0.71\\
+InjunctiveNorms\_02 → InjunctiveNorms\_05 & 0.30 & 0.06 & 0.19 & 0.42 & 0.00\\
+InjunctiveNorms\_02 → DescriptiveNorms\_05 & 0.07 & 0.06 & -0.04 & 0.19 & 0.23\\
+InjunctiveNorms\_02 → FactualBeliefs\_05 & 0.05 & 0.05 & -0.06 & 0.16 & 0.37\\
+InjunctiveNorms\_02 → PersonalBeliefs\_05 & 0.00 & 0.05 & -0.10 & 0.10 & 0.97\\
+InjunctiveNorms\_05 → MaskWearing\_09 & -0.10 & 0.08 & -0.26 & 0.06 & 0.23\\
+InjunctiveNorms\_05 → InjunctiveNorms\_09 & 0.09 & 0.08 & -0.07 & 0.24 & 0.26\\
+InjunctiveNorms\_05 → DescriptiveNorms\_09 & 0.03 & 0.08 & -0.12 & 0.19 & 0.68\\
+InjunctiveNorms\_05 → FactualBeliefs\_09 & 0.03 & 0.09 & -0.16 & 0.21 & 0.78\\
+InjunctiveNorms\_05 → PersonalBeliefs\_09 & -0.06 & 0.09 & -0.23 & 0.11 & 0.51\\
+InjunctiveNorms\_09 → MaskWearing\_11 & 0.07 & 0.10 & -0.13 & 0.27 & 0.51\\
+InjunctiveNorms\_09 → InjunctiveNorms\_11 & 0.13 & 0.08 & -0.03 & 0.28 & 0.12\\
+InjunctiveNorms\_09 → DescriptiveNorms\_11 & -0.02 & 0.07 & -0.17 & 0.12 & 0.77\\
+InjunctiveNorms\_09 → FactualBeliefs\_11 & 0.00 & 0.10 & -0.20 & 0.19 & 0.98\\
+InjunctiveNorms\_09 → PersonalBeliefs\_11 & 0.02 & 0.08 & -0.14 & 0.19 & 0.76\\
+InjunctiveNorms\_11 → MaskWearing\_13 & -0.01 & 0.08 & -0.16 & 0.14 & 0.91\\
+InjunctiveNorms\_11 → InjunctiveNorms\_13 & 0.31 & 0.07 & 0.18 & 0.45 & 0.00\\
+InjunctiveNorms\_11 → DescriptiveNorms\_13 & 0.24 & 0.07 & 0.10 & 0.38 & 0.00\\
+InjunctiveNorms\_11 → FactualBeliefs\_13 & 0.18 & 0.08 & 0.03 & 0.33 & 0.02\\
+InjunctiveNorms\_11 → PersonalBeliefs\_13 & 0.10 & 0.07 & -0.04 & 0.23 & 0.18\\
+InjunctiveNorms\_13 → MaskWearing\_14 & -0.05 & 0.08 & -0.21 & 0.11 & 0.51\\
+InjunctiveNorms\_13 → InjunctiveNorms\_14 & 0.42 & 0.07 & 0.28 & 0.55 & 0.00\\
+InjunctiveNorms\_13 → DescriptiveNorms\_14 & 0.16 & 0.07 & 0.02 & 0.30 & 0.02\\
+InjunctiveNorms\_13 → FactualBeliefs\_14 & 0.03 & 0.08 & -0.12 & 0.18 & 0.68\\
+InjunctiveNorms\_13 → PersonalBeliefs\_14 & 0.10 & 0.07 & -0.03 & 0.22 & 0.15\\
+InjunctiveNorms\_14 → MaskWearing\_15 & 0.07 & 0.08 & -0.09 & 0.23 & 0.40\\
+InjunctiveNorms\_14 → InjunctiveNorms\_15 & 0.45 & 0.07 & 0.31 & 0.59 & 0.00\\
+InjunctiveNorms\_14 → DescriptiveNorms\_15 & 0.28 & 0.07 & 0.16 & 0.41 & 0.00\\
+InjunctiveNorms\_14 → FactualBeliefs\_15 & 0.11 & 0.08 & -0.05 & 0.27 & 0.16\\
+InjunctiveNorms\_14 → PersonalBeliefs\_15 & 0.07 & 0.08 & -0.09 & 0.23 & 0.38\\
+InjunctiveNorms\_15 → MaskWearing\_16 & 0.20 & 0.09 & 0.03 & 0.37 & 0.02\\
+InjunctiveNorms\_15 → InjunctiveNorms\_16 & 0.19 & 0.07 & 0.05 & 0.33 & 0.01\\
+InjunctiveNorms\_15 → DescriptiveNorms\_16 & 0.04 & 0.06 & -0.09 & 0.16 & 0.56\\
+InjunctiveNorms\_15 → FactualBeliefs\_16 & 0.04 & 0.09 & -0.14 & 0.22 & 0.67\\
+InjunctiveNorms\_15 → PersonalBeliefs\_16 & 0.12 & 0.08 & -0.03 & 0.28 & 0.12\\
+InjunctiveNorms\_16 → MaskWearing\_17 & 0.00 & 0.10 & -0.19 & 0.20 & 0.97\\
+InjunctiveNorms\_16 → InjunctiveNorms\_17 & 0.32 & 0.07 & 0.19 & 0.44 & 0.00\\
+InjunctiveNorms\_16 → DescriptiveNorms\_17 & 0.10 & 0.06 & -0.01 & 0.21 & 0.08\\
+InjunctiveNorms\_16 → FactualBeliefs\_17 & 0.00 & 0.11 & -0.21 & 0.21 & 0.97\\
+InjunctiveNorms\_16 → PersonalBeliefs\_17 & -0.04 & 0.09 & -0.22 & 0.14 & 0.68\\
+InjunctiveNorms\_17 → MaskWearing\_18 & 0.00 & 0.10 & -0.21 & 0.20 & 0.98\\
+InjunctiveNorms\_17 → InjunctiveNorms\_18 & 0.37 & 0.06 & 0.25 & 0.50 & 0.00\\
+InjunctiveNorms\_17 → DescriptiveNorms\_18 & 0.18 & 0.06 & 0.07 & 0.29 & 0.00\\
+InjunctiveNorms\_17 → FactualBeliefs\_18 & 0.07 & 0.11 & -0.15 & 0.30 & 0.53\\
+InjunctiveNorms\_17 → PersonalBeliefs\_18 & -0.02 & 0.10 & -0.22 & 0.18 & 0.84\\
+DescriptiveNorms\_02 → MaskWearing\_05 & 0.20 & 0.07 & 0.06 & 0.34 & 0.01\\
+DescriptiveNorms\_02 → InjunctiveNorms\_05 & 0.18 & 0.06 & 0.07 & 0.30 & 0.00\\
+DescriptiveNorms\_02 → DescriptiveNorms\_05 & 0.40 & 0.06 & 0.29 & 0.52 & 0.00\\
+DescriptiveNorms\_02 → FactualBeliefs\_05 & 0.05 & 0.06 & -0.06 & 0.16 & 0.39\\
+DescriptiveNorms\_02 → PersonalBeliefs\_05 & 0.02 & 0.05 & -0.08 & 0.12 & 0.71\\
+DescriptiveNorms\_05 → MaskWearing\_09 & 0.26 & 0.08 & 0.10 & 0.42 & 0.00\\
+DescriptiveNorms\_05 → InjunctiveNorms\_09 & 0.26 & 0.08 & 0.11 & 0.41 & 0.00\\
+DescriptiveNorms\_05 → DescriptiveNorms\_09 & 0.29 & 0.08 & 0.13 & 0.45 & 0.00\\
+DescriptiveNorms\_05 → FactualBeliefs\_09 & 0.19 & 0.09 & 0.01 & 0.37 & 0.04\\
+DescriptiveNorms\_05 → PersonalBeliefs\_09 & 0.29 & 0.09 & 0.12 & 0.46 & 0.00\\
+DescriptiveNorms\_09 → MaskWearing\_11 & 0.08 & 0.09 & -0.09 & 0.25 & 0.36\\
+DescriptiveNorms\_09 → InjunctiveNorms\_11 & 0.19 & 0.07 & 0.06 & 0.33 & 0.01\\
+DescriptiveNorms\_09 → DescriptiveNorms\_11 & 0.24 & 0.07 & 0.11 & 0.37 & 0.00\\
+DescriptiveNorms\_09 → FactualBeliefs\_11 & 0.05 & 0.09 & -0.12 & 0.22 & 0.59\\
+DescriptiveNorms\_09 → PersonalBeliefs\_11 & 0.07 & 0.07 & -0.07 & 0.22 & 0.31\\
+DescriptiveNorms\_11 → MaskWearing\_13 & 0.14 & 0.08 & -0.02 & 0.30 & 0.09\\
+DescriptiveNorms\_11 → InjunctiveNorms\_13 & 0.01 & 0.07 & -0.13 & 0.16 & 0.86\\
+DescriptiveNorms\_11 → DescriptiveNorms\_13 & 0.31 & 0.08 & 0.15 & 0.46 & 0.00\\
+DescriptiveNorms\_11 → FactualBeliefs\_13 & 0.11 & 0.08 & -0.05 & 0.28 & 0.17\\
+DescriptiveNorms\_11 → PersonalBeliefs\_13 & 0.07 & 0.08 & -0.08 & 0.23 & 0.34\\
+DescriptiveNorms\_13 → MaskWearing\_14 & 0.15 & 0.07 & 0.01 & 0.29 & 0.03\\
+DescriptiveNorms\_13 → InjunctiveNorms\_14 & 0.21 & 0.06 & 0.09 & 0.33 & 0.00\\
+DescriptiveNorms\_13 → DescriptiveNorms\_14 & 0.41 & 0.06 & 0.29 & 0.53 & 0.00\\
+DescriptiveNorms\_13 → FactualBeliefs\_14 & 0.04 & 0.07 & -0.09 & 0.17 & 0.54\\
+DescriptiveNorms\_13 → PersonalBeliefs\_14 & 0.01 & 0.06 & -0.10 & 0.12 & 0.85\\
+DescriptiveNorms\_14 → MaskWearing\_15 & 0.09 & 0.09 & -0.08 & 0.26 & 0.31\\
+DescriptiveNorms\_14 → InjunctiveNorms\_15 & -0.03 & 0.08 & -0.18 & 0.11 & 0.65\\
+DescriptiveNorms\_14 → DescriptiveNorms\_15 & 0.34 & 0.07 & 0.20 & 0.47 & 0.00\\
+DescriptiveNorms\_14 → FactualBeliefs\_15 & 0.13 & 0.09 & -0.04 & 0.30 & 0.13\\
+DescriptiveNorms\_14 → PersonalBeliefs\_15 & 0.07 & 0.08 & -0.09 & 0.24 & 0.38\\
+DescriptiveNorms\_15 → MaskWearing\_16 & 0.03 & 0.09 & -0.14 & 0.20 & 0.69\\
+DescriptiveNorms\_15 → InjunctiveNorms\_16 & 0.21 & 0.07 & 0.06 & 0.35 & 0.00\\
+DescriptiveNorms\_15 → DescriptiveNorms\_16 & 0.25 & 0.06 & 0.13 & 0.37 & 0.00\\
+DescriptiveNorms\_15 → FactualBeliefs\_16 & 0.20 & 0.09 & 0.02 & 0.38 & 0.03\\
+DescriptiveNorms\_15 → PersonalBeliefs\_16 & 0.03 & 0.08 & -0.13 & 0.18 & 0.72\\
+DescriptiveNorms\_16 → MaskWearing\_17 & 0.12 & 0.12 & -0.12 & 0.36 & 0.33\\
+DescriptiveNorms\_16 → InjunctiveNorms\_17 & 0.22 & 0.08 & 0.07 & 0.37 & 0.01\\
+DescriptiveNorms\_16 → DescriptiveNorms\_17 & 0.47 & 0.07 & 0.34 & 0.60 & 0.00\\
+DescriptiveNorms\_16 → FactualBeliefs\_17 & 0.14 & 0.13 & -0.10 & 0.39 & 0.26\\
+DescriptiveNorms\_16 → PersonalBeliefs\_17 & 0.08 & 0.11 & -0.13 & 0.30 & 0.45\\
+DescriptiveNorms\_17 → MaskWearing\_18 & 0.03 & 0.12 & -0.20 & 0.26 & 0.81\\
+DescriptiveNorms\_17 → InjunctiveNorms\_18 & 0.28 & 0.07 & 0.14 & 0.42 & 0.00\\
+DescriptiveNorms\_17 → DescriptiveNorms\_18 & 0.41 & 0.07 & 0.29 & 0.54 & 0.00\\
+DescriptiveNorms\_17 → FactualBeliefs\_18 & 0.16 & 0.13 & -0.09 & 0.41 & 0.22\\
+DescriptiveNorms\_17 → PersonalBeliefs\_18 & 0.13 & 0.11 & -0.09 & 0.36 & 0.24\\
+FactualBeliefs\_02 → MaskWearing\_05 & 0.07 & 0.10 & -0.13 & 0.26 & 0.50\\
+FactualBeliefs\_02 → InjunctiveNorms\_05 & -0.15 & 0.08 & -0.30 & 0.01 & 0.06\\
+FactualBeliefs\_02 → DescriptiveNorms\_05 & -0.04 & 0.08 & -0.20 & 0.12 & 0.62\\
+FactualBeliefs\_02 → FactualBeliefs\_05 & 0.20 & 0.07 & 0.05 & 0.34 & 0.01\\
+FactualBeliefs\_02 → PersonalBeliefs\_05 & -0.09 & 0.07 & -0.22 & 0.05 & 0.20\\
+FactualBeliefs\_05 → MaskWearing\_09 & 0.29 & 0.12 & 0.05 & 0.53 & 0.02\\
+FactualBeliefs\_05 → InjunctiveNorms\_09 & -0.06 & 0.11 & -0.27 & 0.16 & 0.60\\
+FactualBeliefs\_05 → DescriptiveNorms\_09 & -0.06 & 0.12 & -0.29 & 0.17 & 0.60\\
+FactualBeliefs\_05 → FactualBeliefs\_09 & 0.21 & 0.14 & -0.05 & 0.48 & 0.12\\
+FactualBeliefs\_05 → PersonalBeliefs\_09 & 0.03 & 0.13 & -0.21 & 0.28 & 0.80\\
+FactualBeliefs\_09 → MaskWearing\_11 & 0.16 & 0.09 & -0.02 & 0.35 & 0.08\\
+FactualBeliefs\_09 → InjunctiveNorms\_11 & 0.03 & 0.07 & -0.11 & 0.18 & 0.63\\
+FactualBeliefs\_09 → DescriptiveNorms\_11 & 0.08 & 0.07 & -0.06 & 0.21 & 0.25\\
+FactualBeliefs\_09 → FactualBeliefs\_11 & 0.35 & 0.09 & 0.17 & 0.53 & 0.00\\
+FactualBeliefs\_09 → PersonalBeliefs\_11 & 0.15 & 0.08 & 0.01 & 0.30 & 0.04\\
+FactualBeliefs\_11 → MaskWearing\_13 & 0.12 & 0.07 & -0.01 & 0.26 & 0.07\\
+FactualBeliefs\_11 → InjunctiveNorms\_13 & 0.03 & 0.06 & -0.09 & 0.15 & 0.59\\
+FactualBeliefs\_11 → DescriptiveNorms\_13 & -0.09 & 0.06 & -0.22 & 0.03 & 0.14\\
+FactualBeliefs\_11 → FactualBeliefs\_13 & 0.13 & 0.07 & -0.01 & 0.26 & 0.06\\
+FactualBeliefs\_11 → PersonalBeliefs\_13 & 0.10 & 0.06 & -0.03 & 0.22 & 0.12\\
+FactualBeliefs\_13 → MaskWearing\_14 & 0.03 & 0.08 & -0.13 & 0.19 & 0.71\\
+FactualBeliefs\_13 → InjunctiveNorms\_14 & 0.05 & 0.07 & -0.09 & 0.18 & 0.49\\
+FactualBeliefs\_13 → DescriptiveNorms\_14 & 0.00 & 0.07 & -0.14 & 0.13 & 0.96\\
+FactualBeliefs\_13 → FactualBeliefs\_14 & 0.27 & 0.08 & 0.12 & 0.42 & 0.00\\
+FactualBeliefs\_13 → PersonalBeliefs\_14 & 0.16 & 0.07 & 0.03 & 0.29 & 0.02\\
+FactualBeliefs\_14 → MaskWearing\_15 & 0.30 & 0.08 & 0.14 & 0.45 & 0.00\\
+FactualBeliefs\_14 → InjunctiveNorms\_15 & -0.03 & 0.07 & -0.17 & 0.11 & 0.67\\
+FactualBeliefs\_14 → DescriptiveNorms\_15 & 0.17 & 0.07 & 0.04 & 0.30 & 0.01\\
+FactualBeliefs\_14 → FactualBeliefs\_15 & 0.53 & 0.08 & 0.37 & 0.68 & 0.00\\
+FactualBeliefs\_14 → PersonalBeliefs\_15 & 0.34 & 0.08 & 0.18 & 0.49 & 0.00\\
+FactualBeliefs\_15 → MaskWearing\_16 & 0.08 & 0.09 & -0.09 & 0.26 & 0.36\\
+FactualBeliefs\_15 → InjunctiveNorms\_16 & 0.08 & 0.08 & -0.06 & 0.23 & 0.27\\
+FactualBeliefs\_15 → DescriptiveNorms\_16 & 0.09 & 0.07 & -0.04 & 0.22 & 0.19\\
+FactualBeliefs\_15 → FactualBeliefs\_16 & 0.42 & 0.09 & 0.23 & 0.61 & 0.00\\
+FactualBeliefs\_15 → PersonalBeliefs\_16 & 0.12 & 0.08 & -0.04 & 0.29 & 0.13\\
+FactualBeliefs\_16 → MaskWearing\_17 & 0.18 & 0.09 & 0.01 & 0.35 & 0.04\\
+FactualBeliefs\_16 → InjunctiveNorms\_17 & 0.02 & 0.06 & -0.09 & 0.13 & 0.78\\
+FactualBeliefs\_16 → DescriptiveNorms\_17 & -0.01 & 0.05 & -0.11 & 0.09 & 0.82\\
+FactualBeliefs\_16 → FactualBeliefs\_17 & 0.24 & 0.09 & 0.06 & 0.42 & 0.01\\
+FactualBeliefs\_16 → PersonalBeliefs\_17 & 0.08 & 0.08 & -0.08 & 0.24 & 0.33\\
+FactualBeliefs\_17 → MaskWearing\_18 & 0.08 & 0.08 & -0.08 & 0.23 & 0.34\\
+FactualBeliefs\_17 → InjunctiveNorms\_18 & -0.05 & 0.05 & -0.15 & 0.04 & 0.26\\
+FactualBeliefs\_17 → DescriptiveNorms\_18 & 0.04 & 0.04 & -0.05 & 0.13 & 0.38\\
+FactualBeliefs\_17 → FactualBeliefs\_18 & 0.36 & 0.09 & 0.19 & 0.53 & 0.00\\
+FactualBeliefs\_17 → PersonalBeliefs\_18 & 0.40 & 0.08 & 0.25 & 0.55 & 0.00\\
+PersonalBeliefs\_02 → MaskWearing\_05 & 0.06 & 0.11 & -0.16 & 0.28 & 0.58\\
+PersonalBeliefs\_02 → InjunctiveNorms\_05 & 0.09 & 0.09 & -0.09 & 0.26 & 0.33\\
+PersonalBeliefs\_02 → DescriptiveNorms\_05 & 0.07 & 0.09 & -0.11 & 0.24 & 0.45\\
+PersonalBeliefs\_02 → FactualBeliefs\_05 & 0.16 & 0.08 & -0.01 & 0.32 & 0.06\\
+PersonalBeliefs\_02 → PersonalBeliefs\_05 & 0.41 & 0.08 & 0.25 & 0.57 & 0.00\\
+PersonalBeliefs\_05 → MaskWearing\_09 & -0.45 & 0.13 & -0.70 & -0.19 & 0.00\\
+PersonalBeliefs\_05 → InjunctiveNorms\_09 & -0.26 & 0.12 & -0.49 & -0.03 & 0.03\\
+PersonalBeliefs\_05 → DescriptiveNorms\_09 & -0.09 & 0.13 & -0.33 & 0.16 & 0.50\\
+PersonalBeliefs\_05 → FactualBeliefs\_09 & -0.40 & 0.14 & -0.68 & -0.12 & 0.00\\
+PersonalBeliefs\_05 → PersonalBeliefs\_09 & -0.34 & 0.13 & -0.60 & -0.07 & 0.01\\
+PersonalBeliefs\_09 → MaskWearing\_11 & 0.08 & 0.10 & -0.12 & 0.27 & 0.45\\
+PersonalBeliefs\_09 → InjunctiveNorms\_11 & 0.06 & 0.08 & -0.09 & 0.22 & 0.44\\
+PersonalBeliefs\_09 → DescriptiveNorms\_11 & 0.04 & 0.07 & -0.11 & 0.18 & 0.60\\
+PersonalBeliefs\_09 → FactualBeliefs\_11 & 0.10 & 0.10 & -0.10 & 0.29 & 0.32\\
+PersonalBeliefs\_09 → PersonalBeliefs\_11 & 0.19 & 0.08 & 0.02 & 0.35 & 0.02\\
+PersonalBeliefs\_11 → MaskWearing\_13 & 0.06 & 0.08 & -0.09 & 0.21 & 0.41\\
+PersonalBeliefs\_11 → InjunctiveNorms\_13 & 0.07 & 0.07 & -0.06 & 0.19 & 0.33\\
+PersonalBeliefs\_11 → DescriptiveNorms\_13 & 0.11 & 0.07 & -0.03 & 0.25 & 0.13\\
+PersonalBeliefs\_11 → FactualBeliefs\_13 & 0.18 & 0.07 & 0.03 & 0.33 & 0.02\\
+PersonalBeliefs\_11 → PersonalBeliefs\_13 & 0.17 & 0.07 & 0.04 & 0.31 & 0.01\\
+PersonalBeliefs\_13 → MaskWearing\_14 & 0.27 & 0.09 & 0.09 & 0.44 & 0.00\\
+PersonalBeliefs\_13 → InjunctiveNorms\_14 & -0.10 & 0.08 & -0.24 & 0.05 & 0.20\\
+PersonalBeliefs\_13 → DescriptiveNorms\_14 & -0.02 & 0.08 & -0.17 & 0.13 & 0.80\\
+PersonalBeliefs\_13 → FactualBeliefs\_14 & 0.36 & 0.08 & 0.20 & 0.53 & 0.00\\
+PersonalBeliefs\_13 → PersonalBeliefs\_14 & 0.44 & 0.07 & 0.29 & 0.58 & 0.00\\
+PersonalBeliefs\_14 → MaskWearing\_15 & -0.06 & 0.09 & -0.24 & 0.12 & 0.53\\
+PersonalBeliefs\_14 → InjunctiveNorms\_15 & 0.12 & 0.08 & -0.05 & 0.28 & 0.17\\
+PersonalBeliefs\_14 → DescriptiveNorms\_15 & -0.12 & 0.07 & -0.26 & 0.03 & 0.12\\
+PersonalBeliefs\_14 → FactualBeliefs\_15 & 0.01 & 0.09 & -0.17 & 0.19 & 0.89\\
+PersonalBeliefs\_14 → PersonalBeliefs\_15 & 0.14 & 0.09 & -0.04 & 0.32 & 0.14\\
+PersonalBeliefs\_15 → MaskWearing\_16 & 0.09 & 0.09 & -0.08 & 0.27 & 0.31\\
+PersonalBeliefs\_15 → InjunctiveNorms\_16 & 0.05 & 0.08 & -0.10 & 0.20 & 0.52\\
+PersonalBeliefs\_15 → DescriptiveNorms\_16 & 0.10 & 0.07 & -0.03 & 0.23 & 0.11\\
+PersonalBeliefs\_15 → FactualBeliefs\_16 & 0.13 & 0.10 & -0.06 & 0.32 & 0.17\\
+PersonalBeliefs\_15 → PersonalBeliefs\_16 & 0.41 & 0.08 & 0.24 & 0.57 & 0.00\\
+PersonalBeliefs\_16 → MaskWearing\_17 & -0.02 & 0.09 & -0.20 & 0.16 & 0.83\\
+PersonalBeliefs\_16 → InjunctiveNorms\_17 & 0.00 & 0.06 & -0.12 & 0.12 & 0.98\\
+PersonalBeliefs\_16 → DescriptiveNorms\_17 & -0.02 & 0.05 & -0.12 & 0.08 & 0.73\\
+PersonalBeliefs\_16 → FactualBeliefs\_17 & 0.28 & 0.10 & 0.08 & 0.47 & 0.00\\
+PersonalBeliefs\_16 → PersonalBeliefs\_17 & 0.58 & 0.08 & 0.42 & 0.75 & 0.00\\
+PersonalBeliefs\_17 → MaskWearing\_18 & 0.09 & 0.08 & -0.07 & 0.24 & 0.26\\
+PersonalBeliefs\_17 → InjunctiveNorms\_18 & 0.00 & 0.05 & -0.10 & 0.09 & 0.96\\
+PersonalBeliefs\_17 → DescriptiveNorms\_18 & -0.04 & 0.04 & -0.12 & 0.05 & 0.41\\
+PersonalBeliefs\_17 → FactualBeliefs\_18 & 0.16 & 0.09 & -0.01 & 0.33 & 0.07\\
+PersonalBeliefs\_17 → PersonalBeliefs\_18 & 0.10 & 0.08 & -0.05 & 0.25 & 0.21\\
 \bottomrule
@@
 
-\hypertarget{supplementary-references}{%
-\subsection{Supplementary References}\label{supplementary-references}}
-
-Pearl, J. (1995). Causal diagrams for empirical research. \emph{Biometrika}, \emph{82}(4), 669--688, \url{https://doi.org/10.1093/biomet/82.4.669}
+\begin{center}
+\begin{ThreePartTable}
+
+\footnotesize{
+
+\begin{longtable}{lccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:unconstrainedTable2}Unstandardized autoregressive and cross-lagged parameters from time-varying random-intercept cross-lagged panel model (with factual beliefs and personal normative beliefs removed).
+\emph{Variable name suffixes indicate time points.
+Arrows indicate the direction of prediction.}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:unconstrainedTable2} continued}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endhead
+MaskWearing\_02 → MaskWearing\_05 & 0.25 & 0.06 & 0.13 & 0.37 & 0.00\\
+MaskWearing\_02 → InjunctiveNorms\_05 & 0.04 & 0.05 & -0.06 & 0.13 & 0.46\\
+MaskWearing\_02 → DescriptiveNorms\_05 & 0.03 & 0.05 & -0.07 & 0.12 & 0.60\\
+MaskWearing\_05 → MaskWearing\_09 & 0.15 & 0.07 & 0.03 & 0.28 & 0.02\\
+MaskWearing\_05 → InjunctiveNorms\_09 & 0.09 & 0.06 & -0.02 & 0.21 & 0.11\\
+MaskWearing\_05 → DescriptiveNorms\_09 & -0.03 & 0.06 & -0.15 & 0.09 & 0.61\\
+MaskWearing\_09 → MaskWearing\_11 & -0.08 & 0.08 & -0.24 & 0.08 & 0.31\\
+MaskWearing\_09 → InjunctiveNorms\_11 & 0.03 & 0.06 & -0.09 & 0.15 & 0.61\\
+MaskWearing\_09 → DescriptiveNorms\_11 & 0.09 & 0.06 & -0.02 & 0.21 & 0.11\\
+MaskWearing\_11 → MaskWearing\_13 & 0.07 & 0.06 & -0.04 & 0.18 & 0.22\\
+MaskWearing\_11 → InjunctiveNorms\_13 & 0.04 & 0.05 & -0.06 & 0.13 & 0.44\\
+MaskWearing\_11 → DescriptiveNorms\_13 & 0.02 & 0.05 & -0.08 & 0.12 & 0.71\\
+MaskWearing\_13 → MaskWearing\_14 & 0.20 & 0.07 & 0.08 & 0.33 & 0.00\\
+MaskWearing\_13 → InjunctiveNorms\_14 & 0.05 & 0.05 & -0.06 & 0.16 & 0.36\\
+MaskWearing\_13 → DescriptiveNorms\_14 & 0.08 & 0.06 & -0.03 & 0.19 & 0.13\\
+MaskWearing\_14 → MaskWearing\_15 & 0.25 & 0.07 & 0.12 & 0.38 & 0.00\\
+MaskWearing\_14 → InjunctiveNorms\_15 & 0.05 & 0.06 & -0.06 & 0.16 & 0.38\\
+MaskWearing\_14 → DescriptiveNorms\_15 & 0.09 & 0.05 & -0.01 & 0.19 & 0.07\\
+MaskWearing\_15 → MaskWearing\_16 & 0.32 & 0.07 & 0.18 & 0.46 & 0.00\\
+MaskWearing\_15 → InjunctiveNorms\_16 & 0.04 & 0.06 & -0.08 & 0.16 & 0.52\\
+MaskWearing\_15 → DescriptiveNorms\_16 & 0.05 & 0.05 & -0.06 & 0.15 & 0.36\\
+MaskWearing\_16 → MaskWearing\_17 & 0.42 & 0.07 & 0.29 & 0.56 & 0.00\\
+MaskWearing\_16 → InjunctiveNorms\_17 & 0.00 & 0.04 & -0.08 & 0.09 & 0.98\\
+MaskWearing\_16 → DescriptiveNorms\_17 & 0.08 & 0.04 & 0.01 & 0.15 & 0.03\\
+MaskWearing\_17 → MaskWearing\_18 & 0.43 & 0.05 & 0.33 & 0.54 & 0.00\\
+MaskWearing\_17 → InjunctiveNorms\_18 & -0.08 & 0.03 & -0.15 & -0.01 & 0.02\\
+MaskWearing\_17 → DescriptiveNorms\_18 & 0.03 & 0.03 & -0.03 & 0.09 & 0.37\\
+InjunctiveNorms\_02 → MaskWearing\_05 & 0.01 & 0.07 & -0.13 & 0.15 & 0.92\\
+InjunctiveNorms\_02 → InjunctiveNorms\_05 & 0.31 & 0.06 & 0.20 & 0.43 & 0.00\\
+InjunctiveNorms\_02 → DescriptiveNorms\_05 & 0.07 & 0.06 & -0.04 & 0.19 & 0.22\\
+InjunctiveNorms\_05 → MaskWearing\_09 & -0.10 & 0.08 & -0.26 & 0.06 & 0.23\\
+InjunctiveNorms\_05 → InjunctiveNorms\_09 & 0.07 & 0.08 & -0.08 & 0.22 & 0.38\\
+InjunctiveNorms\_05 → DescriptiveNorms\_09 & 0.00 & 0.08 & -0.15 & 0.16 & 0.96\\
+InjunctiveNorms\_09 → MaskWearing\_11 & 0.10 & 0.10 & -0.09 & 0.29 & 0.29\\
+InjunctiveNorms\_09 → InjunctiveNorms\_11 & 0.14 & 0.08 & -0.01 & 0.29 & 0.07\\
+InjunctiveNorms\_09 → DescriptiveNorms\_11 & -0.01 & 0.07 & -0.15 & 0.13 & 0.92\\
+InjunctiveNorms\_11 → MaskWearing\_13 & 0.02 & 0.07 & -0.12 & 0.17 & 0.75\\
+InjunctiveNorms\_11 → InjunctiveNorms\_13 & 0.32 & 0.07 & 0.19 & 0.45 & 0.00\\
+InjunctiveNorms\_11 → DescriptiveNorms\_13 & 0.22 & 0.07 & 0.08 & 0.36 & 0.00\\
+InjunctiveNorms\_13 → MaskWearing\_14 & 0.00 & 0.08 & -0.16 & 0.16 & 0.97\\
+InjunctiveNorms\_13 → InjunctiveNorms\_14 & 0.39 & 0.07 & 0.26 & 0.53 & 0.00\\
+InjunctiveNorms\_13 → DescriptiveNorms\_14 & 0.16 & 0.07 & 0.02 & 0.29 & 0.02\\
+InjunctiveNorms\_14 → MaskWearing\_15 & 0.06 & 0.08 & -0.11 & 0.22 & 0.50\\
+InjunctiveNorms\_14 → InjunctiveNorms\_15 & 0.43 & 0.07 & 0.29 & 0.58 & 0.00\\
+InjunctiveNorms\_14 → DescriptiveNorms\_15 & 0.30 & 0.07 & 0.17 & 0.43 & 0.00\\
+InjunctiveNorms\_15 → MaskWearing\_16 & 0.22 & 0.09 & 0.05 & 0.39 & 0.01\\
+InjunctiveNorms\_15 → InjunctiveNorms\_16 & 0.20 & 0.07 & 0.06 & 0.35 & 0.01\\
+InjunctiveNorms\_15 → DescriptiveNorms\_16 & 0.08 & 0.06 & -0.05 & 0.21 & 0.22\\
+InjunctiveNorms\_16 → MaskWearing\_17 & 0.03 & 0.10 & -0.17 & 0.22 & 0.80\\
+InjunctiveNorms\_16 → InjunctiveNorms\_17 & 0.32 & 0.06 & 0.19 & 0.44 & 0.00\\
+InjunctiveNorms\_16 → DescriptiveNorms\_17 & 0.10 & 0.05 & 0.00 & 0.21 & 0.06\\
+InjunctiveNorms\_17 → MaskWearing\_18 & 0.01 & 0.10 & -0.20 & 0.21 & 0.93\\
+InjunctiveNorms\_17 → InjunctiveNorms\_18 & 0.37 & 0.06 & 0.24 & 0.49 & 0.00\\
+InjunctiveNorms\_17 → DescriptiveNorms\_18 & 0.19 & 0.06 & 0.08 & 0.30 & 0.00\\
+DescriptiveNorms\_02 → MaskWearing\_05 & 0.19 & 0.07 & 0.05 & 0.33 & 0.01\\
+DescriptiveNorms\_02 → InjunctiveNorms\_05 & 0.14 & 0.06 & 0.03 & 0.26 & 0.01\\
+DescriptiveNorms\_02 → DescriptiveNorms\_05 & 0.40 & 0.06 & 0.29 & 0.52 & 0.00\\
+DescriptiveNorms\_05 → MaskWearing\_09 & 0.21 & 0.08 & 0.05 & 0.36 & 0.01\\
+DescriptiveNorms\_05 → InjunctiveNorms\_09 & 0.19 & 0.07 & 0.05 & 0.33 & 0.01\\
+DescriptiveNorms\_05 → DescriptiveNorms\_09 & 0.27 & 0.08 & 0.12 & 0.42 & 0.00\\
+DescriptiveNorms\_09 → MaskWearing\_11 & 0.11 & 0.09 & -0.06 & 0.28 & 0.22\\
+DescriptiveNorms\_09 → InjunctiveNorms\_11 & 0.20 & 0.07 & 0.06 & 0.33 & 0.00\\
+DescriptiveNorms\_09 → DescriptiveNorms\_11 & 0.26 & 0.07 & 0.13 & 0.38 & 0.00\\
+DescriptiveNorms\_11 → MaskWearing\_13 & 0.18 & 0.08 & 0.03 & 0.34 & 0.02\\
+DescriptiveNorms\_11 → InjunctiveNorms\_13 & 0.03 & 0.07 & -0.10 & 0.17 & 0.63\\
+DescriptiveNorms\_11 → DescriptiveNorms\_13 & 0.29 & 0.08 & 0.14 & 0.44 & 0.00\\
+DescriptiveNorms\_13 → MaskWearing\_14 & 0.17 & 0.07 & 0.03 & 0.31 & 0.02\\
+DescriptiveNorms\_13 → InjunctiveNorms\_14 & 0.21 & 0.06 & 0.09 & 0.33 & 0.00\\
+DescriptiveNorms\_13 → DescriptiveNorms\_14 & 0.41 & 0.06 & 0.29 & 0.53 & 0.00\\
+DescriptiveNorms\_14 → MaskWearing\_15 & 0.13 & 0.09 & -0.04 & 0.30 & 0.15\\
+DescriptiveNorms\_14 → InjunctiveNorms\_15 & -0.01 & 0.08 & -0.15 & 0.14 & 0.94\\
+DescriptiveNorms\_14 → DescriptiveNorms\_15 & 0.34 & 0.07 & 0.21 & 0.47 & 0.00\\
+DescriptiveNorms\_15 → MaskWearing\_16 & 0.07 & 0.09 & -0.10 & 0.24 & 0.44\\
+DescriptiveNorms\_15 → InjunctiveNorms\_16 & 0.25 & 0.07 & 0.11 & 0.40 & 0.00\\
+DescriptiveNorms\_15 → DescriptiveNorms\_16 & 0.28 & 0.06 & 0.15 & 0.41 & 0.00\\
+DescriptiveNorms\_16 → MaskWearing\_17 & 0.18 & 0.12 & -0.06 & 0.41 & 0.14\\
+DescriptiveNorms\_16 → InjunctiveNorms\_17 & 0.24 & 0.08 & 0.09 & 0.39 & 0.00\\
+DescriptiveNorms\_16 → DescriptiveNorms\_17 & 0.45 & 0.06 & 0.33 & 0.58 & 0.00\\
+DescriptiveNorms\_17 → MaskWearing\_18 & 0.10 & 0.12 & -0.12 & 0.33 & 0.37\\
+DescriptiveNorms\_17 → InjunctiveNorms\_18 & 0.28 & 0.07 & 0.14 & 0.42 & 0.00\\
+DescriptiveNorms\_17 → DescriptiveNorms\_18 & 0.42 & 0.06 & 0.30 & 0.55 & 0.00\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
+
+\begin{center}
+\begin{ThreePartTable}
+
+\footnotesize{
+
+\begin{longtable}{lccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:unconstrainedTable3}Unstandardized autoregressive and cross-lagged parameters from time-varying random-intercept cross-lagged panel model (with factual beliefs, personal normative beliefs, and all exogenous covariates removed).
+\emph{Variable name suffixes indicate time points.
+Arrows indicate the direction of prediction.}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:unconstrainedTable3} continued}}\\
+\toprule
+Parameter & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endhead
+MaskWearing\_02 → MaskWearing\_05 & 0.24 & 0.06 & 0.12 & 0.36 & 0.00\\
+MaskWearing\_02 → InjunctiveNorms\_05 & 0.08 & 0.05 & -0.02 & 0.18 & 0.13\\
+MaskWearing\_02 → DescriptiveNorms\_05 & 0.04 & 0.05 & -0.06 & 0.14 & 0.46\\
+MaskWearing\_05 → MaskWearing\_09 & 0.16 & 0.07 & 0.03 & 0.29 & 0.02\\
+MaskWearing\_05 → InjunctiveNorms\_09 & 0.09 & 0.06 & -0.03 & 0.20 & 0.15\\
+MaskWearing\_05 → DescriptiveNorms\_09 & 0.00 & 0.06 & -0.13 & 0.12 & 0.98\\
+MaskWearing\_09 → MaskWearing\_11 & -0.07 & 0.08 & -0.23 & 0.09 & 0.38\\
+MaskWearing\_09 → InjunctiveNorms\_11 & 0.03 & 0.06 & -0.09 & 0.15 & 0.58\\
+MaskWearing\_09 → DescriptiveNorms\_11 & 0.13 & 0.06 & 0.02 & 0.24 & 0.02\\
+MaskWearing\_11 → MaskWearing\_13 & 0.11 & 0.05 & 0.00 & 0.22 & 0.05\\
+MaskWearing\_11 → InjunctiveNorms\_13 & 0.08 & 0.05 & -0.01 & 0.18 & 0.07\\
+MaskWearing\_11 → DescriptiveNorms\_13 & 0.07 & 0.05 & -0.03 & 0.16 & 0.18\\
+MaskWearing\_13 → MaskWearing\_14 & 0.23 & 0.06 & 0.11 & 0.36 & 0.00\\
+MaskWearing\_13 → InjunctiveNorms\_14 & 0.06 & 0.05 & -0.04 & 0.17 & 0.23\\
+MaskWearing\_13 → DescriptiveNorms\_14 & 0.13 & 0.05 & 0.02 & 0.24 & 0.02\\
+MaskWearing\_14 → MaskWearing\_15 & 0.27 & 0.07 & 0.13 & 0.40 & 0.00\\
+MaskWearing\_14 → InjunctiveNorms\_15 & 0.05 & 0.06 & -0.06 & 0.16 & 0.36\\
+MaskWearing\_14 → DescriptiveNorms\_15 & 0.12 & 0.05 & 0.02 & 0.22 & 0.02\\
+MaskWearing\_15 → MaskWearing\_16 & 0.36 & 0.07 & 0.22 & 0.49 & 0.00\\
+MaskWearing\_15 → InjunctiveNorms\_16 & 0.06 & 0.06 & -0.05 & 0.18 & 0.28\\
+MaskWearing\_15 → DescriptiveNorms\_16 & 0.09 & 0.05 & -0.01 & 0.19 & 0.10\\
+MaskWearing\_16 → MaskWearing\_17 & 0.42 & 0.06 & 0.29 & 0.54 & 0.00\\
+MaskWearing\_16 → InjunctiveNorms\_17 & -0.02 & 0.04 & -0.10 & 0.06 & 0.60\\
+MaskWearing\_16 → DescriptiveNorms\_17 & 0.08 & 0.04 & 0.01 & 0.15 & 0.03\\
+MaskWearing\_17 → MaskWearing\_18 & 0.46 & 0.05 & 0.35 & 0.56 & 0.00\\
+MaskWearing\_17 → InjunctiveNorms\_18 & -0.06 & 0.03 & -0.12 & 0.00 & 0.07\\
+MaskWearing\_17 → DescriptiveNorms\_18 & 0.02 & 0.03 & -0.03 & 0.08 & 0.42\\
+InjunctiveNorms\_02 → MaskWearing\_05 & -0.02 & 0.07 & -0.17 & 0.12 & 0.73\\
+InjunctiveNorms\_02 → InjunctiveNorms\_05 & 0.29 & 0.06 & 0.18 & 0.41 & 0.00\\
+InjunctiveNorms\_02 → DescriptiveNorms\_05 & 0.08 & 0.06 & -0.04 & 0.19 & 0.20\\
+InjunctiveNorms\_05 → MaskWearing\_09 & -0.13 & 0.08 & -0.28 & 0.03 & 0.11\\
+InjunctiveNorms\_05 → InjunctiveNorms\_09 & 0.07 & 0.08 & -0.07 & 0.22 & 0.32\\
+InjunctiveNorms\_05 → DescriptiveNorms\_09 & -0.03 & 0.08 & -0.18 & 0.13 & 0.74\\
+InjunctiveNorms\_09 → MaskWearing\_11 & 0.16 & 0.10 & -0.03 & 0.35 & 0.10\\
+InjunctiveNorms\_09 → InjunctiveNorms\_11 & 0.16 & 0.07 & 0.01 & 0.30 & 0.03\\
+InjunctiveNorms\_09 → DescriptiveNorms\_11 & 0.01 & 0.07 & -0.12 & 0.15 & 0.84\\
+InjunctiveNorms\_11 → MaskWearing\_13 & 0.05 & 0.07 & -0.09 & 0.20 & 0.48\\
+InjunctiveNorms\_11 → InjunctiveNorms\_13 & 0.31 & 0.07 & 0.18 & 0.44 & 0.00\\
+InjunctiveNorms\_11 → DescriptiveNorms\_13 & 0.22 & 0.07 & 0.09 & 0.36 & 0.00\\
+InjunctiveNorms\_13 → MaskWearing\_14 & 0.01 & 0.08 & -0.14 & 0.17 & 0.87\\
+InjunctiveNorms\_13 → InjunctiveNorms\_14 & 0.45 & 0.07 & 0.32 & 0.58 & 0.00\\
+InjunctiveNorms\_13 → DescriptiveNorms\_14 & 0.21 & 0.07 & 0.07 & 0.34 & 0.00\\
+InjunctiveNorms\_14 → MaskWearing\_15 & 0.11 & 0.08 & -0.05 & 0.27 & 0.19\\
+InjunctiveNorms\_14 → InjunctiveNorms\_15 & 0.45 & 0.07 & 0.32 & 0.59 & 0.00\\
+InjunctiveNorms\_14 → DescriptiveNorms\_15 & 0.33 & 0.06 & 0.20 & 0.45 & 0.00\\
+InjunctiveNorms\_15 → MaskWearing\_16 & 0.22 & 0.09 & 0.05 & 0.39 & 0.01\\
+InjunctiveNorms\_15 → InjunctiveNorms\_16 & 0.23 & 0.07 & 0.09 & 0.38 & 0.00\\
+InjunctiveNorms\_15 → DescriptiveNorms\_16 & 0.12 & 0.06 & -0.01 & 0.25 & 0.06\\
+InjunctiveNorms\_16 → MaskWearing\_17 & 0.02 & 0.10 & -0.17 & 0.21 & 0.83\\
+InjunctiveNorms\_16 → InjunctiveNorms\_17 & 0.34 & 0.06 & 0.22 & 0.46 & 0.00\\
+InjunctiveNorms\_16 → DescriptiveNorms\_17 & 0.11 & 0.05 & 0.00 & 0.21 & 0.05\\
+InjunctiveNorms\_17 → MaskWearing\_18 & 0.02 & 0.10 & -0.18 & 0.21 & 0.87\\
+InjunctiveNorms\_17 → InjunctiveNorms\_18 & 0.42 & 0.06 & 0.31 & 0.54 & 0.00\\
+InjunctiveNorms\_17 → DescriptiveNorms\_18 & 0.19 & 0.05 & 0.08 & 0.29 & 0.00\\
+DescriptiveNorms\_02 → MaskWearing\_05 & 0.22 & 0.07 & 0.09 & 0.36 & 0.00\\
+DescriptiveNorms\_02 → InjunctiveNorms\_05 & 0.16 & 0.06 & 0.04 & 0.27 & 0.01\\
+DescriptiveNorms\_02 → DescriptiveNorms\_05 & 0.40 & 0.06 & 0.29 & 0.51 & 0.00\\
+DescriptiveNorms\_05 → MaskWearing\_09 & 0.23 & 0.08 & 0.07 & 0.38 & 0.00\\
+DescriptiveNorms\_05 → InjunctiveNorms\_09 & 0.20 & 0.07 & 0.06 & 0.35 & 0.00\\
+DescriptiveNorms\_05 → DescriptiveNorms\_09 & 0.30 & 0.08 & 0.14 & 0.45 & 0.00\\
+DescriptiveNorms\_09 → MaskWearing\_11 & 0.14 & 0.09 & -0.04 & 0.31 & 0.12\\
+DescriptiveNorms\_09 → InjunctiveNorms\_11 & 0.23 & 0.07 & 0.10 & 0.36 & 0.00\\
+DescriptiveNorms\_09 → DescriptiveNorms\_11 & 0.29 & 0.06 & 0.16 & 0.41 & 0.00\\
+DescriptiveNorms\_11 → MaskWearing\_13 & 0.24 & 0.08 & 0.08 & 0.39 & 0.00\\
+DescriptiveNorms\_11 → InjunctiveNorms\_13 & 0.07 & 0.07 & -0.06 & 0.21 & 0.29\\
+DescriptiveNorms\_11 → DescriptiveNorms\_13 & 0.30 & 0.07 & 0.16 & 0.44 & 0.00\\
+DescriptiveNorms\_13 → MaskWearing\_14 & 0.18 & 0.07 & 0.04 & 0.32 & 0.01\\
+DescriptiveNorms\_13 → InjunctiveNorms\_14 & 0.21 & 0.06 & 0.09 & 0.33 & 0.00\\
+DescriptiveNorms\_13 → DescriptiveNorms\_14 & 0.38 & 0.06 & 0.26 & 0.51 & 0.00\\
+DescriptiveNorms\_14 → MaskWearing\_15 & 0.13 & 0.09 & -0.04 & 0.30 & 0.14\\
+DescriptiveNorms\_14 → InjunctiveNorms\_15 & 0.04 & 0.07 & -0.10 & 0.18 & 0.61\\
+DescriptiveNorms\_14 → DescriptiveNorms\_15 & 0.36 & 0.07 & 0.23 & 0.49 & 0.00\\
+DescriptiveNorms\_15 → MaskWearing\_16 & 0.10 & 0.09 & -0.06 & 0.27 & 0.22\\
+DescriptiveNorms\_15 → InjunctiveNorms\_16 & 0.27 & 0.07 & 0.13 & 0.41 & 0.00\\
+DescriptiveNorms\_15 → DescriptiveNorms\_16 & 0.29 & 0.06 & 0.17 & 0.42 & 0.00\\
+DescriptiveNorms\_16 → MaskWearing\_17 & 0.16 & 0.12 & -0.07 & 0.39 & 0.17\\
+DescriptiveNorms\_16 → InjunctiveNorms\_17 & 0.26 & 0.07 & 0.12 & 0.40 & 0.00\\
+DescriptiveNorms\_16 → DescriptiveNorms\_17 & 0.47 & 0.06 & 0.35 & 0.60 & 0.00\\
+DescriptiveNorms\_17 → MaskWearing\_18 & 0.21 & 0.11 & -0.01 & 0.42 & 0.06\\
+DescriptiveNorms\_17 → InjunctiveNorms\_18 & 0.30 & 0.06 & 0.17 & 0.42 & 0.00\\
+DescriptiveNorms\_17 → DescriptiveNorms\_18 & 0.46 & 0.06 & 0.34 & 0.58 & 0.00\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
+
+\begin{lltable}
+
+\begin{longtable}{m{2.5cm}m{2cm}m{12.5cm}m{1cm}m{1cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:itemTable}List of norm interpretation questions asked at Time 7.
+\emph{These questions were preceded by the following text:} ``There may or may not be a difference between what people around you are doing and what they should be doing.
+You can learn about what people are doing and what they should be doing in different ways.
+For each of the following information sources, we want to know if you can learn from it what people are doing, what people should be doing, or both''.
+\emph{Participants answered all questions on a 7-point Likert scale, from (1) Not At All to (7) Very Strongly.}}\\
+\toprule
+Interpretation & \multicolumn{1}{c}{Item} & \multicolumn{1}{c}{Question} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:itemTable} continued}}\\
+\toprule
+Interpretation & \multicolumn{1}{c}{Item} & \multicolumn{1}{c}{Question} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD}\\
+\midrule
+\endhead
+Provides descriptive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone is doing?
+& 4.65 & 1.68\\
+ &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone is doing?
+& 4.86 & 1.65\\
+ & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone is doing? & 4.35 & 1.89\\
+ &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone is doing? & 4.16 & 1.88\\
+Provides injunctive information & Descriptive & Does noticing the proportion of people in your area that wear a mask while doing recreational/social activities indoors (e.g., going to the gym, eating at a restaurant, attending a party) tell you what everyone should be doing?
+& 5.09 & 1.90\\
+ &  & Does noticing the proportion of people in your area that wear a mask while doing routine activities indoors (e.g., running errands, shopping, going to work) tell you what everyone should be doing?
+& 5.14 & 1.87\\
+ & Injunctive & Do mask-wearing rules encouraged in your area (e.g., by local or state government officials, businesses, etc.) tell you what everyone should be doing? & 5.70 & 1.73\\
+ &  & Does how often you see people that you respect and trust wearing a mask (e.g., on tv, news, etc.) tell you what everyone should be doing? & 5.38 & 1.75\\
+\bottomrule
+\end{longtable}
+
+\end{lltable}
+
+\newpage
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{longtable}{lllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:constructValTable}Pairwise comparisons from a multilevel model testing the construct validity of the four self-reported social norm items.
+\emph{Data were analysed in long format, with the self-report social norm item and the type of normative information provided (descriptive or injunctive) as predictors and a random intercept for participant.
+All pairwise comparisons account for multiple comparisons using Tukey adjustment.}}\\
+\toprule
+Type & \multicolumn{1}{c}{Pairwise contrast} & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{p}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:constructValTable} continued}}\\
+\toprule
+Type & \multicolumn{1}{c}{Pairwise contrast} & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{p}\\
+\midrule
+\endhead
+Descriptive & MaskEncouraged - MaskRespect & 0.19 & 0.09 & 0.11\\
+Descriptive & MaskEncouraged - NeighborMask1 & -0.51 & 0.09 & 0.00\\
+Descriptive & MaskEncouraged - NeighborMask2 & -0.30 & 0.09 & 0.00\\
+Descriptive & MaskRespect - NeighborMask1 & -0.70 & 0.09 & 0.00\\
+Descriptive & MaskRespect - NeighborMask2 & -0.49 & 0.09 & 0.00\\
+Descriptive & NeighborMask1 - NeighborMask2 & 0.21 & 0.09 & 0.07\\
+Injunctive & MaskEncouraged - MaskRespect & 0.32 & 0.09 & 0.00\\
+Injunctive & MaskEncouraged - NeighborMask1 & 0.55 & 0.09 & 0.00\\
+Injunctive & MaskEncouraged - NeighborMask2 & 0.61 & 0.09 & 0.00\\
+Injunctive & MaskRespect - NeighborMask1 & 0.24 & 0.09 & 0.03\\
+Injunctive & MaskRespect - NeighborMask2 & 0.29 & 0.09 & 0.00\\
+Injunctive & NeighborMask1 - NeighborMask2 & 0.05 & 0.09 & 0.93\\
+\bottomrule
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
+
+\begin{center}
+\begin{ThreePartTable}
+
+\begin{longtable}{llccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:constrainedTable4}Unstandardized autoregressive and cross-lagged parameters from multi-group time-invariant random-intercept cross-lagged panel model, split by majority-Democrat and majority-Republican states.
+\emph{The model contains all covariates.
+Arrows indicate the direction of prediction.}}\\
+\toprule
+Group & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:constrainedTable4} continued}}\\
+\toprule
+Group & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{SE} & \multicolumn{1}{c}{2.5\%} & \multicolumn{1}{c}{97.5\%} & \multicolumn{1}{c}{p}\\
+\midrule
+\endhead
+Democrat & Mask wearing → Mask wearing & 0.20 & 0.04 & 0.13 & 0.27 & 0.00\\
+Democrat & Mask wearing → Injunctive norms & 0.01 & 0.03 & -0.04 & 0.06 & 0.81\\
+Democrat & Mask wearing → Descriptive norms & 0.01 & 0.03 & -0.04 & 0.06 & 0.68\\
+Democrat & Injunctive norms → Mask wearing & -0.02 & 0.04 & -0.11 & 0.06 & 0.59\\
+Democrat & Injunctive norms → Injunctive norms & 0.32 & 0.04 & 0.24 & 0.39 & 0.00\\
+Democrat & Injunctive norms → Descriptive norms & 0.21 & 0.04 & 0.14 & 0.28 & 0.00\\
+Democrat & Descriptive norms → Mask wearing & 0.08 & 0.04 & 0.00 & 0.17 & 0.06\\
+Democrat & Descriptive norms → Injunctive norms & 0.24 & 0.03 & 0.17 & 0.30 & 0.00\\
+Democrat & Descriptive norms → Descriptive norms & 0.33 & 0.04 & 0.25 & 0.40 & 0.00\\
+Republican & Mask wearing → Mask wearing & 0.19 & 0.04 & 0.11 & 0.26 & 0.00\\
+Republican & Mask wearing → Injunctive norms & 0.05 & 0.03 & -0.01 & 0.10 & 0.10\\
+Republican & Mask wearing → Descriptive norms & 0.07 & 0.03 & 0.02 & 0.12 & 0.01\\
+Republican & Injunctive norms → Mask wearing & 0.00 & 0.04 & -0.08 & 0.09 & 0.91\\
+Republican & Injunctive norms → Injunctive norms & 0.24 & 0.04 & 0.16 & 0.31 & 0.00\\
+Republican & Injunctive norms → Descriptive norms & 0.05 & 0.03 & -0.01 & 0.11 & 0.12\\
+Republican & Descriptive norms → Mask wearing & 0.14 & 0.05 & 0.05 & 0.24 & 0.00\\
+Republican & Descriptive norms → Injunctive norms & 0.11 & 0.04 & 0.04 & 0.19 & 0.00\\
+Republican & Descriptive norms → Descriptive norms & 0.32 & 0.04 & 0.24 & 0.39 & 0.00\\
+\bottomrule
+\end{longtable}
+
+\end{ThreePartTable}
+\end{center}
 

### 2023-07-06 R/functions.R

@@ -81,14 +81,94 @@ loadOwid <- function(fileOwid) {
     mutate(date = dmy(date))
 }
 
-# t-test: construct validity
-testConstructValidity <- function(d, type) {
-  # how much information do the descriptive items give about X type of norms?
-  y1 <- (d[,paste0(type, "Learning1.7")] + d[,paste0(type, "Learning2.7")]) / 2
-  # how much information do the injunctive items give about X type of norms?
-  y2 <- (d[,paste0(type, "Learning3.7")] + d[,paste0(type, "Learning4.7")]) / 2
-  # paired t-test
-  t.test(y1[,1], y2[,1], paired = TRUE)
+# construct validity
+fitConstructValidityModel <- function(d) {
+  # vector of items
+  items <- c(
+    "1.7" = "MaskEncouraged",
+    "2.7" = "MaskRespect",
+    "3.7" = "NeighborMask2",
+    "4.7" = "NeighborMask1"
+  )
+  # modify data
+  out <-
+    d %>%
+    mutate(id = 1:nrow(d)) %>%
+    dplyr::select(id, starts_with("DescriptiveLearning") | starts_with("InjunctiveLearning")) %>%
+    pivot_longer(cols = !id) %>%
+    mutate(
+      value = value,
+      Type = ifelse(str_starts(name, "Descriptive"), "Descriptive", "Injunctive"),
+      Item = as.character(items[str_sub(name, -3, -1)])
+    ) %>%
+    drop_na() %>%
+    # fit model
+    lmer(value ~ Type*Item + (1 | id), data = .) %>%
+    emmeans(., ~ Item | Type) %>%
+    pairs()
+  return(out)
+}
+
+# interaction frequency
+fitIntFreqModel <- function(d, pred = "") {
+  # modify data
+  out <-
+    d %>%
+    mutate(
+      # unique participant ids
+      id = 1:nrow(d),
+      # characters to numeric for analysis
+      Contacts24hrs.17 = as.numeric(Contacts24hrs.17),
+      Contacts24hrs.18 = as.numeric(Contacts24hrs.18),
[…]

### 2023-07-06 _targets.R

@@ -3,12 +3,12 @@ library(targets)
 library(tarchetypes)
 source("R/functions.R")
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("arm", "broom.mixed", "cowplot", "dagitty", "ggarchery",
-                            "ggdag", "ggeffects", "ggraph", "huxtable", "jtools", 
-                            "knitr","kableExtra", "lavaan", "lme4", "lmerTest", 
-                            "lubridate", "MuMIn", "papaja", "rnaturalearth", 
-                            "rnaturalearthdata", "scales", "semTools", "sf", 
-                            "zipcodeR"))
+tar_option_set(packages = c("arm", "broom.mixed", "cowplot", "dagitty", "emmeans",
+                            "ggarchery", "ggdag", "ggeffects", "ggraph", "huxtable",
+                            "jtools", "knitr","kableExtra", "lavaan", "lme4", 
+                            "lmerTest", "lubridate", "MuMIn", "ordinal", "papaja", 
+                            "psych", "rnaturalearth", "rnaturalearthdata", "scales",
+                            "semTools", "sf", "zipcodeR"))
 # workflow
 list(
   # files
@@ -19,9 +19,11 @@ list(
   tar_target(d, loadData(fileData, fileStates)),
   # load US covid counts from our world in data
   tar_target(owid, loadOwid(fileOwid)),
-  # t-tests: construct validity
-  tar_target(constructValDes, testConstructValidity(d, type = "Descriptive")),
-  tar_target(constructValInj, testConstructValidity(d, type = "Injunctive")),
+  # construct validity
+  tar_target(constructVal, fitConstructValidityModel(d)),
+  # interaction frequency
+  tar_target(intFreq24hrs, fitIntFreqModel(d, pred = "Contacts24hrs")),
+  tar_target(intFreq7days, fitIntFreqModel(d, pred = "Contacts7days")),
   # multilevel modelling
   # correlations between behaviour and norms
   tar_target(m1.1, fitCorBehNorm(d, predictor = "DesNorms")),
@@ -42,31 +44,52 @@ list(
   tar_target(pars2.3, getChangePointPars(m2.3, outcome = "Injunctive norms")),
[…]

### 2023-07-24 appendix.Rmd
update supp materials

@@
-\renewcommand{\appendixname}{\textbf{Supplementary Material}}
 \renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
@@
 \renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
-\pagenumbering{gobble}
 
-#
+# **Supplementary Material**
+
+\setcounter{page}{1}
+\small
+\centering
+\setlength\parindent{8mm}
+Descriptive norms caused increases in mask wearing during the COVID-19 pandemic \newline
+\hspace*{8mm} Samantha L. Heiman, Scott Claessens, Jessica D. Ayers, Diego Guevara Beltrán,\newline
+\hspace*{8mm} Andrew Van Horn, Edward R. Hirt, Athena Aktipis & Peter M. Todd \newline
+\normalsize
+\raggedright
+\setlength{\parindent}{40pt}
+\newpage
 
@@
 Finally, demographic variables such as gender, age, ethnicity, socioeconomic status, and political orientation (summarised as Demographics) are exogenous variables that are common causes of all other variables.
-Using the backdoor criterion[@Pearl1995], this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived descriptive norms on mask wearing.
 Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived injunctive norms on mask wearing.
@@
 ```
+
+\newpage
+\setlength\parindent{0pt}
+
+## Supplementary References
+
+Pearl, J. (1995). Causal diagrams for empirical research. _Biometrika_, _82_(4), 669-688.

### 2023-07-24 manuscript.tex
update supp materials

@@
 
+% \usepackage{etoolbox}
 \makeatletter
-\usepackage{etoolbox}
-\patchcmd{\maketitle}
+\patchcmd{\HyOrg@maketitle}
   {\section{\normalfont\normalsize\abstractname}}
@@
   {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\maketitle}
+\patchcmd{\HyOrg@maketitle}
   {\section{\protect\normalfont{\@title}}}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Pearl1995}{}}%
-\CSLLeftMargin{52. }%
-\CSLRightInline{Pearl, J. \href{https://doi.org/10.2307/2337329}{Causal diagrams for empirical research}. \emph{Biometrika} \textbf{82}, 669--688 (1995).}
-
 \end{CSLReferences}
@@
 \newpage
-
-\hypertarget{appendix-appendix}{%
-\appendix}
-
-\renewcommand{\appendixname}{\textbf{Supplementary Material}}
+\vspace*{60mm}
+
 \renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
@@
 \renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
-\pagenumbering{gobble}
-
-\hypertarget{section}{%
-\section{}\label{section}}
+
+\hypertarget{supplementary-material}{%
+\section{\texorpdfstring{\textbf{Supplementary Material}}{Supplementary Material}}\label{supplementary-material}}
+
+\setcounter{page}{1}
+\small
+\centering
+\setlength\parindent{8mm}
+
+Descriptive norms caused increases in mask wearing during the COVID-19 pandemic \newline
+\hspace*{8mm} Samantha L. Heiman, Scott Claessens, Jessica D. Ayers, Diego Guevara Beltrán,\newline
+\hspace*{8mm} Andrew Van Horn, Edward R. Hirt, Athena Aktipis \& Peter M. Todd \newline
+\normalsize
+
+\raggedright
+\setlength{\parindent}{40pt}
+\newpage
 
@@
 Finally, demographic variables such as gender, age, ethnicity, socioeconomic status, and political orientation (summarised as Demographics) are exogenous variables that are common causes of all other variables.
-Using the backdoor criterion\textsuperscript{52}, this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived descriptive norms on mask wearing.
+Using the backdoor criterion (Pearl, 1995), this causal model implies that it is necessary to control for perceptions of injunctive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived descriptive norms on mask wearing.
 Similarly, it is necessary to control for perceptions of descriptive norms, factual beliefs, personal normative beliefs, and all demographics to estimate the direct causal effect of perceived injunctive norms on mask wearing.}
@@
 
+\newpage
+\setlength\parindent{0pt}
+
+\hypertarget{supplementary-references}{%
+\subsection{Supplementary References}\label{supplementary-references}}
+
+Pearl, J. (1995). Causal diagrams for empirical research. \emph{Biometrika}, \emph{82}(4), 669-688.
+
 \end{document}
